### 2014-08-21 chapters/MetaTracts.tex
MetaTracts working copy

%\chapter {MetaTracts - Robust Extraction and Visualization of Fiber Bundles in Fiber Reinforced Composites}
%\label{metaTracts.ch}

\section{Introduction and Motivation}
\label{sec:intro}

% Modern industry is constantly driving their research towards novel lightweight components. For these high-tech applications, the industry unites behind the mantra of function orientation, high integration as well as cost and eco-efficiency. A promising type of material, showing high specific stiffness, high specific strength, and high corrosion resistance, allows to integrate today's demands of industry. Advanced composites and in particular carbon fiber reinforced polymers (CFRP) were successfully introduced in aeronautic and automotive applications and within the recent years their share is rapidly rising. To increase the usage, not only structural but also complex primary structures and  highly loaded components have been manufactured from CFRP. As for today,  besides commercial aerospace and automotive industry, CFRP is used in wide range of industries, e.g., space/satellite, marine, sporting goods, automotive, civil engineering or wind energy.

%REPLACED BY 
%Carbon fiber reinforced polymers (CFRP) are promising materials for future components as they allow integrating the continuously rising demands of various industrial  branches.

Carbon fiber reinforced polymers (CFRP) are bundles of individual carbon fibers
woven into sheets of carbon fiber cloth which are layered
and reinforced by an epoxy material.
To manufacture CFRP's, bundles of individual carbon fibers are woven into sheets of carbon fiber cloth which are layered and reinforced by an epoxy matrix. The alignment and weave pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. The resulting material shows high specific stiffness, high specific strength, and high corrosion resistance. 
Increasingly, complex primary structures and highly loaded components are being manufactured from CFRP.
Advanced composites and in particular carbon fiber reinforced polymers have thus successfully been introduced in aeronautic and automotive applications. Besides commercial aerospace and automotive industry, CFRP is also being used
in a wide range of applications such as aeronautical, marine and civil engineering, automotive and wind turbine design,
and sporting goods manufacturing.

The increasing share of fiber reinforced polymers also generates a strong demand for non-destructive testing (NDT) techniques~\cite{Red2012} of such materials and components. Ultrasonic testing is the most commonly used method for non-destructive testing of fiber reinforce polymers.
It provides a quick, cost-efficient overview, but has low resolution and is imprecise. Recently, industrial 3D X-Ray computed tomography (XCT) has been applied to non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, taken throughout a full rotation of the specimen.

In XCT, the specimen is  placed on the rotary table between X-ray source and detector and penetrated by an incident X-rays of the source.
The X-rays pass through the specimen and are attenuated by the material in the specimen.
The detector transfers the X-rays in its scintillator layer into visible light. The light is then recorded in a 2D projection image. 
The table is rotated by a small amount and the the process is repeated~\cite{heinzl-2008-thesis}.

In state of the art devices, XCT can reach voxel sizes below $500 nm$,
it generates high resolution XCT volume data for comprehensive and detailed analyses of the test specimens. However, there is still a trade-off between viewport and image resolution. The magnification reached within a cone beam XCT scan is determined by the specified distances
between source and specimen as well as source and detector. The magnification directly influences both resolution and viewport: higher resolutions decreases the viewport or how much of the structure can be imaged while lower resolutions allow for larger viewports.
The domain experts are increasingly interested in visualizing the internal geometric structures present in the  weaving pattern of the fiber bundles themselves instead of high resolution studies of individual fibers.

We focus on datasets with larger viewports and lower resolutions where the single fibers are barely visible. 
Figure~\ref{fig:data-char} provides an example of our dataset. It also shows an example of the recurring bundle pattern called an \textit{unit cell}.

 This work is thus motivated from the recent progress in two interrelated fields: On the one hand, fiber reinforced polymer (FRP) components has gained wide application in a variety of industries  because of FRP's superior material and physical properties in comparison to conventional materials \cite{Karpat2012}. On the other hand the use of industrial 3D X-Ray computed tomography (XCT) based non destructive testing of fiber reinforced components has also gained popularity~\cite{Schilling2005}.

% Fiber bundles determine to a large extent the properties in terms of specific stiffness, specific strength, or durability of the final component. However, when scanning reinforced polymers using modern XCT devices, there is still a trade-off between view port and image resolution. Due to the intrinsic concept of cone beam and fan beam XCT setups, the magnification reached within a scan is determined by the used distances between source and specimen as well as source and detector. Furthermore the magnification directly influences both resolution and viewport: Higher resolutions decrease the viewport or how much of the structure can be imaged while lower resolutions allow for larger view ports.
%\begin{figure}
%\begin{center}
%\includegraphics[width=0.3\textwidth]{imagesMetaTracts/crop-13-total}
%\caption{Volume rendering of a part of a dataset. Fig~\ref{fig:slice_data:a} and ~\ref{fig:slice_data:b} show image slices along the blue and red plane respectively.  }\label{fig:orig_data}
%\end{center}
%\end{figure}
%!!!!! TODO REFINE DTI PARAGRAPH !!!!!
%Diffusion Tensor Imaging tractography clustering and visualization.
%The topic of DTI visualization has seen an extensive amount of investigation by researchers.
%Advantage of two phase clustering 
%What is the  motivation for fiber-lets
%Seed point selection 
%distance between tracks 
%how well do we do in noise , may be data set 1

%!!!!! I have replaced this with the following para !!!!!
%In this work, we present a framework which interprets and advances these techniques in order to extract and visualize fiber bundles in industrial 3D X-Ray computed tomography data of woven fiber reinforced composites. The main goal of this work is to expand the state of the art in nondestructive testing of composite structures towards lower resolutions and thus larger areas of woven fabrics.

%While the importance of fiber bundles in determining the final properties of the components have  become evident,  the tools for visualization of the same have not developed at the same pace.

While fiber bundles are now understood as increasingly important in determining component properties,
the tools for visualization of the internal structure have not developed at the same pace.
To the best of our knowledge, there is no current work that can resolve simple queries such as:
\begin{itemize}
\item{ How to visualize the geometric structure of a  particular bundle?}
\item{ Which is the largest bundle?}
\item{ Which bundles are in a particular orientation?}
\item{ How to visualize the interaction between a particular pair (crossovers/braiding) or a unit cell?}
\end{itemize}

Extracting the answer to the above queries from the volume rendering of datasets, or from looking at  particular (2D) slices of the data is non-trivial even for experts. We detail and evaluate a method which uses visualization to quickly gain insight into our data. We demarcate the above queries into two, "geometric structure" and "spatial context". Geometric structure refers to the shape, size, orientation of a bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other.

We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites.
The main goal of this work is to expand the state of the art in non destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.

The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  Section \ref{sec:char_data} describes the   characteristics of our data. Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provides details of our approach. Section \ref{sec:results} details experiments and Section~\ref{sec:param_choices} explains our  parameter choices. We end with conclusions.

\section {Related work}
\label{sec:prev_work}
%In order to analyze fiber bundles, whose individual fibers are barely visible, we studied a branch of visualization which seems to be completely off-topic at first sight: Diffusion Tensor Imaging (DTI). Tractography methods to reconstruct fibrous structures such as neural fibers in the human brain from magnetic resonance (MR) images has seen extensive research in recent years. Visualization techniques to reduce the geometric complexity and cluttering caused by dense fiber models of single fibers and emphasizing higher level structures called the fiber bundles have also gained popularity in recent years. In general these bundles are related anatomically or spatially.
%We broadly divide the related work on DTI  into two parts related to our method: fiber tracking and fiber clustering. Furthermore, methods analyzing second order local structure are considered as related work. Finally, we review the field of visual analysis of fiber reinforced composites in this section.

Diffusion Tensor Imaging (DTI) refers to fiber tractography using diffusion  magnetic resonance (MR) images. It is a promising method for reconstructing the pathways of white matter in the human brain.
A variety of algorithms have been proposed aiming to generate fiber-tract trajectories. In general these reconstructions are then clustered in to bundles which are expected to be related anatomically or spatially. 
We broadly divide the related work on DTI into two parts which are of immediate relevance to our proposed solution: fiber tracking and fiber clustering. Furthermore methods analyzing the second order local structure are also discussed. Finally, we review the current state of the art in the visual analysis of fiber reinforced composites.

\subsection {Fiber Tracking in Diffusion Tensor Imaging}
\label{subsec:fiberEx}
Diffusion Magnetic Resonance Imagining has gained popularity in recent years as a means of visualizing the organization of white fibers in human brain.
In detecting fiber tracts as shown by Basser et al.~\cite{Basser2000},~\cite{Basser2002} and Mori et al.~\cite{Mori2002},~\cite{Mori1999}, the key idea is based on the assumption that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel. In these techniques, continuous tracts are created by propagating a virtual particle from a starting point until they reach some termination criterion. This is usually done by solving a second-order Runge-Kutta integration. The assumption is that the principal diffusion direction at each discrete location can be interpolated to from a continuous velocity field. One of the short comings of the process is that because of the local decision-making process, these methods perform poorly in noisy regions and often generate small fibers. Basser et al.~\cite{Basser2000},~\cite{Basser2002} proposed that white matter tracts could be represented as 3D curves in space. They showed that numerical method could be used to follow the fiber tracts and generate tracts in human brain data.

Mori et al.~\cite{Mori2002} showed that tract reconstruction techniques could be broadly divided into line-propagation or energy minimization techniques. In line propagation approaches, trajectories are computed based on local neighborhoods. While, energy minimization approaches find the most favorable trajectory connecting two given endpoints.

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
%!!!!! TODO REFINE FIBER CLUSTERING !!!!!

%The next step is to identify  anatomically relevant fiber bundles. This is achieved through fiber clustering.
The general assumption of fiber clustering in diffusion tensor imaging is that a similarity measure based on factors such as proximity between fibers can be used to interpret a collection of fiber tracts as bundles.  Extensive research has been done on automatic DTI fiber clustering methods (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}). We limit ourselves on advancement of interest to our current work.
 %Zhang et al.~\cite{Zhang2008}, O'Donnell et al.~\cite{westinMEDIA02}. 
 The clustering approach divides the problem into two parts. The first part deals with finding a suitable proximity measure and the second deals with the choice of the clustering method. Pairwise proximity measures include, endpoint distances~\cite{Brun2003},  mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of the above distance so that fibers that are close for certain distance and then  diverge are clustered separately. Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors. The choice for the clustering algorithm can be broadly divided into those using hierarchical clustering (see ~\cite{Moberts2005},~\cite{Zhang2008}) and those using spectral clustering approach (see~\cite{jonasson2005},~\cite{ODonnell2007},~\cite{Brun2004}).

Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps proposed by Belkin and Niyogi~\cite{Belkin01} can be applied to the problem of organizing fiber trace data. The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph. Each node represents a data point and the edges connect neighboring data points. An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure. In the case of fiber bundles, the individual points are fiber traces. In the ideal case fiber traces which belong to the same bundle must remain ``close" to each other in the lower dimensional space. ~\cite{westinMEDIA02} also uses spectral clustering on a Hausdorff distance measure defined as the maximum of pointwise minimum distances between two fibers. Jonasson et al.~\cite{jonasson2005} runs k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.

%The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}.

 The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}. 
 These approaches build on the assumption that proximity measure that compare DTI fiber trajectories can also represent anatomical relationship. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are joined. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results.

The difficulties in hierarchical clustering includes; first, computing all pairs distances for tracts to generate the distance matrix (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times). Second, choosing the ``correct'' distance measure to compare tracts. Thirdly, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

\subsection{Second order local structure}
%!!!!! TODO REFINE SECTION !!!!!

Unlike DTI, we do not have diffusion tensor data. Instead, we have a scalar volume with tubular structures embedded in them. Analyzing curvilinear structures in volumetric images has been utilized for a variety of purposes including center line extraction~\cite{Bouix2005} and vascular image enhancement Frangi et al.~\cite{Frangi1998}, Sato et al.~\cite{Sato1997}. Frangi et al.~\cite{Frangi1998} introduced a method based on studying the eigenvalues of the local second order structure namely the Hessian matrix, specifically for the purposes of developing vessel enhancement filters.

\subsection {Visual Analysis of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores. Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. This approach allows analyzing fiber orientations based on direction transfer functions. Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites. They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations. Reh et al.~\cite{PMI_AR_2012} introduced an approach to explore pores of carbon fiber reinforced composites. Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for  interactive exploration of fibers in reinforced polymers.

Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes has received extensive research and recently the application of similar techniques to fiber tracing in composite materials has gained momentum. To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers, which is the main scope of this work.

%To the best of our knowledge there is no approach focusing extraction of fiber bundles structures of low resolution XCT  of composite fibers, which is the main scope of this work.

%\begin{figure}
%  \begin{center}
%    \includegraphics[width=0.25\textwidth,clip=true, trim=2cm 2cm 0cm 2cm]{imagesMetaTracts/crop-13-total}
%  \end{center}
% \caption{Volume rendering of a part of a dataset. Fig.~\ref{fig:slice_data:a} and ~\ref{fig:slice_data:b} show image slices along the blue and red plane respectively. }
% \label{fig:orig_data}
%\end{figure}
%\begin{figure}[h]
%  \centering
%              \subfloat [][A single slice along blue plane in fig~\ref{fig:orig_data}.]
%                        {\includegraphics[width=0.22\textwidth]{imagesMetaTracts/crop-13-1c-anno.png}\label{fig:slice_data:a}}\quad 
%                        \subfloat [][A single slice along the red plane in fig~\ref{fig:orig_data}]
%                        {\includegraphics[width=0.22\textwidth]{{imagesMetaTracts/crop-13-1b-anno}}\label{fig:slice_data:b}}
%  \caption{Single slices along different orientations of the data set. In Fig.~\ref{fig:slice_data:b}, the red annotation demarcates the boundary between two different bundles. The green squares show that the cross sections of the bundles also widely vary.}
%  \label{fig:slice_data}
%\end{figure}

\begin{figure}
\includegraphics[width=0.5\textwidth,clip=true, trim= 0cm 0cm 0cm 0cm]{imagesMT2014/MT_data_4_2.png}
\caption{Data Characteristics. A: Shows a rendering of the data. B: Shows 2D slices along the X (Right) and Z (Left) axis. C: Shows a zoom of the yellow square, multiple bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}

Figure \ref{fig:data-char} shows one of our data sets with woven fiber bundles. The dataset is 450x300x500 voxels in size and clearly shows the weaving structure of the composite unit cell (a regional subset of the recurring fiber bundle weaving pattern). The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weave pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}A shows a volume rendering of the dataset. Figure~\ref{fig:data-char}B shows 2D slices along the X(right) and Z(left) axis. Figure~\ref{fig:data-char}C shows the zoomed in version of the yellow square. At isotropic resolution of 2 micrometers, the individual carbon fibers are hard to resolve from the 2D slices. In Figure~\ref{fig:data-char}C contains multiple bundles going in opposite directions. The separation between two bundles is barely discriminable.
Carbon fibers themselves do not differ as much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. 
The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}A shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the bundles cross each other in different orientations. 
In general the number of orientations is limited by the weaving pattern. In all our data sets the number of separate orientations was two. Nevertheless, the weaving pattern may cause the individual bundles to be curved. 
In consequence of the weaving pattern of the fiber bundles, fibers may be adjacent in euclidean distance but belong to different bundles.

We make the following assumptions on our data.
\begin{itemize}
\item The structure embedded in the data contains fiber bundles  of indiscernible fibers.
\item (Local orientation)Each point within a fiber  bundle has a local orientation which is parallel to the bundle.
\item The local orientation changes gradually inside the bundle.
\item Local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

%\begin{figure}[b]
%\centering
%\includegraphics[width=\linewidth]{imagesMetaTracts/crop-13-3.eps}
%\caption{Volume rendering of the scalar volume of one of our data set. Cut-out A shows a slice along the X axis, cut-out B shows a slice along the Z axis.}\label{fig:orig_data}
%\end{figure}

\section {Extracting MetaTracts }
\label{sec:approach}

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{imagesMT2014/algo_A.png}
\caption{Flow chart of our approach}
\label{fig:flowchart}
\end{figure}
 Figure \ref{fig:flowchart} shows our pipeline. The approach consists of multiple steps which we divide into two broad phases, extracting the MetaTracts  (this section) and  finding plausible fiber bundles (Section~\ref{subsec:fiber-bundles}).
%\subsection {Extracting MetaTracts}
%\label{subsec:ExFiberCyl}
%We explain the different steps of MetaTracts extraction.
\subsection {Reliable Hessians}
\label{subsec:rh}

The input to the first stage of our pipeline, is the original scalar dataset  in a uniform lattice grid in $\mathbb{R}^3$. As output, each grid location becomes associated with a vector representing the local orientation and a real value [0,1] which represents a measure of the reliability of the local orientation. 
We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel. The principal directions, in which the local second order structure of the image can be decomposed are given by the eigen decomposition of the Hessian matrix which describes the local curvature.

The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute a similar metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigen values of the Hessian matrix. Specifically, a  pixel belonging to vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$).  The following measures are defined in ~\cite{Frangi1998}.  
\begin{equation}\label{RA}
\mathcal{R_{A}}=\frac{\textrm{Largest  Cross Section}\big/ \pi}{{\textrm{Largest Axis Semi-length}}^{2}}=\frac{|\lambda_{2}|}{|\lambda_{3}|}
\end{equation}
\begin{equation}\label{RB}
\mathcal{R_{B}}=\frac{\textrm{Volume}\big/ (4\pi \big/ 3)}{{(\textrm{Largest Cross Section Area}\big/ \pi)}^{\frac{3}{2}}}=\frac{|\lambda_{1}|}{\sqrt{|\lambda_{2}\lambda_{3}|}}
\end{equation}
In equation~\ref{RB}, $\mathcal{R_{B}}$ provides a measure of deviation from blob like structure while in equation~\ref{RA}, $\mathcal{R_{A}}$ distinguishes between plate-like and line-like structure. Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors. Thus we compute a measure, ``equal" to the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
$$
R_{H} = \left\{ \begin{array}{ccc}
 0 & \mbox{ if $\lambda_{2}>0$ or $\lambda_{3}>0$} \\
  (1-e^{\frac{\mathcal{-R_{A}}^{2}}{2\alpha^{2}}})
  (e^{\frac{\mathcal{-R_{B}}^{2}}{2\beta^{2}}}) (1-e^{\frac{-s^2}{2c^2}}) &\mbox{ otherwise}
       \end{array} \right.
$$

Variable $s$ is the Frobenius norm of the Hessian matrix. The value  of $(1-e^{\frac{-s^2}{2c^2}})$ will be low in regions with no structure.

The utility of the vesselness is a little different in our framework than Frangi et al.~\cite{Frangi1998}.
First, in biology vesselness is computed for different scales because the vessels can be of different sizes. In our case, usually the width of individual fibers are known a priori.
Second, and more importantly, we do not have clear tubular structures embedded in a dark contrast matrix such as in blood vessels.
Instead, we are trying to associate each grid location  with a probable orientation based on its local second order structure. The $R_{H}$ is interpreted as a measure reliability of the local orientation.

Grid locations where the $R_{H}$ is above a certain threshold are marked as regions with reliable orientations.
We use 0.3 as the cutoff for all our test cases. (See Section~\ref{sec:param_choices} on the choice of 0.3)

%\begin{figure}[h]
%	\centering
% \subfloat[]{ \includegraphics[width=0.2\textwidth,clip=true,trim=0cm 1cm 0cm 1cm]{imagesMetaTracts/crop-13-tracks-hess.eps}\label{fig:r_h:a}}
%   \subfloat[]{\includegraphics[width=0.25\textwidth,scale=0.2]{imagesMetaTracts/rh-1c-anno.png}\label{fig:r_h:b}}
%  \caption{Reliable Hessians: (a) Shows $R_{H}$ colored according to the absolute value of the normalized major eigenvector (local orientation) mapped to RGB color space. Fig~\ref{fig:r_h:b} shows part of a single slice where two bundles along Z(blue)axis and X(red)axis meet. Spurious specks shows that the local orientation is still noisy. Black denotes area below the $R_H$ threshold.}\label{fig:r_h}
%\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=0.4\textwidth,scale=0.1,clip=true,trim=0cm 1cm 0cm 1cm]{imagesMetaTracts/crop-13-hess-2}
% \caption{Reliable Hessians: Left, shows $R_{H}$ colored according to the absolute value of the normalized major eigenvector (local orientation) mapped to RGB color space. Right, shows a single slice along the X axis. Spurious specks shows that the local orientation is still noisy. Black denotes area below the $R_H$ threshold.}\label{fig:r_h}
%\end{figure}
%

\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
\caption{Reliable Hessians. A: Colored according to Orientation vector mapped to RGB. B: Slice along Z axis. C: Slice along X axis. D: close up of the region marked in yellow in C.}
\label{fig:reliable_hessian}
\end{figure}
 In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the threshold. The unit vector representing the principal direction has been mapped to RGB color space. Figure~\ref{fig:reliable_hessian}A shows the entire data set. Regions where the principal direction is parallel to the X axis are red in color and those parallel to the Z axis are blue. Figure~\ref{fig:reliable_hessian}B shows 2D slices along the Z and X axis respectively. Figure~\ref{fig:reliable_hessian}C shows a close up (region similar to Figure~\ref{fig:data-char}C). Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold. The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.

% We note some intrinsic differences between the DTI and our XCT data. The Hessian computation works best when the tubular structures are well separated from the surrounding, unfortunately, this is not the case for our data (Fig.~\ref{fig:r_h}). Thus the local orientation at each voxel for XCT data is inherently more noisy (Fig.~\ref{fig:r_h} Right) than those computed from eigen analysis of diffusion tensors in DTI datasets.

 We note some intrinsic differences between the DTI and our XCT data. Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}. The principal direction based on the hessian matrix works best when the tubular structures are well separated from the surrounding, this is not the case for our data. The local orientaion at each voxel is inherently more noisy. Thus techniques popular in DTI cannot be directly applied in XCT data.

\subsection {MetaTracts properties and description} 
\label{subsec:mtprop}
%\begin{figure*}
%  \centering
%  \subfloat[Meta tract generation in 2D]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1b.eps}}
%  \subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified examples of meta tract generation process and orientation based similarity measure}
%  \label{fig:algo}
%\end{figure*}
%\begin{figure*}
%  \centering
%  \subfloat[]{\includegraphics[width=0.2\textwidth]{imagesMetaTracts/algo-1d}}
%  %\subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified example of meta tract generation process.}
%  \label{fig:algo}
%\end{figure*}
The input to this step is a grid in $\mathbb{R}^3$ where each grid point is associated with a normalized local orientation and a corresponding $R_H$ value.
Traditional integral curve based techniques cannot be used to extract fiber bundle traces from the reliable Hessians because of the spurious nature of the Hessian based local orientation.  Thus instead of building fiber traces, we wish to find an abstract representation of the fibers. We  build on the key assumptions on the data, namely `local orientation' and `connectivity' while taking into account the noise and lack of resolution. We do this by interpreting the underlying geometric structure of the fibers as a set of polylines(cylinders).

 MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of  cylindrical tubes in $\mathbb{R}^3$ traversing the bundles embedded in the input data. MetaTracts share the following characteristics:

\begin{enumerate}
\item{Each MetaTract is associated with a continuous set of cylinders.}
\item{Each MetaTract is associated with a start point which is a grid vertex.}
\item{Each  individual cylinder in a MetaTract has a constant length,  constant radius and a start point (which is also a grid vertex).}
\item{Each cylinder in a MetaTract (except for the first one)is connected to the previous cylinder at the start point.}
\item{Each cylinder is locally parallel to the local orientation at it's start point.}
\end{enumerate}

% $N_{p}$ (Fig.~\ref{fig:algo} provides an overview). Let $C_{p}$ be the grid point associated with the current cylinder. We insert a subset of the grid points $G_{q}$ that lie within the cylinder into a priority queue. The priority is defined by equation \ref{eqn:algo_1}. In Equation \ref{eqn:algo_1}, $angle$ refers to the angle between the local orientation at the current point $C_{p}$ and the grid point $G_{q}$. $PerpDist$ refers to the perpendicular distance between the $G_{q}$ and $C_{p}$ projected along the direction of the local orientation $N_{p}$. We only insert a grid point $G_{q}$ in the priority queue if the $angle$ is less than 20 degrees and the $PerpDist$ is at least 0.5. This ensures that points with obvious spurious local orientation or points which are very close to  $C_{p}$ are not put in to the queue. This simple check ensures our memory requirements are not excessive.

\subsection {MetaTracts generation}

In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles. We explain the process of MetaTracts generation in $\mathbb{R}^2$. The procedure trivially extends to $\mathbb{R}^3$. The reliable Hessian procedure  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1]. In Figure \ref{fig:algo} all regions which have reliable hessian less than the threshold have been marked in green. Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the blue arrow (Figure~\ref{fig:algo}C). We generate a rectangle of length L and radius R, the blue rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}).

\begin{figure}[h]
  \begin{center}
  \includegraphics[ width=0.9\linewidth]{imagesMT2014/MT_abstract_A2} %algo-1d}
  \end{center}
 \caption{MetaTracts generation. B. MetaTract generation, region with $R_H$ less than threshold. A: close up of marked region in A. C: Parts of a single MetaTract.}
  \label{fig:algo}

\end{figure}

\textit{Candidate vertices}: The set of vertices which are in the blue region but not overlapped by green are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices candidate vertices. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
The order of the candidate vertices is based on the following characteristics.
\begin{itemize}
\item \textit{Orientation Similarity}: We want the orientation of the start points for the consecutive cylinders to have a similar orientations ($N_p$).

\item \textit{Large distance}: We want the MetaTracts to traverse the data using as few cylinders as possible. Thus, we want the distance between $C_p$ and the start point for the next cylinder to be as large as possible. We measure distance of a grid vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$. For example the Euclidean distance between the blue and the orange vertices projected onto $N_P$ gives distance (Figure~\ref{fig:algo}B). We refer to this perpendicular projection distance as \textit{dist}. 
\end{itemize}
% \begin{equation}\label{eqn:algo_1}
% d=\frac{1}{3}e^{(\frac{-angle^{2}}{\alpha^{2}})}+\frac{2}{3}(1-e^{\frac{-{PerpDist}^{2}}{\beta^{2}}})
% \end{equation}
We define a priority  for each of the candidate vertices based on the above factors.
For each cylinder in a MetaTracts, we put the \textit{candidate vertices} of it in a priority queue based on Equation \ref{eqn:algo_1}.
\begin{equation}
Priority = \gamma_1 e^{(-angle^2 / \alpha^2)} + \gamma_2e^{(-dist^2 / \beta^2)}
\label{eqn:algo_1}
\end{equation}
 $\gamma_1$, $\gamma_2$ are the  weights ($\mathbb{R}_{\ge 0}$)  which decide how the priority depends on the affine combination of the two factors. For all our cases we use ($\gamma_1$ = $1 / 3 $ and $\gamma_2$ = $2 / 3$). In general we suggest $\gamma_1 +\gamma_2 = 1 $ and $\gamma_1 \leq \gamma_2$). At each iteration we pick the top element in the priority queue generate the corresponding cylinder and repeat the steps. Essentially, eqn~\ref{eqn:algo_1}, selects a grid point which is the furthest from the current point and going in a similar direction. This gives it the advantage of tackling noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation (for orientation) in an already noisy environment.

 In this particular example Figure~\ref{fig:algo}B we select the orange grid vertex next and repeat the process. We do not select the purple vertex (Figure~\ref{fig:algo}C) because it is not a candidate vertex. If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices inside the cylinder and will be of short length. Short tracts are then removed using a threshold on MetaTracts  length. The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}(a). The MetaTracts are colored with the mean orientation direction mapped to the RGB space. That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.

% \begin{figure}[h]
% \centering
% 	\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.3\textwidth]{imagesMT2014/MT_crop16_MT.png}
% 	\caption{MetaTracts, colored according to mean orientation mapped to RGB space}
% \label{fig:meta-tract}
% \end{figure}

%\begin{figure}
%  \centering
%  {\includegraphics[width=0.2\textwidth]{imagesMetaTracts/crop-13-tracks-1.eps}}
%  \caption{MetaTracts generated from one of our data sets. Each MetaTracts consists of cylinders. The mean local orientation, computed as the mean of the local orientations at the start point of each cylinder composing a particular MetaTract is mapped to the RGB space }\label{fig:meta-tracts}
%\end{figure}
%

%\begin{figure*}[htp]
%  \centering
%  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/rplot-crop13.eps}}
% % \subfloat[]{\includegraphics[width=0.28\linewidth]{imagesMetaTracts/crop-13-tracks-clus-b.eps}}
%    \subfloat[]{\includegraphics[width=0.22\linewidth]{imagesMetaTracts/crop-13-tracks-clus-b-anno.png}}
%  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/crop-13-tracks-clus-a.eps}}
%  \caption{Results from orientation based and hierarchical clustering. (a) shows the result of the K-means clustering with the data points projected to the top three Eigenvectors as the major axes, (b) shows the MetaTracts colored according to clustering results in (a), the corresponding clusters divide the MetaTracts according to their major orientation. (c) shows the result of hierarchical clustering of the results in (b), each orientation cluster (in this case 2) was divided into 7 classes.}\label{fig:clustering}
%\end{figure*}

%\begin{figure}
%  \centering
%  \subfloat[]{\includegraphics[width=0.22\textwidth,clip=true,trim=1cm 0cm 1cm 2cm]{imagesMetaTracts/rplot-crop13.eps}}
% % \subfloat[]{\includegraphics[width=0.28\linewidth]{imagesMetaTracts/crop-13-tracks-clus-b.eps}}
%    \subfloat[]{\includegraphics[width=0.3\linewidth,,clip=true,trim=1cm 0cm 2cm 0cm]{imagesMetaTracts/crop-13-tracks-clus-b-anno.png}}
%  \subfloat[]{\includegraphics[width=0.4\linewidth]{imagesMetaTracts/crop-13-tracks-clus-a.eps}}
%  \caption{Results from orientation based and hierarchical clustering. (a) The result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes. (b) MetaTracts colored according to clustering results in (a). Corresponding clusters dividing the MetaTracts according to their two major orientations. (c) The result of hierarchical clustering of the results in (b), The two orientation clusters are divided into seven classes.}\label{fig:clustering}
%\end{figure}

%\subsection {Seeding}

%We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.
%For our result we add a constraint. Only those grid vertices which have reliable hessians (Section~\ref{subsec:rh}). This restriction limits MetaTracts to  start from regions which have some underlying geometric structure. Regions that do not have reliable Hessians have poor local orientation and will not generate long MetaTracts. We also allow the user to selectively seed particular regions more densely.

\begin{figure}[h]
\centering
	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/MT_crop16_MT.png}}
	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/length_distribution.PNG}}
	\caption{(a) MetaTracts colored according to mean orientations. (b) length distribution of individual MetaTracts for a particular bundle.}
	\label{fig:length_distribution}
\end{figure}

\section {Fiber Bundle Generation}
\label{subsec:fiber-bundles}
The output of the above step is a set of MetaTracts. In this step we focus on clustering the MetaTracts into plausible bundles. We use hierarchical clustering which has enjoyed considerable attention and success in DTI field.

However, the MetaTracts produced as described in Section 4 often do not extend the full length of the fiber bundle.  Thus, MetaTracts
in a given fiber bundle may have very different lengths and may only partially overlap (figure~\ref{fig:length_distribution}(b)).

Moreover, Carbon fibers have both orientation and proximity information which can be used for clustering. Instead of creating some formula
which artificially combined the two measures,  we first cluster based on orientation and then further subdivide each orientation cluster using proximity information. We found that different clustering methods were preferable
for the different measures. We use dimension reduction followed by K-means clustering
for orientation clustering while we use hierarchical clustering to further subdivide each cluster
based on proximity measures. (See Section~\ref{sec:results} for experimental results.) Breaking clustering into two steps also helps us and potential users
to better understand the results of the clustering.

%However, in approaches developed for DTI fibers such as proximity based distance measures and clustering, the short comings of the fiber tracking techniques are generally not discussed and it is assumed that fibers in each bundle are of similar lengths and which traverse the data end to end. This is not the case for noisy data such as ours.

%In Figure~\ref{fig:hclust_issue_a} proximity based distance, for example mean of the minimum distances between points on MetaTracts might falsely find the red MetaTract being closer to the blue, than the orange MetaTract. 
%We also note from Figure~\ref{fig:meta-tract} and from domain knowledge that all our fibers are well separated by orientation. Thus instead of directly applying hierarchical clustering to our data,
%we first cluster based on orientation using a simple measure.

%We use domain knowledge to divide the clustering into two intuitive parts:because of the woven nature of fiber bundles in our application domain we first cluster in terms of orientation. This has the advantage of being able to separate tracts which might be close in an Euclidean space but traveling in different direction. We introduce a simple orientation based similarity measure (sec~\ref{subsec:ori-sim-mes}). Finally, for each of these clusters we re-cluster in terms of distances measured in a Euclidean space (sec~\ref{subsec:dist_clustering}).

\subsection {Orientation bundles}
%We use the inherent features expressed as weaving pattern of our data to pre-cluster the metaTracts into the number of classes based on the major directions of the pattern.

%It is important to note that this step broadly divides the MetaTracts into classes based on the major orientations and not individual fiber bundles. 
This step broadly divides the MetaTracts into classes based on the major orientations.
In order to cluster MetaTracts going in same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01} and later used in DTI fiber coloring~\cite{Brun2003} and fiber clustering Brun et al.~\cite{Brun2004}. The key notion is to find a suitable similarity measure to define the weights of an undirected weighted graph. The nodes being the data points and the edges being the weights. Then solving an eigenvalue problem which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
This lower dimensional space can then be clustered more efficiently than the higher dimensional space. 
Let G be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem:
\begin{equation}\label{equn:eigenMaps}
L\textbf{f}=\lambda D\textbf{f}
\end{equation}
where D is the diagonal weight matrix and L is the Laplacian matrix. The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an m-dimensional space. K-means clustering can then be used to cluster in this lower dimensional space.

\subsubsection{Orientation based similarity measure}
\label{subsec:ori-sim-mes}
Brun et al.~\cite{Brun2003} based their measure on the simple assumption that two traces with similar endpoints should 
be considered similar. Brun et al. ~\cite{Brun2004} based their measure on a 9-D fiber descriptors.
In our case, each MetaTract is a data point. We introduce a simple \textit{orientation based similarity measure}. There is no spatial information involved, just a partition of the MetaTracts based on orientation which is already inherent in the data.
Following this, given a pair of  MetaTracts, we define the edge weights between two tracts as the cosine of the maximum angle between the local orientations ($N_P$) of all pairs of start points ($C_P$) between the two MetaTracts.
The edge weights give a distance matrix representing the distance between each pair of points.
Using the Belkin and Niyogi algorithm we "embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix.

%August 13 2014
%For all our tests the fibers were then embedded in a $\eta$ low dimensional space. The number of dimensions in the high dimensional space is equal to the number of datapoints or the number of MetaTracts in our case.
%For all our test cases we set $\eta$ to be five (see sec~\ref{sec:param_choices} for parameter choices for $\eta$.).

  \begin{figure}
  \centering
  	%\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}~
  	%	\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}
%  	\captionsetup[subfigure]{labelformat=empty}
%  	\subfloat[1]{\includegraphics[width=0.2\linewidth]{imagesMT2014/orientation_clustering_A.png}}
%  	\subfloat[2]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_1.PNG}}
% 	\subfloat[3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_2.PNG}}
% 	\subfloat[4]{\includegraphics[width=0.2\linewidth]{imagesMT2014/curveD.PNG}}
%  	
 \includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/Capture4.PNG}
  	\caption{(B)Shows the result of K means clustering with data points projected to the top three eigenvectors as major axes. (A,C) show the MetaTracts colored according to the clustering results. (The white to black tracts are only to show context.)  }
  \label{fig:orientation_clustering}
  \end{figure}

   We used conventional K-means clustering to cluster the lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
For our test case there are two major directions of the fiber bundles. So $K$ was set to 2. It is important to note that due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern. Figure~\ref{fig:orientation_clustering}B shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes. As is expected there is a clear distinction based on fiber orientation. Figure~\ref{fig:orientation_clustering}(A,C) shows the MetaTracts colored according to clustering results.
Orientation clustering uses only orientation information. To subdivide the oriented clusters into fiber bundles,
we need to include information about the geometric proximity between MetaTracts.

%\subsection {Distance based clustering}
%\label{subsec:dist_clustering}
%We now apply "single-linkage" hierarchical clustering~\cite{Moberts2005} to the results of the orientation clustering results.
%If we assume each fiber is represented as a set of points($C_P$) then the distance between two MetaTracts is computed as the minimum of the directed Hausdorff (maximum of the minimum euclidean) distances between points on the MetaTracts. Hierarchical clustering has one parameter, which is the number of clusters (h). Single linkage clustering inspite of it's advantages and  wide spread use, though suffers acutely in the presence of outliers.
%
%\begin{figure}[h]
%  \centering
%  	\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.3\textwidth]{imagesMT2014/clustering_probs_B.png}
%  	\caption{test}
%  \label{fig:cluster_prob_B}
%  \end{figure}
%For example Figure~\ref{fig:cluster_prob_B}B shows two main bundles (blue, orange), one outlier which has close proximity to both bundles (green) and another outlier far away from both the bundles (purple).
%Figure~\ref{fig:cluster_prob_B}A shows some sample MetaTracts from the bundles as points in a euclidean space.
%Setting h to 2 and performing single linkage hierarchical clustering would give a wrong result (Figure~\ref{fig:cluster_prob_B}C) It would create the outlier as a separate class instead of breaking the larger group. As a robust solution to the problem we perform hierarchical clustering iteratively. At each step removing clusters which have low cardinality(this is a common practice). However, we also remove elements from each cluster which have length lower than the median length of the fibers in the bundle. After a few iterations a stable state is reached where the outliers are removed and the clustering into bundles is robust(Figure~\ref{fig:cluster_prob_B}D).
%\begin{figure}[h]
%  \centering
%  	\includegraphics[ trim = 1cm 1cm 1cm 1cm, clip=true, width=0.3\textwidth]{imagesMT2014/final_clustering_A.png}
%  	\caption{Results of MetaTracts clustering. A: shows the X orientation cluster, clustered into 10 clusters. B: shows the Z orientation cluster, clustered into 10 clusters. Clusters with cardinality less than 10 are not shown. C: shows a close of the curved bundle extracted from the data. D: shows the original data for comparison.}
%  \label{fig:cluster_final_A}
%  \end{figure}
%  Figure~\ref{fig:cluster_final_A}(A,B) show the results of clustering each orientation bundle into ten clusters. Clusters with cardinality less than ten are not shown. Figure~\ref{fig:cluster_final_A}C shows the closeup of the curved bundle and Figure~\ref{fig:cluster_final_A}D shows the original data for reference. Single linkage clustering is very robust to the over-estimation of "h". A very large value of "h" would still keep the main clusters intact and create smaller  clusters with low cardinality which qualitatively does not degrade the results at all.

\subsection{Distance based clustering}
 \label{subsec:dist_clustering}
 Each MetaTract is represented as a set of points ($C_P$). 
 Formally, the directed Hausdorff distance from point set P to point set Q is defined as 
 $H_{dir}(P,Q) = max_{p \in P} min_{q \in Q} d(p,q).$ The Hausdorff distance is defined as $H(P,Q) = max(H_{dir}(P,Q),H_{dir}(Q,P))$.

 The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$
 but the directed Hausdorff is not.
 Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially cover the bundle (figure~\ref{fig:length_distribution}(b)).
 If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$,
 then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
 Thus, $H(P,Q)$ will be large, even though $P$ and $Q$ are in the same
 fiber bundle.

 Instead of using the Hausdorff distance, $\max(H_{dir}(P,Q),H_{dir}(Q,P)$, we use $\min(H_{dir}(P,Q),H_{dir}(Q,P))$. If $P$ covers only part of the fiber bundle covered by $Q$, then $\min(H_{dir}(P,Q),H_{dir}(Q,P))$ is very small.
 Note that if $P$ and $Q$ overlap but do not cover the same parts of the fiber bundle, then $H_{dir}(P,Q)$ and $H_{dir}(Q,P)$ and $\min(H_{dir}(P,Q),H_{dir}(Q,P))$ will be large.

 The directed Hausdorff distance is very sensitive to outliers in the data.
 However, because MetaTracts are constructed using cylinders with similar orientations, they are not plagued by outliers.
 To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
 Hierarchical clustering has a single parameter $h$, the desired number of clusters.
 In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
 Clustered are merged until there are only $h$ clusters left.
 Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
 Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.

 We found that single linkage hierarchical clustering had two major drawbacks.
 First, the clustering would produce some small clusters of just a few MetaTracts.
 These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
 Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same
 fiber bundle.
 A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles,
 causing them to be clustered into a single bundle.

 To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
 We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.

 To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.

We repeated both until a steady state of clusters is reached and no new small clusters are produced.

% We in our approach use Zhang et al.~\cite{Zhang2008} Hierarchical clustering approach. Zhang et al.\cite{Zhang2008} define the distance between two DTI fibers Q and R as the larger mean of thresholded closest distances (equation \ref{eqn:dist}).
%In the above discussion,DTI fibers are represented as points in $\mathbb{R}^3$. MetaTracts can be correspondingly represented as points in  $\mathbb{R}^3$ in terms of the start points of the cylinders. In our implementation Q and R are sets of `$C_{p}$'s. We in our implementation set $t$ to be same radius of the cylinders for all our tests. If two MetaTracts are less than a width of the cylinder apart then they are describing the same region. We use $D_{Lt}$ as our proximity measure.
%\begin{equation}\label{eqn:dist}
%D_{Lt}(Q,R,t)= max (d_{t}(Q,R,t),d_{t}(R,Q,t))
%\end{equation} 
%where, 
%\begin{equation}\label{eqn:dist-2}
%d_{t}(Q,R,t)= mean_{a\in Q,(min_{b\in R})\parallel a-b\parallel \geq t } min_{b\in R}
%{\parallel a-b\parallel}
%\end{equation} 
%The input to this step are the results of orientation based clustering (fig.~\ref{fig:clustering}b), the figure along with our data assumptions  provides some reasoning why $D_{Lt}$ is effective in our case. The MetaTracts donot necessarily run from boundary to boundary of our data. In fig.~\ref{fig:clustering}b, "Orange", MetaTracts within the  yellow rectangle might be "diverging" (running parallel for a distance and then diverge~\cite{Zhang2008}). Also, while the MetaTracts are close at some points, they belong to different bundles. So unlike the orientation measure,  mean of closest distance $d_{t}(Q,R,t)$ is a better fit than single point comparison.
%
%Hierarchical clustering approaches have one main parameter, which is the number of clusters $h$. For each cluster extracted based on the orientation phase we cluster the result into $h$ clusters. Fig. \ref{fig:clustering} (c) shows the result from hierarchical clustering.
%
%\subsection {Notes on Clustering techniques}
%\label{subsec:other_clustering}
% The current two phase clustering technique was chosen for its simplicity, relevance to the domain understanding, intuitive nature of the parameters ($K$ and $h$, see sec~\ref{sec:param_choices}). Moberts et al.~\cite{Moberts2005} also remarks that specifying number of clusters is more intuitive than other criterias such as  number of neighbors and edge thresholds. Secondly, on average we have around ten thousand individual tracts. Solving the eigen decomposition problem for such large number of tracts in time consuming. Once the orientation clustering is completed, for each particular orientation the number of fibers being compared in less, and  hierarchical  clustering can be done in a fraction of time compared to the eigen based approach this makes the choosing of $h$ an interactive process.

\section {Voxelization and surface extraction}
\label{sec:vis}
Apart from direct visualization of the MetaTracts, we show two simple extensions of MetaTracts. The first way is to voxelize the original volume according to the clusters each voxel is associated with.
The second is to extract the corresponding surfaces from the voxel data by binarizing the volume per cluster and extracting the isosurface from the binary volume by marching cubes algorithm. To voxelize the space,  we compute a neighborhood around each voxel. We then enumerate the number of voxels of each class (cluster) in this neighborhood. The voxel is then assigned to the class with the maximum number of elements in the neighborhood.  Intuitively this is a modification of the flood-fill algorithm. Figure~\ref{fig:mesh} and ~\ref{fig:prepreg}D shows examples od extracted meshes.

\begin{figure}[h]
  \centering
   \includegraphics[width=0.9\linewidth]{imagesMT2014/mesh2.png}
   \caption{Two of the extracted triangle meshes.}
   \label{fig:mesh}
   \end{figure}

%The MetaTracts and the clustering provide an abstraction of the fiber bundles and are used as a visualization tool. Through visualization MetaTracts we can answer all the queries mentioned in Section~\ref{sec:intro}. 
%
%In this Section we show two simple extensions to the MetaTracts and discuss how the end user might benefit from each. One way to visualize  the bundles is to color map the original volume according to the clusters each voxels are associated with (Section~\ref{subsec:voxel}). The second is to extract the surfaces from the voxel data (Section~\ref{subsec:surf_ext}). The visualizations can be used together, for example Fig.~\ref{fig:crop-13-mesh}e shows the crossover bundle as a mesh overlay on the MetaTracts.
%
%\subsection{Voxelization}
%\label{subsec:voxel}
%
%We take the following approach for associating a voxel $v$ with one of the cluster classes. We compute a neighborhood around each voxel. We enumerate the number of voxels of each class in this neighborhood. The voxel $v$ is then assigned to the class with the maximum number of elements inside the neighborhood. In this particular implementation, we look for neighbors in a ball of radius $r$ around the voxel. Intuitively  this is a modification of the flood-fill algorithm. In practice, we observed that a suitable value of the radius is half the distance used for seeding the volume.
%More sophisticated approaches could be employed depending upon the error-tolerance of the application used. Fig.~\ref{fig:crop-13-mesh}a shows the results of applying the voxelization step to the clustering result shown in Fig.~\ref{fig:clustering}c. This adds a level of abstraction over the MetaTracts visualization and brings us closer to the declared goal of visualizing the bundles directly at individual voxel level. Clipping planes through the voxel data can then clearly show the changing sizes of cross-sections of the bundles.
%
%\begin{figure}
%  \centering
%   \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-13voxel.eps}}
%  \subfloat[]{\includegraphics[width=0.22\linewidth]{imagesMetaTracts/crop-13-mesh-b.eps}}
%  \subfloat[]{\includegraphics[width=0.18\linewidth]{imagesMetaTracts/crop-13-mesh-a.eps}}
%  \subfloat[]{\includegraphics[width=0.21\linewidth]{imagesMetaTracts/crop-13-mesh-c.eps}}
%  \subfloat[]{\includegraphics[width=0.24\linewidth]{imagesMetaTracts/mtSurf-2}}
%  \caption{Visualization: (a) shows the voxelization. (b) shows the result of extracting surfaces from the result of the voxelization step (Clusters with very few elements have been discarded as a post processing step.), (c) and (d) shows the clusters in particular orientation direction.(e)shows the cross-over bundle as a mesh overlay on the MetaTracts. }\label{fig:crop-13-mesh}
%\end{figure}
%
%
%\subsection{Surface Extraction}
%\label{subsec:surf_ext}
%The voxelization step divides the image volume into coherent clusters. Where each voxel is associated with the cluster index of the corresponding fiber bundles. We can then use isosurfacing algorithms to find the boundaries of separation of the clustered classes. This step produces a triangular mesh following the popular marching cubes algorithm \cite{Lorensen1987}. The surfaces are computed for each of the clusters based on the voxelized grid. Our rather simple voxelization criterion means we cannot guarantee any convex or manifold criteria without adopting a more sophisticated material boundary separation algorithm.
%
%In a post processing step we remove clusters whose number of elements are far less than the mean number of elements per cluster. This simply removes outliers generated by the hierarchical clustering of classes with very few members. Fig.~\ref{fig:crop-13-mesh} shows the results of surface extraction. Fig.~\ref{fig:crop-13-mesh}b shows all the triangle meshes. Fig.~\ref{fig:crop-13-mesh}c and Fig.~\ref{fig:crop-13-mesh}d shows the meshes from two different orientation clusters. 
%Fig.~\ref{fig:crop-6} shows more results. Further processing techniques such as finding the medial axis might be applied to the extracted surface. Other NDT methods which takes meshes as input can also be utilized.

\section {Experimental Results}
\label {sec:results}

%\begin{figure*}
%\centering
%\captionsetup[subfigure]{labelformat=empty}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-G.PNG}}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-H.PNG}}
%	%\subfloat[]{\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6-I.PNG}}
%	
%	\subfloat[(a) Dat set 2]{\includegraphics[width=0.7\linewidth]{imagesMT2014/prepreg/Capture8.PNG}}
%	\subfloat[(b) Data set 3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/dataset203.png}}	
%	\caption{(a) Shows in order, the volume rendering of the second data set, a single (2D) slice, the orientation clustering bundles and the final bundles followed by proximity clustering. (b) Shows in order the third data set, a single (2D) slice and finally the extracted fiber bundles.}
%	\label{fig:crop-6}
%\end{figure*}

\begin{figure}[h]
\centering
	\includegraphics[width=0.5\textwidth]{imagesMT2014/all_clusters.PNG}
	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.}
	\label{fig:crop-16-decomp}
\end{figure}
\begin{figure}
\centering
	\includegraphics[width=0.5\textwidth]{imagesMT2014/crop-6-J.PNG}
	\caption{Data set with flat thin and compact bundles. A shows the volume rendering and a 2D slice, B shows the clusters according to individual orientation.C shows the complete result. D shows the cropped part of the mesh generated from the marked bundle.}
	\label{fig:prepreg}
\end{figure}
\begin{figure}
\centering
	\includegraphics[width=0.5\textwidth]{imagesMT2014/Graph_crop16_2.PNG}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Figure~\ref{fig:orientation_clustering}C clustered into 10 clusters. The unit for length is the grid cube length. Figure~\ref{fig:length_distribution}(b) shows the length of individual MetaTracts in cluster 6.}
	\label{fig:len_dist_crop16}
\end{figure}

%Fig~\ref{fig:kull} (b) shows the results after the MetaTracts extraction. Fig.~\ref{fig:kull} (c) shows the results of the orientation clustering with $K$ set to 2. Note, that the curved fiber bundle has also been detected with the correct orientation. Fig. \ref{fig:kull} (d) and (e) shows the results of hierarchical clustering of \ref{fig:kull} (c) using proximity based clustering. Both the orientation were clustered into 14 classes.

 %glass

%\begin{figure}
%\centering
%\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/glass-scalar.eps}}
%  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/glass-scalar-slice.eps}}
%  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/glass-cluster-b.eps}}
%  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/glass-cluster-a.eps}}
%  \caption{Our algorithm on a data set made of glass fibers. (a) The scalar dataset, (b) Depicts a slice from the data set, the  fibers are clearly demarcated from the rest of the matrix as oppose to the type of data  where it is hard to distinguish the fibers (Fig. \ref{fig:orig_data}). (c) The result of the MetaTracts process, the colors are to help visualization. (d) The result of the orientation based clustering, the two clusters are marked red and yellow.}\label{fig:glass}
%\end{figure}

We tested our method on multiple datasets with varying characteristics. 
Figure~\ref{fig:crop-16-decomp} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into 10 clusters (the ground truth is 6 and 5 respectively). The hierarchical clustering is robust to over segmentation.
Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster. We generate the correct clusters and outlier bundles with very few elements which can be discarded. As noted in section~\ref{subsec:dist_clustering} the tracts donot run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length). Fig.~\ref{fig:length_distribution}(b) shows the length per MetaTract distribution.

Fig.~\ref{fig:prepreg} shows the results on a dense data set with \textit{flat and thin} bundles. The bundles are indistinguishable in the original data set (Fig.~\ref{fig:prepreg}A).
The fiber bundle extraction using MetaTracts was implemented in ITK and C++. All clustering was done in R. The preprocessing consists of metaTract generation the distance computation and the orientation clustering. The hierarchical clustering separately,  on the orientation clusters can be done in a few minutes on a Intel Xeon E5-2667 workstation PC according to user input `h'(section~\ref{sec:param_choices}). We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without extra computation.

%\textit{Glass Fibers}, While the core advantages of our approach are in cases when the individual fibers are not discernible, our approach also works on data where the fibers are relatively well separated. Fig. \ref{fig:glass} shows our framework applied on a glass  data set (350x350x300 in size) with an isotropic resolution of $2 \mu m$. Fig. \ref{fig:glass}b shows a slice from the original dataset, (comparison with Fig. \ref{fig:orig_data} shows the individual fiber separation is easier in this case.) 
% Fig.~\ref{fig:glass}c shows the result of the MetaTracts extraction process. Fig~\ref{fig:glass}d shows the results of the orientation based clustering into two clusters (reds and yellows). TODO: with this as the ground truth, add noise and repeat experiment, to show robustness.

\begin{figure}[h]
\centering
\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/compare-clus-1.eps}}
\subfloat[]{\includegraphics[width=0.33\linewidth]{imagesMT2014/compare-clus-3.png}}
\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/Fiber_extraction_for_bundles_1.eps}}
\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/Fiber_extraction_for_bundles_2.eps}}

   \caption{\textit{Comparison of clustering approaches}.(a)Our approach,  Orientation clustering into 2 classes followed by hierarchical clustering into 3 classes each. (b) Hierarchical clustering into 6 classes based on proximity alone.\\
   \textit{Comparison of fiber extraction approaches}. (c) Salaberger et al. ~\cite{Salaberger2011}, they generate numerous small length fibers and miss the curved bundles. (d) close up of the medial axis generated by their approach.}\label{fig:compare-2}
\end{figure}

\subsection {Comparison with hierarchical fiber clustering}
\label{subsec:compare}
Here we compare our framework against the single link clustering based on proximity alone as used by Zhang et al.\cite{Zhang2008}, Moberts et al.\cite{Moberts2005}. In Fig.~\ref{fig:compare-2}(a) we see the results of our approach. 169 fibers were first clustered based on orientation into 2 clusters and then each cluster was classified into 3 classes. In Fig.~\ref{fig:compare-2}(b) we see the results of direct hierarchical clustering of the distances into 6 classes. Direct hierarchical clustering based on proximity alone is not able to partition well, especially in the presence of partially overlapping tracts (Fig.~\ref{fig:length_distribution}(b)). Without the iterative approach (section~\ref{subsec:dist_clustering}) single linkage clustering tends to produce large clusters( `green' in Fig.~\ref{fig:compare-2}(b)) and outliers as separate cluster. The green cluster incorrectly contains MetaTracts from three different fiber bundles (A,D,B). In our approach all major fiber bundles are classified separately.

\subsection{Comparison with individual fiber extraction method}
 Recently  Salaberger et al.~\cite{Salaberger2011} provided a method to study the fiber length distribution in XCT data by extracting individual fibers. They employ the Hessian followed by a medial axis detection approach to follow  individual fibers. We apply their algorithm to show the short comings of applying current fiber detection approaches on our datasets. Fig.~\ref{fig:compare-2}(c-d) shows the comparison between our approaches. Fig.~\ref{fig:compare-2}(c) shows the medial axis extracted by their algorithm. Fig.~\ref{fig:compare-2}(d) shows a close up view. The width of the line is not important because cylinders could be drawn around (c) to produce the same effect as MetaTracts. Their ~\cite{Salaberger2011} approach is designed for well separated fibers, so in our data it produces numerous small length fibers, more over due to the high noise in the Hessians around the curved bundles and the lack of well separated fibers, their approach completely misses the curved bundle.

%\begin{figure}
%  \centering
%\includegraphics[width=0.2\linewidth]{imagesMetaTracts/compare-clus-1.eps}
%\includegraphics[width=0.2\linewidth]{imagesMetaTracts/compare-clus-2.eps}
%  \caption{Comparison of clustering approaches.(a) Orientation clustering into 2 classes followed by hierarchical clustering into 3 classes each. (b)Hierarchical clustering into 6 classes.}
%  
% \label{fig:fiber-clus-2}
%\end{figure}

%\subsection {Proximity measures}
%\textbf{weighted undirected graph}
%In DT-MRI, each voxel is assigned a tensor that describes local water diffusion. In our case the Hessian is error prone

\section {User Evaluation}
\label {sec:evaluation}
\begin{figure}[h]
\centering
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
(b)Shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
\label{fig:context}
\end{figure}

\section{Parameter choices}
\label{sec:param_choices}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions. Which is generally known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Section~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and thus small clusters can be removed if they have too few elements. This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.

The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests. A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection. On the other hand a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the width parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and breadth, respectively for all tests. $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays, we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited, thus even for small $\eta$, the distinction between  the orientation clusters is preserved quite well. We compared  $\eta=3-7$ experimentally without any dramatic change in results. 
%\begin{figure}
%\centering
% \subfloat[$\eta=7$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim3k2}}
%  \subfloat[$\eta=3$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim7k2}}
%  \caption{Embedding from $\mathbb{R}^m\to \mathbb{R}^\eta$, here m is the number of MetaTracts, 580 for this example. Followed by clustering into K=2 clusters.}
%  \label{fig:eta}
%\end{figure}
%\begin{figure}[htp]
%\centering
% \includegraphics[width=\linewidth]{imagesMetaTracts/kul-results.eps}
%  \caption{Results from orientation based dimensionality reduction and clustering. A is the original data set, B shows the results of fiber tracking. C shows the results of orientation based clustering and D shows the results of the proximity based clustering. }\label{fig:kull-full}
%\end{figure}

%\begin{figure}[htp]
%\centering
% \includegraphics[width=\linewidth]{imagesMetaTracts/glass-clus-2.eps}
%  \caption{ Fig  A and C shows the results fiber extraction. B and D shows the results of clustering with clipping planes.}\label{fig:closeup-full}
%\end{figure}

\section{Conclusion and future work}
In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials. We show that our framework works at comparatively low resolution and with dense fiber arrangements (when extracting single fibers might not be possible). It handles complex fiber patterns such as ``cross overs''/``braiding''. The presented techniques are attractive and novel features for a industrial fiber bundle visualization tool.

%
%\begin{figure}
%  \centering
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-scalar.eps}}
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-tracks-1.eps}}
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-tracks-clus-a.eps}}
%    \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-tracks-clus-b.eps}}
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-labels-all.eps}}
%  \caption{Dense data set with bundles which are flat and thin. (a) shows the results original dataset, (b) shows the results of metaTract extraction, (c) shows the results of our two step clustering  into 14 classes. (d) shows the results of voxelization  and (e) shows the results of surface extraction. }\label{fig:crop-6}
%\end{figure}
%% if specified like this the section will be ommitted in review mode
% %\acknowledgements{
% The research leading to these results has received funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM}

### 2014-08-21 metaTracts-pvis2014.tex
MetaTracts working copy

% $Id: template.tex 11 2007-04-03 22:25:53Z jpeltier $

\documentclass{vgtc}                          % final (conference style)
%\documentclass[review]{vgtc}                 % review
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint]{vgtc}               % preprint
%\documentclass[electronic]{vgtc}             % electronic version

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Figures should be in CMYK or Grey scale format, otherwise, colour 
%% shifting may occur during the printing process.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1 
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\let\ifpdf\relax
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{amssymb}
\usepackage{subfig}
\usepackage{wrapfig}
\usepackage{comment}
\usepackage{epstopdf}
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites}

%% This is how authors are specified in the conference style

%% Author and Affiliation (single author).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}}
%%\affiliation{\scriptsize Allied Widgets Research}

%% Author and Affiliation (multiple authors with single affiliations).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
%%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
%%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
%%\affiliation{\scriptsize Martha Stewart Enterprises \\ Microsoft Research}

%% Author and Affiliation (multiple authors with multiple affiliations)
\author{Arindam Bhattacharya\thanks{e-mail: arindamb86@gmail.com}\\ %
        \scriptsize The Ohio State University %
\and Christoph Heinzl\thanks{e-mail:christoph.heinzl@fh-wels.at}\\ %
     \scriptsize University of Applied Sciences Upper Austria. %
\and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
    \scriptsize The Ohio State University}

%% A teaser figure can be included as follows, but is not recommended since
%% the space is now taken up by a full width abstract.
%\teaser{
%  \includegraphics[width=1.5in]{sample.eps}
%  \caption{Lookit! Lookit!}
%}

%% Abstract section.
\abstract{
Carbon fiber reinforced polymers are bundles of individual carbon fibers woven into sheets of carbon fiber cloth which are layered
and reinforced by an epoxy material. Industrial 3D X-ray computed tomography (XCT) generates a 3D volumetric representation of a scanned object, reconstructed from a series of 2D penetration images,
taken throughout a full rotation of the object.	We describe a technique for extracting and visualizing fiber bundles in fiber reinforced composites scanned with X-ray computed tomography.
Our goal is to visualize the geometric structures of the weaving pattern by analyzing large  areas of  composite woven fabrics using XCT scans of low resolution. Individual fibers are barely visible in such scans. We apply and extend methods originally developed for diffusion tensor imaging (DTI) on the generated XCT data. In the first step a coarse version of integral curves is used to extract sections of the fiber bundles in the woven fabrics. We call these sections MetaTracts.
In the next step the extracted fiber bundle sections are clustered using a two phase approach, wherein we first cluster according to orientation and then  cluster by proximity. We test our algorithm on a number of data sets and show that it effectively identifies and separates different fiber bundles. This work will provide a stepping stone to introduce novel non destructive testing techniques for reinforced fiber bundles analysis.
} % end of abstract

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

\CCScatlist{ 
  \CCScat{K.6.1}{Management of Computing and Information Systems}%
{Project and People Management}{Life Cycle};
  \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
%\firstsection{Introduction}

\maketitle

\input {./chapters/MetaTracts.tex}

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{xBundle}
\end{document}

### 2014-08-22 chapters/MetaTracts.tex
updated images

@@
 Intuitively this is a modification of the flood-fill algorithm.
-Figure~\ref{fig:mesh} and ~\ref{fig:prepreg}D shows examples od extracted meshes.
+Figure~\ref{fig:mesh} and Fig.~\ref{fig:prepreg}D shows examples od extracted meshes.
 
@@
 \centering
-	\includegraphics[width=0.5\textwidth]{imagesMT2014/Graph_crop16_2.PNG}
-	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Figure~\ref{fig:orientation_clustering}C clustered into 10 clusters.
+	\includegraphics[width=0.5\textwidth]{imagesMT2014/Graph_crop16.PNG}
+	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}C, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top).
 The unit for length is the grid cube length.

### 2014-08-26 chapters/MetaTracts.tex
updated copy added a section

@@
 \centering
-	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/MT_crop16_MT.png}}
-	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/length_distribution.PNG}}
+%	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/MT_crop16_MT_2.png}}
+	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/MT_crop16_MT_2.PNG}} 
+	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/length_distribution.PNG}}\\
+%	\subfloat[]{\includegraphics[width=0.9\linewidth]{imagesMT2014/mt_crop_16_orient.PNG}}
 	\caption{(a) MetaTracts colored according to mean orientations. (b) length distribution of individual MetaTracts for a particular bundle.}
@@
 \label{sec:vis}
-Apart from direct visualization of the MetaTracts, we show two simple extensions of MetaTracts.
-The first way is to voxelize the original volume according to the clusters each voxel is associated with.
-The second is to extract the corresponding surfaces from the voxel data by binarizing the volume per cluster and extracting the isosurface from the binary volume by marching cubes algorithm.
+Apart from direct visualization of the MetaTracts, we show two simple extensions. The first way is to voxelize the original volume according to the clusters each voxel is associated with.
+The second is to extract the corresponding surfaces from the voxel data by binarizing the volume per cluster and extracting the isosurface of the largest connected component from the binary volume.
 To voxelize the space,  we compute a neighborhood around each voxel.
 We then enumerate the number of voxels of each class (cluster) in this neighborhood.
-The voxel is then assigned to the class with the maximum number of elements in the neighborhood.
-Intuitively this is a modification of the flood-fill algorithm.
-Figure~\ref{fig:mesh} and Fig.~\ref{fig:prepreg}D shows examples od extracted meshes.
-
-\begin{figure}[h]
-  \centering
-   \includegraphics[width=0.9\linewidth]{imagesMT2014/mesh2.png}
-   \caption{Two of the extracted triangle meshes.}
-   \label{fig:mesh}
-   \end{figure}
-
+The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Fig.~\ref{fig:prepreg}D shows the result of voxelization{\tiny }.
+Figure~\ref{fig:mesh} shows examples of extracted meshes.
 %The MetaTracts and the clustering provide an abstraction of the fiber bundles and are used as a visualization tool.
@@
 \end{figure}
+\begin{figure}[h]
+  \centering
+   \includegraphics[width=0.9\linewidth]{imagesMT2014/mesh.png}
+   \caption{Extracted meshes. (A) shows two of the extracted meshes separately. (B) Shows two different views of the combined meshes.}
+   \label{fig:mesh}
+   \end{figure}
 \begin{figure}
 \centering
-	\includegraphics[width=0.5\textwidth]{imagesMT2014/crop-6-J.PNG}
+	\includegraphics[width=0.5\textwidth]{imagesMT2014/crop-6-J_2.PNG}
 	\caption{Data set with flat thin and compact bundles.
-A shows the volume rendering and a 2D slice, B shows the clusters according to individual orientation.C shows the complete result.
-D shows the cropped part of the mesh generated from the marked bundle.}
+A shows the volume rendering and a 2D slice with one of the boundaries marked in red, B shows the clusters according to individual orientation.C shows the complete result.
+D shows the voxelization of C.}
 	\label{fig:prepreg}
@@
 	\includegraphics[width=0.5\textwidth]{imagesMT2014/Graph_crop16.PNG}
+	%\includegraphics[width=0.23\textwidth]{imagesMT2014/Graph_crop16_3.PNG}
 	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}C, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top).
@@
  %glass
-
-%\begin{figure}
-%\centering
-%\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/glass-scalar.eps}}
-%  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/glass-scalar-slice.eps}}
-%  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/glass-cluster-b.eps}}
-%  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMetaTracts/glass-cluster-a.eps}}
-%  \caption{Our algorithm on a data set made of glass fibers.
-(a) The scalar dataset, (b) Depicts a slice from the data set, the  fibers are clearly demarcated from the rest of the matrix as oppose to the type of data  where it is hard to distinguish the fibers (Fig.
-\ref{fig:orig_data}).
-(c) The result of the MetaTracts process, the colors are to help visualization.
-(d) The result of the orientation based clustering, the two clusters are marked red and yellow.}\label{fig:glass}
-%\end{figure}
 
@@
 We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without extra computation.
-
-%\textit{Glass Fibers}, While the core advantages of our approach are in cases when the individual fibers are not discernible, our approach also works on data where the fibers are relatively well separated.
-Fig.
-\ref{fig:glass} shows our framework applied on a glass  data set (350x350x300 in size) with an isotropic resolution of $2 \mu m$.
-Fig.
-\ref{fig:glass}b shows a slice from the original dataset, (comparison with Fig.
-\ref{fig:orig_data} shows the individual fiber separation is easier in this case.)
-% Fig.~\ref{fig:glass}c shows the result of the MetaTracts extraction process.
-Fig~\ref{fig:glass}d shows the results of the orientation based clustering into two clusters (reds and yellows).
-TODO: with this as the ground truth, add noise and repeat experiment, to show robustness.
 
@@
 
-%\begin{figure}
-%  \centering
-%\includegraphics[width=0.2\linewidth]{imagesMetaTracts/compare-clus-1.eps}
-%\includegraphics[width=0.2\linewidth]{imagesMetaTracts/compare-clus-2.eps}
-%  \caption{Comparison of clustering approaches.(a) Orientation clustering into 2 classes followed by hierarchical clustering into 3 classes each. (b)Hierarchical clustering into 6 classes.}
-%  
-% \label{fig:fiber-clus-2}
-%\end{figure}
-
-%\subsection {Proximity measures}
-%\textbf{weighted undirected graph}
-%In DT-MRI, each voxel is assigned a tensor that describes local water diffusion. In our case the Hessian is error prone
-
 \section {User Evaluation}
@@
 Large $h$ keeps the major bundles intact.
-For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and thus small clusters can be removed if they have too few elements.
+For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4 and 5 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements).
 This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.

### 2014-08-26 chapters/supplementaryMaterial.tex
updated copy added a section

\pagebreak
\clearpage
\section{Supplementary Material}
\label{sec:sm}

\begin{figure}
\centering
\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/glass-scalar.eps}}
  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/glass-scalar-slice}}
  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/glass-cluster-b.eps}}
  \subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/glass-cluster-a.eps}}
  \caption{Our algorithm on a data set made of glass fibers. (a) The scalar dataset, (b) Depicts a slice from the data set, the  fibers are clearly demarcated from the rest of the matrix as oppose to the type of data  where it is hard to distinguish the fibers (Fig.~\ref{fig:data-char} or Fig.~\ref{fig:prepreg}A). (c) The result of the MetaTracts process, the colors are to help visualization. (d) The result of the orientation based clustering, the two clusters are marked red and yellow.}\label{fig:glass}
\end{figure}
\subsection{Glass Fibers}
One simple way we used to validate the accuracy of shape and size of the extracted MetaTracts, is to run the MetaTract technique on Glass fibers.
Unlike the presented data, in glass the fibers are clearly visible (Fig.~\ref{fig:glass}(a)). We noted that the extracted MetaTracts closely followed the fibers in the data.
\subsection{Voxelization and surface extraction Extended}
The voxelization results and some of the meshes extracted from our second data set (Fig.~\ref{fig:prepreg}) are shown in Fig.~\ref{fig:vox_extended}.
\begin{figure}[h]
\centering

\subfloat[]{\includegraphics[width=0.2\textwidth]{imagesMT2014/crop-6-C2-vol}}
\subfloat[]{\includegraphics[width=0.2\textwidth]{imagesMT2014/mesh_prepreg_A}}
\\
\caption{(a) Voxelization of Fig.~\ref{fig:prepreg}C, (b) Shows some of the extracted meshes.}
\label{fig:vox_extended}
\end{figure}
\subsection{Robustness to parameter changes}
Figure~\ref{fig:radius_3} shows the results of data set in Fig.~\ref{fig:data-char} clustered into 2 orientation cluster, and each orientation further clustered into 15 clusters (instead of 10 in Fig.~\ref{fig:len_dist_crop16}). Fig.~\ref{fig:radius_3}(a) shows the number of tracts, the median, minimum and maximum length of the MetaTracts in orientation cluster Fig.~\ref{fig:orientation_clustering}C.
The ground truth is 5 bundles. The radius of the individual MetaTract has been set to 3. We see that similar to Fig~\ref{fig:len_dist_crop16} the correct bundles are identified even with overestimated `h'. The median lengths of the extracted bundles are also similar. Fig~\ref{fig:radius_3}(c, d) shows the individual orientation bundles and Fig~\ref{fig:radius_3}(b) shows the combined bundles.
\begin{figure}[h]
\centering
\subfloat[]{\includegraphics[width=0.6\textwidth]{imagesMT2014/Graph_crop16_3}}\\
\subfloat[]{\includegraphics[width=0.3\textwidth]{imagesMT2014/crop-16/radius_3_orientC}}\\
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-16/radius_3_orientA}}
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-16/radius_3_orientB}}
\caption{(a) Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}C, clustered into 15 clusters (Fig.~\ref{fig:crop-16-decomp}A top). The unit for length is the grid cube length.(b) shows the combined clusters from both the orientation, (c),(d) show the clustering of each orientation.}
\label{fig:radius_3}
\end{figure}

### 2014-08-28 chapters/MetaTracts.tex
current copy.
Added the supplementary  section

@@
 We detail and evaluate a method which uses visualization to quickly gain insight into our data.
-We demarcate the above queries into two, "geometric structure" and "spatial context".
+We demarcate the above queries into two, ``geometric structure" and ``spatial context".
 Geometric structure refers to the shape, size, orientation of a bundle or a group of bundles.
@@
 The general assumption of fiber clustering in diffusion tensor imaging is that a similarity measure based on factors such as proximity between fibers can be used to interpret a collection of fiber tracts as bundles.
-Extensive research has been done on automatic DTI fiber clustering methods (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}).
-We limit ourselves on advancement of interest to our current work.
- %Zhang et al.~\cite{Zhang2008}, O'Donnell et al.~\cite{westinMEDIA02}. 
+Extensive research has been done on automatic DTI fiber clustering methods (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}.
  The clustering approach divides the problem into two parts.
@@
 In the case of fiber bundles, the individual points are fiber traces.
-In the ideal case fiber traces which belong to the same bundle must remain ``close" to each other in the lower dimensional space.
-~\cite{westinMEDIA02} also uses spectral clustering on a Hausdorff distance measure defined as the maximum of pointwise minimum distances between two fibers.
+In the ideal case fiber traces which belong to the same bundle must remain ``close" to each other in the lower dimensional space.Westin et al.~\cite{westinMEDIA02} also uses spectral clustering on a Hausdorff distance measure defined as the maximum of pointwise minimum distances between two fibers.
 Jonasson et al.~\cite{jonasson2005} runs k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
@@
 Figure~\ref{fig:data-char}C shows the zoomed in version of the yellow square.
-At isotropic resolution of 2 micrometers, the individual carbon fibers are hard to resolve from the 2D slices.
-In Figure~\ref{fig:data-char}C contains multiple bundles going in opposite directions.
+At isotropic resolution of 2 micrometers, the individual carbon fibers in a fiber bundle are hard to resolve from the 2D slices.
+Figure~\ref{fig:data-char}C contains multiple bundles going in opposite directions.
 The separation between two bundles is barely discriminable.
@@
 Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}.
-The principal direction based on the hessian matrix works best when the tubular structures are well separated from the surrounding, this is not the case for our data.
+The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding, this is not the case for our data.
 The local orientaion at each voxel is inherently more noisy.
@@
 The reliable Hessian procedure  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1].
-In Figure \ref{fig:algo} all regions which have reliable hessian less than the threshold have been marked in green.
+In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in green.
 Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}).
@@
 At each iteration we pick the top element in the priority queue generate the corresponding cylinder and repeat the steps.
-Essentially, eqn~\ref{eqn:algo_1}, selects a grid point which is the furthest from the current point and going in a similar direction.
+Essentially, Eqn.~\ref{eqn:algo_1}, selects a grid point which is the furthest from the current point and going in a similar direction.
 This gives it the advantage of tackling noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation (for orientation) in an already noisy environment.
@@
 We do not select the purple vertex (Figure~\ref{fig:algo}C) because it is not a candidate vertex.
-If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices inside the cylinder and will be of short length.
-Short tracts are then removed using a threshold on MetaTracts  length.
+If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length.
+Short tracts are then remove.
 The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}(a).
@@
 However, the MetaTracts produced as described in Section 4 often do not extend the full length of the fiber bundle.  Thus, MetaTracts
-in a given fiber bundle may have very different lengths and may only partially overlap (figure~\ref{fig:length_distribution}(b)).
+in a given fiber bundle may have very different lengths and may only partially overlap Figure~\ref{fig:length_distribution}(b).
 
@@
 The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an m-dimensional space.
-K-means clustering can then be used to cluster in this lower dimensional space.
+K-means clustering is then used to cluster in this lower dimensional space.
 
@@
 The edge weights give a distance matrix representing the distance between each pair of points.
-Using the Belkin and Niyogi algorithm we "embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix.
+Using the Belkin and Niyogi algorithm we ``embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix.
 
@@
  \label{subsec:dist_clustering}
+ We use the directed Hausdorff distance for distance based clustering.
  Each MetaTract is represented as a set of points ($C_P$). 
@@
 
-We repeated both until a steady state of clusters is reached and no new small clusters are produced.
+We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters are shown in Fig.~\ref{fig:crop-16-decomp}.
 
@@
 We then enumerate the number of voxels of each class (cluster) in this neighborhood.
-The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Fig.~\ref{fig:prepreg}D shows the result of voxelization{\tiny }.
-Figure~\ref{fig:mesh} shows examples of extracted meshes.
+The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Fig.~\ref{fig:mesh}A shows the result of voxelization.
+Fig.~\ref{fig:mesh}(B, C) show examples of extracted meshes.
 %The MetaTracts and the clustering provide an abstraction of the fiber bundles and are used as a visualization tool.
@@
   \centering
-   \includegraphics[width=0.9\linewidth]{imagesMT2014/mesh.png}
-   \caption{Extracted meshes. (A) shows two of the extracted meshes separately. (B) Shows two different views of the combined meshes.}
+   \includegraphics[width=0.5\textwidth]{imagesMT2014/mesh_2_C.png}
+   \caption{Volumes and  meshes. (A) Shows the voxelized volume. (B) Shows two different extracted meshes (C) shows the combined meshes.}
    \label{fig:mesh}
@@
  %glass
-
 We tested our method on multiple datasets with varying characteristics. 
@@
 As noted in section~\ref{subsec:dist_clustering} the tracts donot run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length).
-Fig.~\ref{fig:length_distribution}(b) shows the length per MetaTract distribution.
+Fig.~\ref{fig:length_distribution}(b) shows the length per MetaTract distribution for a particular fiber bundle.
 
-Fig.~\ref{fig:prepreg} shows the results on a dense data set with \textit{flat and thin} bundles. The bundles are indistinguishable in the original data set (Fig.~\ref{fig:prepreg}A).
+Fig.~\ref{fig:prepreg} shows the results on a dense data set with \textit{flat and thin} bundles.
+The bundles are indistinguishable in the original data set (Fig.~\ref{fig:prepreg}A).
+A red dotted line shows one of the fiber bundle boundary.
+Fig.~\ref{fig:prepreg}B shows the fiber bundles for the two orientation.
+Fig.~\ref{fig:prepreg}C shows the combined results.Fig.~\ref{fig:prepreg}D shows the volume after voxelization.
 The fiber bundle extraction using MetaTracts was implemented in ITK and C++.
@@
 The preprocessing consists of metaTract generation the distance computation and the orientation clustering.
-The hierarchical clustering separately,  on the orientation clusters can be done in a few minutes on a Intel Xeon E5-2667 workstation PC according to user input `h'(section~\ref{sec:param_choices}).
+The hierarchical clustering separately, on the results of the orientation clusters can be done  quickly in a few minutes on a Intel Xeon E5-2667 workstation PC according to user input `h'(section~\ref{sec:param_choices}).
 We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without extra computation.
-
-\begin{figure}[h]
-\centering
-\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/compare-clus-1.eps}}
-\subfloat[]{\includegraphics[width=0.33\linewidth]{imagesMT2014/compare-clus-3.png}}
-\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/Fiber_extraction_for_bundles_1.eps}}
-\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/Fiber_extraction_for_bundles_2.eps}}
-
-   \caption{\textit{Comparison of clustering approaches}.(a)Our approach,  Orientation clustering into 2 classes followed by hierarchical clustering into 3 classes each.
-(b) Hierarchical clustering into 6 classes based on proximity alone.\\
-   \textit{Comparison of fiber extraction approaches}.
-(c) Salaberger et al.
-~\cite{Salaberger2011}, they generate numerous small length fibers and miss the curved bundles.
-(d) close up of the medial axis generated by their approach.}\label{fig:compare-2}
-\end{figure}
-
-\subsection {Comparison with hierarchical fiber clustering}
-\label{subsec:compare}
-Here we compare our framework against the single link clustering based on proximity alone as used by Zhang et al.\cite{Zhang2008}, Moberts et al.\cite{Moberts2005}.
-In Fig.~\ref{fig:compare-2}(a) we see the results of our approach.
-169 fibers were first clustered based on orientation into 2 clusters and then each cluster was classified into 3 classes.
-In Fig.~\ref{fig:compare-2}(b) we see the results of direct hierarchical clustering of the distances into 6 classes.
-Direct hierarchical clustering based on proximity alone is not able to partition well, especially in the presence of partially overlapping tracts (Fig.~\ref{fig:length_distribution}(b)).
-Without the iterative approach (section~\ref{subsec:dist_clustering}) single linkage clustering tends to produce large clusters( `green' in Fig.~\ref{fig:compare-2}(b)) and outliers as separate cluster.
-The green cluster incorrectly contains MetaTracts from three different fiber bundles (A,D,B).
-In our approach all major fiber bundles are classified separately.
-
-\subsection{Comparison with individual fiber extraction method}
- Recently  Salaberger et al.~\cite{Salaberger2011} provided a method to study the fiber length distribution in XCT data by extracting individual fibers.
-They employ the Hessian followed by a medial axis detection approach to follow  individual fibers.
-We apply their algorithm to show the short comings of applying current fiber detection approaches on our datasets.
-Fig.~\ref{fig:compare-2}(c-d) shows the comparison between our approaches.
-Fig.~\ref{fig:compare-2}(c) shows the medial axis extracted by their algorithm.
-Fig.~\ref{fig:compare-2}(d) shows a close up view.
-The width of the line is not important because cylinders could be drawn around (c) to produce the same effect as MetaTracts.
-Their ~\cite{Salaberger2011} approach is designed for well separated fibers, so in our data it produces numerous small length fibers, more over due to the high noise in the Hessians around the curved bundles and the lack of well separated fibers, their approach completely misses the curved bundle.
 

### 2014-08-28 chapters/supplementaryMaterial.tex
current copy.
Added the supplementary  section

@@
 \label{sec:sm}
+ In this section we show the results of some extended tests.
+ section~\ref{subsec:glass} shows our results on glass fibers.
+Section~\ref{subsec:vox_ex} shows the some more results of voxelization and mesh generation.
+Section~\ref{subsec:compare} shows comparisons to hierarchical clustering and with individual fiber extraction technique.
+Finally, section~\ref{subsec:robustness} shows another result with different parameter to show robustness of parameter selection.
 
@@
 \subsection{Glass Fibers}
+\label{subsec:glass}
 One simple way we used to validate the accuracy of shape and size of the extracted MetaTracts, is to run the MetaTract technique on Glass fibers.
@@
 \subsection{Voxelization and surface extraction Extended}
+\label{subsec:vox_ex}
 The voxelization results and some of the meshes extracted from our second data set (Fig.~\ref{fig:prepreg}) are shown in Fig.~\ref{fig:vox_extended}.
@@
 \centering
-
 \subfloat[]{\includegraphics[width=0.2\textwidth]{imagesMT2014/crop-6-C2-vol}}
@@
 \end{figure}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{imagesMT2014/UserStudy/vol_crop16}
+\\
+\caption{ A shows the Voxelization of Fig.~\ref{fig:data-char}. B shows a single slice. C the slice of original data. }
+\label{fig:vox_extended_crop16}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/compare-clus-1.eps}}
+\subfloat[]{\includegraphics[width=0.33\linewidth]{imagesMT2014/compare-clus-3.png}}
+\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/Fiber_extraction_for_bundles_1.eps}}
+\subfloat[]{\includegraphics[width=0.25\linewidth]{imagesMT2014/Fiber_extraction_for_bundles_2.eps}}
+
+   \caption{\textit{Comparison of clustering approaches}.(a)Our approach,  Orientation clustering into 2 classes followed by hierarchical clustering into 3 classes each.
+(b) Hierarchical clustering into 6 classes based on proximity alone.\\
+   \textit{Comparison of fiber extraction approaches}.
+(c) Salaberger et al.
+~\cite{Salaberger2011}, they generate numerous small length fibers and miss the curved bundles.
+(d) close up of the medial axis generated by their approach.}\label{fig:compare-3}
+\end{figure}
+\subsection {Comparison with hierarchical fiber clustering}
+\label{subsec:compare}
+Here we compare our framework against the single link clustering based on proximity alone as used by Zhang et al.\cite{Zhang2008}, Moberts et al.\cite{Moberts2005}.
+In Fig.~\ref{fig:compare-3}(a) we see the results of our approach.
+169 fibers were first clustered based on orientation into 2 clusters and then each cluster was classified into 3 classes.
+In Fig.~\ref{fig:compare-3}(b) we see the results of direct hierarchical clustering of the distances into 6 classes.
+Direct hierarchical clustering based on proximity alone is not able to partition well, especially in the presence of partially overlapping tracts (Fig.~\ref{fig:length_distribution}(b)).
+Without the iterative approach (section~\ref{subsec:dist_clustering}) single linkage clustering tends to produce large clusters( `green' in Fig.~\ref{fig:compare-3}(b)) and outliers as separate cluster.
+The green cluster incorrectly contains MetaTracts from three different fiber bundles (A,D,B).
+In our approach all major fiber bundles are classified separately.
+\subsection{Comparison with individual fiber extraction method}
+ Recently  Salaberger et al.~\cite{Salaberger2011} provided a method to study the fiber length distribution in XCT data by extracting individual fibers.
+They employ the Hessian followed by a medial axis detection approach to follow  individual fibers.
+We apply their algorithm to show the short comings of applying current fiber detection approaches on our datasets.
+Fig.~\ref{fig:compare-3}(c-d) shows the comparison between our approaches.
+Fig.~\ref{fig:compare-3}(c) shows the medial axis extracted by their algorithm.
+Fig.~\ref{fig:compare-3}(d) shows a close up view.
+The width of the line is not important because cylinders could be drawn around (c) to produce the same effect as MetaTracts.
+Their ~\cite{Salaberger2011} approach is designed for well separated fibers, so in our data it produces numerous small length fibers, more over due to the high noise in the Hessians around the curved bundles and the lack of well separated fibers, their approach completely misses the curved bundle.
+
 \subsection{Robustness to parameter changes}
+\label{subsec:robustness}
 Figure~\ref{fig:radius_3} shows the results of data set in Fig.~\ref{fig:data-char} clustered into 2 orientation cluster, and each orientation further clustered into 15 clusters (instead of 10 in Fig.~\ref{fig:len_dist_crop16}).
@@
 Fig~\ref{fig:radius_3}(c, d) shows the individual orientation bundles and Fig~\ref{fig:radius_3}(b) shows the combined bundles.
-\begin{figure}[h]
+\begin{figure*}
 \centering
 \subfloat[]{\includegraphics[width=0.6\textwidth]{imagesMT2014/Graph_crop16_3}}\\
-\subfloat[]{\includegraphics[width=0.3\textwidth]{imagesMT2014/crop-16/radius_3_orientC}}\\
+\subfloat[]{\includegraphics[width=0.3\textwidth]{imagesMT2014/crop-16/radius_3_orientC}}
 \subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-16/radius_3_orientA}}
@@
 \label{fig:radius_3}
-\end{figure}
+\end{figure*}

### 2014-08-28 metaTracts-pvis2014.tex
updated copy added a section
Added the supplementary  section

@@
 \abstract{
-Carbon fiber reinforced polymers are bundles of individual carbon fibers woven into sheets of carbon fiber cloth which are layered
+Carbon fiber reinforced polymers are bundles of individual carbon fibers (henceforth referred to as fiber bundle) woven into sheets of carbon fiber cloth which are layered
 and reinforced by an epoxy material. Industrial 3D X-ray computed tomography (XCT) generates a 3D volumetric representation of a scanned object, reconstructed from a series of 2D penetration images,
 taken throughout a full rotation of the object.	We describe a technique for extracting and visualizing fiber bundles in fiber reinforced composites scanned with X-ray computed tomography.
+Current approaches focus on identifying and extracting individual fibers.
 Our goal is to visualize the geometric structures of the weaving pattern by analyzing large  areas of  composite woven fabrics using XCT scans of low resolution.
@@
 We call these sections MetaTracts.
-In the next step the extracted fiber bundle sections are clustered using a two phase approach, wherein we first cluster according to orientation and then  cluster by proximity.
+Next the extracted fiber bundle sections are clustered using a two step approach, wherein we first cluster according to orientation and then  cluster by proximity.
 We test our algorithm on a number of data sets and show that it effectively identifies and separates different fiber bundles.
@@
 \bibliography{xBundle}
+\input {./chapters/supplementaryMaterial.tex}
 \end{document}

### 2014-09-01 chapters/MetaTracts.tex
modified abstract + intro

@@
 %REPLACED BY 
-%Carbon fiber reinforced polymers (CFRP) are promising materials for future components as they allow integrating the continuously rising demands of various industrial  branches.
-
-Carbon fiber reinforced polymers (CFRP) are bundles of individual carbon fibers
-woven into sheets of carbon fiber cloth which are layered
-and reinforced by an epoxy material.
-To manufacture CFRP's, bundles of individual carbon fibers are woven into sheets of carbon fiber cloth which are layered and reinforced by an epoxy matrix.
-The alignment and weave pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
+Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and future components. 
+The materials of choice of a strongly growing number of industrial branches are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.
+Carbon fiber reinforced polymers, more specifically CFRPs with endless carbon fibers which we focus on in this work, typically consist of two main components: 
+The matrix, which acts as a glue and the reinforcements, which allows for achieve the desired characteristics. 
+Resins such as epoxy are typically used in a CFRP material system with endless carbon fibers for the matrix. 
+For special applications additional filler materials, such as colors or other particles, may be integrated. 
+However, the most important component is the reinforcement component, which consists in our application case of bundles of endless carbon fibers. 
+The carbon fiber bundles are woven into sheets of carbon fiber cloth, which are layered and reinforced by an epoxy material to manufacture a CFRP laminate.
+The alignment and weaving pattern of a CFRP laminate strongly influences the strength and stiffness properties of the material.
 The resulting material shows high specific stiffness, high specific strength, and high corrosion resistance.

### 2014-09-01 metaTracts-pvis2014.tex
modified abstract + intro

--- 2014-08-21 metaTracts-pvis2014.tex
@@
 \and Christoph Heinzl\thanks{e-mail:christoph.heinzl@fh-wels.at}\\ %
-     \scriptsize University of Applied Sciences Upper Austria. %
+     \scriptsize University of Applied Sciences - Upper Austria. %
 \and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
@@
 \abstract{
-Carbon fiber reinforced polymers are bundles of individual carbon fibers woven into sheets of carbon fiber cloth which are layered
-and reinforced by an epoxy material. Industrial 3D X-ray computed tomography (XCT) generates a 3D volumetric representation of a scanned object, reconstructed from a series of 2D penetration images,
-taken throughout a full rotation of the object.	We describe a technique for extracting and visualizing fiber bundles in fiber reinforced composites scanned with X-ray computed tomography.
-Our goal is to visualize the geometric structures of the weaving pattern by analyzing large  areas of  composite woven fabrics using XCT scans of low resolution.
-Individual fibers are barely visible in such scans.
-We apply and extend methods originally developed for diffusion tensor imaging (DTI) on the generated XCT data.
-In the first step a coarse version of integral curves is used to extract sections of the fiber bundles in the woven fabrics.
-We call these sections MetaTracts.
-In the next step the extracted fiber bundle sections are clustered using a two phase approach, wherein we first cluster according to orientation and then  cluster by proximity.
-We test our algorithm on a number of data sets and show that it effectively identifies and separates different fiber bundles.
+Carbon fiber reinforced polymers (CFRP) consist of two components: a matrix and a reinforcement. 
+The base material for the matrix is typically a resin such as epoxy. 
+The reinforcement provides the desired material characteristics. 
+In case of CFRPs the reinforcements are typically bundles of individual carbon fibers (henceforth referred to as fiber bundle) woven into sheets of carbon fiber cloth which are layered and reinforced by an epoxy material.
+For quality control of CFRP materials and components industrial 3D X-ray computed tomography (XCT) is increasingly applied as it allows detailed characterizations of the material in a non-destructive manner.
+
+This work contributes to the state of the art of CFRP characterization by the introduction of MetaTracts, a technique for extracting and visualizing fiber bundles and the geometric structure of the weaving pattern in fiber reinforced composites scanned with X-ray computed tomography.
+The declared goal of this work is to analyzing large areas of CFRPs using XCT scans of low resolution, in which individual fibers are barely visible. 
+To reach this goal we apply and extend methods originally developed for diffusion tensor imaging (DTI) on the generated XCT data. 
+In the first step a coarse version of integral curves is used to extract sections of the fiber bundles in the woven fabrics. We call these sections MetaTracts.
+Next, the extracted fiber bundle sections are clustered using a two step approach, wherein we first cluster according to orientation and then  cluster by proximity. 
+We evaluate our algorithm on a number of data sets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 
+%mention user study here?%
 This work will provide a stepping stone to introduce novel non destructive testing techniques for reinforced fiber bundles analysis.

### 2014-09-01 metaTracts-pvis2014.tex
no message

@@
 \abstract{
-Carbon fiber reinforced polymers (CFRP) consist of two components: a matrix and a reinforcement. 
-The base material for the matrix is typically a resin such as epoxy. 
-The reinforcement provides the desired material characteristics. 
-In case of CFRPs the reinforcements are typically bundles of individual carbon fibers (henceforth referred to as fiber bundle) woven into sheets of carbon fiber cloth which are layered and reinforced by an epoxy material.
+Carbon fiber reinforced polymers (CFRP) typically consist of two main components: the matrix and the reinforcement. 
+Most important in order to achieve the desired characteristics of a CFRP is the reinforcement component. 
+In case of CFRPs the reinforcements typically consist of bundles of endless carbon fibers woven into sheets of carbon fiber cloth, which are then layered and combined to a CFRP by the matrix component.
 For quality control of CFRP materials and components industrial 3D X-ray computed tomography (XCT) is increasingly applied as it allows detailed characterizations of the material in a non-destructive manner.
 
-This work contributes to the state of the art of CFRP characterization by the introduction of MetaTracts, a technique for extracting and visualizing fiber bundles and the geometric structure of the weaving pattern in fiber reinforced composites scanned with X-ray computed tomography.
-The declared goal of this work is to analyzing large areas of CFRPs using XCT scans of low resolution, in which individual fibers are barely visible. 
+The presented work contributes to the state of the art of CFRP characterization by the introduction of MetaTracts, a technique for extracting and visualizing fiber bundles and the geometric structure of the weaving pattern in fiber reinforced composites scanned with X-ray computed tomography.
+The declared goal of this work is to analyze large areas of CFRPs using XCT scans of low resolution, in which individual fibers are barely visible. 
 To reach this goal we apply and extend methods originally developed for diffusion tensor imaging (DTI) on the generated XCT data. 
@@
 Next, the extracted fiber bundle sections are clustered using a two step approach, wherein we first cluster according to orientation and then  cluster by proximity. 
-We evaluate our algorithm on a number of data sets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 
+We evaluate our algorithm on a number of datasets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 
 %mention user study here?%

### 2014-09-04 chapters/MetaTracts.tex
updated image,

@@
 B: Shows 2D slices along the X (Right) and Z (Left) axis.
-C: Shows a zoom of the yellow square, multiple bundles cross and are indistinguishable by visual inspection alone.
+C: Shows a zoom of the green square, multiple bundles cross and are indistinguishable by visual inspection alone.
 }
@@
 \centering
-\includegraphics[width=0.45\textwidth]{imagesMT2014/algo_A.png}
+\includegraphics[width=0.5\textwidth]{imagesMT2014/algo.png}
 \caption{Flow chart of our approach}
@@
   \begin{center}
-  \includegraphics[ width=0.9\linewidth]{imagesMT2014/MT_abstract_A2} %algo-1d}
+  \includegraphics[ width=0.4\textwidth]{imagesMT2014/MT_abstract} 
   \end{center}
@@
 Short tracts are then remove.
-The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}(a).
+The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}(A).
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
@@
 \centering
-%	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/MT_crop16_MT_2.png}}
-	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/MT_crop16_MT_2.PNG}} 
-	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/length_distribution.PNG}}\\
-%	\subfloat[]{\includegraphics[width=0.9\linewidth]{imagesMT2014/mt_crop_16_orient.PNG}}
-	\caption{(a) MetaTracts colored according to mean orientations. (b) length distribution of individual MetaTracts for a particular bundle.}
+	\includegraphics[width=0.4\textwidth]{imagesMT2014/crop-16/length_dist.png}
+	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.}
 	\label{fig:length_distribution}
@@
 However, the MetaTracts produced as described in Section 4 often do not extend the full length of the fiber bundle.  Thus, MetaTracts
-in a given fiber bundle may have very different lengths and may only partially overlap Figure~\ref{fig:length_distribution}(b).
+in a given fiber bundle may have very different lengths and may only partially overlap Figure~\ref{fig:length_distribution}(B).
 
@@
 %  	
- \includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/Capture4.PNG}
-  	\caption{(B)Shows the result of K means clustering with data points projected to the top three eigenvectors as major axes.
-(A,C) show the MetaTracts colored according to the clustering results.
-(The white to black tracts are only to show context.)  }
+ \includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
+  	\caption{(C) Shows the result of K means clustering with data points projected to the top three eigenvectors as major axes.
+(A) shows the MetaTracts projected on top three eigenvectors.
+(B) shows 2D slices of C.
+}
   \label{fig:orientation_clustering}
@@
  but the directed Hausdorff is not.
- Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially cover the bundle (figure~\ref{fig:length_distribution}(b)).
+ Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially cover the bundle (figure~\ref{fig:length_distribution}(B)).
  If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$,
@@
 We then enumerate the number of voxels of each class (cluster) in this neighborhood.
-The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Fig.~\ref{fig:mesh}A shows the result of voxelization.
-Fig.~\ref{fig:mesh}(B, C) show examples of extracted meshes.
+The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Fig.~\ref{fig:crop-16-decomp}C and Fig~\ref{fig:prepreg}D shows the result of voxelization.
+Fig.~\ref{fig:crop-16-decomp}(E, F) show examples of extracted meshes.
 %The MetaTracts and the clustering provide an abstraction of the fiber bundles and are used as a visualization tool.
@@
 \label {sec:results}
-
 %\begin{figure*}
@@
 %\end{figure*}
-
-\begin{figure}[h]
+\begin{figure*}[t]
 \centering
-	\includegraphics[width=0.5\textwidth]{imagesMT2014/all_clusters.PNG}
-	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.}
+	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
+	\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
+	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of B.
+(D) shows a 2D slice of the volume.
+(E,F) shows the extracted meshes.}
 	\label{fig:crop-16-decomp}
-\end{figure}
-\begin{figure}[h]
-  \centering
-   \includegraphics[width=0.5\textwidth]{imagesMT2014/mesh_2_C.png}
-   \caption{Volumes and  meshes. (A) Shows the voxelized volume. (B) Shows two different extracted meshes (C) shows the combined meshes.}
-   \label{fig:mesh}
-   \end{figure}
+\end{figure*}
 \begin{figure}
 \centering
-	\includegraphics[width=0.5\textwidth]{imagesMT2014/crop-6-J_2.PNG}
+	\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6.PNG}
 	\caption{Data set with flat thin and compact bundles.
@@
 	\label{fig:prepreg}
-\end{figure}
-\begin{figure}
-\centering
-	\includegraphics[width=0.5\textwidth]{imagesMT2014/Graph_crop16.PNG}
-	%\includegraphics[width=0.23\textwidth]{imagesMT2014/Graph_crop16_3.PNG}
-	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}C, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top).
-The unit for length is the grid cube length.
-Figure~\ref{fig:length_distribution}(b) shows the length of individual MetaTracts in cluster 6.}
-	\label{fig:len_dist_crop16}
 \end{figure}
@@
 As noted in section~\ref{subsec:dist_clustering} the tracts donot run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length).
-Fig.~\ref{fig:length_distribution}(b) shows the length per MetaTract distribution for a particular fiber bundle.
+Fig.~\ref{fig:length_distribution}(B) shows the length per MetaTract distribution for a particular fiber bundle.
 
@@
 \end{figure}
-
 \section{Parameter choices}
 \label{sec:param_choices}
+\begin{figure}
+\centering
+	\includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/histogram_10_A.PNG}
+	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top).
+The unit for length is the grid cube length.
+Figure~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6.}
+	\label{fig:len_dist_crop16}
+\end{figure}
 The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering.

### 2014-09-05 chapters/MetaTracts.tex
updated image locations
shortened abstract, red thread in introduction

@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and future components.
-The materials of choice of a strongly growing number of industrial branches are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.
-Carbon fiber reinforced polymers, more specifically CFRPs with endless carbon fibers which we focus on in this work, typically consist of two main components: 
-The matrix, which acts as a glue and the reinforcements, which allows for achieve the desired characteristics. 
-Resins such as epoxy are typically used in a CFRP material system with endless carbon fibers for the matrix. 
-For special applications additional filler materials, such as colors or other particles, may be integrated. 
-However, the most important component is the reinforcement component, which consists in our application case of bundles of endless carbon fibers. 
-The carbon fiber bundles are woven into sheets of carbon fiber cloth, which are layered and reinforced by an epoxy material to manufacture a CFRP laminate.
-The alignment and weaving pattern of a CFRP laminate strongly influences the strength and stiffness properties of the material.
-The resulting material shows high specific stiffness, high specific strength, and high corrosion resistance.
-Increasingly, complex primary structures and highly loaded components are being manufactured from CFRP.
-Advanced composites and in particular carbon fiber reinforced polymers have thus successfully been introduced in aeronautic and automotive applications.
-Besides commercial aerospace and automotive industry, CFRP is also being used
-in a wide range of applications such as aeronautical, marine and civil engineering, automotive and wind turbine design,
-and sporting goods manufacturing.
+The material of choice of a strongly growing number of industrial branches are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.
 
-The increasing share of fiber reinforced polymers also generates a strong demand for non-destructive testing (NDT) techniques~\cite{Red2012} of such materials and components.
-Ultrasonic testing is the most commonly used method for non-destructive testing of fiber reinforce polymers.
-It provides a quick, cost-efficient overview, but has low resolution and is imprecise.
-Recently, industrial 3D X-Ray computed tomography (XCT) has been applied to non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
+CFRP materials show characteristics as high specific stiffness, high specific strength, high corrosion resistance at low weight in comparison to conventional materials.
+At the same time using CFRPs highly complex and integrated components may be produced, which were previously impossible to manufacture.
+Increasingly, primary structures and highly loaded components are being manufactured from CFRP.
+Thus CFRPs have successfully been introduced in aeronautic and automotive applications.
+Besides aerospace and automotive industry, CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.
+
+Carbon fiber reinforced polymer components, more specifically CFRP laminates with endless carbon fibers which we focus on in this work, typically consist of two main components: 
+The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics e.g.
+in terms of stiffness, strength or durability.
+To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use. 
+Most of these processes start with reinforcement component, weaving individual carbon fiber bundles into sheets of carbon fiber cloth in a predefined pattern. 
+These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also refereed to as fabric.
+Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. 
+Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength and stiffness properties. 
+Next, resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. 
+The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles. 
+For the matrix of CFRP material systems with endless carbon fibers typically resins such as epoxy are frequently used. 
+In special applications additional filler materials, such as colors or other particles, may be integrated. 
+After curing the resin the production process of the CFRP laminate is finished.
+
+The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
+Ultrasonic testing is yet the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
+On the one hand ultrasonic testing provides a quick and cost-efficient overview.
+On the other hand it shows low resolution and may generate arbitrary results, e.g.
+due to the geometry of the final component.
+
+Aside ultrasonic testing, industrial 3D X-Ray computed tomography (XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. 
 XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, taken throughout a full rotation of the specimen. 
-
-In XCT, the specimen is  placed on the rotary table between X-ray source and detector and penetrated by an incident X-rays of the source.
+In industrial XCT, the specimen is placed on the rotary table between X-ray source and detector and penetrated by incident X-rays of the source.
 The X-rays pass through the specimen and are attenuated by the material in the specimen.
-The detector transfers the X-rays in its scintillator layer into visible light. The light is then recorded in a 2D projection image. 
-The table is rotated by a small amount and the the process is repeated~\cite{heinzl-2008-thesis}.
-
-In state of the art devices, XCT can reach voxel sizes below $500 nm$,
-it generates high resolution XCT volume data for comprehensive and detailed analyses of the test specimens.
-However, there is still a trade-off between viewport and image resolution.
+The detector records 2D projection image depending on its principle, e.g. by transferring the X-rays in a scintillator layer into visible light.  
+The specimen is rotated by a predefined angular step and the process is repeated~\cite{heinzl-2008-thesis}. 
+In state of the art devices, XCT can reach voxel sizes below $500 nm$.
+XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. 
+Unfortunately, there is still a trade-off between viewport and image resolution. 
 The magnification reached within a cone beam XCT scan is determined by the specified distances
 between source and specimen as well as source and detector.
-The magnification directly influences both resolution and viewport: higher resolutions decreases the viewport or how much of the structure can be imaged while lower resolutions allow for larger viewports.
-The domain experts are increasingly interested in visualizing the internal geometric structures present in the  weaving pattern of the fiber bundles themselves instead of high resolution studies of individual fibers.
+The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport showing more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.
 
-We focus on datasets with larger viewports and lower resolutions where the single fibers are barely visible. 
-Figure~\ref{fig:data-char} provides an example of our dataset. It also shows an example of the recurring bundle pattern called an \textit{unit cell}.
-
- This work is thus motivated from the recent progress in two interrelated fields: On the one hand, fiber reinforced polymer (FRP) components has gained wide application in a variety of industries  because of FRP's superior material and physical properties in comparison to conventional materials \cite{Karpat2012}.
+In this work we focus on datasets with larger viewports and lower resolutions where the single fibers are not or barely visible.
+Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of the fiber bundles themselves instead of high resolution studies of individual fibers.
+Figure~\ref{fig:data-char} provides an example of our dataset. It also shows an example of the recurring fiber bundle pattern, called a \textit{unit cell}.
+Our work is thus motivated from the recent progress in two interrelated fields: On the one hand, fiber reinforced polymer (FRP) components has gained wide application in a variety of industries  because of FRP's superior material and physical properties in comparison to conventional materials \cite{Karpat2012}.
 On the other hand the use of industrial 3D X-Ray computed tomography (XCT) based non destructive testing of fiber reinforced components has also gained popularity~\cite{Schilling2005}.
@@
 %\end{figure*}
-\begin{figure*}[t]
+\begin{figure*}
 \centering

### 2014-09-05 metaTracts-pvis2014.tex
updated image,
shortened abstract, red thread in introduction

--- 2014-08-28 metaTracts-pvis2014.tex
@@
 \and Christoph Heinzl\thanks{e-mail:christoph.heinzl@fh-wels.at}\\ %
-     \scriptsize University of Applied Sciences Upper Austria. %
+     \scriptsize University of Applied Sciences - Upper Austria%
 \and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
@@
 \abstract{
-Carbon fiber reinforced polymers are bundles of individual carbon fibers (henceforth referred to as fiber bundle) woven into sheets of carbon fiber cloth which are layered
-and reinforced by an epoxy material. Industrial 3D X-ray computed tomography (XCT) generates a 3D volumetric representation of a scanned object, reconstructed from a series of 2D penetration images,
-taken throughout a full rotation of the object.	We describe a technique for extracting and visualizing fiber bundles in fiber reinforced composites scanned with X-ray computed tomography.
-Current approaches focus on identifying and extracting individual fibers.
-Our goal is to visualize the geometric structures of the weaving pattern by analyzing large  areas of  composite woven fabrics using XCT scans of low resolution.
-Individual fibers are barely visible in such scans.
-We apply and extend methods originally developed for diffusion tensor imaging (DTI) on the generated XCT data.
-In the first step a coarse version of integral curves is used to extract sections of the fiber bundles in the woven fabrics.
-We call these sections MetaTracts.
-Next the extracted fiber bundle sections are clustered using a two step approach, wherein we first cluster according to orientation and then  cluster by proximity.
-We test our algorithm on a number of data sets and show that it effectively identifies and separates different fiber bundles.
-This work will provide a stepping stone to introduce novel non destructive testing techniques for reinforced fiber bundles analysis.
+%Carbon fiber reinforced polymers (CFRP) typically consist of two main components: the matrix and the reinforcement. 
+%Most important in order to achieve the desired characteristics of a CFRP is the reinforcement component. 
+%In case of CFRPs the reinforcements typically consist of bundles of endless carbon fibers woven into sheets of carbon fiber %cloth, which are then layered and combined to a CFRP by the matrix component.
+%For quality control of CFRP materials and components industrial 3D X-ray computed tomography (XCT) is increasingly applied %as it allows detailed characterizations of the material in a non-destructive manner.
+
+This work introduces MetaTracts, a novel method for extracting and visualizing individual fiber bundles as well as their weaving pattern in endless carbon fiber reinforced polymer (CFRP) materials, scanned by X-ray computed tomography (XCT).
+
+For the purpose of characterizing larger areas of CFRP materials such as complete unit cells integrating the recurring weaving pattern, the proposed workflow is designed to handle XCT scans of low resolution, in which the individual fibers of the bundles to be extracted are not or just barely visible.
+The proposed workflow applies and extends methods originally presented for the field of diffusion tensor imaging in order to cope with the low data quality: 
+First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
+In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach, wherein the first step clusters according to orientation and the second step cluster by proximity of the MetaTracts.
+Finally, the tool generates volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
+
+We evaluate the proposed workflow on a number of real world datasets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 
+%mention user study here?%
+%This work will provide a stepping stone to introduce novel non destructive testing techniques for reinforced fiber bundles analysis.
 } % end of abstract
@@
 \bibliography{xBundle}
-\input {./chapters/supplementaryMaterial.tex}
+%\input {./chapters/supplementaryMaterial.tex}
 \end{document}

### 2014-09-08 chapters/MetaTracts.tex
changed images sizes
changes

@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and future components.
-The material of choice of a strongly growing number of industrial branches are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.
+The material of choice of a strongly growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry.
+CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing more and more parts made of conventional materials such as aluminum or steel.
+But besides the aerospace and automotive industry, CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.
 
-CFRP materials show characteristics as high specific stiffness, high specific strength, high corrosion resistance at low weight in comparison to conventional materials.
+In general, CFRP materials show highly desired characteristics as high specific stiffness, high specific strength, high corrosion resistance at considerably lower weight compared to conventional materials.
 At the same time using CFRPs highly complex and integrated components may be produced, which were previously impossible to manufacture.
-Increasingly, primary structures and highly loaded components are being manufactured from CFRP.
-Thus CFRPs have successfully been introduced in aeronautic and automotive applications.
-Besides aerospace and automotive industry, CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.
+For example even primary structures and highly loaded components have been increasingly manufactured from CFRP in recent years.
 
 Carbon fiber reinforced polymer components, more specifically CFRP laminates with endless carbon fibers which we focus on in this work, typically consist of two main components: 
-The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics e.g.
-in terms of stiffness, strength or durability.
+The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics.
 To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use.
@@
 Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth.
-Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength and stiffness properties. 
+Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties.
 Next, resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
@@
 In special applications additional filler materials, such as colors or other particles, may be integrated.
-After curing the resin the production process of the CFRP laminate is finished.
+The production process CFRP laminate is finished after curing the resin and the CFRP laminate is ready for further processing.
 
@@
 
-Aside ultrasonic testing, industrial 3D X-Ray computed tomography (XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. 
-XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, taken throughout a full rotation of the specimen. 
-In industrial XCT, the specimen is placed on the rotary table between X-ray source and detector and penetrated by incident X-rays of the source.
-The X-rays pass through the specimen and are attenuated by the material in the specimen.
-The detector records 2D projection image depending on its principle, e.g. by transferring the X-rays in a scintillator layer into visible light.  
-The specimen is rotated by a predefined angular step and the process is repeated~\cite{heinzl-2008-thesis}. 
+Aside ultrasonic testing, industrial 3D X-Ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
+XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which taken throughout a full rotation of the specimen.
+In industrial XCT, the typical scanning geometry foresees the specimen to be placed on the rotary table between X-ray source and detector.
+X-rays are emitted by the source in the form of a cone beam.
+When passing through the specimen the X-rays get attenuated by the material in the specimen.
+The detector records the corresponding 2D attenuation image (penetration image), depending on its principle e.g.
+by transferring the X-rays in a scintillator layer into visible light.
+To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}.
+Finally the series of penetration images is reconstructed to a 3D volumetric dataset of the specimen.
 In state of the art devices, XCT can reach voxel sizes below $500 nm$.
 XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens.
-Unfortunately, there is still a trade-off between viewport and image resolution. 
-The magnification reached within a cone beam XCT scan is determined by the specified distances
-between source and specimen as well as source and detector.
-The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport showing more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.
+Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry.
+The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector.
+The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but showing more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.
 
 In this work we focus on datasets with larger viewports and lower resolutions where the single fibers are not or barely visible.
-Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of the fiber bundles themselves instead of high resolution studies of individual fibers.
-Figure~\ref{fig:data-char} provides an example of our dataset. It also shows an example of the recurring fiber bundle pattern, called a \textit{unit cell}.
-Our work is thus motivated from the recent progress in two interrelated fields: On the one hand, fiber reinforced polymer (FRP) components has gained wide application in a variety of industries  because of FRP's superior material and physical properties in comparison to conventional materials \cite{Karpat2012}.
-On the other hand the use of industrial 3D X-Ray computed tomography (XCT) based non destructive testing of fiber reinforced components has also gained popularity~\cite{Schilling2005}.
+Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of the individual fibers.
+%short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
+Figure~\ref{fig:data-char} provides an example of our targeted datasets. It also shows an example of the recurring fiber bundle pattern, called a \textit{unit cell}.
+Our work is motivated from the recent progress in two interrelated fields: On the one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials \cite{Karpat2012}.
+On the other hand the use of industrial 3D X-Ray computed tomography (XCT) based non destructive testing of fiber reinforced components has increasingly gained popularity~\cite{Schilling2005}.
 
@@
 
-While fiber bundles are now understood as increasingly important in determining component properties,
-the tools for visualization of the internal structure have not developed at the same pace.
+While fiber bundles are now understood as highly important in determining component properties, the tools for visualization of the internal structure have not developed at the same pace.
 To the best of our knowledge, there is no current work that can resolve simple queries such as:
 \begin{itemize}
-\item{ How to visualize the geometric structure of a  particular bundle?}
-\item{ Which is the largest bundle?}
+\item{ How to extract and visualize the geometric structure of a particular bundle?}
+\item{ How to visualize the interaction between a particular pair (crossovers/braiding) or a unit cell?}
 \item{ Which bundles are in a particular orientation?}
-\item{ How to visualize the interaction between a particular pair (crossovers/braiding) or a unit cell?}
+\item{ Which is the largest or smallest bundle?}
 \end{itemize}
@@
 We detail and evaluate a method which uses visualization to quickly gain insight into our data.
-We demarcate the above queries into two, ``geometric structure" and ``spatial context".
+We demarcate the above queries into two parts, ``geometric structure" and ``spatial context".
 Geometric structure refers to the shape, size, orientation of a bundle or a group of bundles.
 Spatial context refers to how two or more bundles interact with each other.
-
 We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites.
@@
 \centering
-	\includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/histogram_10_A.PNG}
+	\includegraphics[width=0.4\textwidth]{imagesMT2014/crop-16/histogram_10_A.PNG}
 	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top).
 The unit for length is the grid cube length.
-Figure~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6.}
+Figure~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6.
+Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10.
+}
 	\label{fig:len_dist_crop16}

### 2014-09-09 chapters/MetaTracts.tex
2nd pass

@@
 \label{sec:intro}
-
 % Modern industry is constantly driving their research towards novel lightweight components.
@@
 In general, CFRP materials show highly desired characteristics as high specific stiffness, high specific strength, high corrosion resistance at considerably lower weight compared to conventional materials.
-At the same time using CFRPs highly complex and integrated components may be produced, which were previously impossible to manufacture.
-For example even primary structures and highly loaded components have been increasingly manufactured from CFRP in recent years.
+At the same time highly complex and integrated components may be produced from CFRPs, which were previously impossible to manufacture.
+For example even primary structures and highly loaded components in aeronautics are been increasingly manufactured from CFRP.
 
@@
 To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use.
-Most of these processes start with reinforcement component, weaving individual carbon fiber bundles into sheets of carbon fiber cloth in a predefined pattern.
+Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern.
 These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also refereed to as fabric.
@@
 
-The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
+The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
 Ultrasonic testing is yet the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
 On the one hand ultrasonic testing provides a quick and cost-efficient overview.
-On the other hand it shows low resolution and may generate arbitrary results, e.g.
+But on the other hand it shows low resolution and may generate arbitrary results, e.g.
 due to the geometry of the final component.
@@
 In industrial XCT, the typical scanning geometry foresees the specimen to be placed on the rotary table between X-ray source and detector.
-X-rays are emitted by the source in the form of a cone beam.
+X-rays are emitted by the source in the form of a conic beam.
 When passing through the specimen the X-rays get attenuated by the material in the specimen.
@@
 To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}.
-Finally the series of penetration images is reconstructed to a 3D volumetric dataset of the specimen.
+Finally the series of penetration images is finally reconstructed to a 3D volumetric dataset of the specimen.
 In state of the art devices, XCT can reach voxel sizes below $500 nm$.
@@
 
-In this work we focus on datasets with larger viewports and lower resolutions where the single fibers are not or barely visible.
+In this work we focus on datasets with larger viewports and lower resolutions where the single carbon fibers (filaments) are not or barely visible.
 Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of the individual fibers.
-%short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Figure~\ref{fig:data-char} provides an example of our targeted datasets. It also shows an example of the recurring fiber bundle pattern, called a \textit{unit cell}.
+%of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
+Figure~\ref{fig:data-char} provides an example of our targeted datasets. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
 Our work is motivated from the recent progress in two interrelated fields: On the one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials \cite{Karpat2012}.
-On the other hand the use of industrial 3D X-Ray computed tomography (XCT) based non destructive testing of fiber reinforced components has increasingly gained popularity~\cite{Schilling2005}.
+On the other hand, recent developments of industrial 3D X-Ray computed tomography (XCT) with regard to larger detectors, larger field of views and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
 
@@
 \begin{itemize}
-\item{ How to extract and visualize the geometric structure of a particular bundle?}
-\item{ How to visualize the interaction between a particular pair (crossovers/braiding) or a unit cell?}
-\item{ Which bundles are in a particular orientation?}
-\item{ Which is the largest or smallest bundle?}
+\item{ How to extract and visualize the geometric structure of a particular fiber bundle?}
+\item{ How to visualize the interaction between a particular pair of fiber bundles (crossovers/braiding) or a unit cell?}
+\item{ Which fiber bundles show a particular orientation, how may different orientations are prevalent in a unit cell? }
+\item{ Which fiber bundles are of the same type of yarn (i.e. which bundles show similar sizes or diameters), which is the largest or smallest fiber bundle?}
 \end{itemize}
 
-Extracting the answer to the above queries from the volume rendering of datasets, or from looking at  particular (2D) slices of the data is non-trivial even for experts.
-We detail and evaluate a method which uses visualization to quickly gain insight into our data.
+Extracting the answer to the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial even for experts.
+We detail and evaluate a method which uses visualization techniques to gain insight into our data.
 We demarcate the above queries into two parts, ``geometric structure" and ``spatial context".
-Geometric structure refers to the shape, size, orientation of a bundle or a group of bundles.
+Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles.
 Spatial context refers to how two or more bundles interact with each other.
@@
 
-Diffusion Tensor Imaging (DTI) refers to fiber tractography using diffusion  magnetic resonance (MR) images.
-It is a promising method for reconstructing the pathways of white matter in the human brain.
+Diffusion Magnetic Resonance Imagining (dMRI, also referred to as Diffusion Tensor Imaging (DTI)) has gained popularity in medical diagnosis within recent years.
+Its main clinical application is found in the study and treatment of neurological disorders.
+DTI may reveal abnormalities in white matter fiber structure and allows for visualizing the organization of white fibers in the human brain as well as the brain connectivity.
 A variety of algorithms have been proposed aiming to generate fiber-tract trajectories.
-In general these reconstructions are then clustered in to bundles which are expected to be related anatomically or spatially.
+In general, these reconstructions of fiber trajectories are then clustered in to bundles which are expected to be related anatomically or spatially.
 We broadly divide the related work on DTI into two parts which are of immediate relevance to our proposed solution: fiber tracking and fiber clustering.
@@
 \label{subsec:fiberEx} 
-Diffusion Magnetic Resonance Imagining has gained popularity in recent years as a means of visualizing the organization of white fibers in human brain.
 In detecting fiber tracts as shown by Basser et al.~\cite{Basser2000},~\cite{Basser2002} and Mori et al.~\cite{Mori2002},~\cite{Mori1999}, the key idea is based on the assumption that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel.
-In these techniques, continuous tracts are created by propagating a virtual particle from a starting point until they reach some termination criterion.
+In these techniques, continuous tracts are created by propagating a virtual particle from a starting point until they reach a termination criterion.
 This is usually done by solving a second-order Runge-Kutta integration.
@@
 Basser et al.~\cite{Basser2000},~\cite{Basser2002} proposed that white matter tracts could be represented as 3D curves in space.
-They showed that numerical method could be used to follow the fiber tracts and generate tracts in human brain data.
-
+They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data.
 Mori et al.~\cite{Mori2002} showed that tract reconstruction techniques could be broadly divided into line-propagation or energy minimization techniques.
@@
 %!!!!! TODO REFINE FIBER CLUSTERING !!!!!
-
 %The next step is to identify  anatomically relevant fiber bundles. This is achieved through fiber clustering.
 The general assumption of fiber clustering in diffusion tensor imaging is that a similarity measure based on factors such as proximity between fibers can be used to interpret a collection of fiber tracts as bundles.
-Extensive research has been done on automatic DTI fiber clustering methods (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}.
+Extensive research has been done on automatic DTI fiber clustering methods (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}).
+
 The clustering approach divides the problem into two parts.
 The first part deals with finding a suitable proximity measure and the second deals with the choice of the clustering method.
-Pairwise proximity measures include, endpoint distances~\cite{Brun2003},  mean of the closest distances between points on two fibers~\cite{Corouge2004}.
+Pairwise proximity measures include, endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}.
 Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of the above distance so that fibers that are close for certain distance and then diverge are clustered separately.
 Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors.
-The choice for the clustering algorithm can be broadly divided into those using hierarchical clustering (see ~\cite{Moberts2005},~\cite{Zhang2008}) and those using spectral clustering approach (see~\cite{jonasson2005},~\cite{ODonnell2007},~\cite{Brun2004}).
+The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering (see ~\cite{Moberts2005},~\cite{Zhang2008}) and those using spectral clustering (see~\cite{jonasson2005},~\cite{ODonnell2007},~\cite{Brun2004})
 
@@
 Jonasson et al.~\cite{jonasson2005} runs k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
-
 %The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}.
@@
 
-The difficulties in hierarchical clustering includes; first, computing all pairs distances for tracts to generate the distance matrix (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times).
+The difficulties in hierarchical clustering include: first, computing all pairs distances for tracts to generate the distance matrix (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times).
 Second, choosing the ``correct'' distance measure to compare tracts.
@@
 %!!!!! TODO REFINE SECTION !!!!!
-
 Unlike DTI, we do not have diffusion tensor data.
 Instead, we have a scalar volume with tubular structures embedded in them.
-Analyzing curvilinear structures in volumetric images has been utilized for a variety of purposes including center line extraction~\cite{Bouix2005} and vascular image enhancement Frangi et al.~\cite{Frangi1998}, Sato et al.~\cite{Sato1997}.
+Analyzing curvilinear structures in volumetric images has been utilized for a variety of purposes including center line extraction~\cite{Bouix2005} and vascular image enhancement as proposed by Frangi et al.~\cite{Frangi1998} and Sato et al.~\cite{Sato1997}.
 Frangi et al.~\cite{Frangi1998} introduced a method based on studying the eigenvalues of the local second order structure namely the Hessian matrix, specifically for the purposes of developing vessel enhancement filters.
@@
 \subsection {Visual Analysis of Fiber Reinforced Composites}
-The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores.
-Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
+The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores: Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
 This approach allows analyzing fiber orientations based on direction transfer functions.
@@
 Reh et al.~\cite{PMI_AR_2012} introduced an approach to explore pores of carbon fiber reinforced composites.
-Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for  interactive exploration of fibers in reinforced polymers.
+Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers.
 
@@
 
-\begin{figure}
-\includegraphics[width=0.5\textwidth,clip=true, trim= 0cm 0cm 0cm 0cm]{imagesMT2014/MT_data_4_2.png}
+\begin{figure}[tb]
+\centering
+\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
 \caption{Data Characteristics.
-A: Shows a rendering of the data.
-B: Shows 2D slices along the X (Right) and Z (Left) axis.
-C: Shows a zoom of the green square, multiple bundles cross and are indistinguishable by visual inspection alone.
+(A) shows a rendering of the data.
+(B) shows 2D slices along the X and (C) along the Z axis.
+(D) close up of the region marked in green in (C), multiple fiber bundles cross and are indistinguishable by visual inspection alone.
 }
@@
 
-Figure \ref{fig:data-char} shows one of our data sets with woven fiber bundles.
-The dataset is 450x300x500 voxels in size and clearly shows the weaving structure of the composite unit cell (a regional subset of the recurring fiber bundle weaving pattern).
+Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles.
+Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$.
+It clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
 The unit cell is used as a basis structure for CFRP manufacturing.
-The alignment and weave pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
+The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
 Figure~\ref{fig:data-char}A shows a volume rendering of the dataset.
-Figure~\ref{fig:data-char}B shows 2D slices along the X(right) and Z(left) axis.
-Figure~\ref{fig:data-char}C shows the zoomed in version of the yellow square.
-At isotropic resolution of 2 micrometers, the individual carbon fibers in a fiber bundle are hard to resolve from the 2D slices.
-Figure~\ref{fig:data-char}C contains multiple bundles going in opposite directions.
-The separation between two bundles is barely discriminable.
-Carbon fibers themselves do not differ as much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. 
+B shows 2D slices along the X(right) and C in Z(left) axis.
+Figure~\ref{fig:data-char}D shows the zoomed in version of the green square.
+The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only.
+Figure~\ref{fig:data-char}D contains multiple bundles going in opposite directions.
+However, the separation between two fiber bundles is barely visible.
+The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization.
 The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}A shows the large variation in cross section sizes among the bundles.
-Depending on the weaving pattern the bundles cross each other in different orientations. 
+Depending on the weaving pattern the fiber bundles cross each other in different orientations.
 In general the number of orientations is limited by the weaving pattern.
@@
 Nevertheless, the weaving pattern may cause the individual bundles to be curved.
-In consequence of the weaving pattern of the fiber bundles, fibers may be adjacent in euclidean distance but belong to different bundles.
-
-We make the following assumptions on our data.
+In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in euclidean distance but belong to different bundles.
+We make the following assumptions on our data:
 \begin{itemize}
 \item The structure embedded in the data contains fiber bundles of indiscernible fibers.
-\item (Local orientation)Each point within a fiber  bundle has a local orientation which is parallel to the bundle.
-\item The local orientation changes gradually inside the bundle.
-\item Local orientation may be noisy and not reliable.
+\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
+\item The local orientation may change gradually inside the fiber bundle.
+\item The local orientation may be noisy and not reliable.
 \item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
-\item Bundles going in different directions only interact near the surface of contact.
+\item Fiber bundles going in different directions only interact near the surface of contact.
 \end{itemize}
@@
 
-\begin{figure}
+\begin{figure}[tb]
 \centering
-\includegraphics[width=0.5\textwidth]{imagesMT2014/algo.png}
-\caption{Flow chart of our approach}
+\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
+\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
 \label{fig:flowchart}
 \end{figure}
+
 Figure \ref{fig:flowchart} shows our pipeline.
@@
 %We explain the different steps of MetaTracts extraction.
+
 \subsection {Reliable Hessians}
@@
 The principal directions, in which the local second order structure of the image can be decomposed are given by the eigen decomposition of the Hessian matrix which describes the local curvature.
-
 The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.
@@
 Here $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigen values of the Hessian matrix.
-Specifically, a  pixel belonging to vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$).
+Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$).
 The sign indicates if the vessel is bright in a dark background or dark in a bright background.
 In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$).
-The following measures are defined in ~\cite{Frangi1998}.
+The following measures are defined in their work~\cite{Frangi1998}.
 \begin{equation}\label{RA}
@@
 \end{equation}
-In equation~\ref{RB}, $\mathcal{R_{B}}$ provides a measure of deviation from blob like structure while in equation~\ref{RA}, $\mathcal{R_{A}}$ distinguishes between plate-like and line-like structure.
+In equation~\ref{RB}, $\mathcal{R_{B}}$ provides a measure of deviation from a blob like structure while in equation~\ref{RA}, $\mathcal{R_{A}}$ distinguishes between plate-like and line-like structure.
 Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors.
@@
 $$
-
-Variable $s$ is the Frobenius norm of the Hessian matrix. The value  of $(1-e^{\frac{-s^2}{2c^2}})$ will be low in regions with no structure.
-
+Variable $s$ is the Frobenius norm of the Hessian matrix.
+The value  of $(1-e^{\frac{-s^2}{2c^2}})$ will be low in regions with no structure.
 The utility of the vesselness is a little different in our framework than Frangi et al.~\cite{Frangi1998}.
@@
 Instead, we are trying to associate each grid location  with a probable orientation based on its local second order structure.
-The $R_{H}$ is interpreted as a measure reliability of the local orientation.
-
+The $R_{H}$ is interpreted as a reliability measure of the local orientation.
 Grid locations where the $R_{H}$ is above a certain threshold are marked as regions with reliable orientations.
-We use 0.3 as the cutoff for all our test cases. (See Section~\ref{sec:param_choices} on the choice of 0.3)
-
+We use 0.3 as the cutoff for all our test cases. (See Section~\ref{sec:param_choices} on the choice of 0.3).
 %\begin{figure}[h]
@@
 %\end{figure}
-
 %\begin{figure}
@@
 %\end{figure}
-%
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
-\caption{Reliable Hessians. A: Colored according to Orientation vector mapped to RGB. B: Slice along Z axis. C: Slice along X axis. D: close up of the region marked in yellow in C.}
-\label{fig:reliable_hessian}
-\end{figure}
- In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the threshold.
+In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold.
 The unit vector representing the principal direction has been mapped to RGB color space.
 Figure~\ref{fig:reliable_hessian}A shows the entire data set.
-Regions where the principal direction is parallel to the X axis are red in color and those parallel to the Z axis are blue.
-Figure~\ref{fig:reliable_hessian}B shows 2D slices along the Z and X axis respectively.
-Figure~\ref{fig:reliable_hessian}C shows a close up (region similar to Figure~\ref{fig:data-char}C).
+Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue.
+Figure~\ref{fig:reliable_hessian}B and C show 2D slices along the Z and X axis respectively.
+Figure~\ref{fig:reliable_hessian}D shows a close up (region similar to Figure~\ref{fig:data-char}D).
 Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold.
@@
 Thus techniques popular in DTI cannot be directly applied in XCT data.
+
+%
+\begin{figure}[tb]
+\centering
+\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
+\caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) close up of the region marked in green in C.}
+\label{fig:reliable_hessian}
+\end{figure}
 
@@
 We build on the key assumptions on the data, namely `local orientation' and `connectivity' while taking into account the noise and lack of resolution.
-We do this by interpreting the underlying geometric structure of the fibers as a set of polylines(cylinders).
-
- MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of  cylindrical tubes in $\mathbb{R}^3$ traversing the bundles embedded in the input data.
-MetaTracts share the following characteristics:
+We do this by interpreting the underlying geometric structure of the fibers as a set of polylines (cylinders).
+MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data.
+MetaTracts share the following \textit{properties}:
 
@@
 \item{Each MetaTract is associated with a start point which is a grid vertex.}
-\item{Each  individual cylinder in a MetaTract has a constant length,  constant radius and a start point (which is also a grid vertex).}
-\item{Each cylinder in a MetaTract (except for the first one)is connected to the previous cylinder at the start point.}
+\item{Each cylinder in a MetaTract has a constant length, a constant radius and a start point (which is also a grid vertex).}
+\item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
 \item{Each cylinder is locally parallel to the local orientation at it's start point.}
@@
 \subsection {MetaTracts generation}
-
+\label{subsec:MetaTracts-generation}
 In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles.
@@
 The procedure trivially extends to $\mathbb{R}^3$.
-The reliable Hessian procedure  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1].
+The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1].
 In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in green.
@@
 The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the blue arrow (Figure~\ref{fig:algo}C).
-We generate a rectangle of length L and radius R, the blue rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}).
-
-\begin{figure}[h]
+We generate a rectangle of length L and radius R, the blue dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}).
+\begin{figure}[tb]
   \begin{center}
-  \includegraphics[ width=0.4\textwidth]{imagesMT2014/MT_abstract} 
+  \includegraphics[ width=0.45\textwidth]{imagesMT2014/MT_abstract} 
   \end{center}
- \caption{MetaTracts generation. B. MetaTract generation, region with $R_H$ less than threshold. A: close up of marked region in A. C: Parts of a single MetaTract.}
+ \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
   \label{fig:algo}
-
 \end{figure}
-
-\textit{Candidate vertices}: The set of vertices which are in the blue region but not overlapped by green are possible candidates for the start point of the next cylinder in a particular MetaTract.
-We call these vertices candidate vertices.
+The set of vertices which are in the blue region but not overlapped by green are possible candidates for the start point of the next cylinder in a particular MetaTract.
+We call these vertices \textit{``candidate vertices"}.
 From these we select another grid vertex which will be the start point for the next cylinder.
@@
 %\end{figure}
-The order of the candidate vertices is based on the following characteristics.
+The order of the candidate vertices is based on the following characteristics:
+
 \begin{itemize}
 \item \textit{Orientation Similarity}: We want the orientation of the start points for the consecutive cylinders to have a similar orientations ($N_p$). 
-
 \item \textit{Large distance}: We want the MetaTracts to traverse the data using as few cylinders as possible.
@@
 For example the Euclidean distance between the blue and the orange vertices projected onto $N_P$ gives distance (Figure~\ref{fig:algo}B).
-We refer to this perpendicular projection distance as \textit{dist}.
+We refer to this perpendicular projection distance as \textit{``dist"}.
 \end{itemize}
@@
 % \end{equation}
-We define a priority  for each of the candidate vertices based on the above factors.
-For each cylinder in a MetaTracts, we put the \textit{candidate vertices} of it in a priority queue based on Equation \ref{eqn:algo_1}.
+We further define a priority for each of the candidate vertices based on the above factors.
+For each cylinder in a MetaTract, we put the \textit{candidate vertices} of it in a priority queue based on Equation \ref{eqn:algo_1}.
 \begin{equation}
@@
 At each iteration we pick the top element in the priority queue generate the corresponding cylinder and repeat the steps.
-Essentially, Eqn.~\ref{eqn:algo_1}, selects a grid point which is the furthest from the current point and going in a similar direction.
+Essentially, Equation~\ref{eqn:algo_1}, selects a grid point which is the furthest from the current point and going in a similar direction.
 This gives it the advantage of tackling noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation (for orientation) in an already noisy environment.
 
- In this particular example Figure~\ref{fig:algo}B we select the orange grid vertex next and repeat the process.
+In this particular example in Figure~\ref{fig:algo}B we select the orange grid vertex next and repeat the process.
 We do not select the purple vertex (Figure~\ref{fig:algo}C) because it is not a candidate vertex.
 If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length.
-Short tracts are then remove.
-The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}(A).
+Short tracts are then removed.
+The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}A.
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
@@
 
-\begin{figure}[h]
+\begin{figure}[tb]
 \centering
-	\includegraphics[width=0.4\textwidth]{imagesMT2014/crop-16/length_dist.png}
+	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
 	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.}
@@
 We use hierarchical clustering which has enjoyed considerable attention and success in DTI field.
-
-However, the MetaTracts produced as described in Section 4 often do not extend the full length of the fiber bundle.  Thus, MetaTracts
-in a given fiber bundle may have very different lengths and may only partially overlap Figure~\ref{fig:length_distribution}(B).
-
-Moreover, Carbon fibers have both orientation and proximity information which can be used for clustering. Instead of creating some formula
-which artificially combined the two measures,  we first cluster based on orientation and then further subdivide each orientation cluster using proximity information.
-We found that different clustering methods were preferable
-for the different measures. We use dimension reduction followed by K-means clustering
-for orientation clustering while we use hierarchical clustering to further subdivide each cluster
-based on proximity measures. (See Section~\ref{sec:results} for experimental results.) Breaking clustering into two steps also helps us and potential users
-to better understand the results of the clustering.
+However, often the MetaTracts produced in Section~\ref{sec:approach} do not extend the full length of the fiber bundle.
+Thus, MetaTracts in a given fiber bundle may have very different lengths and may only partially overlap as in Figure~\ref{fig:length_distribution}B.
+Moreover, carbon fibers have both orientation and proximity information which can be used for clustering.
+Instead of creating an heuristic, which artificially combines the two measures, we first cluster based on orientation and then further subdivide each orientation cluster using proximity information.
+We found that different clustering methods were preferable for the different measures.
+We use dimension reduction followed by K-means clustering for orientation clustering while we use hierarchical clustering to further subdivide each cluster based on proximity measures (see Section~\ref{sec:results} for experimental results).
+Breaking clustering into two steps also helps us and potential users to better understand the results of the clustering.
 
@@
 This is not the case for noisy data such as ours.
-
 %In Figure~\ref{fig:hclust_issue_a} proximity based distance, for example mean of the minimum distances between points on MetaTracts might falsely find the red MetaTract being closer to the blue, than the orange MetaTract.
@@
 %we first cluster based on orientation using a simple measure.
-
 %We use domain knowledge to divide the clustering into two intuitive parts:because of the woven nature of fiber bundles in our application domain we first cluster in terms of orientation.
@@
 
-\subsection {Orientation bundles}
+\subsection {Orientation based clustering}
 %We use the inherent features expressed as weaving pattern of our data to pre-cluster the metaTracts into the number of classes based on the major directions of the pattern.
-
 %It is important to note that this step broadly divides the MetaTracts into classes based on the major orientations and not individual fiber bundles. 
-This step broadly divides the MetaTracts into classes based on the major orientations.
-In order to cluster MetaTracts going in same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01} and later used in DTI fiber coloring~\cite{Brun2003} and fiber clustering Brun et al.~\cite{Brun2004}.
+This step divides the individual MetaTracts into classes based on their major orientations.
+In order to cluster MetaTracts going in same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01} and later used in DTI fiber coloring~\cite{Brun2003} and fiber clustering~\cite{Brun2004}.
 The key notion is to find a suitable similarity measure to define the weights of an undirected weighted graph.
 The nodes being the data points and the edges being the weights.
-Then solving an eigenvalue problem which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
+Then, an eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
 This lower dimensional space can then be clustered more efficiently than the higher dimensional space. 
@@
 
-\subsubsection{Orientation based similarity measure}
-\label{subsec:ori-sim-mes}
-Brun et al.~\cite{Brun2003} based their measure on the simple assumption that two traces with similar endpoints should 
-be considered similar. Brun et al. ~\cite{Brun2004} based their measure on a 9-D fiber descriptors.
+Brun et al.~\cite{Brun2003} based their measure on the simple assumption that two traces with similar endpoints should be considered similar.
+Brun et al.
+~\cite{Brun2004} based their measure on a 9-D fiber descriptors.
 In our case, each MetaTract is a data point.
@@
 Using the Belkin and Niyogi algorithm we ``embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix.
-
+%
 %August 13 2014
@@
 %For all our test cases we set $\eta$ to be five (see sec~\ref{sec:param_choices} for parameter choices for $\eta$.).
+%  
+We used conventional K-means for clustering this lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
+For our test case there are two major directions of the fiber bundles.
+So $K$ was set to 2.
+It is important to note that due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
+Figure~\ref{fig:orientation_clustering}B shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
+As is expected there is a clear distinction based on fiber orientation.
+Figure~\ref{fig:orientation_clustering}(A,C) shows the MetaTracts colored according to clustering results.
 
-  \begin{figure}
+\begin{figure}[tb]
   \centering
@@
 %  	
- \includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
-  	\caption{(C) Shows the result of K means clustering with data points projected to the top three eigenvectors as major axes.
-(A) shows the MetaTracts projected on top three eigenvectors.
-(B) shows 2D slices of C.
-}
+ \includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
+  	\caption{(C) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts projected on top three eigenvectors. (B) 2D slices of (C). }
   \label{fig:orientation_clustering}
   \end{figure}
-
-   We used conventional K-means clustering to cluster the lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
-For our test case there are two major directions of the fiber bundles.
-So $K$ was set to 2.
-It is important to note that due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
-Figure~\ref{fig:orientation_clustering}B shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
-As is expected there is a clear distinction based on fiber orientation.
-Figure~\ref{fig:orientation_clustering}(A,C) shows the MetaTracts colored according to clustering results.
-Orientation clustering uses only orientation information. To subdivide the oriented clusters into fiber bundles,
-we need to include information about the geometric proximity between MetaTracts.
 
@@
  \label{subsec:dist_clustering}
+Orientation clustering uses only orientation information.
+To subdivide the oriented clusters into fiber bundles, we need to include further information about the geometric proximity between MetaTracts.
 We use the directed Hausdorff distance for distance based clustering.
- Each MetaTract is represented as a set of points ($C_P$). 
- Formally, the directed Hausdorff distance from point set P to point set Q is defined as 
- $H_{dir}(P,Q) = max_{p \in P} min_{q \in Q} d(p,q).$ The Hausdorff distance is defined as $H(P,Q) = max(H_{dir}(P,Q),H_{dir}(Q,P))$.
-
- The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$
- but the directed Hausdorff is not.
- Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially cover the bundle (figure~\ref{fig:length_distribution}(B)).
- If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$,
- then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
- Thus, $H(P,Q)$ will be large, even though $P$ and $Q$ are in the same
- fiber bundle.
-
+Each MetaTract is represented as a set of points ($C_P$). Formally, the directed Hausdorff distance from point set P to point set Q is defined as 
+$H_{dir}(P,Q) = max_{p \in P} min_{q \in Q} d(p,q)$ .
+The Hausdorff distance is defined as $H(P,Q) = max(H_{dir}(P,Q),H_{dir}(Q,P))$.
+The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$ but the directed Hausdorff is not.
+Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially covers the bundle (Figure~\ref{fig:length_distribution}(B)).
+If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$, then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
+Thus, $H(P,Q)$ will be large, even though $P$ and $Q$ are in the same fiber bundle.
 Instead of using the Hausdorff distance, $\max(H_{dir}(P,Q),H_{dir}(Q,P)$, we use $\min(H_{dir}(P,Q),H_{dir}(Q,P))$.
@@
 In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
- Clustered are merged until there are only $h$ clusters left.
+Clusters are merged until there are only $h$ clusters left.
 Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
 Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
-
 We found that single linkage hierarchical clustering had two major drawbacks.
@@
 These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
- Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same
- fiber bundle.
- A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles,
- causing them to be clustered into a single bundle.
-
+Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
+A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
 To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
 We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
-
 To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
 shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
-
 We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters are shown in Fig.~\ref{fig:crop-16-decomp}.
@@
 \label{sec:vis}
-Apart from direct visualization of the MetaTracts, we show two simple extensions. The first way is to voxelize the original volume according to the clusters each voxel is associated with.
+Apart from direct visualization of the MetaTracts, we show two simple but yet highly important extensions for our domain specialist.
+The first one is to voxelize the original volume according to the clusters each voxel is associated with.
 The second is to extract the corresponding surfaces from the voxel data by binarizing the volume per cluster and extracting the isosurface of the largest connected component from the binary volume.
+Both methods are required for further analysis of the data, e.g.
+in simulations.
 To voxelize the space,  we compute a neighborhood around each voxel.
 We then enumerate the number of voxels of each class (cluster) in this neighborhood.
-The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Fig.~\ref{fig:crop-16-decomp}C and Fig~\ref{fig:prepreg}D shows the result of voxelization.
-Fig.~\ref{fig:crop-16-decomp}(E, F) show examples of extracted meshes.
+The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Figure~\ref{fig:crop-16-decomp}C and Figure~\ref{fig:prepreg}D shows the result of voxelization.
+Figure~\ref{fig:crop-16-decomp}(E, F) show examples of extracted meshes.
 %The MetaTracts and the clustering provide an abstraction of the fiber bundles and are used as a visualization tool.
@@
 	\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
-	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of B.
+	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of (B).
 (D) shows a 2D slice of the volume.
@@
 \end{figure*}
+
 \begin{figure}
 \centering
-	\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6.PNG}
+	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-6.PNG}
 	\caption{Data set with flat thin and compact bundles.
-A shows the volume rendering and a 2D slice with one of the boundaries marked in red, B shows the clusters according to individual orientation.C shows the complete result.
-D shows the voxelization of C.}
+(A) shows the volume rendering and a 2D slice with one of the boundaries marked in red, (B) shows the clusters according to individual orientation.
+(C) shows the complete result.
+(D) shows the voxelization of (C).}
 	\label{fig:prepreg}
@@
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
-As noted in section~\ref{subsec:dist_clustering} the tracts donot run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length).
+As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length).
 Fig.~\ref{fig:length_distribution}(B) shows the length per MetaTract distribution for a particular fiber bundle.
-
 Fig.~\ref{fig:prepreg} shows the results on a dense data set with \textit{flat and thin} bundles.
@@
 Fig.~\ref{fig:prepreg}C shows the combined results.Fig.~\ref{fig:prepreg}D shows the volume after voxelization.
-The fiber bundle extraction using MetaTracts was implemented in ITK and C++.
+The fiber bundle extraction using MetaTracts was implemented in C++ using ITK.
 All clustering was done in R.
-The preprocessing consists of metaTract generation the distance computation and the orientation clustering.
-The hierarchical clustering separately, on the results of the orientation clusters can be done  quickly in a few minutes on a Intel Xeon E5-2667 workstation PC according to user input `h'(section~\ref{sec:param_choices}).
+The preprocessing consists of MetaTract generation the distance computation and the orientation clustering.
+The hierarchical clustering separately, on the results of the orientation clusters can be done quickly in a few minutes on a Intel Xeon E5-2667 workstation according to user input `h'(section~\ref{sec:param_choices}).
 We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without extra computation.
@@
 \caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
-(b)Shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
+(b) shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
 \label{fig:context}
 \end{figure}
+
 \section{Parameter choices}
@@
 The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering.
-K denotes the number of major directions.
-Which is generally known a priori or can be easily estimated by looking at the weaving pattern.
+K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern.
 Our framework is robust to the choice of $h$ (Section~\ref{sec:results}).
@@
 This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
-
 The following parameters were fixed for all the tests.
@@
 We show that our framework works at comparatively low resolution and with dense fiber arrangements (when extracting single fibers might not be possible).
-It handles complex fiber patterns such as ``cross overs''/``braiding''.
+It handles complex fiber patterns such as ``cross overs'' and ``braiding''.
 The presented techniques are attractive and novel features for a industrial fiber bundle visualization tool.
-
 %
@@
 %% if specified like this the section will be ommitted in review mode
-% %\acknowledgements{
-% The research leading to these results has received funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM}
+
+%\acknowledgements{
+% The research leading to these results has received funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM.
+Our special thanks go to Ilya Ilya Straumit (KU Leuven) who provided us with dataset1.
+%}

### 2014-09-11 chapters/MetaTracts.tex
dell8 local changes
Updated images

@@
 \centering
-\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
+%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
+\includegraphics[width=0.5\textwidth]{imagesMT2014/image1}
 \caption{Data Characteristics.
@@
 Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles.
-Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$.
+Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the datatype is uint8.
 It clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
@@
 \begin{figure}[tb]
-\centering
-\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
+%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
+\includegraphics[width=0.5\textwidth]{imagesMT2014/image3.png}
 \caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
 \centering
-\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
+%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
+\includegraphics[width=0.5\textwidth]{imagesMT2014/image2}
 \caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) close up of the region marked in green in C.}
@@
   \begin{center}
-  \includegraphics[ width=0.45\textwidth]{imagesMT2014/MT_abstract} 
+   \includegraphics[ width=0.45\textwidth]{imagesMT2014/image4}
+ % \includegraphics[ width=0.45\textwidth]{imagesMT2014/MT_abstract} 
   \end{center}
@@
 \centering
-	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
+	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
+		\includegraphics[width=0.45\textwidth]{imagesMT2014/image5}
 	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.}
@@
 This lower dimensional space can then be clustered more efficiently than the higher dimensional space. 
-Let G be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem:
-\begin{equation}\label{equn:eigenMaps}
-L\textbf{f}=\lambda D\textbf{f}
-\end{equation}
+Let G be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem:$L\textbf{f}=\lambda D\textbf{f}$,
+%\begin{equation}\label{equn:eigenMaps}
+%L\textbf{f}=\lambda D\textbf{f}
+%\end{equation}
 where D is the diagonal weight matrix and L is the Laplacian matrix.
@@
   \centering  	
-  	%\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}~
-  	%	\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}
-%  	\captionsetup[subfigure]{labelformat=empty}
-%  	\subfloat[1]{\includegraphics[width=0.2\linewidth]{imagesMT2014/orientation_clustering_A.png}}
-%  	\subfloat[2]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_1.PNG}}
-% 	\subfloat[3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_2.PNG}}
-% 	\subfloat[4]{\includegraphics[width=0.2\linewidth]{imagesMT2014/curveD.PNG}}
-%  	
- \includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
-  	\caption{(C) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts projected on top three eigenvectors. (B) 2D slices of (C). }
+ %\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
+  \includegraphics[width=0.45\textwidth]{imagesMT2014/image6}
+  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes.
+(A) MetaTracts belonging to orientation cluster 1.
+(B) MetaTracts belonging to orientation cluster 2.
+MetaTracts in gray(A,C) show context.
+}
   \label{fig:orientation_clustering}
@@
 We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters are shown in Fig.~\ref{fig:crop-16-decomp}.
+\begin{figure*}
+\centering
+		\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
+		\includegraphics[width=0.4\textwidth]{imagesMT2014/image7b}
+	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
+	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
+	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of (B).
+(D) shows a 2D slice of the volume.
+(E,F) shows the extracted meshes.}
+	\label{fig:crop-16-decomp}
+\end{figure*}
 
@@
 %\end{figure*}
-\begin{figure*}
+
+\begin{figure}[t]
 \centering
-	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
-	\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
-	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of (B).
-(D) shows a 2D slice of the volume.
-(E,F) shows the extracted meshes.}
-	\label{fig:crop-16-decomp}
-\end{figure*}
-
-\begin{figure}
-\centering
-	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-6.PNG}
+	\includegraphics[width=0.45\textwidth]{imagesMT2014/image8.PNG}
 	\caption{Data set with flat thin and compact bundles.
-(A) shows the volume rendering and a 2D slice with one of the boundaries marked in red, (B) shows the clusters according to individual orientation.
+(A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation.
 (C) shows the complete result.
@@
 Fig.~\ref{fig:length_distribution}(B) shows the length per MetaTract distribution for a particular fiber bundle.
-Fig.~\ref{fig:prepreg} shows the results on a dense data set with \textit{flat and thin} bundles.
-The bundles are indistinguishable in the original data set (Fig.~\ref{fig:prepreg}A).
-A red dotted line shows one of the fiber bundle boundary.
-Fig.~\ref{fig:prepreg}B shows the fiber bundles for the two orientation.
-Fig.~\ref{fig:prepreg}C shows the combined results.Fig.~\ref{fig:prepreg}D shows the volume after voxelization.
+Fig.~\ref{fig:prepreg} shows the results on our dataset2.
+This dataset is dense with \textit{flat and thin} bundles.
+The dimensions are 300x350x300 and the datatype is uint16.
+The bundles are indistinguishable in the original data set Fig.~\ref{fig:prepreg}(A).
+A green dotted line shows one of the fiber bundle boundary.
+Fig.~\ref{fig:prepreg}(B) shows the fiber bundles for the two orientation.
+Fig.~\ref{fig:prepreg}(C) shows the combined results.Fig.~\ref{fig:prepreg}(D) shows the volume after voxelization.
 The fiber bundle extraction using MetaTracts was implemented in C++ using ITK.
@@
 \label {sec:evaluation}
-\begin{figure}[h]
-\centering
-\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
-\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
-\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
-(b) shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
-\label{fig:context}
-\end{figure}
+%\begin{figure}[h]
+%\centering
+%\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
+%\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
+%\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
+%(b) shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
+%\label{fig:context}
+%\end{figure}
 

### 2014-09-15 chapters/MetaTracts.tex
minor changes
updated images
updated text

@@
 
-Carbon fiber reinforced polymer components, more specifically CFRP laminates with endless carbon fibers which we focus on in this work, typically consist of two main components: 
+Typically,  carbon fiber reinforced polymer components, more specifically CFRP laminates consists of two main components:
+
 The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics.
@@
 In special applications additional filler materials, such as colors or other particles, may be integrated.
-The production process CFRP laminate is finished after curing the resin and the CFRP laminate is ready for further processing.
+%The production process CFRP laminate is finished after curing the resin and the CFRP laminate is ready for further processing.
 
@@
 \item{ How to visualize the interaction between a particular pair of fiber bundles (crossovers/braiding) or a unit cell?}
-\item{ Which fiber bundles show a particular orientation, how may different orientations are prevalent in a unit cell? }
+\item{ Which fiber bundles show a particular orientation? }
 \item{ Which fiber bundles are of the same type of yarn (i.e. which bundles show similar sizes or diameters), which is the largest or smallest fiber bundle?}
@@
 The general assumption of fiber clustering in diffusion tensor imaging is that a similarity measure based on factors such as proximity between fibers can be used to interpret a collection of fiber tracts as bundles.
-Extensive research has been done on automatic DTI fiber clustering methods (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}).
+Extensive research has been done on automatic DTI fiber clustering methods
+% (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}).
 
@@
 The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
-Figure~\ref{fig:data-char}A shows a volume rendering of the dataset.
-B shows 2D slices along the X(right) and C in Z(left) axis.
-Figure~\ref{fig:data-char}D shows the zoomed in version of the green square.
+Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset.
+Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis.
+Figure~\ref{fig:data-char}(D) shows the zoomed in version of the green square.
 The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only.
-Figure~\ref{fig:data-char}D contains multiple bundles going in opposite directions.
+Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions.
 However, the separation between two fiber bundles is barely visible.
 The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization.
-The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}A shows the large variation in cross section sizes among the bundles.
+The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
 Depending on the weaving pattern the fiber bundles cross each other in different orientations.
@@
 \begin{figure}[tb]
+\centering
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
-\includegraphics[width=0.5\textwidth]{imagesMT2014/image3.png}
+\includegraphics[width=0.5\textwidth]{imagesMT2014/image3b.png}
 \caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
 Figure \ref{fig:flowchart} shows our pipeline.
-The approach consists of multiple steps which we divide into two broad phases, extracting the MetaTracts  (this section) and  finding plausible fiber bundles (Section~\ref{subsec:fiber-bundles}).
+The approach consists of multiple steps.
+MetaTracts generations Sec.~\ref{sec:approach} and  finding plausible fiber bundles (Section~\ref{subsec:fiber-bundles}).
 %\subsection {Extracting MetaTracts}
@@
 The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1].
-In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in green.
+In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue.
 Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}).
-The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the blue arrow (Figure~\ref{fig:algo}C).
-We generate a rectangle of length L and radius R, the blue dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}).
+The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow (Figure~\ref{fig:algo}C).
+We generate a rectangle of length L and radius R, the green dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}).
 \begin{figure}[tb]
@@
   \end{center}
+  \vspace{-1em}
  \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
   \label{fig:algo}
+    \vspace{-1.5em}
 \end{figure}
-The set of vertices which are in the blue region but not overlapped by green are possible candidates for the start point of the next cylinder in a particular MetaTract.
+The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract.
 We call these vertices \textit{``candidate vertices"}.
@@
 
-In this particular example in Figure~\ref{fig:algo}B we select the orange grid vertex next and repeat the process.
-We do not select the purple vertex (Figure~\ref{fig:algo}C) because it is not a candidate vertex.
+In this particular example in Figure~\ref{fig:algo}(B) we select the orange grid vertex next and repeat the process.
+We do not select the purple vertex (Figure~\ref{fig:algo}(C)) because it is not a candidate vertex.
 If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length.
@@
  %\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
-  \includegraphics[width=0.45\textwidth]{imagesMT2014/image6}
+  \includegraphics[width=0.55\textwidth]{imagesMT2014/garb2}
   	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes.
@@
 The directed Hausdorff distance is very sensitive to outliers in the data.
-However, because MetaTracts are constructed using cylinders with similar orientations, they are not plagued by outliers.
+However, because MetaTracts after orientation clustering are constructed using cylinders with similar orientations, they are not plagued by outliers.
 To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
@@
 shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
-We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters are shown in Fig.~\ref{fig:crop-16-decomp}.
-\begin{figure*}
+We repeated both until a steady state of clusters is reached and no new small fibers can be removed.
+The results of clustering the orientation clusters Fig.~\ref{fig:orientation_clustering}(A,C) are shown in Fig.~\ref{fig:orientation_clustering}(D,E) respectively.
+Note how the purple cluster Fig.~\ref{fig:orientation_clustering}(E) bundle ``7'' is extracted well.
+\begin{figure}
 \centering
-		\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
-		\includegraphics[width=0.4\textwidth]{imagesMT2014/image7b}
+		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
+		\includegraphics[width=0.5\textwidth]{imagesMT2014/vol}
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
@@
 	\label{fig:crop-16-decomp}
-\end{figure*}
+\end{figure}
 

### 2014-09-15 metaTracts-pvis2014.tex
2nd pass
updated text

@@
 First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
-In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach, wherein the first step clusters according to orientation and the second step cluster by proximity of the MetaTracts.
-Finally, the tool generates volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
+In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach, wherein the first step clusters according to orientation and the second step clusters by proximity of the MetaTracts.
+Finally, the tool can also  generate volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
 

### 2014-09-15 chapters/MetaTracts.tex
updated text

@@
 		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
-		\includegraphics[width=0.5\textwidth]{imagesMT2014/vol}
+		\includegraphics[width=0.5\textwidth]{imagesMT2014/image7b}
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
 	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
-	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of (B).
-(D) shows a 2D slice of the volume.
-(E,F) shows the extracted meshes.}
+	\caption{(A) Volume rendering of data set Figure~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
 	\label{fig:crop-16-decomp}
@@
 \label{sec:vis}
-Apart from direct visualization of the MetaTracts, we show two simple but yet highly important extensions for our domain specialist.
+Apart from direct visualization of the MetaTracts, we show two simple yet highly important extensions for our domain specialist.
 The first one is to voxelize the original volume according to the clusters each voxel is associated with.
 The second is to extract the corresponding surfaces from the voxel data by binarizing the volume per cluster and extracting the isosurface of the largest connected component from the binary volume.
-Both methods are required for further analysis of the data, e.g.
+Both methods can possibly be used for further analysis of the data, e.g.
 in simulations.
@@
 We then enumerate the number of voxels of each class (cluster) in this neighborhood.
-The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Figure~\ref{fig:crop-16-decomp}C and Figure~\ref{fig:prepreg}D shows the result of voxelization.
-Figure~\ref{fig:crop-16-decomp}(E, F) show examples of extracted meshes.
+The voxel is then assigned to the class with the maximum number of elements in the neighborhood.
+Figure~\ref{fig:crop-16-decomp}A and Figure~\ref{fig:prepreg}D shows the result of voxelization.
+Figure~\ref{fig:crop-16-decomp}(C, D) show examples of extracted meshes.
 %The MetaTracts and the clustering provide an abstraction of the fiber bundles and are used as a visualization tool.
@@
 We tested our method on multiple datasets with varying characteristics. 
-Figure~\ref{fig:crop-16-decomp} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into 10 clusters (the ground truth is 6 and 5 respectively).
+Figure~\ref{fig:orientation_clustering} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into 10 clusters (the ground truth is 6 and 5 respectively).
 The hierarchical clustering is robust to over segmentation.
@@
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
-As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length).
+As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 Fig.~\ref{fig:length_distribution} is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length).
 Fig.~\ref{fig:length_distribution}(B) shows the length per MetaTract distribution for a particular fiber bundle.
@@
 This dataset is dense with \textit{flat and thin} bundles.
-The dimensions are 300x350x300 and the datatype is uint16.
+The dimensions are 300x350x300 and the datatype is uint16 (compared to dataset1 which was only uint8).
 The bundles are indistinguishable in the original data set Fig.~\ref{fig:prepreg}(A).
@@
 \label {sec:evaluation}
+We integrated MetaTracts into a tool and created a user evaluation questionnaire which we distributed.
 %\begin{figure}[h]
@@
 Large $h$ keeps the major bundles intact.
-For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4 and 5 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements).
+For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements).
 This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.

### 2014-09-18 chapters/MetaTracts.tex
updating images
Updated with Artem's images
revise update
revise update
Minor local commit
Minor local commit
local commit
Added new images

@@
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
-\includegraphics[width=0.5\textwidth]{imagesMT2014/image1}
+\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{imagesMT2014/image1B.pdf}
 \caption{Data Characteristics.
@@
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
-\includegraphics[width=0.5\textwidth]{imagesMT2014/image2}
+\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{imagesMT2014/image2B.pdf}
 \caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) close up of the region marked in green in C.}
@@
  %\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
-  \includegraphics[width=0.55\textwidth]{imagesMT2014/garb2}
+ \includegraphics[width=0.5\textwidth]{imagesMT2014/image_clustering.png}
   	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes.
 (A) MetaTracts belonging to orientation cluster 1.
-(B) MetaTracts belonging to orientation cluster 2.
+(C) MetaTracts belonging to orientation cluster 2.
 MetaTracts in gray(A,C) show context.
-}
+  	(D,E) shows the result of distance based clustering on the orientation cluster. (F) shows the combined result. }
   \label{fig:orientation_clustering}
@@
 		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
-		\includegraphics[width=0.5\textwidth]{imagesMT2014/image7b}
+		\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 15mm, clip]{imagesMT2014/image7b.pdf}
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
@@
 \centering
-	\includegraphics[width=0.45\textwidth]{imagesMT2014/image8.PNG}
+%	\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
 	\caption{Data set with flat thin and compact bundles.

### 2014-09-22 chapters/MetaTracts.tex

@@
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
-		\includegraphics[width=0.45\textwidth]{imagesMT2014/image5}
-	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.}
+		\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
+	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length)}
 	\label{fig:length_distribution}
@@
 		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
-		\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 15mm, clip]{imagesMT2014/image7b.pdf}
+		\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
@@
 \centering
-	\includegraphics[width=0.4\textwidth]{imagesMT2014/crop-16/histogram_10_A.PNG}
+	\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
 	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top).

### 2014-09-22 metaTracts-pvis2014.tex

@@
 \abstract{
-%Carbon fiber reinforced polymers (CFRP) typically consist of two main components: the matrix and the reinforcement. 
-%Most important in order to achieve the desired characteristics of a CFRP is the reinforcement component. 
-%In case of CFRPs the reinforcements typically consist of bundles of endless carbon fibers woven into sheets of carbon fiber %cloth, which are then layered and combined to a CFRP by the matrix component.
-%For quality control of CFRP materials and components industrial 3D X-ray computed tomography (XCT) is increasingly applied %as it allows detailed characterizations of the material in a non-destructive manner.
-
 This work introduces MetaTracts, a novel method for extracting and visualizing individual fiber bundles as well as their weaving pattern in endless carbon fiber reinforced polymer (CFRP) materials, scanned by X-ray computed tomography (XCT).

### 2014-09-22 chapters/MetaTracts.tex

@@
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
-\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{imagesMT2014/image1B.pdf}
+\includegraphics[width=0.45\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{imagesMT2014/image1B.pdf}
 \caption{Data Characteristics.
@@
 
-\begin{figure}[tb]
+\begin{figure*}[tb]
 \centering
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
-\includegraphics[width=0.5\textwidth]{imagesMT2014/image3b.png}
+\includegraphics[width=0.95\textwidth]{imagesMT2014/workflow.pdf}
 \caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
 \label{fig:flowchart}
-\end{figure}
+\end{figure*}
 
@@
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
-\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{imagesMT2014/image2B.pdf}
+\includegraphics[width=0.45\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{imagesMT2014/image2B.pdf}
 \caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) close up of the region marked in green in C.}
@@
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
-		\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
+		\includegraphics[width=0.45\textwidth]{imagesMT2014/figure5.pdf}
 	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length)}
@@
  %\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
- \includegraphics[width=0.5\textwidth]{imagesMT2014/image_clustering.png}
+ \includegraphics[width=0.45\textwidth]{imagesMT2014/figure6.pdf}
   	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes.
@@
 		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
-		\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
+		\includegraphics[width=0.5\textwidth]{imagesMT2014/figure7.pdf}
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 

### 2014-09-22 chapters/MetaTracts.tex
Updated text

--- 2014-09-22 chapters/MetaTracts.tex
@@
 \label{sec:intro}
-% Modern industry is constantly driving their research towards novel lightweight components.
-For these high-tech applications, the industry unites behind the mantra of function orientation, high integration as well as cost and eco-efficiency.
-A promising type of material, showing high specific stiffness, high specific strength, and high corrosion resistance, allows to integrate today's demands of industry.
-Advanced composites and in particular carbon fiber reinforced polymers (CFRP) were successfully introduced in aeronautic and automotive applications and within the recent years their share is rapidly rising.
-To increase the usage, not only structural but also complex primary structures and  highly loaded components have been manufactured from CFRP.
-As for today,  besides commercial aerospace and automotive industry, CFRP is used in wide range of industries, e.g., space/satellite, marine, sporting goods, automotive, civil engineering or wind energy.
+Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components.
+The material of choice of a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry.
+CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing  parts made of conventional materials such as aluminum or steel.
+CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.
 
-%REPLACED BY 
-Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and future components.
-The material of choice of a strongly growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry.
-CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing more and more parts made of conventional materials such as aluminum or steel.
-But besides the aerospace and automotive industry, CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.
+In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength, high corrosion resistance.
+CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
+At the same time highly complex and integrated components which were previously  impossible to manufacture may be produced from CFRP's.
+For example even primary structures and highly loaded components in aeronautics are being increasingly manufactured from CFRP.
 
-In general, CFRP materials show highly desired characteristics as high specific stiffness, high specific strength, high corrosion resistance at considerably lower weight compared to conventional materials.
-At the same time highly complex and integrated components may be produced from CFRPs, which were previously impossible to manufacture.
-For example even primary structures and highly loaded components in aeronautics are been increasingly manufactured from CFRP.
-
-Typically,  carbon fiber reinforced polymer components, more specifically CFRP laminates consists of two main components:
-
-The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics.
-To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use.
+Typically,  carbon fiber reinforced polymer components, more specifically CFRP laminates consists of two main components.
+A matrix on the one side, which acts as a bonding component and  reinforcements on the other side, which allow for achieving the desired characteristics.
+Various production processes are used to manufacture CFRP laminates with endless carbon fibers.
 Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern.
-These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also refereed to as fabric.
+These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric.
 Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth.
 Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties.
-Next, resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
+Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
 The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles.
-For the matrix of CFRP material systems with endless carbon fibers typically resins such as epoxy are frequently used.
-In special applications additional filler materials, such as colors or other particles, may be integrated.
-%The production process CFRP laminate is finished after curing the resin and the CFRP laminate is ready for further processing.
-
-The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
-Ultrasonic testing is yet the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
-On the one hand ultrasonic testing provides a quick and cost-efficient overview.
-But on the other hand it shows low resolution and may generate arbitrary results, e.g.
+The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
+Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
+On the one hand ultrasonic testing provides a quick and cost-efficient overview of the material.
+On the other hand, it shows low resolution and may generate arbitrary results, e.g.
 due to the geometry of the final component.
 
-Aside ultrasonic testing, industrial 3D X-Ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
-XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which taken throughout a full rotation of the specimen.
-In industrial XCT, the typical scanning geometry foresees the specimen to be placed on the rotary table between X-ray source and detector.
+Industrial 3D X-Ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly being applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
+XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen.
+In industrial XCT, the typical scanning geometry places the specimen on a rotary table between X-ray source and detector.
 X-rays are emitted by the source in the form of a conic beam.
 When passing through the specimen the X-rays get attenuated by the material in the specimen.
-The detector records the corresponding 2D attenuation image (penetration image), depending on its principle e.g.
-by transferring the X-rays in a scintillator layer into visible light.
+The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light.
 To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}.
-Finally the series of penetration images is finally reconstructed to a 3D volumetric dataset of the specimen.
+Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen.
 In state of the art devices, XCT can reach voxel sizes below $500 nm$.
@@
 The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector.
-The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but showing more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.
+The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.
 
-In this work we focus on datasets with larger viewports and lower resolutions where the single carbon fibers (filaments) are not or barely visible.
-Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of the individual fibers.
+In this work, we focus on datasets with larger viewports and lower resolutions where the single carbon fibers (filaments) are not or are  barely visible.
+Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of  individual fibers.
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
@@
 
-Extracting the answer to the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial even for experts.
+Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts.
 We detail and evaluate a method which uses visualization techniques to gain insight into our data.
@@
 We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites.
-The main goal of this work is to expand the state of the art in non destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.
+The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.
 
@@
 Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provides details of our approach.
-Section \ref{sec:results} details experiments and Section~\ref{sec:param_choices} explains our  parameter choices.
-We end with conclusions.
+Section \ref{sec:results} details experiments, section~\ref{sec:user_eval} summarizes our user evaluation results.
+Section~\ref{sec:param_choices} explains our  parameter choices.
+We end with conclusions in section~\ref{sec:conclusions}.
 
@@
 Its main clinical application is found in the study and treatment of neurological disorders.
-DTI may reveal abnormalities in white matter fiber structure and allows for visualizing the organization of white fibers in the human brain as well as the brain connectivity.
-A variety of algorithms have been proposed aiming to generate fiber-tract trajectories.
-In general, these reconstructions of fiber trajectories are then clustered in to bundles which are expected to be related anatomically or spatially.
-We broadly divide the related work on DTI into two parts which are of immediate relevance to our proposed solution: fiber tracking and fiber clustering.
-Furthermore methods analyzing the second order local structure are also discussed.
+DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of white fibers in the human brain and  brain connectivity.
+A variety of algorithms have been proposed aiming for generating fiber-tract trajectories.
+In general, these reconstructions of fiber trajectories are  clustered in bundles which are expected to be related anatomically or spatially.
+We broadly divide the related work on DTI into two parts that are of immediate relevance to our proposed solution: fiber tracking and fiber clustering.
+We discuss  methods for analyzing the second order local structure.
 Finally, we review the current state of the art in the visual analysis of fiber reinforced composites.
@@
 \label{subsec:fiberEx} 
-In detecting fiber tracts as shown by Basser et al.~\cite{Basser2000},~\cite{Basser2002} and Mori et al.~\cite{Mori2002},~\cite{Mori1999}, the key idea is based on the assumption that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel.
-In these techniques, continuous tracts are created by propagating a virtual particle from a starting point until they reach a termination criterion.
-This is usually done by solving a second-order Runge-Kutta integration.
-The assumption is that the principal diffusion direction at each discrete location can be interpolated to from a continuous velocity field.
-One of the short comings of the process is that because of the local decision-making process, these methods perform poorly in noisy regions and often generate small fibers.
+A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Basser2000, Mori1999, Mori2002}.
+Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion.
+ This is usually done by solving a second-order Runge-Kutta integration. The principal diffusion direction at each discrete location is interpolated to from a continuous velocity field. 
+ Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers.
 Basser et al.~\cite{Basser2000},~\cite{Basser2002} proposed that white matter tracts could be represented as 3D curves in space.
 They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data.
-Mori et al.~\cite{Mori2002} showed that tract reconstruction techniques could be broadly divided into line-propagation or energy minimization techniques.
-In line propagation approaches, trajectories are computed based on local neighborhoods.
-While, energy minimization approaches find the most favorable trajectory connecting two given endpoints.
+Mori et al.~\cite{Mori1999},~\cite{Mori2002} showed that tract reconstruction techniques could be broadly divided into line-propagation or energy minimization techniques.
+In line propagation approaches, trajectories are computed based on local neighborhoods and in energy minimization approaches the most  favorable trajectory connecting two given endpoints is selected.
 
@@
 \label {subsec:fiberClus}
-%!!!!! TODO REFINE FIBER CLUSTERING !!!!!
-%The next step is to identify  anatomically relevant fiber bundles. This is achieved through fiber clustering.
-The general assumption of fiber clustering in diffusion tensor imaging is that a similarity measure based on factors such as proximity between fibers can be used to interpret a collection of fiber tracts as bundles.
-Extensive research has been done on automatic DTI fiber clustering methods
-% (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}).
+In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles.
+Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Corouge2004, Brun2003, Brun2004, Zhang2008, westinMEDIA02}.
+Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
+Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}.
+Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge are clustered separately.
+Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors.
+The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering ~\cite{Moberts2005, Zhang2008} and those using spectral clustering ~\cite{jonasson2005, ODonnell2007, Brun2004}.
 
-The clustering approach divides the problem into two parts.
-The first part deals with finding a suitable proximity measure and the second deals with the choice of the clustering method.
-Pairwise proximity measures include, endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}.
-Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of the above distance so that fibers that are close for certain distance and then diverge are clustered separately.
-Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors.
-The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering (see ~\cite{Moberts2005},~\cite{Zhang2008}) and those using spectral clustering (see~\cite{jonasson2005},~\cite{ODonnell2007},~\cite{Brun2004})
-
-Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps proposed by Belkin and Niyogi~\cite{Belkin01} can be applied to the problem of organizing fiber trace data.
+Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}) can be applied to the problem of organizing fiber trace data.
 The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph.
@@
 In the case of fiber bundles, the individual points are fiber traces.
-In the ideal case fiber traces which belong to the same bundle must remain ``close" to each other in the lower dimensional space.Westin et al.~\cite{westinMEDIA02} also uses spectral clustering on a Hausdorff distance measure defined as the maximum of pointwise minimum distances between two fibers.
-Jonasson et al.~\cite{jonasson2005} runs k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
+In the ideal case fiber traces that belong to the same bundle must remain ``close" to each other in the lower dimensional space.
+Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of pointwise minimum distances between two fibers.
+Jonasson et al.~\cite{jonasson2005} run k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
 %The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}.
@@
 The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}. 
-These approaches build on the assumption that proximity measure that compare DTI fiber trajectories can also represent anatomical relationship.
+These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationship.
 An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster.
@@
 
-The difficulties in hierarchical clustering include: first, computing all pairs distances for tracts to generate the distance matrix (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times).
-Second, choosing the ``correct'' distance measure to compare tracts.
+There are a number of difficulties in hierarchical clustering.
+First, computing all pairs distances for tracts to generate the distance matrix is time consuming (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times).
+Second, a ``correct'' distance measure to compare tracts must be chosen.
 Thirdly, hierarchical clustering is best suited for similar length fibers.
@@
 Analyzing curvilinear structures in volumetric images has been utilized for a variety of purposes including center line extraction~\cite{Bouix2005} and vascular image enhancement as proposed by Frangi et al.~\cite{Frangi1998} and Sato et al.~\cite{Sato1997}.
-Frangi et al.~\cite{Frangi1998} introduced a method based on studying the eigenvalues of the local second order structure namely the Hessian matrix, specifically for the purposes of developing vessel enhancement filters.
+Frangi et al.~\cite{Frangi1998} introduced a method based on studying the eigenvalues of the  Hessian matrix specifically for the purposes of developing vessel enhancement filters.
 
@@
 Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles.
-Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the datatype is uint8.
-It clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
+Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the data type is unsigned 8 bit integer.
+Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
 The unit cell is used as a basis structure for CFRP manufacturing.
@@
 Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis.
-Figure~\ref{fig:data-char}(D) shows the zoomed in version of the green square.
+Figure~\ref{fig:data-char}(D) shows the magnified version of the green square.
 The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only.
@@
 The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization.
-The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
+The fiber bundles themselves may differ in terms of the amount of fibers in the bundle.
+Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
 Depending on the weaving pattern the fiber bundles cross each other in different orientations.
@@
 Nevertheless, the weaving pattern may cause the individual bundles to be curved.
-In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in euclidean distance but belong to different bundles.
+In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in Euclidean distance but belong to different bundles.
 We make the following assumptions on our data:
@@
 \begin{figure}[tb]
-\centering
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
-\includegraphics[width=0.5\textwidth]{imagesMT2014/image3b.png}
+\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
 \caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
 The approach consists of multiple steps.
-MetaTracts generations Sec.~\ref{sec:approach} and  finding plausible fiber bundles (Section~\ref{subsec:fiber-bundles}).
+MetaTracts generations  is explained in Section.~\ref{sec:approach} and  finding plausible fiber bundles in Section~\ref{subsec:fiber-bundles}, Section~\ref{sec:vis} talks about visualization extensions.
 %\subsection {Extracting MetaTracts}
@@
 
-The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
+The input to the first stage of our pipelines, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
 As output, each grid location becomes associated with a vector representing the local orientation and a real value [0,1] which represents a measure of the reliability of the local orientation.
 We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
-The principal directions, in which the local second order structure of the image can be decomposed are given by the eigen decomposition of the Hessian matrix which describes the local curvature.
+The principal directions, in which the local second order structure of the image can be decomposed, are given by the eigen decomposition of the Hessian matrix which describes the local curvature.
 The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.
@@
 \item{Each MetaTract is associated with a start point which is a grid vertex.}
-\item{Each cylinder in a MetaTract has a constant length, a constant radius and a start point (which is also a grid vertex).}
+\item{Each individual cylinder in a MetaTract has a constant length, a constant radius and a start point (which is also a grid vertex).}
 \item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
@@
 The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1].
-In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue.
+In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue(Fig.~\ref{fig:algo}B).
 Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}).
-The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow (Figure~\ref{fig:algo}C).
+The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow (Fig.~\ref{fig:algo}C).
 We generate a rectangle of length L and radius R, the green dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}).
@@
 Thus, we want the distance between $C_p$ and the start point for the next cylinder to be as large as possible.
-We measure distance of a grid vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$.
+We measure distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$.
 For example the Euclidean distance between the blue and the orange vertices projected onto $N_P$ gives distance (Figure~\ref{fig:algo}B).
-We refer to this perpendicular projection distance as \textit{``dist"}.
+We refer to this perpendicular projection distance as \textit{``projected\_dist"}.
 \end{itemize}
@@
 \begin{equation}
-Priority = \gamma_1 e^{(-angle^2 / \alpha^2)} + \gamma_2e^{(-dist^2 / \beta^2)}
+Priority = \gamma_1 e^{(-angle^2 / \alpha^2)} + \gamma_2e^{(-projected\_dist^2 / \beta^2)}
 \label{eqn:algo_1}
@@
 Essentially, Equation~\ref{eqn:algo_1}, selects a grid point which is the furthest from the current point and going in a similar direction.
-This gives it the advantage of tackling noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation (for orientation) in an already noisy environment.
+This gives it the advantage of tackling noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation in an already noisy environment.
 
@@
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.
+Fig.~\ref{fig:length_distribution}(B) shows the MetaTracts of a particular bundle colored according to their length.
+MetaTracts in a given fiber bundle do not extend the full length of the bundle. MetaTracts may have different lengths and may partially overlap.
 
@@
 		\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
-	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length)}
+	\caption{(A) MetaTracts colored according to mean orientations.
+(B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length).
+A single fiber bundle is made of MetaTracts of varying length}
 	\label{fig:length_distribution}
@@
 \label{subsec:fiber-bundles}
-The output of the above step is a set of MetaTracts.
-In this step we focus on clustering the MetaTracts into plausible bundles.
-We use hierarchical clustering which has enjoyed considerable attention and success in DTI field.
-However, often the MetaTracts produced in Section~\ref{sec:approach} do not extend the full length of the fiber bundle.
-Thus, MetaTracts in a given fiber bundle may have very different lengths and may only partially overlap as in Figure~\ref{fig:length_distribution}B.
-Moreover, carbon fibers have both orientation and proximity information which can be used for clustering.
-Instead of creating an heuristic, which artificially combines the two measures, we first cluster based on orientation and then further subdivide each orientation cluster using proximity information.
-We found that different clustering methods were preferable for the different measures.
-We use dimension reduction followed by K-means clustering for orientation clustering while we use hierarchical clustering to further subdivide each cluster based on proximity measures (see Section~\ref{sec:results} for experimental results).
-Breaking clustering into two steps also helps us and potential users to better understand the results of the clustering.
+
+The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles. 
+Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering.
+We have found that orientation measure was very reliable while measures of geometric proximity had problems with fibers which partially overlap (Fig.~\ref{fig:length_distribution}B).
+Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
+We found that different clustering techniques performed preferably for different measures.
+The two step clustering also made the problem more tractable and  the parameters intuitive for the end users.
+For the orientation clustering, we use dimension reduction followed by K-means clustering.
+We then use hierarchical clustering to further subdivide each cluster based on the geometric proximity measures.
 
@@
 \centering
-%	\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
+	\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
 	\caption{Data set with flat thin and compact bundles.
@@
 \section {User Evaluation}
-\label {sec:evaluation}
+\label {sec:user_eval}
 We integrated MetaTracts into a tool and created a user evaluation questionnaire which we distributed.
@@
 \section{Conclusion and future work}
+\label{sec:conclusions}
 In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials.

### 2014-09-22 metaTracts-pvis2014.tex
Updated text

@@
 \abstract{
-This work introduces MetaTracts, a novel method for extracting and visualizing individual fiber bundles as well as their weaving pattern in endless carbon fiber reinforced polymer (CFRP) materials, scanned by X-ray computed tomography (XCT).
-
-For the purpose of characterizing larger areas of CFRP materials such as complete unit cells integrating the recurring weaving pattern, the proposed workflow is designed to handle XCT scans of low resolution, in which the individual fibers of the bundles to be extracted are not or just barely visible.
-The proposed workflow applies and extends methods originally presented for the field of diffusion tensor imaging in order to cope with the low data quality: 
+This work introduces MetaTracts, a novel method  for extracting and visualizing individual fiber bundles as well as their weaving patterns from X-ray computed tomography (XCT) scans of  endless carbon fiber reinforced  polymers (CFRP).
+The  proposed work flow is designed  to analyze large areas of CFRP material  such as complete unit cells integrating the recurring weaving pattern.
+It is designed to handle XCT scans of low resolution, in which  individual  fibers  are not  visible or are barely visible.
 First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
-In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach, wherein the first step clusters according to orientation and the second step clusters by proximity of the MetaTracts.
-Finally, the tool can also  generate volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
-
-We evaluate the proposed workflow on a number of real world datasets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 
+In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach: first by orientation then by proximity.
+The tool can also  generate volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
+We evaluate the proposed work flow on a number of real world datasets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 
 %mention user study here?%

### 2014-09-22 chapters/MetaTracts.tex
Local old commit
Updated text
local text commit

@@
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
-\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
+%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
+\includegraphics[width=0.5\textwidth]{images_pvis/workflow}
 \caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
   \begin{center}
-   \includegraphics[ width=0.45\textwidth]{imagesMT2014/image4}
- % \includegraphics[ width=0.45\textwidth]{imagesMT2014/MT_abstract} 
+ \includegraphics[ width=0.4\textwidth]{imagesMT2014/image4}
   \end{center}
@@
 \centering
-	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
-		\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
+	\includegraphics[width=0.45\textwidth]{images_pvis/figure5.png}
+	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
 	\caption{(A) MetaTracts colored according to mean orientations.
@@
   \centering  	
- %\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
- \includegraphics[width=0.5\textwidth]{imagesMT2014/image_clustering.png}
+  %\includegraphics[width=0.45\textwidth]{images_pvis/figure6.png}
+% \includegraphics[width=0.5\textwidth]{imagesMT2014/image_clustering.png}
   	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes.
@@
 \centering
-		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
-		\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
+		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
+		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
@@
 \centering
-	\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
+	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
+	\includegraphics[width=0.4\textwidth, trim = 0mm 30mm 0mm 0mm, clip,]{images_pvis/figure8}
 	\caption{Data set with flat thin and compact bundles.

### 2014-09-23 chapters/MetaTracts.tex
local commit problems with figure 6
Figure 6 problems
local commit

@@
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
-\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{imagesMT2014/image1B.pdf}
+\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
 \caption{Data Characteristics.
@@
 
-\begin{figure}[tb]
+\begin{figure}[t]
+	\centering
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
 %\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
-\includegraphics[width=0.5\textwidth]{images_pvis/workflow}
+\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
 \caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
 \centering
-%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
-\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{imagesMT2014/image2B.pdf}
+%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
+\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
 \caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) close up of the region marked in green in C.}
@@
   \centering  	
-  %\includegraphics[width=0.45\textwidth]{images_pvis/figure6.png}
-% \includegraphics[width=0.5\textwidth]{imagesMT2014/image_clustering.png}
+  \includegraphics[width=0.5\textwidth]{images_pvis/figure6.png}
+  %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
   	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes.
@@
 \centering
-	\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
+	\includegraphics[width=0.5\textwidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
+	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
 	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top).

### 2014-09-23 chapters/MetaTracts.tex
local commit

@@
   \begin{center}
- \includegraphics[ width=0.4\textwidth]{imagesMT2014/image4}
+ \includegraphics[ width=0.3\textwidth]{imagesMT2014/image4}
   \end{center}
@@
 We compared  $\eta=3-7$ experimentally without any dramatic change in results.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth, trim = 10mm 100mm 0mm 170mm, clip]{images_pvis/figure10}
+	%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
+	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
+	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
+	\caption{(A) Volume rendering of data set Figure~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
+	\label{fig:comparison}
+\end{figure} 
 %\begin{figure}

### 2014-09-23 chapters/MetaTracts.tex
Updated text ready for review.

--- 2014-09-23 chapters/MetaTracts.tex
@@
 In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles.
-Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Corouge2004, Brun2003, Brun2004, Zhang2008, westinMEDIA02}.
+Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}.
 Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
@@
 Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors.
-The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering ~\cite{Moberts2005, Zhang2008} and those using spectral clustering ~\cite{jonasson2005, ODonnell2007, Brun2004}.
+The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering ~\cite{Moberts2005, Zhang2008} and those using spectral clustering ~\cite{ Brun2004,jonasson2005, ODonnell2007}.
 
@@
 An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster.
-At each stage of the algorithm the two most similar clusters are joined.
+At each stage of the algorithm the two most similar clusters are merged based on some criterion.
 The two basic cluster similarity measures are single-link and complete-link.
@@
 Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers.
+Like~\cite{Weissenbock2014}, most carbon fiber extraction literature deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.
 
@@
 In general the number of orientations is limited by the weaving pattern.
-In all our data sets the number of separate orientations was two.
 Nevertheless, the weaving pattern may cause the individual bundles to be curved.
@@
 Figure \ref{fig:flowchart} shows our pipeline.
-The approach consists of multiple steps.
 MetaTracts generations  is explained in Section.~\ref{sec:approach} and  finding plausible fiber bundles in Section~\ref{subsec:fiber-bundles}, Section~\ref{sec:vis} talks about visualization extensions.
@@
 Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude.
-Here $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigen values of the Hessian matrix.
+Here $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix.
 Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$).
@@
 Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors.
-Thus we compute a measure, ``equal" to the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
+Thus we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
 $$
@@
 Figure~\ref{fig:reliable_hessian}B and C show 2D slices along the Z and X axis respectively.
-Figure~\ref{fig:reliable_hessian}D shows a close up (region similar to Figure~\ref{fig:data-char}D).
-Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold.
+Figure~\ref{fig:reliable_hessian}D shows a magnified region similar to Figure~\ref{fig:data-char}D.
+Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable.
 The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.
@@
 \includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
-\caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) close up of the region marked in green in C.}
+\caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) Magnified region marked in green square (in C).}
 \label{fig:reliable_hessian}
@@
 \label{subsec:MetaTracts-generation}
+ In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$.
+The procedure trivially extends to $\mathbb{R}^3$.
 In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles.
-We explain the process of MetaTracts generation in $\mathbb{R}^2$.
-The procedure trivially extends to $\mathbb{R}^3$.
 The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1].
-In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue(Fig.~\ref{fig:algo}B).
+In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue (Fig.~\ref{fig:algo}B).
+Fig.~\ref{fig:algo}(C) shows a single cylinder of a MetaTracts.
 Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}).
-The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow (Fig.~\ref{fig:algo}C).
+The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow.
 We generate a rectangle of length L and radius R, the green dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}).
@@
 % \end{equation}
-We further define a priority for each of the candidate vertices based on the above factors.
+We define a priority for each of the candidate vertices based on the above factors.
 For each cylinder in a MetaTract, we put the \textit{candidate vertices} of it in a priority queue based on Equation \ref{eqn:algo_1}.
@@
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
-That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
+Fig.\ref{fig:length_distribution}A also shows a single slice of YZ plane comparable to Fig.~\ref{fig:data-char}C and Fig.~\ref{fig:reliable_hessian}C.That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.
-Fig.~\ref{fig:length_distribution}(B) shows the MetaTracts of a particular bundle colored according to their length.
+Fig.~\ref{fig:length_distribution}(B) shows the MetaTracts of a particular bundle colored according to their individual length.
 MetaTracts in a given fiber bundle do not extend the full length of the bundle. MetaTracts may have different lengths and may partially overlap.
@@
 	\caption{(A) MetaTracts colored according to mean orientations.
-(B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length).
+(B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length:  $2\mu m$).
 A single fiber bundle is made of MetaTracts of varying length}
@@
 Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering.
-We have found that orientation measure was very reliable while measures of geometric proximity had problems with fibers which partially overlap (Fig.~\ref{fig:length_distribution}B).
+Experimentally we found that orientation measure was very reliable while measures of geometric proximity had problems with fibers which partially overlap (Fig.~\ref{fig:length_distribution}B).
 Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
@@
 We used conventional K-means for clustering this lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
-For our test case there are two major directions of the fiber bundles.
-So $K$ was set to 2.
-It is important to note that due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
+For our test case there are two major directions of the fiber bundles and $K$ was set to 2.
+Due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
 Figure~\ref{fig:orientation_clustering}B shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
@@
 Hierarchical clustering has a single parameter $h$, the desired number of clusters.
-In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
 Clusters are merged until there are only $h$ clusters left.
@@
 First, the clustering would produce some small clusters of just a few MetaTracts.
-These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
+These MetaTracts were anomalies caused by overlapping fibers and did not represent true fiber bundles.
 Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
@@
 The results of clustering the orientation clusters Fig.~\ref{fig:orientation_clustering}(A,C) are shown in Fig.~\ref{fig:orientation_clustering}(D,E) respectively.
-Note how the purple cluster Fig.~\ref{fig:orientation_clustering}(E) bundle ``7'' is extracted well.
+Note how the thin and curved purple cluster Fig.~\ref{fig:orientation_clustering}(E) bundle ``7'' is extracted well.
+
+In comparison Fig.~\ref{fig:comparison} shows the MetaTracts in Fig.~\ref{fig:length_distribution}(A) being hierarchically clustered directly using proximity alone into 10,15 and 20 clusters (ground truth being 11).
+As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low cardinality outliers.
+In Fig.~\ref{fig:comparison} when ``h" is 10 two large multi-bundle incorrect cluster is generated and the rest are outliers.
+As ``h" increases, some appropriate bundles start to form.
+Even at ``h" equals 20, the top bundles all clusters together into one large incorrectly identified bundle.
 \begin{figure}
@@
 We tested our method on multiple datasets with varying characteristics. 
-Figure~\ref{fig:orientation_clustering} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into 10 clusters (the ground truth is 6 and 5 respectively).
+Figure~\ref{fig:orientation_clustering} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into ``h" equlas 10 clusters (the ground truth is 6 and 5 respectively).
 The hierarchical clustering is robust to over segmentation.
@@
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
-As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 Fig.~\ref{fig:length_distribution} is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length).
+As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 (Fig.~\ref{fig:length_distribution}) is 200 while the maximum MetaTract length in the bundle is about 500 (measured in units of grid cube length: $2\mu m$).
 Fig.~\ref{fig:length_distribution}(B) shows the length per MetaTract distribution for a particular fiber bundle.
@@
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
-	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top).
+	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}A, clustered into 10 clusters.
 The unit for length is the grid cube length.
@@
 Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber.
-We used 10.00 and 2.00 for length and breadth, respectively for all tests.
+We used 10.00 and 2.00 for length and breadth(measured in grid cube length), respectively for all tests.
 $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays, we have used [7-10] and half the length of cylinder (5.0).
-Our number of fiber bundle directions are limited, thus even for small $\eta$, the distinction between the orientation clusters is preserved quite well.
-We compared  $\eta=3-7$ experimentally without any dramatic change in results.
+Our number of fiber bundle directions are limited, thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well.
+We compared  $m=3-7$ experimentally without any dramatic change in results.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
+	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
+	\label{fig:comparison}
+\end{figure} 
 %\begin{figure}

### 2014-09-24 chapters/MetaTracts.tex
Updated text and images for User Evaluation

@@
 We tested our method on multiple datasets with varying characteristics. 
-Figure~\ref{fig:orientation_clustering} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into ``h" equlas 10 clusters (the ground truth is 6 and 5 respectively).
+Figure~\ref{fig:orientation_clustering} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into ``h" equals 10 clusters (the ground truth is 6 and 5 respectively).
 The hierarchical clustering is robust to over segmentation.
@@
 \label {sec:user_eval}
-We integrated MetaTracts into a tool and created a user evaluation questionnaire which we distributed.
-%\begin{figure}[h]
-%\centering
-%\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
-%\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
-%\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
-%(b) shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
-%\label{fig:context}
-%\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.45\textwidth,  trim = 0mm 60mm 0mm 0mm, clip]{images_pvis/usereval}
+	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
+	\caption{The user evaluation consists of a set of queries on videos of our results, volume rendering and Hessian colormap.
+The questionnaire tests the effectiveness of MetaTracts in visualizing ``geometric structure" and ``spatial context".
+The Likert scale goes from 1-5: strongly disagree, disagree, neutral, agree and strongly agree.
+}
+	\label{fig:userstudy}
+\end{figure}
+We integrated MetaTracts into a tool and created a user evaluation questionnaire.
+We approached four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization to evaluate our work.
+The questionnaire uses Likert scale (LS).
+LS consists of 5 levels from ``Strongly disagree" to ``Strongly agree".
+Figure~\ref{fig:comparison} shows the results.
+Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video of 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A).
+This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion effects the user.
+As expected, Fig.~\ref{fig:comparison} shows that evaluators found it difficult to compute the number of bundles from volume rendering alone.
+Another common approach is to color map the top eigenvector of Hessian matrix to the RGB space.
+We show a video of the color mapped eigenvector (similar to fig.~\ref{fig:reliable_hessian}A) and asked the same query.
+Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed relatively better in this test than the direct volume rendering.
+Majority evaluators agreed or strongly agreed (LS) that Hessian color map was more effective.
+Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately fig~\ref{fig:orientation_clustering}D,E).
+All evaluators could correctly identify the number of fiber bundles.
+All evaluators agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering.
+They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig~\ref{fig:data-char}C).
+The next set of tasks (on the MetaTracts results) tests the spatial context perceived from MetaTracts.
+We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, selecting the smallest from a subset and given a fiber bundle selecting the most similar one from a subset.
+Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering itself, but is straight forward from MetaTracts.
+In the Likert scale,  majority strongly agreed that the tasks are undemanding compared to volume rendering.
+Finally, the majority evaluators agreed that the extracted meshes provides better spatial context than volume rendering.
 
@@
 \centering
-	\includegraphics[width=0.5\textwidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
+	\includegraphics[width=0.45\textwidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}

### 2014-09-24 chapters/MetaTracts.tex
Minor Change

@@
  Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers.
-Basser et al.~\cite{Basser2000},~\cite{Basser2002} proposed that white matter tracts could be represented as 3D curves in space.
+Basser et al.~\cite{Basser2002},~\cite{Basser2000} proposed that white matter tracts could be represented as 3D curves in space.
 They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data.
@@
 Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors.
-The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering ~\cite{Moberts2005, Zhang2008} and those using spectral clustering ~\cite{ Brun2004,jonasson2005, ODonnell2007}.
+The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{ Brun2004,jonasson2005, ODonnell2007}.
 
@@
 
-The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}. 
+The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
 These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationship.
@@
 %\end{figure}
-The order of the candidate vertices is based on the following characteristics:
+Based on the assumptions on our data (Section~\ref{sec:char_data}) local orientation and connectivity, the order of the candidate vertices is based on the following characteristics:
 

### 2014-09-24 chapters/MetaTracts.tex
reviewed user study, modified enums + itemizes

@@
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Figure~\ref{fig:data-char} provides an example of our targeted datasets. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
+Fig.~\ref{fig:data-char} provides an example of our targeted datasets. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
 Our work is motivated from the recent progress in two interrelated fields: On the one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials \cite{Karpat2012}.
@@
 To the best of our knowledge, there is no current work that can resolve simple queries such as:
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item{ How to extract and visualize the geometric structure of a particular fiber bundle?}
@@
 The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
-Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset.
+Fig.~\ref{fig:data-char}(A) shows a volume rendering of the dataset.
 Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis.
-Figure~\ref{fig:data-char}(D) shows the magnified version of the green square.
+Fig.~\ref{fig:data-char}(D) shows the magnified version of the green square.
 The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only.
-Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions.
+Fig.~\ref{fig:data-char}(D) contains two bundles going in opposite directions.
 However, the separation between two fiber bundles is barely visible.
@@
 The fiber bundles themselves may differ in terms of the amount of fibers in the bundle.
-Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
+Fig.~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
 Depending on the weaving pattern the fiber bundles cross each other in different orientations.
@@
 We make the following assumptions on our data:
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item The structure embedded in the data contains fiber bundles of indiscernible fibers.
@@
 
-Figure \ref{fig:flowchart} shows our pipeline.
+Fig.
+\ref{fig:flowchart} shows our pipeline.
 MetaTracts generations  is explained in Section.~\ref{sec:approach} and  finding plausible fiber bundles in Section~\ref{subsec:fiber-bundles}, Section~\ref{sec:vis} talks about visualization extensions.
@@
 %\end{figure}
-In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold.
+In Fig.~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold.
 The unit vector representing the principal direction has been mapped to RGB color space.
-Figure~\ref{fig:reliable_hessian}A shows the entire data set.
+Fig.~\ref{fig:reliable_hessian}A shows the entire data set.
 Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue.
-Figure~\ref{fig:reliable_hessian}B and C show 2D slices along the Z and X axis respectively.
-Figure~\ref{fig:reliable_hessian}D shows a magnified region similar to Figure~\ref{fig:data-char}D.
+Fig.~\ref{fig:reliable_hessian}B and C show 2D slices along the Z and X axis respectively.
+Fig.~\ref{fig:reliable_hessian}D shows a magnified region similar to Fig.~\ref{fig:data-char}D.
 Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable.
@@
 
-\begin{enumerate}
+\begin{enumerate}[noitemsep]
 \item{Each MetaTract is associated with a continuous set of cylinders.}
@@
 The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1].
-In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue (Fig.~\ref{fig:algo}B).
+In Fig.
+\ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue (Fig.~\ref{fig:algo}B).
 Fig.~\ref{fig:algo}(C) shows a single cylinder of a MetaTracts.
@@
 
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item \textit{Orientation Similarity}: We want the orientation of the start points for the consecutive cylinders to have a similar orientations ($N_p$). 
@@
 We measure distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$.
-For example the Euclidean distance between the blue and the orange vertices projected onto $N_P$ gives distance (Figure~\ref{fig:algo}B).
+For example the Euclidean distance between the blue and the orange vertices projected onto $N_P$ gives distance (Fig.~\ref{fig:algo}B).
 We refer to this perpendicular projection distance as \textit{``projected\_dist"}.
@@
 
-In this particular example in Figure~\ref{fig:algo}(B) we select the orange grid vertex next and repeat the process.
-We do not select the purple vertex (Figure~\ref{fig:algo}(C)) because it is not a candidate vertex.
+In this particular example in Fig.~\ref{fig:algo}(B) we select the orange grid vertex next and repeat the process.
+We do not select the purple vertex (Fig.~\ref{fig:algo}(C)) because it is not a candidate vertex.
 If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length.
 Short tracts are then removed.
-The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}A.
+The MetaTracts generated are shown in Fig.~\ref{fig:length_distribution}A.
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
@@
 Due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
-Figure~\ref{fig:orientation_clustering}B shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
+Fig.~\ref{fig:orientation_clustering}B shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
 As is expected there is a clear distinction based on fiber orientation.
-Figure~\ref{fig:orientation_clustering}(A,C) shows the MetaTracts colored according to clustering results.
+Fig.~\ref{fig:orientation_clustering}(A,C) shows the MetaTracts colored according to clustering results.
 
@@
 The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$ but the directed Hausdorff is not.
-Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially covers the bundle (Figure~\ref{fig:length_distribution}(B)).
+Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially covers the bundle (Fig.~\ref{fig:length_distribution}(B)).
 If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$, then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
@@
 Even at ``h" equals 20, the top bundles all clusters together into one large incorrectly identified bundle.
-\begin{figure}
+\begin{figure}[tb]
 \centering
@@
 	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
-	\caption{(A) Volume rendering of data set Figure~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
+	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
 	\label{fig:crop-16-decomp}
@@
 The voxel is then assigned to the class with the maximum number of elements in the neighborhood.
-Figure~\ref{fig:crop-16-decomp}A and Figure~\ref{fig:prepreg}D shows the result of voxelization.
-Figure~\ref{fig:crop-16-decomp}(C, D) show examples of extracted meshes.
+Fig.~\ref{fig:crop-16-decomp}A and Fig.~\ref{fig:prepreg}D shows the result of voxelization.
+Fig.~\ref{fig:crop-16-decomp}(C, D) show examples of extracted meshes.
 %The MetaTracts and the clustering provide an abstraction of the fiber bundles and are used as a visualization tool.
@@
 
-\begin{figure}[t]
+\begin{figure}[tb]
 \centering
@@
 We tested our method on multiple datasets with varying characteristics. 
-Figure~\ref{fig:orientation_clustering} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into ``h" equals 10 clusters (the ground truth is 6 and 5 respectively).
+Fig.~\ref{fig:orientation_clustering} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into ``h" equals 10 clusters (the ground truth is 6 and 5 respectively).
 The hierarchical clustering is robust to over segmentation.
-Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster.
+Fig.~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster.
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
@@
 \label {sec:user_eval}
-\begin{figure}
+\begin{figure}[tb]
 	\centering
@@
 \end{figure}
-We integrated MetaTracts into a tool and created a user evaluation questionnaire.
-We approached four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization to evaluate our work.
-The questionnaire uses Likert scale (LS).
+Four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization participated in the user evaluation of our work.
+We therefore setup a questionnaire with questions on solving given tasks as well as questions using Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques.
 LS consists of 5 levels from ``Strongly disagree" to ``Strongly agree".
-Figure~\ref{fig:comparison} shows the results.
-Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video of 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A).
-This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion effects the user.
-As expected, Fig.~\ref{fig:comparison} shows that evaluators found it difficult to compute the number of bundles from volume rendering alone.
-Another common approach is to color map the top eigenvector of Hessian matrix to the RGB space.
-We show a video of the color mapped eigenvector (similar to fig.~\ref{fig:reliable_hessian}A) and asked the same query.
-Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed relatively better in this test than the direct volume rendering.
-Majority evaluators agreed or strongly agreed (LS) that Hessian color map was more effective.
-Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately fig~\ref{fig:orientation_clustering}D,E).
-All evaluators could correctly identify the number of fiber bundles.
-All evaluators agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering.
-They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig~\ref{fig:data-char}C).
-The next set of tasks (on the MetaTracts results) tests the spatial context perceived from MetaTracts.
-We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, selecting the smallest from a subset and given a fiber bundle selecting the most similar one from a subset.
-Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering itself, but is straight forward from MetaTracts.
-In the Likert scale,  majority strongly agreed that the tasks are undemanding compared to volume rendering.
-Finally, the majority evaluators agreed that the extracted meshes provides better spatial context than volume rendering.
+Fig.~\ref{fig:comparison} shows the results.
+Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A).
+This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user.
+As expected, Fig.~\ref{fig:comparison} shows that the participants found it difficult to determine the number of bundles from volume rendering alone.
+Another common approach is to color map the main eigenvector of Hessian matrix to RGB space.
+We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Fig.~\ref{fig:reliable_hessian}A) and asked the same query.
+Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test before.
+The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective.
+Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Fig.~\ref{fig:orientation_clustering}D,E).
+All participants could correctly identify the number of fiber bundles.
+All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering.
+They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig.~\ref{fig:data-char}C).
+The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts.
+We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset.
+Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts.
+In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
+Finally, the majority participants agreed that the extracted meshes provides better spatial context than volume rendering.
 
@@
 \label{sec:param_choices}
-\begin{figure}
+\begin{figure}[tb]
 \centering
@@
 The unit for length is the grid cube length.
-Figure~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6.
+Fig.~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6.
 Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10.
@@
 We compared  $m=3-7$ experimentally without any dramatic change in results.
-\begin{figure}
+\begin{figure}[tb]
 	\centering

### 2014-09-24 chapters/MetaTracts.tex
corrections intro + section 2

--- 2014-09-24 chapters/MetaTracts.tex
@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components.
-The material of choice of a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry.
+The materials of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry.
 CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing  parts made of conventional materials such as aluminum or steel.
@@
 CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
-At the same time highly complex and integrated components which were previously  impossible to manufacture may be produced from CFRP's.
-For example even primary structures and highly loaded components in aeronautics are being increasingly manufactured from CFRP.
+At the same time highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
+Primary structures and highly loaded components in aeronautics are one example.
 
-Typically,  carbon fiber reinforced polymer components, more specifically CFRP laminates consists of two main components.
-A matrix on the one side, which acts as a bonding component and  reinforcements on the other side, which allow for achieving the desired characteristics.
+Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates, consist of two main components:
+A matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.
 Various production processes are used to manufacture CFRP laminates with endless carbon fibers.
@@
 Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
-The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles.
-The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
+The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles.
+The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
 Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
-On the one hand ultrasonic testing provides a quick and cost-efficient overview of the material.
+On one hand ultrasonic testing provides a quick and cost-efficient overview of the material.
 On the other hand, it shows low resolution and may generate arbitrary results, e.g.
@@
 
-Industrial 3D X-Ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly being applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
+Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
 XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen.
-In industrial XCT, the typical scanning geometry places the specimen on a rotary table between X-ray source and detector.
+Typical scanning geometry in industrial XCT is the specimen placed on a rotary table between X-ray source and detector.
 X-rays are emitted by the source in the form of a conic beam.
@@
 Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen.
-In state of the art devices, XCT can reach voxel sizes below $500 nm$.
+In state of the art devices, XCT can reach voxel sizes below 500 $nm$.
 XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens.
@@
 
-In this work, we focus on datasets with larger viewports and lower resolutions where the single carbon fibers (filaments) are not or are  barely visible.
+In this work, we focus on datasets with larger viewports and lower resolutions where the individual carbon fibers (filaments) are indiscernible or are barely visible.
 Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of individual fibers.
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Figure~\ref{fig:data-char} provides an example of our targeted datasets. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
-Our work is motivated from the recent progress in two interrelated fields: On the one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials \cite{Karpat2012}.
-On the other hand, recent developments of industrial 3D X-Ray computed tomography (XCT) with regard to larger detectors, larger field of views and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
+Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
+Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
+On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
 
@@
 We detail and evaluate a method which uses visualization techniques to gain insight into our data.
-We demarcate the above queries into two parts, ``geometric structure" and ``spatial context".
+We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}.
 Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles.
 Spatial context refers to how two or more bundles interact with each other.
-We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites.
-The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.
+We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites.
+The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
 
-The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  Section \ref{sec:char_data} describes the   characteristics of our data.
-Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provides details of our approach.
-Section \ref{sec:results} details experiments, section~\ref{sec:user_eval} summarizes our user evaluation results.
+The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  section \ref{sec:char_data} describes the   characteristics of our data.
+Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provide details of our approach.
+Section \ref{sec:results} details experiments and section~\ref{sec:user_eval} summarizes our user evaluation results.
 Section~\ref{sec:param_choices} explains our  parameter choices.
@@
 \includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
-\caption{Data Characteristics.
-(A) shows a rendering of the data.
-(B) shows 2D slices along the X and (C) along the Z axis.
+\caption{Data characteristics: (A) shows a rendering of the data.
+(B) shows 2D slices along the X- and (C) along the Z-axis.
 (D) close up of the region marked in green in (C), multiple fiber bundles cross and are indistinguishable by visual inspection alone.
@@
 
-Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles.
-Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the data type is unsigned 8 bit integer.
-Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
+Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
+Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$, the data type is unsigned 8 bit integer.
+Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
 The unit cell is used as a basis structure for CFRP manufacturing.
@@
 Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset.
-Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis.
-Figure~\ref{fig:data-char}(D) shows the magnified version of the green square.
-The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only.
+Figure~\ref{fig:data-char}(B) shows 2D slices along the $X$- and Figure~\ref{fig:data-char}(C) in $Z$-axis.
+Figure~\ref{fig:data-char}(D) shows the magnified version of the green region of interest.
+The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only.
 Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions.
@@
 The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization.
-The fiber bundles themselves may differ in terms of the amount of fibers in the bundle.
+The fiber bundles may differ in terms of the amount of fibers in the bundle.
 Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
@@
 Nevertheless, the weaving pattern may cause the individual bundles to be curved.
-In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in Euclidean distance but belong to different bundles.
+In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.
 We make the following assumptions on our data:
@@
 \item The structure embedded in the data contains fiber bundles of indiscernible fibers.
-\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
+\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
 \item The local orientation may change gradually inside the fiber bundle.
 \item The local orientation may be noisy and not reliable.
-\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
+\item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
 \item Fiber bundles going in different directions only interact near the surface of contact.

### 2014-09-24 chapters/MetaTracts.tex
reviewed until section 5

@@
 
-The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  section \ref{sec:char_data} describes the   characteristics of our data.
+The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  Section \ref{sec:char_data} describes the   characteristics of our data.
 Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provide details of our approach.
-Section \ref{sec:results} details experiments and section~\ref{sec:user_eval} summarizes our user evaluation results.
+Section \ref{sec:results} details experiments and Section~\ref{sec:user_eval} summarizes our user evaluation results.
 Section~\ref{sec:param_choices} explains our  parameter choices.
-We end with conclusions in section~\ref{sec:conclusions}.
+We end with conclusions in Section~\ref{sec:conclusions}.
 
@@
 The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
-Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset.
-Figure~\ref{fig:data-char}(B) shows 2D slices along the $X$- and Figure~\ref{fig:data-char}(C) in $Z$-axis.
-Figure~\ref{fig:data-char}(D) shows the magnified version of the green region of interest.
+Figure~\ref{fig:data-char}a shows a volume rendering of the dataset.
+Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis.
+Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest.
 The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only.
-Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions.
+Figure~\ref{fig:data-char}d contains two bundles going in opposite directions.
 However, the separation between two fiber bundles is barely visible.
@@
 The fiber bundles may differ in terms of the amount of fibers in the bundle.
-Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
+Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
 Depending on the weaving pattern the fiber bundles cross each other in different orientations.
@@
 The unit vector representing the principal direction has been mapped to RGB color space.
-Figure~\ref{fig:reliable_hessian}A shows the entire data set.
+Figure~\ref{fig:reliable_hessian}a shows the entire data set.
 Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue.
-Figure~\ref{fig:reliable_hessian}B and C show 2D slices along the Z and X axis respectively.
-Figure~\ref{fig:reliable_hessian}D shows a magnified region similar to Figure~\ref{fig:data-char}D.
+Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axis respectively.
+Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest.
 Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable.
@@
 The local orientaion at each voxel is inherently more noisy.
-Thus techniques popular in DTI cannot be directly applied in XCT data.
+Thus techniques popular in DTI cannot be directly applied to XCT data.
 
@@
 The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1].
-In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue (Fig.~\ref{fig:algo}B).
-Fig.~\ref{fig:algo}(C) shows a single cylinder of a MetaTracts.
-Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}).
+In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue.
+Figure~\ref{fig:algo}c shows a single cylinder of a MetaTracts.
+Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Section~\ref{subsec:mtprop}).
 The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow.
-We generate a rectangle of length L and radius R, the green dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}).
+We generate a rectangle of length $L$ and radius $R$, the green dashed rectangle shows the rectangle generated at $C_p$ (properties 3 and 5 in Section~\ref{subsec:mtprop}).
 \begin{figure}[tb]
@@
 %\end{figure}
-Based on the assumptions on our data (Section~\ref{sec:char_data}) local orientation and connectivity, the order of the candidate vertices is based on the following characteristics:
+Based on such assumptions on our data as local orientation and connectivity (Section~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:
 
 \begin{itemize}
-\item \textit{Orientation Similarity}: We want the orientation of the start points for the consecutive cylinders to have a similar orientations ($N_p$). 
+\item \textit{Orientation similarity}: We want the orientation of the start points for the consecutive cylinders to have a similar orientations ($N_p$). 
 \item \textit{Large distance}: We want the MetaTracts to traverse the data using as few cylinders as possible.
@@
 For example the Euclidean distance between the blue and the orange vertices projected onto $N_P$ gives distance (Figure~\ref{fig:algo}B).
-We refer to this perpendicular projection distance as \textit{``projected\_dist"}.
+We refer to this perpendicular projection distance as \textit{`projected\_dist'}.
 \end{itemize}
@@
 $\gamma_1$, $\gamma_2$ are the  weights ($\mathbb{R}_{\ge 0}$)  which decide how the priority depends on the affine combination of the two factors.
-For all our cases we use ($\gamma_1$ = $1 / 3 $ and $\gamma_2$ = $2 / 3$).
-In general we suggest $\gamma_1 +\gamma_2 = 1 $ and $\gamma_1 \leq \gamma_2$).
+For all our cases we use $\gamma_1$ = $1 / 3 $ and $\gamma_2$ = $2 / 3$.
+In general we suggest $\gamma_1 +\gamma_2 = 1 $ and $\gamma_1 \leq \gamma_2$.
 At each iteration we pick the top element in the priority queue generate the corresponding cylinder and repeat the steps.
-Essentially, Equation~\ref{eqn:algo_1}, selects a grid point which is the furthest from the current point and going in a similar direction.
+Essentially, Equation~\ref{eqn:algo_1} selects a grid point which is the furthest from the current point and going in a similar direction.
 This gives it the advantage of tackling noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation in an already noisy environment.
 
-In this particular example in Figure~\ref{fig:algo}(B) we select the orange grid vertex next and repeat the process.
-We do not select the purple vertex (Figure~\ref{fig:algo}(C)) because it is not a candidate vertex.
+In this particular example in Figure~\ref{fig:algo}b we select the orange grid vertex next and repeat the process.
+We do not select the purple vertex (Figure~\ref{fig:algo}c) because it is not a candidate vertex.
 If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length.
 Short tracts are then removed.
-The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}A.
+The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}a.
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
-Fig.\ref{fig:length_distribution}A also shows a single slice of YZ plane comparable to Fig.~\ref{fig:data-char}C and Fig.~\ref{fig:reliable_hessian}C.That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
+Figure\ref{fig:length_distribution}a shows a single slice of YZ plane.
+That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.
-Fig.~\ref{fig:length_distribution}(B) shows the MetaTracts of a particular bundle colored according to their individual length.
+Figure~\ref{fig:length_distribution}b shows the MetaTracts of a particular bundle colored according to their individual length.
 MetaTracts in a given fiber bundle do not extend the full length of the bundle. MetaTracts may have different lengths and may partially overlap.

### 2014-09-24 chapters/MetaTracts.tex
reviewed until user eval

@@
 	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
-	\caption{(A) MetaTracts colored according to mean orientations.
-(B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length:  $2\mu m$).
+	\caption{(A) MetaTracts color-coded according to their mean orientations.
+(B) length distribution of individual MetaTracts for a particular bundle (unit for length is grid cube edge length:  $2\mu m$).
 A single fiber bundle is made of MetaTracts of varying length}
@@
 Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering.
-Experimentally we found that orientation measure was very reliable while measures of geometric proximity had problems with fibers which partially overlap (Fig.~\ref{fig:length_distribution}B).
+Experimentally we found that orientation measure was very reliable while measures of geometric proximity had problems with fibers which partially overlap (Figure~\ref{fig:length_distribution}b).
 Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
@@
 This lower dimensional space can then be clustered more efficiently than the higher dimensional space. 
-Let G be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem:$L\textbf{f}=\lambda D\textbf{f}$,
+Let $G$ be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem:$L\textit{\textbf{f}}=\lambda D\textit{\textbf{f}}$,
 %\begin{equation}\label{equn:eigenMaps}
@@
 %\end{equation}
-where D is the diagonal weight matrix and L is the Laplacian matrix.
-The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an m-dimensional space.
+where $D$ is the diagonal weight matrix and $L$ is the Laplacian matrix.
+The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an $m$-dimensional space.
 K-means clustering is then used to cluster in this lower dimensional space.
@@
 Due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
-Figure~\ref{fig:orientation_clustering}B shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
+Figure~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
 As is expected there is a clear distinction based on fiber orientation.
-Figure~\ref{fig:orientation_clustering}(A,C) shows the MetaTracts colored according to clustering results.
+Figure~\ref{fig:orientation_clustering}a,c shows the MetaTracts colored according to clustering results.
 
@@
 We use the directed Hausdorff distance for distance based clustering.
-Each MetaTract is represented as a set of points ($C_P$). Formally, the directed Hausdorff distance from point set P to point set Q is defined as 
+Each MetaTract is represented as a set of points ($C_P$). Formally, the directed Hausdorff distance from point set $P$ to point set $Q$ is defined as 
 $H_{dir}(P,Q) = max_{p \in P} min_{q \in Q} d(p,q)$ .
@@
 The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$ but the directed Hausdorff is not.
-Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially covers the bundle (Figure~\ref{fig:length_distribution}(B)).
+Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially covers the bundle (Figure~\ref{fig:length_distribution}b).
 If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$, then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
@@
 We repeated both until a steady state of clusters is reached and no new small fibers can be removed.
-The results of clustering the orientation clusters Fig.~\ref{fig:orientation_clustering}(A,C) are shown in Fig.~\ref{fig:orientation_clustering}(D,E) respectively.
-Note how the thin and curved purple cluster Fig.~\ref{fig:orientation_clustering}(E) bundle ``7'' is extracted well.
+The results of clustering the orientation clusters (see Figure~\ref{fig:orientation_clustering}a,c) are shown in Figure~\ref{fig:orientation_clustering}d,e respectively.
+Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e).
 
-In comparison Fig.~\ref{fig:comparison} shows the MetaTracts in Fig.~\ref{fig:length_distribution}(A) being hierarchically clustered directly using proximity alone into 10,15 and 20 clusters (ground truth being 11).
-As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low cardinality outliers.
-In Fig.~\ref{fig:comparison} when ``h" is 10 two large multi-bundle incorrect cluster is generated and the rest are outliers.
-As ``h" increases, some appropriate bundles start to form.
-Even at ``h" equals 20, the top bundles all clusters together into one large incorrectly identified bundle.
+In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11).
+As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers.
+When number of clusters ($h$) is 10 two large multi-bundle incorrect clusters are generated and the rest are outliers.
+As $h$ increases, some appropriate bundles start to form.
+Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
 \begin{figure}
 	\centering
-		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
-		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
-	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
-	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
-	\caption{(A) Volume rendering of data set Figure~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
-	\label{fig:crop-16-decomp}
+	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
+	\caption{MetaTracts data (left) and its straight-forward hierarchical clustering for various numbers of clusters.}
+	\label{fig:comparison}
 \end{figure}
@@
 \label{sec:vis}
-Apart from direct visualization of the MetaTracts, we show two simple yet highly important extensions for our domain specialist.
-The first one is to voxelize the original volume according to the clusters each voxel is associated with.
+
+Apart from direct visualization of the MetaTracts, we show two additional extensions which were requested by our domain specialists as highly important.
+The first one is voxelizing the original volume according to the clusters each voxel is associated with.
 The second is to extract the corresponding surfaces from the voxel data by binarizing the volume per cluster and extracting the isosurface of the largest connected component from the binary volume.
@@
 The voxel is then assigned to the class with the maximum number of elements in the neighborhood.
-Figure~\ref{fig:crop-16-decomp}A and Figure~\ref{fig:prepreg}D shows the result of voxelization.
-Figure~\ref{fig:crop-16-decomp}(C, D) show examples of extracted meshes.
+Figures ~\ref{fig:crop-16-decomp}a and~\ref{fig:prepreg}d show the result of voxelization.
+Figure~\ref{fig:crop-16-decomp}c,d shows examples of extracted meshes.
+
+\begin{figure}
+\centering
+		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
+		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
+	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
+	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
+	\caption{(A) Volume rendering of data set Figure~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
+	\label{fig:crop-16-decomp}
+\end{figure}
+
 %The MetaTracts and the clustering provide an abstraction of the fiber bundles and are used as a visualization tool.
@@
 We tested our method on multiple datasets with varying characteristics. 
-Figure~\ref{fig:orientation_clustering} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into ``h" equals 10 clusters (the ground truth is 6 and 5 respectively).
+Figure~\ref{fig:orientation_clustering} shows the result of dataset~\ref{fig:data-char} decomposed into 2 orientation clusters with each orientation orientation cluster decomposed into $h=10$ clusters (the ground truth is 6 and 5 respectively).
 The hierarchical clustering is robust to over segmentation.
@@
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
-As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 (Fig.~\ref{fig:length_distribution}) is 200 while the maximum MetaTract length in the bundle is about 500 (measured in units of grid cube length: $2\mu m$).
-Fig.~\ref{fig:length_distribution}(B) shows the length per MetaTract distribution for a particular fiber bundle.
-Fig.~\ref{fig:prepreg} shows the results on our dataset2.
+As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 (Figure~\ref{fig:length_distribution}) is 200 while the maximum MetaTract length in the bundle is about 500 (measured in units of grid cube length: 2 $\mu m$).
+Figure~\ref{fig:length_distribution}b shows the length per MetaTract distribution for a particular fiber bundle.
+Figure~\ref{fig:prepreg} shows the results on our dataset 2.
 This dataset is dense with \textit{flat and thin} bundles.
-The dimensions are 300x350x300 and the datatype is uint16 (compared to dataset1 which was only uint8).
-The bundles are indistinguishable in the original data set Fig.~\ref{fig:prepreg}(A).
+The dimensions are 300$\times$350$\times$300 and the datatype is uint16 (dataset 1 is uint8).
+The bundles are indistinguishable in the original data set Figure~\ref{fig:prepreg}a.
 A green dotted line shows one of the fiber bundle boundary.
-Fig.~\ref{fig:prepreg}(B) shows the fiber bundles for the two orientation.
-Fig.~\ref{fig:prepreg}(C) shows the combined results.Fig.~\ref{fig:prepreg}(D) shows the volume after voxelization.
+Figure~\ref{fig:prepreg}b shows the fiber bundles for the two orientations, Figure~\ref{fig:prepreg}c shows the combined results, and Figure~\ref{fig:prepreg}d shows the volume after the voxelization.
 The fiber bundle extraction using MetaTracts was implemented in C++ using ITK.
@@
 The preprocessing consists of MetaTract generation the distance computation and the orientation clustering.
-The hierarchical clustering separately, on the results of the orientation clusters can be done quickly in a few minutes on a Intel Xeon E5-2667 workstation according to user input `h'(section~\ref{sec:param_choices}).
-We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without extra computation.
+The hierarchical clustering separately, on the results of the orientation clusters can be done withing an order of a few minutes on Intel Xeon E5-2667 workstation according to user input of $h$ (section~\ref{sec:param_choices}).
+We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without any extra computations.
 
@@
 We compared  $m=3-7$ experimentally without any dramatic change in results.
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
-	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
-	\label{fig:comparison}
-\end{figure} 
+
 %\begin{figure}

### 2014-09-24 chapters/MetaTracts.tex
figure
Updated with orientation comparison text

--- 2014-09-24 chapters/MetaTracts.tex
@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components.
-The materials of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry.
+The material of choice of a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry.
 CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing  parts made of conventional materials such as aluminum or steel.
@@
 CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
-At the same time highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
-Primary structures and highly loaded components in aeronautics are one example.
+At the same time highly complex and integrated components which were previously  impossible to manufacture may be produced from CFRP's.
+For example even primary structures and highly loaded components in aeronautics are being increasingly manufactured from CFRP.
 
-Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates, consist of two main components:
-A matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.
+Typically,  carbon fiber reinforced polymer components, more specifically CFRP laminates consists of two main components.
+A matrix on the one side, which acts as a bonding component and  reinforcements on the other side, which allow for achieving the desired characteristics.
 Various production processes are used to manufacture CFRP laminates with endless carbon fibers.
@@
 Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
-The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles.
-The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
+The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles.
+The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
 Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
-On one hand ultrasonic testing provides a quick and cost-efficient overview of the material.
+On the one hand ultrasonic testing provides a quick and cost-efficient overview of the material.
 On the other hand, it shows low resolution and may generate arbitrary results, e.g.
@@
 
-Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
+Industrial 3D X-Ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly being applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
 XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen.
-Typical scanning geometry in industrial XCT is the specimen placed on a rotary table between X-ray source and detector.
+In industrial XCT, the typical scanning geometry places the specimen on a rotary table between X-ray source and detector.
 X-rays are emitted by the source in the form of a conic beam.
@@
 Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen.
-In state of the art devices, XCT can reach voxel sizes below 500 $nm$.
+In state of the art devices, XCT can reach voxel sizes below $500 nm$.
 XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens.
@@
 
-In this work, we focus on datasets with larger viewports and lower resolutions where the individual carbon fibers (filaments) are indiscernible or are barely visible.
+In this work, we focus on datasets with larger viewports and lower resolutions where the single carbon fibers (filaments) are not or are  barely visible.
 Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of  individual fibers.
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
-Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
-On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
+Fig.~\ref{fig:data-char} provides an example of our targeted datasets. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
+Our work is motivated from the recent progress in two interrelated fields: On the one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials \cite{Karpat2012}.
+On the other hand, recent developments of industrial 3D X-Ray computed tomography (XCT) with regard to larger detectors, larger field of views and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
 
@@
 We detail and evaluate a method which uses visualization techniques to gain insight into our data.
-We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}.
+We demarcate the above queries into two parts, ``geometric structure" and ``spatial context".
 Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles.
 Spatial context refers to how two or more bundles interact with each other.
-We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites.
-The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
+We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites.
+The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.
 
-The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  section \ref{sec:char_data} describes the   characteristics of our data.
-Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provide details of our approach.
-Section \ref{sec:results} details experiments and section~\ref{sec:user_eval} summarizes our user evaluation results.
+The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  Section \ref{sec:char_data} describes the   characteristics of our data.
+Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provides details of our approach.
+Section \ref{sec:results} details experiments, section~\ref{sec:user_eval} summarizes our user evaluation results.
 Section~\ref{sec:param_choices} explains our  parameter choices.
@@
 \includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
-\caption{Data characteristics: (A) shows a rendering of the data.
-(B) shows 2D slices along the X- and (C) along the Z-axis.
+\caption{Data Characteristics.
+(A) shows a rendering of the data.
+(B) shows 2D slices along the X and (C) along the Z axis.
 (D) close up of the region marked in green in (C), multiple fiber bundles cross and are indistinguishable by visual inspection alone.
@@
 
-Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
-Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$, the data type is unsigned 8 bit integer.
-Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
+Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles.
+Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the data type is unsigned 8 bit integer.
+Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
 The unit cell is used as a basis structure for CFRP manufacturing.
 The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
-Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset.
-Figure~\ref{fig:data-char}(B) shows 2D slices along the $X$- and Figure~\ref{fig:data-char}(C) in $Z$-axis.
-Figure~\ref{fig:data-char}(D) shows the magnified version of the green region of interest.
-The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only.
-Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions.
+Fig.~\ref{fig:data-char}(A) shows a volume rendering of the dataset.
+Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis.
+Fig.~\ref{fig:data-char}(D) shows the magnified version of the green square.
+The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only.
+Fig.~\ref{fig:data-char}(D) contains two bundles going in opposite directions.
 However, the separation between two fiber bundles is barely visible.
 The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization.
-The fiber bundles may differ in terms of the amount of fibers in the bundle.
-Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
+The fiber bundles themselves may differ in terms of the amount of fibers in the bundle.
+Fig.~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
 Depending on the weaving pattern the fiber bundles cross each other in different orientations.
@@
 Nevertheless, the weaving pattern may cause the individual bundles to be curved.
-In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.
+In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in Euclidean distance but belong to different bundles.
 We make the following assumptions on our data:
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item The structure embedded in the data contains fiber bundles of indiscernible fibers.
-\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
+\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
 \item The local orientation may change gradually inside the fiber bundle.
 \item The local orientation may be noisy and not reliable.
-\item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
+\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
 \item Fiber bundles going in different directions only interact near the surface of contact.
@@
 %\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
-\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
+\includegraphics[width=0.5\textwidth]{images_pvis/workflow.pdf}
 \caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
   \centering  	
-  \includegraphics[width=0.5\textwidth]{images_pvis/figure6.png}
+  \includegraphics[width=0.5\textwidth]{images_pvis/figure6.pdf}
   %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
@@
 Even at ``h" equals 20, the top bundles all clusters together into one large incorrectly identified bundle.
-\begin{figure}
+
+On the other hand, Fig.~\ref{fig:comparison_orient} shows the MetaTracts in Fig.~\ref{fig:length_distribution}(A) being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone into 10, 15, 20 clusters (ground truth being 11).
+In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
+\begin{figure}[tb]
 \centering
@@
 \label{sec:param_choices}
-\begin{figure}
+\begin{figure}[tb]
 \centering
-	\includegraphics[width=0.45\textwidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
+	\includegraphics[width=0.5\textwidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
@@
 	\label{fig:comparison}
+\end{figure}
+
+\begin{figure}[tb]
+	\centering
+	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/comparison_orient}
+	\caption{Direct K-means clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A), embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
+	\label{fig:comparison_orient}
 \end{figure} 

### 2014-09-25 chapters/MetaTracts.tex

--- 2014-09-24 chapters/MetaTracts.tex
@@
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
+
 Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
@@
 \label{sec:char_data}
-
 Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
@@
 In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.
+
 We make the following assumptions on our data:
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item The structure embedded in the data contains fiber bundles of indiscernible fibers.
@@
 We generate a rectangle of length $L$ and radius $R$, the green dashed rectangle shows the rectangle generated at $C_p$ (properties 3 and 5 in Section~\ref{subsec:mtprop}).
+
 \begin{figure}[tb]
@@
 We refer to this perpendicular projection distance as \textit{`projected\_dist'}.
+
 \end{itemize}
@@
 Figure~\ref{fig:length_distribution}b shows the MetaTracts of a particular bundle colored according to their individual length.
+
 MetaTracts in a given fiber bundle do not extend the full length of the bundle. MetaTracts may have different lengths and may partially overlap.
@@
 The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$ but the directed Hausdorff is not.
+
 Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially covers the bundle (Figure~\ref{fig:length_distribution}b).
 If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$, then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
+
 Thus, $H(P,Q)$ will be large, even though $P$ and $Q$ are in the same fiber bundle.
@@
 
+In comparison Fig.~\ref{fig:comparison} shows the MetaTracts in Fig.~\ref{fig:length_distribution}(A) being hierarchically clustered directly using proximity alone into 10,15 and 20 clusters (ground truth being 11).
+As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low cardinality outliers.
+In Fig.~\ref{fig:comparison} when ``h" is 10 two large multi-bundle incorrect cluster is generated and the rest are outliers.
+As ``h" increases, some appropriate bundles start to form.
+Even at ``h" equals 20, the top bundles all clusters together into one large incorrectly identified bundle.
+
+On the other hand, Fig.~\ref{fig:comparison_orient} shows the MetaTracts in Fig.~\ref{fig:length_distribution}(A) being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone into 10, 15, 20 clusters (ground truth being 11).
+In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
+%\begin{figure}[tb]
+%\centering
+%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
+%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
+%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
+%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
+%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
+%	\label{fig:crop-16-decomp}
+%\end{figure}
+
 In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11).
@@
 	\centering
-	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
+	\includegraphics[width=0.5\textwidth]{images_pvis/figure10}
 	\caption{MetaTracts data (left) and its straight-forward hierarchical clustering for various numbers of clusters.}
@@
 in simulations.
+
 To voxelize the space,  we compute a neighborhood around each voxel.
@@
 We tested our method on multiple datasets with varying characteristics.
+
 Figure~\ref{fig:orientation_clustering} shows the result of dataset~\ref{fig:data-char} decomposed into 2 orientation clusters with each orientation orientation cluster decomposed into $h=10$ clusters (the ground truth is 6 and 5 respectively).
@@
 We compared  $m=3-7$ experimentally without any dramatic change in results.
+\begin{figure}[tb]
+	\centering
+	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
+	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
+	\label{fig:comparison}
+\end{figure}
+
+\begin{figure}[tb]
+	\centering
+	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/comparison_orient}
+	\caption{Direct K-means clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A), embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
+	\label{fig:comparison_orient}
+\end{figure}
 

### 2014-09-25 chapters/MetaTracts.tex
Local text changes

@@
 
-In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength, high corrosion resistance.
+In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
 CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
-At the same time highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
+At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
 Primary structures and highly loaded components in aeronautics are one example.
@@
 Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates, consist of two main components:
-A matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.
+a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.
 Various production processes are used to manufacture CFRP laminates with endless carbon fibers.
@@
 Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
-On one hand ultrasonic testing provides a quick and cost-efficient overview of the material.
+On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material.
 On the other hand, it shows low resolution and may generate arbitrary results, e.g.
@@
 X-rays are emitted by the source in the form of a conic beam.
-When passing through the specimen the X-rays get attenuated by the material in the specimen.
+When passing through the specimen the X-rays are attenuated by the material in the specimen.
 The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light.
@@
 Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion.
- This is usually done by solving a second-order Runge-Kutta integration. The principal diffusion direction at each discrete location is interpolated to from a continuous velocity field. 
+ This is usually done by solving a second-order Runge-Kutta integration. The principal diffusion direction at each discrete location is interpolated to form a continuous velocity field. 
  Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers.
-Basser et al.~\cite{Basser2002},~\cite{Basser2000} proposed that white matter tracts could be represented as 3D curves in space.
+Basser et al.~\cite{Basser2002,Basser2000} proposed that white matter tracts could be represented as 3D curves in space.
 They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data.
-Mori et al.~\cite{Mori1999},~\cite{Mori2002} showed that tract reconstruction techniques could be broadly divided into line-propagation or energy minimization techniques.
+Mori et al.~\cite{Mori1999,Mori2002} divided reconstruction techniques into line-propagation or energy minimization techniques.
 In line propagation approaches, trajectories are computed based on local neighborhoods and in energy minimization approaches the most  favorable trajectory connecting two given endpoints is selected.
@@
 
-Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}) can be applied to the problem of organizing fiber trace data.
+Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}), can be applied to the problem of organizing fiber tracts data.
 The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph.
@@
 An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure.
-In the case of fiber bundles, the individual points are fiber traces.
-In the ideal case fiber traces that belong to the same bundle must remain ``close" to each other in the lower dimensional space.
-Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of pointwise minimum distances between two fibers.
+In the case of fiber bundles, the individual points are fiber tracts.
+In the ideal case fiber tracts that belong to the same bundle must remain ``close" to each other in the lower dimensional space.
+Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of point wise minimum distances between two fibers.
 Jonasson et al.~\cite{jonasson2005} run k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
@@
 The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
-These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationship.
+These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships.
 An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster.
@@
 Second, a ``correct'' distance measure to compare tracts must be chosen.
-Thirdly, hierarchical clustering is best suited for similar length fibers.
+Third, hierarchical clustering is best suited for similar length fibers.
 Spectral methods are also hindered by long matrix computations.
@@
 \subsection {Visual Analysis of Fiber Reinforced Composites}
-The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores: Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
+The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores.
+Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
 This approach allows analyzing fiber orientations based on direction transfer functions.
@@
 
-Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes has received extensive research and recently the application of similar techniques to fiber tracing in composite materials has gained momentum.
+Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research.
+Recently, similar techniques have been applied to fiber tracing in composite materials.
 To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers, which is the main scope of this work.
@@
 \includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
-\caption{Data characteristics: (A) shows a rendering of the data.
-(B) shows 2D slices along the X- and (C) along the Z-axis.
-(D) close up of the region marked in green in (C), multiple fiber bundles cross and are indistinguishable by visual inspection alone.
+\caption{Data characteristics: (a) Rendering of the data.
+(b) 2D slices along the X-axis and (c) 2D slice along the Z-axis.
+(d) Magnified region marked in green square (c).
+Multiple fiber bundles cross and are indistinguishable by visual inspection alone.
 }
@@
 Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
-Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$, the data type is unsigned 8 bit integer.
+Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$.
+The data type is unsigned 8 bit integer.
 Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
@@
 Figure~\ref{fig:data-char}d contains two bundles going in opposite directions.
-However, the separation between two fiber bundles is barely visible.
+The separation between two fiber bundles is barely visible.
 The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization.
@@
 Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
-Depending on the weaving pattern the fiber bundles cross each other in different orientations.
-In general the number of orientations is limited by the weaving pattern.
+Depending on the weaving pattern, the fiber bundles cross each other in different orientations.
+In general, the number of orientations is limited by the weaving pattern.
 Nevertheless, the weaving pattern may cause the individual bundles to be curved.
@@
 We compared  $m=3-7$ experimentally without any dramatic change in results.
-\begin{figure}[tb]
-	\centering
-	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
-	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
-	\label{fig:comparison}
-\end{figure}
+%\begin{figure}[tb]
+%	\centering
+%	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
+%	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
+%	\label{fig:comparison}
+%\end{figure}
+%
 

### 2014-09-25 metaTracts-pvis2014.tex
local commit
reviewed user study, modified enums + itemizes
Local text changes

@@
 
-\documentclass{vgtc}                          % final (conference style)
-%\documentclass[review]{vgtc}                 % review
+%\documentclass{vgtc}                          % final (conference style)
+\documentclass[review]{vgtc}                 % review
 %\documentclass[widereview]{vgtc}             % wide-spaced review
@@
 \usepackage{epstopdf}
+\usepackage{enumitem}
 %% We encourage the use of mathptmx for consistent usage of times font
@@
 \abstract{
-This work introduces MetaTracts, a novel method  for extracting and visualizing individual fiber bundles as well as their weaving patterns from X-ray computed tomography (XCT) scans of  endless carbon fiber reinforced  polymers (CFRP).
+This work introduces MetaTracts, a novel method  for extracting and visualizing individual fiber bundles and weaving patterns from X-ray computed tomography (XCT) scans of  endless carbon fiber reinforced  polymers (CFRP).
 The  proposed work flow is designed  to analyze large areas of CFRP material  such as complete unit cells integrating the recurring weaving pattern.
@@
 First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
-In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach: first by orientation then by proximity.
-The tool can also  generate volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
+In the second step, these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach: first by orientation, then by proximity.
+The tool can generate volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
 We evaluate the proposed work flow on a number of real world datasets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 

### 2014-09-25 chapters/MetaTracts.tex
Updated final ppt with the comparison image

@@
 \centering
-%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
-\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
+%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
+\includegraphics[width=0.5\textwidth]{images_pvis/data-char.pdf}
 \caption{Data characteristics: (a) Rendering of the data.
@@
 MetaTracts generations  is explained in Section.~\ref{sec:approach} and  finding plausible fiber bundles in Section~\ref{subsec:fiber-bundles}, Section~\ref{sec:vis} talks about visualization extensions.
+
+Extracting the MetaTracts consists of two  major steps.
+First we extract a local orientation vector at each grid location (section~\ref{subsec:rh}).
+Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (section~\ref{subsec:mtprop}).
+
 %\subsection {Extracting MetaTracts}
@@
 
-The input to the first stage of our pipelines, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
-As output, each grid location becomes associated with a vector representing the local orientation and a real value [0,1] which represents a measure of the reliability of the local orientation.
+The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
+As output, each grid location becomes associated with a vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation.
 We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
@@
 Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude.
-Here $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix.
+Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix.
 Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$).
@@
 In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$).
-The following measures are defined in their work~\cite{Frangi1998}.
+The following measures are defined in~\cite{Frangi1998}.
 \begin{equation}\label{RA}
@@
 \end{equation}
-In equation~\ref{RB}, $\mathcal{R_{B}}$ provides a measure of deviation from a blob like structure while in equation~\ref{RA}, $\mathcal{R_{A}}$ distinguishes between plate-like and line-like structure.
+In Equation~\ref{RB}, $\mathcal{R_{B}}$ provides a measure of deviation from a blob like structure while in Equation~\ref{RA}, $\mathcal{R_{A}}$ distinguishes between plate-like and line-like structure.
 Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors.
-Thus we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
+Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
 $$
@@
 The utility of the vesselness is a little different in our framework than Frangi et al.~\cite{Frangi1998}.
-First, in biology vesselness is computed for different scales because the vessels can be of different sizes. In our case, usually the width of individual fibers are known a priori.
+First, vesselness in biology is computed for different scales because the vessels can be of different sizes. In our case, usually the width of individual fibers are known a priori.
 Second, and more importantly, we do not have clear tubular structures embedded in a dark contrast matrix such as in blood vessels.
@@
 Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue.
-Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axis respectively.
+Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively.
 Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest.
@@
 Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}.
-The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding, this is not the case for our data.
+The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding.
+This is not the case for our data.
 The local orientaion at each voxel is inherently more noisy.
-Thus techniques popular in DTI cannot be directly applied to XCT data.
+Thus, techniques popular in DTI cannot be directly applied to XCT data.
 
@@
 %\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
-\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
-\caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) Magnified region marked in green square (in C).}
+%\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
+\includegraphics[width=0.5\textwidth]{images_pvis/reliable_hessian.pdf}
+\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square (in c).}
 \label{fig:reliable_hessian}
@@
 Thus instead of building fiber traces, we wish to find an abstract representation of the fibers.
-We build on the key assumptions on the data, namely `local orientation' and `connectivity' while taking into account the noise and lack of resolution.
+We build on the key assumptions on the data, namely ``local orientation" and ``connectivity" while taking into account the noise and lack of resolution.
 We do this by interpreting the underlying geometric structure of the fibers as a set of polylines (cylinders).
@@
 \item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
-\item{Each cylinder is locally parallel to the local orientation at it's start point.}
+\item{Each cylinder is locally parallel to the local orientation at its start point.}
 \end{enumerate}
@@
 \label{subsec:MetaTracts-generation}
- In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$.
-The procedure trivially extends to $\mathbb{R}^3$.
-In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles.
-The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1].
-In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue.
+ In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$.The procedure trivially extends to $\mathbb{R}^3$.
+In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles.
+The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1].
+In Figure \ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue.
 Figure~\ref{fig:algo}c shows a single cylinder of a MetaTracts.
@@
 \begin{itemize}
-\item \textit{Orientation similarity}: We want the orientation of the start points for the consecutive cylinders to have a similar orientations ($N_p$). 
-\item \textit{Large distance}: We want the MetaTracts to traverse the data using as few cylinders as possible.
-Thus, we want the distance between $C_p$ and the start point for the next cylinder to be as large as possible.
-We measure distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$.
-For example the Euclidean distance between the blue and the orange vertices projected onto $N_P$ gives distance (Figure~\ref{fig:algo}B).
+\item \textit{Orientation similarity}: The orientation of the start points ($N_p$) for the consecutive cylinders should be similar. 
+\item \textit{Large distance}: The MetaTracts should traverse the data using as few cylinders as possible.
+Thus, the distance between $C_p$ and the start point for the next cylinder should be as large as possible.
+We measure the distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$.
+
+For example, in Figure~\ref{fig:algo}b the distance is measured as the Euclidean distance between the purple and the orange vertices projected on to $N_p$.
 We refer to this perpendicular projection distance as \textit{`projected\_dist'}.
@@
 We define a priority for each of the candidate vertices based on the above factors.
-For each cylinder in a MetaTract, we put the \textit{candidate vertices} of it in a priority queue based on Equation \ref{eqn:algo_1}.
+For each cylinder in a MetaTract, we put its \textit{candidate vertices} in a priority queue based on Equation \ref{eqn:algo_1}.
 \begin{equation}
@@
 $\gamma_1$, $\gamma_2$ are the  weights ($\mathbb{R}_{\ge 0}$)  which decide how the priority depends on the affine combination of the two factors.
-For all our cases we use $\gamma_1$ = $1 / 3 $ and $\gamma_2$ = $2 / 3$.
-In general we suggest $\gamma_1 +\gamma_2 = 1 $ and $\gamma_1 \leq \gamma_2$.
-At each iteration we pick the top element in the priority queue generate the corresponding cylinder and repeat the steps.
-Essentially, Equation~\ref{eqn:algo_1} selects a grid point which is the furthest from the current point and going in a similar direction.
-This gives it the advantage of tackling noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation in an already noisy environment.
-
-In this particular example in Figure~\ref{fig:algo}b we select the orange grid vertex next and repeat the process.
+For all our cases, we use $\gamma_1$ = $1 / 3 $ and $\gamma_2$ = $2 / 3$.
+In general, we suggest $\gamma_1 +\gamma_2 = 1 $ and $\gamma_1 \leq \gamma_2$.
+At each iteration, we pick the top element in the priority queue to generate the corresponding cylinder, and repeat the steps.
+Essentially, Equation~\ref{eqn:algo_1} selects a grid point which is furthest from the current point and is going in a similar direction.
+This approach tackles noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation in an already noisy environment.
+
+In the example in Figure~\ref{fig:algo}b we select the orange grid vertex next and repeat the process.
 We do not select the purple vertex (Figure~\ref{fig:algo}c) because it is not a candidate vertex.
 If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length.
-Short tracts are then removed.
+Short tracts (less than 0.01 times the smallest axes size) are removed.
 The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}a.
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
-Figure\ref{fig:length_distribution}a shows a single slice of YZ plane.
-That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
+Figure~\ref{fig:length_distribution}a shows a single slice of YZ plane.
+Consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.
@@
 Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering.
-Experimentally we found that orientation measure was very reliable while measures of geometric proximity had problems with fibers which partially overlap (Figure~\ref{fig:length_distribution}b).
+Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with fibers which partially overlap (Figure~\ref{fig:length_distribution}b).
 Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
@@
 This step divides the individual MetaTracts into classes based on their major orientations.
-In order to cluster MetaTracts going in same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01} and later used in DTI fiber coloring~\cite{Brun2003} and fiber clustering~\cite{Brun2004}.
-The key notion is to find a suitable similarity measure to define the weights of an undirected weighted graph.
-The nodes being the data points and the edges being the weights.
-Then, an eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
-This lower dimensional space can then be clustered more efficiently than the higher dimensional space. 
-Let $G$ be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem:$L\textit{\textbf{f}}=\lambda D\textit{\textbf{f}}$,
+In order to cluster MetaTracts going in same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01}.
+An eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
+Let $G$ be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem $L\textit{\textbf{f}}=\lambda D\textit{\textbf{f}}$,
 %\begin{equation}\label{equn:eigenMaps}
@@
 where $D$ is the diagonal weight matrix and $L$ is the Laplacian matrix.
-The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an $m$-dimensional space.
-K-means clustering is then used to cluster in this lower dimensional space.
-
-Brun et al.~\cite{Brun2003} based their measure on the simple assumption that two traces with similar endpoints should be considered similar.
-Brun et al.
-~\cite{Brun2004} based their measure on a 9-D fiber descriptors.
+The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an $m$-dimensional space (see section~\ref{sec:param_choices} for values of m).
  In our case, each MetaTract is a data point.
@@
 
-Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially covers the bundle (Figure~\ref{fig:length_distribution}b).
+Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTracts, which only partially cover the bundle (Figure~\ref{fig:length_distribution}b).
 If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$, then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
@@
 
-In comparison Fig.~\ref{fig:comparison} shows the MetaTracts in Fig.~\ref{fig:length_distribution}(A) being hierarchically clustered directly using proximity alone into 10,15 and 20 clusters (ground truth being 11).
-As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low cardinality outliers.
-In Fig.~\ref{fig:comparison} when ``h" is 10 two large multi-bundle incorrect cluster is generated and the rest are outliers.
-As ``h" increases, some appropriate bundles start to form.
-Even at ``h" equals 20, the top bundles all clusters together into one large incorrectly identified bundle.
-
-On the other hand, Fig.~\ref{fig:comparison_orient} shows the MetaTracts in Fig.~\ref{fig:length_distribution}(A) being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone into 10, 15, 20 clusters (ground truth being 11).
-In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
+\subsection{Choice of clustering techniques}
+As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
 %\begin{figure}[tb]
@@
 
-In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11).
+In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters).
 As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers.
-When number of clusters ($h$) is 10 two large multi-bundle incorrect clusters are generated and the rest are outliers.
+When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers.
 As $h$ increases, some appropriate bundles start to form.
@@
 \end{figure} 
+On the other hand, Figure.~\ref{fig:comparison_orient} shows the MetaTracts in Figure.~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone into 10, 15, 20 clusters.In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
 
@@
 Apart from direct visualization of the MetaTracts, we show two additional extensions which were requested by our domain specialists as highly important.
-The first one is voxelizing the original volume according to the clusters each voxel is associated with.
+The first extension is voxelizing the original volume according to the clusters each voxel is associated with.
 The second is to extract the corresponding surfaces from the voxel data by binarizing the volume per cluster and extracting the isosurface of the largest connected component from the binary volume.
@@
  %glass
-We tested our method on multiple datasets with varying characteristics.
-
-Figure~\ref{fig:orientation_clustering} shows the result of dataset~\ref{fig:data-char} decomposed into 2 orientation clusters with each orientation orientation cluster decomposed into $h=10$ clusters (the ground truth is 6 and 5 respectively).
-The hierarchical clustering is robust to over segmentation.
-Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster.
+We tested our method on three datasets with varying characteristics. We report results on two due to lack of space.  
+Dataset 1 is described in section~\ref{sec:char_data}. Dataset 1 has eleven separate fiber bundles which we identified correctly.
+Figure~\ref{fig:orientation_clustering} shows the result of dataset~\ref{fig:data-char} decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters (the ground truth is 6 and 5 respectively).
+
+The hierarchical clustering is robust against over segmentation. For example,
+Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Figure~\ref{fig:orientation_clustering}d).
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
-As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 (Figure~\ref{fig:length_distribution}) is 200 while the maximum MetaTract length in the bundle is about 500 (measured in units of grid cube length: 2 $\mu m$).
+As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster six (Figure~\ref{fig:len_dist_crop16}) is 200 while the maximum MetaTract length in the bundle is about 500 (measured in units of grid cube length: 2 $\mu m$).
 Figure~\ref{fig:length_distribution}b shows the length per MetaTract distribution for a particular fiber bundle.
+
 Figure~\ref{fig:prepreg} shows the results on our dataset 2.
@@
 Figure~\ref{fig:prepreg}b shows the fiber bundles for the two orientations, Figure~\ref{fig:prepreg}c shows the combined results, and Figure~\ref{fig:prepreg}d shows the volume after the voxelization.
+
 The fiber bundle extraction using MetaTracts was implemented in C++ using ITK.
 All clustering was done in R.
-The preprocessing consists of MetaTract generation the distance computation and the orientation clustering.
-The hierarchical clustering separately, on the results of the orientation clusters can be done withing an order of a few minutes on Intel Xeon E5-2667 workstation according to user input of $h$ (section~\ref{sec:param_choices}).
+The preprocessing is time consuming and consists of MetaTract generation and the distance computation.
+The orientation clustering is robust and is done as a part of preprocessing.
+The hierarchical clustering separately, on the results of the orientation clusters can be done within an order of a few minutes on Intel Xeon E5-2667 workstation and can be repeated according to user input of $h$ (section~\ref{sec:param_choices}).
 We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without any extra computations.
@@
 Large $h$ keeps the major bundles intact.
-For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements).
+For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements).
 This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
@@
 A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.
-On the other hand a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced.
+On the other hand, a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced.
 The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al.
@@
 We used 10.00 and 2.00 for length and breadth(measured in grid cube length), respectively for all tests.
-$\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays, we have used [7-10] and half the length of cylinder (5.0).
-Our number of fiber bundle directions are limited, thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well.
+$\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays; we have used [7-10] and half the length of cylinder (5.0).
+Our number of fiber bundle directions are limited.
+Thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well.
 We compared  $m=3-7$ experimentally without any dramatic change in results.
@@
 It handles complex fiber patterns such as ``cross overs'' and ``braiding''.
-The presented techniques are attractive and novel features for a industrial fiber bundle visualization tool.
+We perform an user evaluation to indicate the usefulness of the method.
+In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up the extraction.
 %

### 2014-09-25 chapters/MetaTracts.tex
Added UserEval2
Added UserEval2 and updated it on pvisImages_final

@@
 The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
-
-The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  Section \ref{sec:char_data} describes the   characteristics of our data.
-Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provide details of our approach.
-Section \ref{sec:results} details experiments and Section~\ref{sec:user_eval} summarizes our user evaluation results.
-Section~\ref{sec:param_choices} explains our  parameter choices.
-We end with conclusions in Section~\ref{sec:conclusions}.
+\begin{figure}[h]
+	\centering
+	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
+	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
+	\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
+	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
+	\label{fig:flowchart}
+\end{figure}
+Figure~\ref{fig:flowchart} shows an over view of our approach.
+Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters.
+For a detailed description of the MetaTracts generation process, see section~\ref{subsec:MetaTracts-generation}.
+For the description of plausible fiber bundle generation see section~\ref{subsec:fiber-bundles}.
+For a discussion of the visualizations we create for our domain experts see section ~\ref{sec:vis}.
+Section~\ref{sec:results} describes our experimental results, Section~\ref{sec:param_choices} presents our parameter choices and  Section~\ref{sec:user_eval} summarizes our user evaluations.
 
@@
 We broadly divide the related work on DTI into two parts that are of immediate relevance to our proposed solution: fiber tracking and fiber clustering.
-We discuss  methods for analyzing the second order local structure.
-Finally, we review the current state of the art in the visual analysis of fiber reinforced composites.
+We discuss  methods for analyzing the second order local structure and review the current state of the art in the visual analysis of fiber reinforced composites.
 
@@
 \label{sec:approach}
-
-\begin{figure}[t]
-	\centering
-%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
-%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
-\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
-\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
-\label{fig:flowchart}
-\end{figure}
-
-Fig.
-\ref{fig:flowchart} shows our pipeline.
-MetaTracts generations  is explained in Section.~\ref{sec:approach} and  finding plausible fiber bundles in Section~\ref{subsec:fiber-bundles}, Section~\ref{sec:vis} talks about visualization extensions.
-
 Extracting the MetaTracts consists of two  major steps.
@@
 
-In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters).
+In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
 As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers.
@@
 Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
-\begin{figure}
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=0.5\textwidth]{images_pvis/figure10}
+	\includegraphics[width=0.5\textwidth,  trim = 0mm 10mm 0mm 10mm, clip]{images_pvis/comparison_proximity.pdf}
 	\caption{MetaTracts data (left) and its straight-forward hierarchical clustering for various numbers of clusters.}
@@
 \end{figure} 
-On the other hand, Figure.~\ref{fig:comparison_orient} shows the MetaTracts in Figure.~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone into 10, 15, 20 clusters.In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth]{images_pvis/comparison_orient.pdf}
+	\caption{Direct K-means clustering of MetaTracts (left) embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
+	\label{fig:comparison_orient}
+\end{figure}
+On the other hand, Figure.~\ref{fig:comparison_orient} shows the MetaTracts in Figure.~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone, without a second hierarchical clustering step.
+We set ``k" to be  10, 15, 20 clusters.
+In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
 
@@
 	\centering
-	\includegraphics[width=0.45\textwidth,  trim = 0mm 60mm 0mm 0mm, clip]{images_pvis/usereval}
+	\includegraphics[width=0.45\textwidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval2}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
@@
 
-\begin{figure}[tb]
-	\centering
-	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/comparison_orient}
-	\caption{Direct K-means clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A), embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
-	\label{fig:comparison_orient}
-\end{figure}
-
 %\begin{figure}

### 2014-09-25 chapters/MetaTracts.tex
top or main eigenvector, consistency!
major figure rework

@@
 The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
-\begin{figure}[h]
+\begin{figure}[tb]
 	\centering
@@
 	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
-	\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
+	\includegraphics[width=0.475\textwidth]{images_pvis/workflow.pdf}
 	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
 %\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
-\includegraphics[width=0.5\textwidth]{images_pvis/data-char.pdf}
+\includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
 \caption{Data characteristics: (a) Rendering of the data.
@@
 %\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
-\includegraphics[width=0.5\textwidth]{images_pvis/reliable_hessian.pdf}
+\includegraphics[width=\linewidth]{images_pvis/reliable_hessian.pdf}
 \caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square (in c).}
@@
   \begin{center}
- \includegraphics[ width=0.3\textwidth]{imagesMT2014/image4}
+ \includegraphics[ width=\linewidth]{images_pvis/mt_gen.pdf}
   \end{center}
@@
   \centering  	
-  \includegraphics[width=0.5\textwidth]{images_pvis/figure6.png}
+  \includegraphics[width=\linewidth]{images_pvis/clustering.pdf}
   %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
@@
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
-	\includegraphics[width=0.4\textwidth, trim = 0mm 30mm 0mm 0mm, clip,]{images_pvis/figure8}
+	\includegraphics[width=\linewidth]{images_pvis/dataset2.pdf}
 	\caption{Data set with flat thin and compact bundles.

### 2014-09-25 chapters/MetaTracts.tex
full paper pass

--- 2014-09-25 chapters/MetaTracts.tex
@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components.
-The materials of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry.
+The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.
 CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing parts made of conventional materials such as aluminum or steel.
-CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.
-
-In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
-CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
+But the application areas of CFRPs are not limited on these two industries.
+For a wide range of different areas ranging from marine and civil engineering, wind energy, to sporting goods manufacturing, CFRP is highly interesting.
+The main reason therefore is found in CFRP materials having desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
+CFRP materials show these characteristics at considerably lower weight compared to conventional materials.
 At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
@@
 
-Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates, consist of two main components:
+Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates we are addressing in this work, consist of two main components:
 a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.
@@
 Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth.
-Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties.
+Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate.
 Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
 The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles.
+After curing, the production process of the CFRP laminate is finished.
+
 The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
@@
 due to the geometry of the final component.
-
 Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
 XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen.
-Typical scanning geometry in industrial XCT is the specimen placed on a rotary table between X-ray source and detector.
+In the typical scanning geometry of industrial XCT, the specimen is placed on a rotary table between X-ray source and detector.
 X-rays are emitted by the source in the form of a conic beam.
@@
 
-In this work, we focus on datasets with larger viewports and lower resolutions where the individual carbon fibers (filaments) are indiscernible or are barely visible.
+In this work, we focus on datasets with larger viewports and lower resolutions where the individual carbon fibers (filaments) are indiscernible or barely visible.
 Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of individual fibers.
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-
-Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
-Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
+Fig.~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
+Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
 On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
-
+%
 % Fiber bundles determine to a large extent the properties in terms of specific stiffness, specific strength, or durability of the final component.
@@
 %how well do we do in noise , may be data set 1
-
+%
 %!!!!! I have replaced this with the following para !!!!!
@@
 The main goal of this work is to expand the state of the art in nondestructive testing of composite structures towards lower resolutions and thus larger areas of woven fabrics.
-
+%
 %While the importance of fiber bundles in determining the final properties of the components have  become evident,  the tools for visualization of the same have not developed at the same pace.
-
+%
 While fiber bundles are now understood as highly important in determining component properties, the tools for visualization of the internal structure have not developed at the same pace.
@@
 	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
-	\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
+	\includegraphics[width=0.45\textwidth]{images_pvis/workflow.png}
 	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
 \end{figure}
-Figure~\ref{fig:flowchart} shows an over view of our approach.
+Fig.~\ref{fig:flowchart} shows an over view of our approach.
 Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters.
-For a detailed description of the MetaTracts generation process, see section~\ref{subsec:MetaTracts-generation}.
-For the description of plausible fiber bundle generation see section~\ref{subsec:fiber-bundles}.
-For a discussion of the visualizations we create for our domain experts see section ~\ref{sec:vis}.
-Section~\ref{sec:results} describes our experimental results, Section~\ref{sec:param_choices} presents our parameter choices and  Section~\ref{sec:user_eval} summarizes our user evaluations.
+For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}.
+For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}.
+For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}.
+Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.
 
@@
 Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}.
-Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge are clustered separately.
+Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge, are clustered separately.
 Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors.
@@
 %Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results. 
-
 There are a number of difficulties in hierarchical clustering.
@@
 \subsection {Visual Analysis of Fiber Reinforced Composites}
-The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores.
+The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores and deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.
 Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
@@
 Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers.
-Like~\cite{Weissenbock2014}, most carbon fiber extraction literature deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.
-
 Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research.
@@
 %\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
-\includegraphics[width=0.5\textwidth]{images_pvis/data-char.pdf}
+\includegraphics[width=0.45\textwidth]{images_pvis/data-char.pdf}
 \caption{Data characteristics: (a) Rendering of the data.
@@
 \label{sec:char_data}
-Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
+Fig.
+\ref{fig:data-char} shows dataset 1 with woven fiber bundles.
 Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$.
 The data type is unsigned 8 bit integer.
-Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
+Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
 The unit cell is used as a basis structure for CFRP manufacturing.
 The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
-Figure~\ref{fig:data-char}a shows a volume rendering of the dataset.
-Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis.
-Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest.
-The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only.
-Figure~\ref{fig:data-char}d contains two bundles going in opposite directions.
+Fig.~\ref{fig:data-char}a shows a volume rendering of the dataset.
+Fig.~\ref{fig:data-char}b shows 2D slices along the $X$- and Fig.~\ref{fig:data-char}c in $Z$-axis.
+Fig.~\ref{fig:data-char}d shows the magnified version of the green region of interest.
+The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images.
+Fig.~\ref{fig:data-char}d contains two bundles going in opposite directions.
 The separation between two fiber bundles is barely visible.
@@
 The fiber bundles may differ in terms of the amount of fibers in the bundle.
-Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
+Fig.~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
 Depending on the weaving pattern, the fiber bundles cross each other in different orientations.
@@
 In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.
-
 We make the following assumptions on our data:
@@
 Extracting the MetaTracts consists of two  major steps.
-First we extract a local orientation vector at each grid location (section~\ref{subsec:rh}).
-Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (section~\ref{subsec:mtprop}).
+First we extract a local orientation vector at each grid location (Sec.~\ref{subsec:rh}).
+Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (Sec.~\ref{subsec:mtprop}).
 
@@
 The $R_{H}$ is interpreted as a reliability measure of the local orientation.
-Grid locations where the $R_{H}$ is above a certain threshold are marked as regions with reliable orientations.
-We use 0.3 as the cutoff for all our test cases. (See Section~\ref{sec:param_choices} on the choice of 0.3).
+Grid locations where the $R_{H}$ is above a cutoff threshold are marked as regions with reliable orientations (see Sec.~\ref{sec:param_choices} on parameter choices).
 %\begin{figure}[h]
@@
 %\end{figure}
-In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold.
+In Fig.~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold.
 The unit vector representing the principal direction has been mapped to RGB color space.
-Figure~\ref{fig:reliable_hessian}a shows the entire data set.
+Fig.~\ref{fig:reliable_hessian}a shows the entire data set.
 Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue.
-Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively.
-Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest.
+Fig.~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively.
+Fig.~\ref{fig:reliable_hessian}d shows a magnified region of interest.
 Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable.
@@
 %\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
-\includegraphics[width=0.5\textwidth]{images_pvis/reliable_hessian.pdf}
+\includegraphics[width=0.45\textwidth]{images_pvis/reliable_hessian.pdf}
 \caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square (in c).}
@@
 The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1].
-In Figure \ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue.
-Figure~\ref{fig:algo}c shows a single cylinder of a MetaTracts.
-Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Section~\ref{subsec:mtprop}).
+In Fig.
+\ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue.
+Fig.~\ref{fig:algo}c shows a single cylinder of a MetaTracts.
+Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}).
 The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow.
 We generate a rectangle of length $L$ and radius $R$, the green dashed rectangle shows the rectangle generated at $C_p$ (properties 3 and 5 in Section~\ref{subsec:mtprop}).
-
+%
 \begin{figure}[tb]
@@
 %\end{figure}
-Based on such assumptions on our data as local orientation and connectivity (Section~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:
-
-\begin{itemize}
+Based on such assumptions on our data as local orientation and connectivity (Sec.~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:
+
+\begin{itemize}[noitemsep]
 \item \textit{Orientation similarity}: The orientation of the start points ($N_p$) for the consecutive cylinders should be similar. 
@@
 We measure the distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$.
-
-For example, in Figure~\ref{fig:algo}b the distance is measured as the Euclidean distance between the purple and the orange vertices projected on to $N_p$.
+For example, in Fig.~\ref{fig:algo}b the distance is measured as the Euclidean distance between the purple and the orange vertices projected on to $N_p$.
 We refer to this perpendicular projection distance as \textit{`projected\_dist'}.
@@
 This approach tackles noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation in an already noisy environment.
-
-In the example in Figure~\ref{fig:algo}b we select the orange grid vertex next and repeat the process.
-We do not select the purple vertex (Figure~\ref{fig:algo}c) because it is not a candidate vertex.
-If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length.
-Short tracts (less than 0.01 times the smallest axes size) are removed.
-The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}a.
+% 
+In the example in Fig.~\ref{fig:algo}b we select the orange grid vertex next and repeat the process.
+We do not select the purple vertex (Fig.~\ref{fig:algo}c) because it is not a candidate vertex.
+If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices.
+Therefore these short tracts are removed.
+The MetaTracts generated are shown in Fig.~\ref{fig:length_distribution}a.
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
-Figure~\ref{fig:length_distribution}a shows a single slice of YZ plane.
+Fig.~\ref{fig:length_distribution}a shows a single slice of YZ plane.
 Consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.
-Figure~\ref{fig:length_distribution}b shows the MetaTracts of a particular bundle colored according to their individual length.
-
+Fig.~\ref{fig:length_distribution}b shows the MetaTracts of a particular bundle colored according to their individual length.
+%
 MetaTracts in a given fiber bundle do not extend the full length of the bundle. MetaTracts may have different lengths and may partially overlap.
@@
 Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering.
-Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with fibers which partially overlap (Figure~\ref{fig:length_distribution}b).
+Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with fibers which partially overlap (Fig.~\ref{fig:length_distribution}b).
 Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
@@
 This step divides the individual MetaTracts into classes based on their major orientations.
-In order to cluster MetaTracts going in same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01}.
+In order to cluster MetaTracts going in the same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01}.
 An eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
@@
 where $D$ is the diagonal weight matrix and $L$ is the Laplacian matrix.
-The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an $m$-dimensional space (see section~\ref{sec:param_choices} for values of m).
+The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an $m$-dimensional space (see Sec.~\ref{sec:param_choices} for values of m).
  In our case, each MetaTract is a data point.
@@
 For our test case there are two major directions of the fiber bundles and $K$ was set to 2.
-Due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
-Figure~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
+Due to the advantages offered by the dimensionality reduction, even if there are curved fiber bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
+Fig.~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
 As is expected there is a clear distinction based on fiber orientation.
-Figure~\ref{fig:orientation_clustering}a,c shows the MetaTracts colored according to clustering results.
+Fig.~\ref{fig:orientation_clustering}a,c shows the MetaTracts colored according to clustering results.
 
@@
   \centering  	
-  \includegraphics[width=0.5\textwidth]{images_pvis/figure6.png}
+  \includegraphics[width=0.45\textwidth]{images_pvis/figure6.png}
   %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
@@
 The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$ but the directed Hausdorff is not.
-
-Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTracts, which only partially cover the bundle (Figure~\ref{fig:length_distribution}b).
+%
+Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTracts, which only partially cover the bundle (Fig.~\ref{fig:length_distribution}b).
 If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$, then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
-
+%
 Thus, $H(P,Q)$ will be large, even though $P$ and $Q$ are in the same fiber bundle.
@@
 Note that if $P$ and $Q$ overlap but do not cover the same parts of the fiber bundle, then $H_{dir}(P,Q)$ and $H_{dir}(Q,P)$ and $\min(H_{dir}(P,Q),H_{dir}(Q,P))$ will be large.
-
+% 
 The directed Hausdorff distance is very sensitive to outliers in the data.
@@
 We repeated both until a steady state of clusters is reached and no new small fibers can be removed.
-The results of clustering the orientation clusters (see Figure~\ref{fig:orientation_clustering}a,c) are shown in Figure~\ref{fig:orientation_clustering}d,e respectively.
-Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e).
+The results of clustering the orientation clusters (see Fig.~\ref{fig:orientation_clustering}a,c) are shown in Fig.~\ref{fig:orientation_clustering}d,e respectively.
+Note how the thin and curved purple cluster bundle 7 is extracted well (Fig.~\ref{fig:orientation_clustering}e).
 
@@
 %\end{figure}  
-
-In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
+%  
+In comparison Fig.~\ref{fig:comparison} shows the MetaTracts in Fig.~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
 As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers.
@@
 Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
-\begin{figure}[h]
+%
+On the other hand, Fig..~\ref{fig:comparison_orient} shows the MetaTracts in Fig..~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone, without a second hierarchical clustering step.
+We set ``k" to be  10, 15, 20 clusters.
+In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
+\begin{figure}[hb]
 	\centering
-	\includegraphics[width=0.5\textwidth,  trim = 0mm 10mm 0mm 10mm, clip]{images_pvis/comparison_proximity.pdf}
+	\includegraphics[width=0.45\textwidth,  trim = 0mm 10mm 0mm 10mm, clip]{images_pvis/comparison_proximity.pdf}
 	\caption{MetaTracts data (left) and its straight-forward hierarchical clustering for various numbers of clusters.}
@@
 \end{figure} 
-\begin{figure}[h]
+\begin{figure}[hb]
 	\centering
-	\includegraphics[width=0.5\textwidth]{images_pvis/comparison_orient.pdf}
+	\includegraphics[width=0.45\textwidth]{images_pvis/comparison_orient.pdf}
 	\caption{Direct K-means clustering of MetaTracts (left) embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
@@
 \end{figure}
-On the other hand, Figure.~\ref{fig:comparison_orient} shows the MetaTracts in Figure.~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone, without a second hierarchical clustering step.
-We set ``k" to be  10, 15, 20 clusters.
-In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
 
@@
 in simulations.
-
+%
 To voxelize the space,  we compute a neighborhood around each voxel.
@@
 The voxel is then assigned to the class with the maximum number of elements in the neighborhood.
-Figures ~\ref{fig:crop-16-decomp}a and~\ref{fig:prepreg}d show the result of voxelization.
-Figure~\ref{fig:crop-16-decomp}c,d shows examples of extracted meshes.
+Fig.~\ref{fig:crop-16-decomp}a and~\ref{fig:prepreg}d show the result of voxelization.
+Fig.~\ref{fig:crop-16-decomp}c,d shows examples of extracted meshes.
 
@@
 \centering
-		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
+		\includegraphics[width=0.45\textwidth]{images_pvis/figure7}
 		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
@@
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
-	\includegraphics[width=0.4\textwidth, trim = 0mm 30mm 0mm 0mm, clip,]{images_pvis/figure8}
+	\includegraphics[width=0.45\textwidth, trim = 0mm 30mm 0mm 0mm, clip,]{images_pvis/figure8}
 	\caption{Data set with flat thin and compact bundles.
@@
 We tested our method on three datasets with varying characteristics. We report results on two due to lack of space.  
-Dataset 1 is described in section~\ref{sec:char_data}. Dataset 1 has eleven separate fiber bundles which we identified correctly.
-Figure~\ref{fig:orientation_clustering} shows the result of dataset~\ref{fig:data-char} decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters (the ground truth is 6 and 5 respectively).
+Dataset 1 is described in Sec.~\ref{sec:char_data}. Dataset 1 has eleven separate fiber bundles which were identified correctly by MetaTracts.
+Fig.~\ref{fig:orientation_clustering} shows the result of dataset~\ref{fig:data-char} decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters (the ground truth is 6 and 5 respectively).
 
 The hierarchical clustering is robust against over segmentation. For example,
-Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Figure~\ref{fig:orientation_clustering}d).
+Fig.~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Fig.~\ref{fig:orientation_clustering}d).
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
-As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster six (Figure~\ref{fig:len_dist_crop16}) is 200 while the maximum MetaTract length in the bundle is about 500 (measured in units of grid cube length: 2 $\mu m$).
-Figure~\ref{fig:length_distribution}b shows the length per MetaTract distribution for a particular fiber bundle.
-
-Figure~\ref{fig:prepreg} shows the results on our dataset 2.
+As noted in Sec.~\ref{subsec:dist_clustering}, the MetaTracts do not run the entire length of the dataset, the median length of MetaTracts in cluster six (Fig.~\ref{fig:len_dist_crop16}) is 200 while the maximum MetaTract length in the bundle is about 500 (measured in units of grid cube length: 2 $\mu m$).
+Fig.~\ref{fig:length_distribution}b shows the length per MetaTract distribution for a particular fiber bundle.
+
+Fig.~\ref{fig:prepreg} shows the results on our dataset 2.
 This dataset is dense with \textit{flat and thin} bundles.
 The dimensions are 300$\times$350$\times$300 and the datatype is uint16 (dataset 1 is uint8).
-The bundles are indistinguishable in the original data set Figure~\ref{fig:prepreg}a.
+The bundles are indistinguishable in the original data set Fig.~\ref{fig:prepreg}a.
 A green dotted line shows one of the fiber bundle boundary.
-Figure~\ref{fig:prepreg}b shows the fiber bundles for the two orientations, Figure~\ref{fig:prepreg}c shows the combined results, and Figure~\ref{fig:prepreg}d shows the volume after the voxelization.
+Fig.~\ref{fig:prepreg}b shows the fiber bundles for the two orientations, Fig.~\ref{fig:prepreg}c shows the combined results, and Fig.~\ref{fig:prepreg}d shows the volume after the voxelization.
 
@@
 All clustering was done in R.
-The preprocessing is time consuming and consists of MetaTract generation and the distance computation.
-The orientation clustering is robust and is done as a part of preprocessing.
-The hierarchical clustering separately, on the results of the orientation clusters can be done within an order of a few minutes on Intel Xeon E5-2667 workstation and can be repeated according to user input of $h$ (section~\ref{sec:param_choices}).
+The preprocessing is time consuming and may take several hours.
+It consists of MetaTract generation and the distance computation.
+The orientation clustering is robust and done as a part of preprocessing.
+The hierarchical clustering itself, on the results of the orientation clusters can be done within an order of a few minutes on Intel Xeon E5-2667 workstation and can be repeated according to user input of $h$ (Sec.~\ref{sec:param_choices}).
 We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without any extra computations.
@@
 As expected, Fig.~\ref{fig:comparison} shows that the participants found it difficult to determine the number of bundles from volume rendering alone.
-Another common approach is to color map the main eigenvector of Hessian matrix to RGB space.
+Another common approach is to color map the top eigenvector of Hessian matrix to RGB space.
 We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Fig.~\ref{fig:reliable_hessian}A) and asked the same query.
@@
 K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern.
-Our framework is robust to the choice of $h$ (Section~\ref{sec:results}).
+Our framework is robust to the choice of $h$ (Sec.~\ref{sec:results}).
 Large $h$ keeps the major bundles intact.
@@
 We used 10.00 and 2.00 for length and breadth(measured in grid cube length), respectively for all tests.
+Short tracts of less than 0.01 times the smallest axes size are prone to have erroneous local orientation and are therefore removed.
 $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays; we have used [7-10] and half the length of cylinder (5.0).

### 2014-09-25 chapters/MetaTracts.tex
figures again

--- 2014-09-25 chapters/MetaTracts.tex
@@
 
-In comparison Figure~\ref{fig:comparison} shows the MetaTracts in Figure~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
+Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
 As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers.
@@
 Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
-\begin{figure}[h]
+
+\begin{figure}[tb]
 	\centering
-	\includegraphics[width=0.5\textwidth,  trim = 0mm 10mm 0mm 10mm, clip]{images_pvis/comparison_proximity.pdf}
-	\caption{MetaTracts data (left) and its straight-forward hierarchical clustering for various numbers of clusters.}
+	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
+	\caption{Applying only hierarchical (a) and K-means (b) clustering methods of MetaTracts for various numbers of clusters.}
 	\label{fig:comparison}
 \end{figure}
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth]{images_pvis/comparison_orient.pdf}
-	\caption{Direct K-means clustering of MetaTracts (left) embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
-	\label{fig:comparison_orient}
-\end{figure}
-On the other hand, Figure.~\ref{fig:comparison_orient} shows the MetaTracts in Figure.~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone, without a second hierarchical clustering step.
-We set ``k" to be  10, 15, 20 clusters.
+
+On the other hand, Figure.~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step for 10, 15, and 20 clusters.
 In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.

### 2014-09-25 chapters/MetaTracts.tex

--- 2014-09-25 chapters/MetaTracts.tex
@@
 	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
-	\includegraphics[width=0.45\textwidth]{images_pvis/workflow.png}
+	\includegraphics[width=0.475\textwidth]{images_pvis/workflow.pdf}
 	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
   \centering  	
-  \includegraphics[width=0.45\textwidth]{images_pvis/figure6.png}
+  \includegraphics[width=\linewidth]{images_pvis/clustering.pdf}
   %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
@@
 	\centering
-	\includegraphics[width=0.45\textwidth,  trim = 0mm 10mm 0mm 10mm, clip]{images_pvis/comparison_proximity.pdf}
-	\caption{MetaTracts data (left) and its straight-forward hierarchical clustering for various numbers of clusters.}
+	\includegraphics[width=0.45\linewidth]{images_pvis/comparison_all.pdf}
+	\caption{Applying only hierarchical (a) and K-means (b) clustering methods of MetaTracts for various numbers of clusters.}
 	\label{fig:comparison}
-\end{figure} 
-\begin{figure}[hb]
-	\centering
-	\includegraphics[width=0.45\textwidth]{images_pvis/comparison_orient.pdf}
-	\caption{Direct K-means clustering of MetaTracts (left) embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
-	\label{fig:comparison_orient}
 \end{figure}
@@
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
-	\includegraphics[width=0.45\textwidth, trim = 0mm 30mm 0mm 0mm, clip,]{images_pvis/figure8}
+	\includegraphics[width=0.45\linewidth]{images_pvis/dataset2.pdf}
 	\caption{Data set with flat thin and compact bundles.

### 2014-09-25 chapters/MetaTracts.tex
small fixes

@@
 	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
-	\includegraphics[width=0.475\textwidth]{images_pvis/workflow.pdf}
+	\includegraphics[width=\linewidth]{images_pvis/workflow.pdf}
 	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
@@
 %\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
-\includegraphics[width=0.45\textwidth]{images_pvis/data-char.pdf}
+\includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
 \caption{Data characteristics: (a) Rendering of the data.
@@
 %\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
-\includegraphics[width=0.45\textwidth]{images_pvis/reliable_hessian.pdf}
-\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square (in c).}
+\includegraphics[width=\linewidth]{images_pvis/reliable_hessian.pdf}
+\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square.}
 \label{fig:reliable_hessian}
@@
   \vspace{-1em}
- \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
+ \caption{MetaTracts generation: (b) MetaTract generation, region with $R_H$ less than threshold. (a) close up of marked region in (b). (c) Parts of a single MetaTract.}
   \label{fig:algo}
@@
 \centering
-	\includegraphics[width=0.45\textwidth]{images_pvis/figure5.png}
+	\includegraphics[width=\linewidth]{images_pvis/figure5.png}
 	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
-	\caption{(A) MetaTracts color-coded according to their mean orientations.
-(B) length distribution of individual MetaTracts for a particular bundle (unit for length is grid cube edge length:  $2\mu m$).
+	\caption{(a) MetaTracts color-coded according to their mean orientations.
+(b) length distribution of individual MetaTracts for a particular bundle (unit for length is grid cube edge length:  $2\mu m$).
 A single fiber bundle is made of MetaTracts of varying length}
@@
   %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
-  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes.
-(A) MetaTracts belonging to orientation cluster 1.
-(C) MetaTracts belonging to orientation cluster 2.
-MetaTracts in gray(A,C) show context.
-  	(D,E) shows the result of distance based clustering on the orientation cluster. (F) shows the combined result. }
+  	\caption{(b) Result of K means clustering with data points projected to the top three eigenvectors as major axes.
+(a) MetaTracts belonging to orientation cluster 1.
+(c) MetaTracts belonging to orientation cluster 2.
+MetaTracts in gray (a,c) show context.
+  	(d,e) shows the result of distance based clustering on the orientation cluster. (f) shows the combined result. }
   \label{fig:orientation_clustering}
@@
 %  
-In comparison Fig.~\ref{fig:comparison} shows the MetaTracts in Fig.~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
+Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
 As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers.
@@
 %
-On the other hand, Fig..~\ref{fig:comparison_orient} shows the MetaTracts in Fig..~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone, without a second hierarchical clustering step.
-We set ``k" to be  10, 15, 20 clusters.
+On the other hand, Figure.~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step for 10, 15, and 20 clusters.
 In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
-\begin{figure}[hb]
+\begin{figure}[tb]
 	\centering
-	\includegraphics[width=0.45\linewidth]{images_pvis/comparison_all.pdf}
+	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
 	\caption{Applying only hierarchical (a) and K-means (b) clustering methods of MetaTracts for various numbers of clusters.}
@@
 \centering
-		\includegraphics[width=0.45\textwidth]{images_pvis/figure7}
+		\includegraphics[width=\linewidth]{images_pvis/figure7}
 		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
@@
 	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
-	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
+	\caption{(a) Voxelization of data set 1; (b) a single slice of the data set; (c) two of the extracted meshes together; (d) the meshes rendered separately.}
 	\label{fig:crop-16-decomp}
@@
 	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
-	\includegraphics[width=0.45\linewidth]{images_pvis/dataset2.pdf}
+	\includegraphics[width=\linewidth]{images_pvis/dataset2.pdf}
 	\caption{Data set with flat thin and compact bundles.
-(A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation.
-(C) shows the complete result.
-(D) shows the voxelization of (C).}
+(a) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (b) shows the clusters according to individual orientation.
+(c) shows the complete result.
+(d) shows the voxelization of (c).}
 	\label{fig:prepreg}
@@
 	\centering
-	\includegraphics[width=0.45\textwidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval2}
+	\includegraphics[width=\linewidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval2}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
@@
 \centering
-	\includegraphics[width=0.5\textwidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
+	\includegraphics[width=\linewidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}

### 2014-09-25 chapters/MetaTracts.tex
figures done
Local text commit

@@
 In the typical scanning geometry of industrial XCT, the specimen is placed on a rotary table between X-ray source and detector.
-X-rays are emitted by the source in the form of a conic beam.
 When passing through the specimen the X-rays are attenuated by the material in the specimen.
@@
 Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion.
-This is usually done by solving a second-order Runge-Kutta integration. The principal diffusion direction at each discrete location is interpolated to form a continuous velocity field. 
+This is usually done by solving a Runge-Kutta integration (typically second or fourth order).
+The principal diffusion direction at each discrete location is interpolated to form a continuous velocity field.
 Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers.
@@
 \centering
-	\includegraphics[width=\linewidth]{images_pvis/figure5.png}
+	\includegraphics[width=\linewidth]{images_pvis/figure5.pdf}
 	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
@@
 	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
-	\caption{Applying only hierarchical (a) and K-means (b) clustering methods of MetaTracts for various numbers of clusters.}
+	\caption{Applying only hierarchical (a) and K-means (b) clustering methods to MetaTracts for various numbers of clusters.}
 	\label{fig:comparison}
@@
 \centering
-		\includegraphics[width=\linewidth]{images_pvis/figure7}
+		\includegraphics[width=\linewidth]{images_pvis/figure8.pdf}
 		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}

### 2014-09-25 chapters/MetaTracts.tex
Added Figure 9, updated figure 8
local update histogram pdf

@@
 	\centering
-	\includegraphics[width=\linewidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval2}
+	\includegraphics[width=\linewidth,  trim = 0mm 30mm 0mm 0mm, clip]{images_pvis/usereval.pdf}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
@@
 \centering
-	\includegraphics[width=\linewidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
+	\includegraphics[width=\linewidth,  trim = 0mm 0mm 10mm 00mm, clip]{images_pvis/histogram.pdf}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}

### 2014-09-26 chapters/MetaTracts.tex
supplemental folders

@@
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Fig.~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
+Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
 Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
@@
 \item{ How to extract and visualize the geometric structure of a particular fiber bundle?}
-\item{ How to visualize the interaction between a particular pair of fiber bundles (crossovers/braiding) or a unit cell?}
+\item{ How to visualize the interaction between a particular pair of fiber bundles (weaving/braiding) or a unit cell?}
 \item{ Which fiber bundles show a particular orientation? }
@@
 
-Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts.
-We detail and evaluate a method which uses visualization techniques to gain insight into our data.
  We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}.
@@
 Spatial context refers to how two or more bundles interact with each other.
+For example, spatial context answers questions such as: Are these bundles in contact at a particular position in the data?
+What is the relative orientations of the contacting bundles?
+
+Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts.
+We detail and evaluate a method which uses visualization techniques to gain insight into our data.
 We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites.
@@
 \end{figure}
-Fig.~\ref{fig:flowchart} shows an over view of our approach.
+Figure~\ref{fig:flowchart} shows an over view of our approach.
 Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters.
@@
 
-\subsection {Visual Analysis of Fiber Reinforced Composites}
+\subsection {Visual Analysis and Modeling of Fiber Reinforced Composites}
 The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores and deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.
@@
 They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
-Reh et al.~\cite{PMI_AR_2012} introduced an approach to explore pores of carbon fiber reinforced composites.
 Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers.
-Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research.
+
+Lomov et al.~\cite{Lomov2010} in a recent paper, discussed the current available solutions in geometric modeling of three dimensional composites.
+Modeling of the composites starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position?
+The second builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.
+
+%Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research.
 Recently, similar techniques have been applied to fiber tracing in composite materials.
@@
 \label{sec:char_data}
-Fig.
-\ref{fig:data-char} shows dataset 1 with woven fiber bundles.
+Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
 Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$.
 The data type is unsigned 8 bit integer.
-Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
+Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
 The unit cell is used as a basis structure for CFRP manufacturing.
 The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
-Fig.~\ref{fig:data-char}a shows a volume rendering of the dataset.
-Fig.~\ref{fig:data-char}b shows 2D slices along the $X$- and Fig.~\ref{fig:data-char}c in $Z$-axis.
-Fig.~\ref{fig:data-char}d shows the magnified version of the green region of interest.
+Figure~\ref{fig:data-char}a shows a volume rendering of the dataset.
+Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis.
+Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest.
 The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images.
-Fig.~\ref{fig:data-char}d contains two bundles going in opposite directions.
+Figure~\ref{fig:data-char}d contains two bundles going in opposite directions.
 The separation between two fiber bundles is barely visible.
@@
 The fiber bundles may differ in terms of the amount of fibers in the bundle.
-Fig.~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
+Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
 Depending on the weaving pattern, the fiber bundles cross each other in different orientations.
@@
 %\end{figure}
-In Fig.~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold.
+In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold.
 The unit vector representing the principal direction has been mapped to RGB color space.
-Fig.~\ref{fig:reliable_hessian}a shows the entire data set.
+Figure~\ref{fig:reliable_hessian}a shows the entire data set.
 Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue.
-Fig.~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively.
-Fig.~\ref{fig:reliable_hessian}d shows a magnified region of interest.
+Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively.
+Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest.
 Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable.
@@
 \includegraphics[width=\linewidth]{images_pvis/reliable_hessian.pdf}
+ \vspace{-1.5em}
 \caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square.}
@@
 The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1].
-In Fig.
-\ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue.
-Fig.~\ref{fig:algo}c shows a single cylinder of a MetaTracts.
+In Figure~\ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue.
+Figure~\ref{fig:algo}c shows a single cylinder of a MetaTracts.
 Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}).
@@
 We measure the distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$.
-For example, in Fig.~\ref{fig:algo}b the distance is measured as the Euclidean distance between the purple and the orange vertices projected on to $N_p$.
+For example, in Figure~\ref{fig:algo}b the distance is measured as the Euclidean distance between the purple and the orange vertices projected on to $N_p$.
 We refer to this perpendicular projection distance as \textit{`projected\_dist'}.
@@
 % 
-In the example in Fig.~\ref{fig:algo}b we select the orange grid vertex next and repeat the process.
-We do not select the purple vertex (Fig.~\ref{fig:algo}c) because it is not a candidate vertex.
+In the example in Figure~\ref{fig:algo}b we select the orange grid vertex next and repeat the process.
+We do not select the purple vertex (Figure~\ref{fig:algo}c) because it is not a candidate vertex.
 If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices.
 Therefore these short tracts are removed.
-The MetaTracts generated are shown in Fig.~\ref{fig:length_distribution}a.
+The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}a.
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
-Fig.~\ref{fig:length_distribution}a shows a single slice of YZ plane.
+Figure~\ref{fig:length_distribution}a shows a single slice of YZ plane.
 Consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.
-Fig.~\ref{fig:length_distribution}b shows the MetaTracts of a particular bundle colored according to their individual length.
+Figure~\ref{fig:length_distribution}b shows the MetaTracts of a particular bundle colored according to their individual length.
 %
@@
 	\caption{(a) MetaTracts color-coded according to their mean orientations.
-(b) length distribution of individual MetaTracts for a particular bundle (unit for length is grid cube edge length:  $2\mu m$).
-A single fiber bundle is made of MetaTracts of varying length}
+(b) length distribution of individual MetaTracts for a particular bundle (unit for length is grid cube edge length:  $2\mu m$).}
 	\label{fig:length_distribution}
@@
 Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering.
-Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with fibers which partially overlap (Fig.~\ref{fig:length_distribution}b).
+Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with fibers which partially overlap (Figure~\ref{fig:length_distribution}b).
 Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
@@
 Due to the advantages offered by the dimensionality reduction, even if there are curved fiber bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
-Fig.~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
+Figrue.~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
 As is expected there is a clear distinction based on fiber orientation.
-Fig.~\ref{fig:orientation_clustering}a,c shows the MetaTracts colored according to clustering results.
+Figure~\ref{fig:orientation_clustering}a,c shows the MetaTracts colored according to clustering results.
 
@@
   \includegraphics[width=\linewidth]{images_pvis/clustering.pdf}
+   \vspace{-1.5em}
   %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
@@
 %
-Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTracts, which only partially cover the bundle (Fig.~\ref{fig:length_distribution}b).
+Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTracts, which only partially cover the bundle (Figure~\ref{fig:length_distribution}b).
 If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$, then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
@@
 To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
-shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
+shortest tracts (tracts less than $\eta$ times the median length, set to 0.6) in each bundle. We then reapplied hierarchical clustering.
 We repeated both until a steady state of clusters is reached and no new small fibers can be removed.
-The results of clustering the orientation clusters (see Fig.~\ref{fig:orientation_clustering}a,c) are shown in Fig.~\ref{fig:orientation_clustering}d,e respectively.
-Note how the thin and curved purple cluster bundle 7 is extracted well (Fig.~\ref{fig:orientation_clustering}e).
+The results of clustering the orientation clusters (see Figure~\ref{fig:orientation_clustering}a,c) are shown in Figure~\ref{fig:orientation_clustering}d,e respectively.
 
@@
 %
-On the other hand, Figure.~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step for 10, 15, and 20 clusters.
+On the other hand, Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step for 10, 15, and 20 clusters.
 In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
@@
 The voxel is then assigned to the class with the maximum number of elements in the neighborhood.
-Fig.~\ref{fig:crop-16-decomp}a and~\ref{fig:prepreg}d show the result of voxelization.
-Fig.~\ref{fig:crop-16-decomp}c,d shows examples of extracted meshes.
+Figure~\ref{fig:crop-16-decomp}a and~\ref{fig:prepreg}d show the result of voxelization.
+Figure~\ref{fig:crop-16-decomp}c,d shows examples of extracted meshes.
 
@@
 		\includegraphics[width=\linewidth]{images_pvis/figure8.pdf}
+		 \vspace{-1.5em}
 		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
@@
 We tested our method on three datasets with varying characteristics. We report results on two due to lack of space.  
-Dataset 1 is described in Sec.~\ref{sec:char_data}. Dataset 1 has eleven separate fiber bundles which were identified correctly by MetaTracts.
-Fig.~\ref{fig:orientation_clustering} shows the result of dataset~\ref{fig:data-char} decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters (the ground truth is 6 and 5 respectively).
+Dataset 1 is described in Sec.~\ref{sec:char_data} (Figure~\ref{fig:data-char}).
+Dataset 1 has eleven separate fiber bundles (5 along one orientation and 6 along the other).
+All of them were identified correctly by MetaTracts.
+With different cross-section sizes and varying degree of curvature of the bundles, dataset 1 is a complicated dataset.
+Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of dataset 1 decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization mesh extraction.
+MetaTracts correctly identified the fiber bundles.
+Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e, Figure~\ref{fig:crop-16-decomp}d).
 
 The hierarchical clustering is robust against over segmentation. For example,
-Fig.~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Fig.~\ref{fig:orientation_clustering}d).
+Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Figure~\ref{fig:orientation_clustering}d).
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
-As noted in Sec.~\ref{subsec:dist_clustering}, the MetaTracts do not run the entire length of the dataset, the median length of MetaTracts in cluster six (Fig.~\ref{fig:len_dist_crop16}) is 200 while the maximum MetaTract length in the bundle is about 500 (measured in units of grid cube length: 2 $\mu m$).
-Fig.~\ref{fig:length_distribution}b shows the length per MetaTract distribution for a particular fiber bundle.
-
-Fig.~\ref{fig:prepreg} shows the results on our dataset 2.
-This dataset is dense with \textit{flat and thin} bundles.
+As noted in Sec.~\ref{subsec:dist_clustering}, the MetaTracts do not run the entire length of the dataset, the median length of MetaTracts in cluster six (Figure~\ref{fig:len_dist_crop16}) is 200 while the maximum MetaTract length in the bundle is about 500 (measured in units of grid cube length: 2 $\mu m$).
+Figure~\ref{fig:length_distribution}b shows the length per MetaTract distribution for a particular fiber bundle.
+
+Figure~\ref{fig:prepreg} shows the results on our dataset 2.
+This dataset has different characteristics, and is dense with \textit{flat and thin} bundles, it consists of 4 bundles along one direction and 5 along the other.
 The dimensions are 300$\times$350$\times$300 and the datatype is uint16 (dataset 1 is uint8).
-The bundles are indistinguishable in the original data set Fig.~\ref{fig:prepreg}a.
+The bundles are indistinguishable in the original data set Figure~\ref{fig:prepreg}a.
 A green dotted line shows one of the fiber bundle boundary.
-Fig.~\ref{fig:prepreg}b shows the fiber bundles for the two orientations, Fig.~\ref{fig:prepreg}c shows the combined results, and Fig.~\ref{fig:prepreg}d shows the volume after the voxelization.
+Figure~\ref{fig:prepreg}b shows the fiber bundles for the two orientations, Figure~\ref{fig:prepreg}c shows the combined results, and Figure~\ref{fig:prepreg}d shows the volume after the voxelization.
 
@@
 Four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization participated in the user evaluation of our work.
-We therefore setup a questionnaire with questions on solving given tasks as well as questions using Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques.
-LS consists of 5 levels from ``Strongly disagree" to ``Strongly agree".
-Fig.~\ref{fig:comparison} shows the results.
-Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A).
+We setup a questionnaire with questions on solving given tasks as well as questions using the Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques.
+LS consists of 5 levels from ``strongly disagree" to ``strongly agree".
+Figure~\ref{fig:userstudy} shows the results.
+Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Figure~\ref{fig:data-char}a).
 This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user.
-As expected, Fig.~\ref{fig:comparison} shows that the participants found it difficult to determine the number of bundles from volume rendering alone.
+Figure~\ref{fig:userstudy} shows that, as expected the participants found it difficult to determine the number of bundles from volume rendering alone.
 Another common approach is to color map the top eigenvector of Hessian matrix to RGB space.
-We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Fig.~\ref{fig:reliable_hessian}A) and asked the same query.
-Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test before.
+We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Figure~\ref{fig:reliable_hessian}a) and asked the same query.
+Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test.
 The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective.
-Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Fig.~\ref{fig:orientation_clustering}D,E).
+Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Figure~\ref{fig:orientation_clustering}d,e).
 All participants could correctly identify the number of fiber bundles.
 All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering.
-They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig.~\ref{fig:data-char}C).
+They especially liked the two dimensional slices of the MetaTracts (Figure~\ref{fig:crop-16-decomp}b, compared to Figure~\ref{fig:data-char}c).
 The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts.
@@
 \centering
-	\includegraphics[width=\linewidth,  trim = 0mm 0mm 10mm 00mm, clip]{images_pvis/histogram.pdf}
+	\includegraphics[width=\linewidth,  trim = 0mm 0mm 0mm 00mm, clip]{images_pvis/figure9.pdf}
+	 \vspace{-1.5em}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
-	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}A, clustered into 10 clusters.
+	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Figure~\ref{fig:orientation_clustering}d, clustered into 10 clusters.
 The unit for length is the grid cube length.
-Fig.~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6.
 Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10.
@@
 Large $h$ keeps the major bundles intact.
-For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements).
+For example, in  Figure~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Figure~\ref{fig:len_dist_crop16} all have less than 10 elements).
 This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
@@
 \cite{Frangi1998} and set to 0.5.
-The length and the width parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are.
-Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber.
-We used 10.00 and 2.00 for length and breadth(measured in grid cube length), respectively for all tests.
-Short tracts of less than 0.01 times the smallest axes size are prone to have erroneous local orientation and are therefore removed.
+The length and the radius parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are.
+These are dependent on the underlying fiber characteristic and the weaving pattern.
+Larger cylinders will also handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber.
+We used 10.00 and 2.00 for length and radius (measured in grid voxel size), respectively for all tests.
+Simpler geometry (example dataset 2) was experimentally found to handle larger cylinders better.
+$\eta$ in sec.~\ref{subsec:dist_clustering} decides how quickly the hierarchical clustering converges, experimentally 0.3 to 0.6 gave similar results.
+ %Short tracts of less than 0.01 times the smallest axes size are prone to have erroneous local orientation and are therefore removed. 
  $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays; we have used [7-10] and half the length of cylinder (5.0).

### 2014-09-26 metaTracts_supplymental_pvis2014.tex
supplemental folders

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{pdfpages}
\begin{document}
\input {./supplemental/supplementaryMaterial.tex}
\end{document}

### 2014-09-26 supplemental/supplementaryMaterial.tex
supplemental folders

\section{Supplementary Material}
\label{sec:sm}
Section~\ref{subsec:usereval} contains the our user evaluation questionnaire.
\subsection{User Evaluation Questionnaire}
The questionnaire contains links to videos which were seen by the participants.
\label{subsec:usereval}
\includepdf[pages={1-13}]{./supplemental/MetaTracts_usereval.pdf}

### 2014-09-26 supplemental/supplementaryMaterial.tex
Updated supplemental material

@@
 \label{sec:sm}
-Section~\ref{subsec:usereval} contains the our user evaluation questionnaire.
+Section~\ref{subsec:ext_exp} contains extended experiments.
+Section~\ref{subsec:usereval} contains our user evaluation questionnaire.
+\subsection{Extended Experiments}
+\label{subsec:ext_exp}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\linewidth]{images_pvis/dataset3.pdf}
+	\caption{Results on dataset3.
+(a) Rendering of our dataset.
+(b) MetaTracts embedded on the top 3 eigenvectors and clustered into 2 clusters.
+(c) Each orientation, hierarchically clustered into 10 clusters.}
+	\label{fig:dataset3}
+\end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\linewidth]{images_pvis/dataset1-extended.pdf}
+	\caption{Results on dataset1. (a) MetaTracts. (b) Voxelized MetaTracts. (c) Subset of the fiber bundles are shown to better visualize the interaction between particular bundles.}
+	\label{fig:dataset1-extended}
+\end{figure}
+Figure~\ref{fig:dataset3} shows the results on our third dataset.
+The data set is 600$\times$500$\times$600 in size and the datatype is uint16.
+For testing purposes we have restricted the number of MetaTracts produced to 10000 which generates the empty spaces in Figure~\ref{fig:dataset3}b,c.
+The dataset has 5 major bundles. All the bundles were extracted by MetaTracts.
+
+Figure~\ref{fig:dataset1-extended} further shows how spatial context can be used to answer questions such as  Are these bundles in contact at a particular position in the data?
+What is the relative orientations of the contacting bundles?
+These bundles are occluded in the original volume and no conclusions can be made from the original volume.
+In the voxelized MetaTracts we can clearly see where and how the bundles interact.
+
 \subsection{User Evaluation Questionnaire}
@@
 \label{subsec:usereval}
+\pagebreak
 \includepdf[pages={1-13}]{./supplemental/MetaTracts_usereval.pdf}

### 2014-09-26 chapters/MetaTracts.tex
Minor changes
Updated with Lomov and Straumit

@@
 
-Lomov et al.~\cite{Lomov2010} in a recent paper, discussed the current available solutions in geometric modeling of three dimensional composites.
-Modeling of the composites starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position?
-The second builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.
+Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{StraumitI2014} in recent papers, discussed the problems and current available solutions in geometric modeling of three dimensional composites.
+Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position?
+The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.
 
@@
 	\centering
-	\includegraphics[width=\linewidth,  trim = 0mm 30mm 0mm 0mm, clip]{images_pvis/usereval.pdf}
+	\includegraphics[width=\linewidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval.pdf}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
@@
 \centering
-	\includegraphics[width=\linewidth,  trim = 0mm 0mm 0mm 00mm, clip]{images_pvis/figure9.pdf}
+	\includegraphics[width=\linewidth,  trim = 0mm 0mm 0mm 00mm, clip]{images_pvis/figure9_AMA.pdf}
 	 \vspace{-1.5em}

### 2014-09-26 chapters/MetaTracts.tex
Updated XBundle Bib and  supplemental material

@@
 
-Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{StraumitI2014} in recent papers, discussed the problems and current available solutions in geometric modeling of three dimensional composites.
+Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014} in recent papers, discussed the problems and current available solutions in geometric modeling of three dimensional composites.
 Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position?
@@
 The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
-As output, each grid location becomes associated with a vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation.
+As output, each grid location becomes associated with an unit vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation.
 We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
-The principal directions, in which the local second order structure of the image can be decomposed, are given by the eigen decomposition of the Hessian matrix which describes the local curvature.
+The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.
+
 The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.

### 2014-09-26 supplemental/supplementaryMaterial.tex
Updated XBundle Bib and  supplemental material

@@
 Section~\ref{subsec:usereval} contains our user evaluation questionnaire.
+\subsection{Enlarged Figures}
+\label{subsec:enlarged_fig}
+\begin{figure}[tb]
+	\centering
+	%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
+	\includegraphics[width=\textwidth]{images_pvis/data-char.pdf}
+	\caption{Data characteristics: (a) Rendering of the data.
+(b) 2D slices along the X-axis and (c) 2D slice along the Z-axis.
+(d) Magnified region marked in green square (c).
+Multiple fiber bundles cross and are indistinguishable by visual inspection alone.
+}
+	\label{fig:data-char}
+\end{figure}
 \subsection{Extended Experiments}

### 2014-12-02 Rebuttal.tex
Rebuttals.

\documentclass[]{article}
\usepackage{graphicx}
\usepackage{color}
%opening
\title{MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites: Rebuttal}
\author{}

\begin{document}

\maketitle

\section{Reviewer 1}
\color{red}
Q1. A new variable, reliable Hessian, is defined and it is stated that $R_{H}$ is similar to the "vesselness".  Later on, the authors claim "the utility of the vesselness is a little different from our framework". Matching term by term between the definitions of "vesselness" and $R_{H}$, they are the same. The authors are encouraged to state clearly the originality and contribution of $R_{H}$.
\color{black}

A1. Reliable Hessians is computed exactly equivalently to Vesselness for a single scale (Firangi et al.~\cite{Frangi1998}). A common approach and one utilized by the Frangi et al.~\cite{Frangi1998}, is to consider the Taylor series expansion in the neighborhood of point $x$ in an image as.
\begin{equation}
f(x) = \sum_{i=0}^{\infty}\frac{f^{(i)}(0)}{i!}x^{i}
\end{equation}
The expansion often is approximated till the second order (Hessian). In particular the hessian maps a  neighborhood  centered about $x$ onto an ellipsoid  whose axes are along the directions given by the eigenvectors of the Hessian and the corresponding axis' semi-lengths are the magnitudes of the respective eigenvalues. Vesselness measure as described in the paper takes into account geometric ratios based on this and acts as a filtering process which searches for geometric structure that can be tubular. Figure~\ref{fig:Vesselness} shows results from Frangi et al.~\cite{Frangi1998} paper. Others including Sato et al.~\cite{Sato1997} have used the eigen decomposition to filter images.

As mentioned in the paper, we do not have clear tubular structures embedded in a dark contrast matrix such as in blood vessels ( Figure~\ref{fig:data-char-rebuttal} shows the data characteristics for our case.). Instead  the utility of $R_{H}$ is  to associate each grid location with a reliability measure for the local orientation as given by the  principal eigenvector. We use this orientation vector to create MetaTracts.

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{images_pvis/Vesselness}
\caption{Left: Xray image of vasculature, middle-left: calculated vesselness of the left image, middle-right: calculated vesselness after inversion of the grey-scale map. Right image after subtracting reference from left image.}
\label{fig:Vesselness}
\end{figure}

\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
	\includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
	\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
	\label{fig:data-char-rebuttal}
\end{figure}

\color{red}
Q2. In the examples presented, it seems only one type of composite was studied. For instance, there is no example showing a composite with fiber bundles that cross each other at an angle $<$ 90 degrees. When two bundles are almost parallel, will the current algorithm be sufficient to separate them apart? The authors are encouraged to provide more than one fiber configurations.
\color{black}

A2.

\color{red}
Q3. All clustering was done in R. The letter R is not explained.
\color{black}

A3. R is a free software environment for statistical computing and graphics.

\section{Reviewer 2}
\color{red}
I feel the background section could be improved. Why the orientation and weaving patterns of fibers indicate the strength of the material, and thus why the trouble to scan the materials, extract the fibers and visualize them. The paper gives some relatively high-level description of this but more background would be appreciated.

\color{black}
Answer.

\color{red}
The authors claim that their method can deal with low resolution data.
But this might be in the eyes of the beholders, since in my opinion the example data in the paper all have fairly high resolutions. Perhaps a comparison with a really high resolution data using this method and other existing methods could diffuse the doubt.
\color{black}
Answer.

\section{Reviewer 4}
\color{red}
In Section 5.1, k-means is applied to the m-dimensional space using
Euclidean distance, i.e., each dimension is considered equally important.
Is that really the case when they represent eigenvectors?

\color{black}
Answer. Belkin  and Niyogi's~\cite{Belkin01} laplacian eigenmaps embeds the high dimensional data ( represented as a graph) into a low dimensional space where the euclidean distance approximates the distances in higher dimension space. For such reason, Eigen embedding for dimensionality reduction has become quiet popular and has found favor in tackling a variety of problems. Empirically the K-means for orientation clustering provided us excellent results. The major orientations are limited by the inherent characteristics of our data set and this also plays a role in Laplacian embedding followed by k-means giving us such good results. A formal proof is not provided.

\color{red}
In Section 5.2, the authors mention that they remove short Meta Tracks.
How much data is thrown away here? Can examples be provided to get a
better understanding?
\color{black}

\color{red}

What are the limitations of the approach? Are there data sets where the
method fails?
\color{black}
Answer. A key assumptions is ``connectivity'', which states that moving along the direction of non-noisy local orientations in small increments, we reach another neighborhood with similar local orientation.
If the data is noisy to the point where this does not hold true and in large neighborhoods the Hessian is completely inaccurate then the process will fail.
We also assume the fiber bundles have a minimum width beyond which the data is too low for the clustering process to succeed.

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{xBundle}
\end{document}

### 2014-12-02 chapters/MetaTracts.tex
Rebuttals.

@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components.
-The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.
+The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of these continuously rising demands.
 CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing parts made of conventional materials such as aluminum or steel.
@@
 
-Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates we are addressing in this work, consist of two main components:
+Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates addressed in this work, consist of two main components:
 a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.
@@
 Figure~\ref{fig:flowchart} shows an over view of our approach.
-Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters.
-For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}.
+Starting from the XCT data, we first generate MetaTracts, then we cluster the MetaTracts to create fiber bundles and finally generate visualizations from the clusters.
+For a detailed description of the MetaTracts generation process see Sec.~\ref{sec:approach}.
 For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}.
@@
 In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles.
-Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}.
+Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}.
 Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
@@
 Recently, similar techniques have been applied to fiber tracing in composite materials.
-To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers, which is the main scope of this work.
+To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers.
 
@@
 In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.
-We make the following assumptions on our data:
+We make the following \textit{assumptions} on our data:
 \begin{itemize}[noitemsep]
@@
 
-The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.
+The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
+This direction coincides with the direction of the tubular structure.
 
@@
 Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors.
-Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
+Thus, we compute ``reliable Hessians" ($R_H$) to determine which locations in the volume provide reliable local orientation.
+($R_H$ is similar to the ``vesselness" in~\cite{Frangi1998}).
+%Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
 $$
@@
 The local orientaion at each voxel is inherently more noisy.
-Thus, techniques popular in DTI cannot be directly applied to XCT data.
 
@@
 We build on the key assumptions on the data, namely ``local orientation" and ``connectivity" while taking into account the noise and lack of resolution.
-We do this by interpreting the underlying geometric structure of the fibers as a set of polylines (cylinders).
+We do this by interpreting the underlying geometric structure of the fibers as a set of cylinders.
+
  MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data.
@@
 \item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
-\item{Each cylinder is locally parallel to the local orientation at its start point.}
+\item{Each cylinder is parallel to the local orientation at its start point.}
 \end{enumerate}
@@
 In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles.
-The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1].
+The reliable Hessian step  generates a local orientation vector in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1].
 In Figure~\ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue.
@@
 The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow.
-We generate a rectangle of length $L$ and radius $R$, the green dashed rectangle shows the rectangle generated at $C_p$ (properties 3 and 5 in Section~\ref{subsec:mtprop}).
+We generate a rectangle of length $L$ and radius $R$ (properties 3 and 5 in Section~\ref{subsec:mtprop}).
 %
@@
   \vspace{-1em}
- \caption{MetaTracts generation: (b) MetaTract generation, region with $R_H$ less than threshold. (a) close up of marked region in (b). (c) Parts of a single MetaTract.}
+ \caption{MetaTracts generation: (b) MetaTract generation (a) close up of marked region in (b). (c) Parts of a single MetaTract.}
   \label{fig:algo}
@@
 %\end{figure}
-Based on such assumptions on our data as local orientation and connectivity (Sec.~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:
+Based on assumptions on our data (local orientation and connectivity), the order of the candidate vertices is based on the following \textit{characteristics}:
 
@@
 We measure the distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$.
-For example, in Figure~\ref{fig:algo}b the distance is measured as the Euclidean distance between the purple and the orange vertices projected on to $N_p$.
+For example, in Figure~\ref{fig:algo}b the distance is measured as the Euclidean distance between the green and the orange vertices projected on to $N_p$.
 We refer to this perpendicular projection distance as \textit{`projected\_dist'}.
@@
 If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices.
-Therefore these short tracts are removed.
+These short tracts are removed.
 The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}a.
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
-Figure~\ref{fig:length_distribution}a shows a single slice of YZ plane.
-Consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
+Figure~\ref{fig:length_distribution}a also shows a single slice of YZ plane.
+Consistent orientation is a key intrinsic feature in the data which becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.
@@
 Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
-We found that different clustering techniques performed preferably for different measures.
+We found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).
 The two step clustering also made the problem more tractable and  the parameters intuitive for the end users.
@@
 These MetaTracts were anomalies caused by overlapping fibers and did not represent true fiber bundles.
-Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
+Second, if two MetaTracts were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
 A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
@@
 \subsection{Choice of clustering techniques}
+\label{subsec:clus_choice}
 As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
@@
 Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
-As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers.
+As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large  incorrect clusters and small(low-cardinality) outliers.
 When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers.
 As $h$ increases, some appropriate bundles start to form.
-Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
-%
-On the other hand, Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step for 10, 15, and 20 clusters.
+Even at $h=20$, the top fiber bundles incorrectly cluster together.
+%
+Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step into 10, 15, and 20 clusters.
 In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
+Even at $h=20$, no correct bundles are identified.
 \begin{figure}[tb]
@@
 	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
-	\caption{Applying only hierarchical (a) and K-means (b) clustering methods to MetaTracts for various numbers of clusters.}
+	\caption{Applying only (a)hierarchical and (b)dimensionality reduction followed by K-means clustering methods to MetaTracts for various numbers of clusters.
+Using min of directed Hausdorffs as distance measure.}
 	\label{fig:comparison}
@@
  %glass
-We tested our method on three datasets with varying characteristics. We report results on two due to lack of space.  
+We tested our method on three datasets with varying characteristics. We report results on two here and third in the supplemental section.  
 Dataset 1 is described in Sec.~\ref{sec:char_data} (Figure~\ref{fig:data-char}).
@@
 With different cross-section sizes and varying degree of curvature of the bundles, dataset 1 is a complicated dataset.
-Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of dataset 1 decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization mesh extraction.
-MetaTracts correctly identified the fiber bundles.
+Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of dataset 1 decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization and mesh extraction.
 Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e, Figure~\ref{fig:crop-16-decomp}d).
 
-The hierarchical clustering is robust against over segmentation. For example,
-Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Figure~\ref{fig:orientation_clustering}d).
+Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Figure~\ref{fig:orientation_clustering}d) clustered into 10 clusters, the ground truth being  5 clusters.
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
@@
 The orientation clustering is robust and done as a part of preprocessing.
-The hierarchical clustering itself, on the results of the orientation clusters can be done within an order of a few minutes on Intel Xeon E5-2667 workstation and can be repeated according to user input of $h$ (Sec.~\ref{sec:param_choices}).
+The hierarchical clustering by itself, on the results of the orientation clusters can be done within an order of a few minutes on Intel Xeon E5-2667 workstation and can be repeated according to user input of $h$ (Sec.~\ref{sec:param_choices}).
 We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without any extra computations.
@@
 LS consists of 5 levels from ``strongly disagree" to ``strongly agree".
-Figure~\ref{fig:userstudy} shows the results.
+Figure~\ref{fig:userstudy} shows the results in the sequential order they were asked.
 Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Figure~\ref{fig:data-char}a).
-This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user.
+This task tests the perception of geometric structure and how common problems like uniform grayscale values and occlusion of the different fiber bundles affects the user.
 Figure~\ref{fig:userstudy} shows that, as expected the participants found it difficult to determine the number of bundles from volume rendering alone.
@@
 We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset.
-Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts.
+Due to the problems of occlusions and similar grayscale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts.
 In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
@@
 K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern.
-Our framework is robust to the choice of $h$ (Sec.~\ref{sec:results}).
+Our framework is robust to the choice of $h$.
 Large $h$ keeps the major bundles intact.
 For example, in  Figure~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Figure~\ref{fig:len_dist_crop16} all have less than 10 elements).
-This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
+This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown (avoiding over-segmentation).
 The following parameters were fixed for all the tests.
-We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests.
-A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.
-On the other hand, a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced.
-The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al.
+We set the reliable Hessian threshold $R_{H}$ to be 0.3 for all our tests.
+A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.
+A very high $ R_{H}$ would lead to a decline in number of MetaTracts produced.
+The $\alpha$ and $\beta$ in $R_{H}$ are as explained in Frangi et al.
 \cite{Frangi1998} and set to 0.5.
-The length and the radius parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are.
+The length and the radius parameters for the cylinders of MetaTracts decide how coarse the fiber cylinders are.
 These are dependent on the underlying fiber characteristic and the weaving pattern.
-Larger cylinders will also handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber.
+Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber.
 We used 10.00 and 2.00 for length and radius (measured in grid voxel size), respectively for all tests.
@@
 Thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well.
-We compared  $m=3-7$ experimentally without any dramatic change in results.
+We compared  $m=3-7$ experimentally without any change in results.
 %\begin{figure}[tb]
@@
 We perform an user evaluation to indicate the usefulness of the method.
-In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up the extraction.
+In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.
 %

### 2014-12-02 metaTracts_supplymental_pvis2014.tex
Rebuttals.

@@
 \begin{document}
+	\title{MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites \\
+		Supplementary Material}
+	\author{Online submission ID: \\ 181}
+	\maketitle
+Section~\ref{sec:ext_exp} contains extended experiments including results on our third dataset.
+We also include the user evaluation  questionnaire as a separate file. The file contains links to videos which were seen by the participants. 
 \input {./supplemental/supplementaryMaterial.tex}

### 2014-12-02 supplemental/supplementaryMaterial.tex
Rebuttals.

--- 2014-09-26 supplemental/supplementaryMaterial.tex
@@
-\section{Supplementary Material}
-\label{sec:sm}
-Section~\ref{subsec:ext_exp} contains extended experiments.
-Section~\ref{subsec:usereval} contains our user evaluation questionnaire.
-\subsection{Extended Experiments}
-\label{subsec:ext_exp}
+%Section~\ref{subsec:usereval} contains our user evaluation questionnaire.
+
+\section{Extended Experiments}
+\label{sec:ext_exp}
 
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{images_pvis/dataset3.pdf}
+	\includegraphics[width=\linewidth]{images_pvis/dataset3.pdf}
 	\caption{Results on dataset3.
@@
 (b) MetaTracts embedded on the top 3 eigenvectors and clustered into 2 clusters.
-(c) Each orientation, hierarchically clustered into 10 clusters.}
+(c) Combined results of each orientation, hierarchically clustered into 10 clusters.}
 	\label{fig:dataset3}
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{images_pvis/dataset1-extended.pdf}
+	\includegraphics[width=\linewidth]{images_pvis/dataset1-extended.pdf}
 	\caption{Results on dataset1. (a) MetaTracts. (b) Voxelized MetaTracts. (c) Subset of the fiber bundles are shown to better visualize the interaction between particular bundles.}
@@
 
-Figure~\ref{fig:dataset1-extended} further shows how spatial context can be used to answer questions such as  Are these bundles in contact at a particular position in the data?
+Figure~\ref{fig:dataset1-extended} further shows how spatial context can be used to answer questions such as:  Are these bundles in contact at a particular position in the data?
 What is the relative orientations of the contacting bundles?
-These bundles are occluded in the original volume and no conclusions can be made from the original volume.
+These annotated bundles are occluded in the original volume and no conclusions can be made from the volume rendering.
 In the voxelized MetaTracts we can clearly see where and how the bundles interact.
-
-\subsection{User Evaluation Questionnaire}
-The questionnaire contains links to videos which were seen by the participants.
-\label{subsec:usereval}
-\pagebreak
-\includepdf[pages={1-13}]{./supplemental/MetaTracts_usereval.pdf}

### 2014-12-08 chapters/MetaTracts.tex
Updated  reviews

@@
 They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.
-In order to determine reliable Hessians, we compute a similar metric.
+In order to determine reliable Hessians, we compute
+\color{red}{ the same metric} \color{blue}{a similar metric}\color{black}.
 We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details.
@@
 
-\section{Conclusion and future work}
+\section{Limitations and future work}
 \label{sec:conclusions}
+\color{red}
+ A key assumption of the method is ``connectivity'' (Sec.~\ref{sec:char_data}).
+If the ``connectivity'' is not satisfied due to noise in the image, the then MetaTracts will be inaccurate. 
+The clustering process also assumes that the fiber bundles have a minimum width.
+\color{green}
 In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials.
@@
 We perform an user evaluation to indicate the usefulness of the method.
-In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.
+\color{black} In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.
 %

### 2014-12-08 chapters/MetaTracts.tex
shortened introduction

@@
 The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of these continuously rising demands.
-CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing parts made of conventional materials such as aluminum or steel.
-But the application areas of CFRPs are not limited on these two industries.
-For a wide range of different areas ranging from marine and civil engineering, wind energy, to sporting goods manufacturing, CFRP is highly interesting.
-The main reason therefore is found in CFRP materials having desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
-CFRP materials show these characteristics at considerably lower weight compared to conventional materials.
-At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
+%CFRPs have been successfully introduced in aeronautic and automotive %applications within recent years, replacing parts made of conventional %materials such as aluminum or steel.
+But the application areas of CFRPs %are not limited on these two industries.
+For a wide range of different %areas ranging from marine and civil engineering, wind energy, to sporting %goods manufacturing, CFRP is highly interesting.
+%The main reason therefore is found in CFRP materials having desirable %characteristics such as high specific stiffness, high specific strength %and high corrosion resistance. 
+CFRP materials are increasingly replacing conventional materials such as aluminum or steel, as they have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
+Moreover, CFRP materials show these characteristics at considerably lower weight.
+At the same time, highly complex and integrated components, which were previously impossible to manufacture may be produced from CFRPs.
 Primary structures and highly loaded components in aeronautics are one example.
 
-Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates addressed in this work, consist of two main components:
-a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.
-Various production processes are used to manufacture CFRP laminates with endless carbon fibers.
+Typically, carbon fiber reinforced polymer components and more specifically CFRP laminates with endless carbon fibers as addressed in this work, consist of two main components:
+a matrix, which acts as a bonding component, and the reinforcements, which allow for achieving the desired characteristics.
+Various production processes are used to manufacture CFRP laminates.
 Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern.
-These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric.
-Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth.
+These sheets of woven carbon fiber cloth are also referred to as fabric and form the geometrical structure of the final CFRP materials.
+Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth in similar or different orientation.
 Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate.
 Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
-The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles.
+The main function of the matrix is to act as a bonding between the individual carbon fiber bundles.
 After curing, the production process of the CFRP laminate is finished.
 
-The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
-Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
-On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material.
-On the other hand, it shows low resolution and may generate arbitrary results, e.g.
+The increasing share of CFRPs as well as the complexity of material system and final components has generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
+Ultrasonic testing (UT) is the most commonly used method for this purpose.
+While UT provides a quick and cost-efficient overview of the material, it also shows low resolution and may generate arbitrary results, e.g.
 due to the geometry of the final component.
 Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
-XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen.
-In the typical scanning geometry of industrial XCT, the specimen is placed on a rotary table between X-ray source and detector.
-When passing through the specimen the X-rays are attenuated by the material in the specimen.
-The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light.
-To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}.
-Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen.
-In state of the art devices, XCT can reach voxel sizes below 500 $nm$.
-XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens.
-Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry.
-The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector.
-The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.
+In contrast to UT, XCT generates a highly detailed 3D volumetric representation of the scanned specimen.
+In cone beam XCT geometry the specimen is placed on a rotary table between X-ray source and detector.
+The X-rays passing through the specimen get attenuated by the materials present.
+By transferring the X-rays in a scintillator layer into visible light, the detector records the corresponding 2D attenuation image (penetration image).
+The specimen is rotated stepwise recording at each position a 2D attenuation image to generate the full series of attenuation images of a $360^\circ$ rotation for a complete reconstruction of the data volume ~\cite{heinzl-2008-thesis}.
+Finally the series of penetration images is reconstructed to a 3D volumetric dataset of the specimen.
+While cone beam XCT can reach voxel sizes below 500 $nm$ generating high resolution volume data for comprehensive and detailed analyses, unfortunately there is still a trade-off between viewport and image resolution.
+The magnification reached within an XCT scan is determined by the used distances between source and specimen as well as source and detector.
+The magnification therefore directly influences both resolution and viewport: while higher resolutions decrease the viewport but show more details, lower resolutions allow for larger viewports and thus larger portions of the specimen.
 
@@
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
-Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
-On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
+Figure~\ref{fig:data-char} provides an example of our targeted dataset type. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
+Our work is motivated from the recent progress in two interrelated fields: Firstly, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
+Secondly, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
 %
@@
 Spatial context refers to how two or more bundles interact with each other.
-For example, spatial context answers questions such as: Are these bundles in contact at a particular position in the data?
+It provides answers to questions such as: Are these bundles in contact at a particular position in the data?
 What is the relative orientations of the contacting bundles?
 
-Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts.
-We detail and evaluate a method which uses visualization techniques to gain insight into our data.
-We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites.
+Providing answers to the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts.
+We present and evaluate a method which uses visualization techniques to gain insight into our data.
+We interpret and advance techniques from diffusion tensor imaging to extract and visualize geometric structures from XCT data of the woven carbon fiber reinforced composites.
 The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
@@
 Figure~\ref{fig:flowchart} shows an over view of our approach.
-Starting from the XCT data, we first generate MetaTracts, then we cluster the MetaTracts to create fiber bundles and finally generate visualizations from the clusters.
+Starting from the XCT data, we first generate MetaTracts.
+Then we cluster the MetaTracts to generate and visualize fiber bundles.
 For a detailed description of the MetaTracts generation process see Sec.~\ref{sec:approach}.

### 2014-12-09 chapters/MetaTracts.tex
minor updates
integrated acknowledgements, updated authors list

@@
  \vspace{-1.5em}
-\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square.}
+\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in white square.}
 \label{fig:reliable_hessian}
@@
 
-%\acknowledgements{
-% The research leading to these results has received funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM.
+\acknowledgements{
+The research leading to these results has received funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM.
 Our special thanks go to Ilya Ilya Straumit (KU Leuven) who provided us with dataset1.
-%}
+}

### 2014-12-10 chapters/MetaTracts.tex
paper pass until orientation based clustering

@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components.
-The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of these continuously rising demands.
+The material of choice for a growing number of applications is carbon fiber reinforced polymer (CFRP), which allows an integration of these continuously rising demands and increasingly replaces conventional materials such as aluminum or steel.
 %CFRPs have been successfully introduced in aeronautic and automotive %applications within recent years, replacing parts made of conventional %materials such as aluminum or steel.
@@
 %The main reason therefore is found in CFRP materials having desirable %characteristics such as high specific stiffness, high specific strength %and high corrosion resistance. 
-CFRP materials are increasingly replacing conventional materials such as aluminum or steel, as they have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
+CFRP materials have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
 Moreover, CFRP materials show these characteristics at considerably lower weight.
@@
 Primary structures and highly loaded components in aeronautics are one example.
-
-Typically, carbon fiber reinforced polymer components and more specifically CFRP laminates with endless carbon fibers as addressed in this work, consist of two main components:
+Typically, carbon fiber reinforced polymer components and more specifically CFRP laminates with endless carbon fibers as addressed in this work consist of two main components:
 a matrix, which acts as a bonding component, and the reinforcements, which allow for achieving the desired characteristics.
 Various production processes are used to manufacture CFRP laminates.
-Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern.
+Most of these processes start with the reinforcement component, weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern.
 These sheets of woven carbon fiber cloth are also referred to as fabric and form the geometrical structure of the final CFRP materials.
-Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth in similar or different orientation.
+Depending on the requirements of the final component, fabrics may be stacked in multiple layers in similar or different orientation.
 Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate.
-Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
+Resins are then integrated in the material system to fill the gaps in the fabric forming the matrix component.
 The main function of the matrix is to act as a bonding between the individual carbon fiber bundles.
@@
 
-The increasing share of CFRPs as well as the complexity of material system and final components has generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
+The increasing share of CFRPs as well as the complexity of both material system and final component has generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
 Ultrasonic testing (UT) is the most commonly used method for this purpose.
-While UT provides a quick and cost-efficient overview of the material, it also shows low resolution and may generate arbitrary results, e.g.
-due to the geometry of the final component.
+While UT provides a quick and cost-efficient overview of the material, it also lacks resolution and may generate arbitrary results, e.g., due to the geometry of the component.
 Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
@@
 By transferring the X-rays in a scintillator layer into visible light, the detector records the corresponding 2D attenuation image (penetration image).
-The specimen is rotated stepwise recording at each position a 2D attenuation image to generate the full series of attenuation images of a $360^\circ$ rotation for a complete reconstruction of the data volume ~\cite{heinzl-2008-thesis}.
-Finally the series of penetration images is reconstructed to a 3D volumetric dataset of the specimen.
+The specimen is rotated recording at each angular step a 2D attenuation image to generate the full series of attenuation images of a $360^\circ$ rotation required for a complete reconstruction of the data volume ~\cite{heinzl-2008-thesis}.
 While cone beam XCT can reach voxel sizes below 500 $nm$ generating high resolution volume data for comprehensive and detailed analyses, unfortunately there is still a trade-off between viewport and image resolution.
@@
 
-In this work, we focus on datasets with larger viewports and lower resolutions where the individual carbon fibers (filaments) are indiscernible or barely visible.
+In this work, we focus on datasets with larger viewports but lower resolutions where the individual carbon fibers (filaments) are indiscernible or barely visible.
 Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of individual fibers.
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Figure~\ref{fig:data-char} provides an example of our targeted dataset type. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
+Figure~\ref{fig:data-char} depicts our targeted dataset type. It shows the recurring fiber bundle pattern in the final CFRP laminate, the \textit{unit cell}.
 Our work is motivated from the recent progress in two interrelated fields: Firstly, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
-Secondly, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
+Secondly, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of NDT for fiber reinforced components~\cite{Schilling2005}.
 %
@@
 %
-While fiber bundles are now understood as highly important in determining component properties, the tools for visualization of the internal structure have not developed at the same pace.
+While fiber bundles are now understood as highly important in determining component properties, the tools for visualizing the internal structure have not developed at the same pace.
 To the best of our knowledge, there is no current work that can resolve simple queries such as:
@@
 \item{ Which fiber bundles show a particular orientation? }
-\item{ Which fiber bundles are of the same type of yarn (i.e. which bundles show similar sizes or diameters), which is the largest or smallest fiber bundle?}
+\item{ Which fiber bundles are of the same type of yarn? (i.e. which bundles show similar sizes or diameters, which is the largest or smallest fiber bundle?)}
 \end{itemize}
@@
 We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}.
-Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles.
+Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of them.
 Spatial context refers to how two or more bundles interact with each other.
@@
 What is the relative orientations of the contacting bundles?
-
-Providing answers to the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts.
+Providing answers to the above queries from the volume renderings of XCT datasets or from visual inspection of particular 2D slices is non-trivial, even for experts.
 We present and evaluate a method which uses visualization techniques to gain insight into our data.
 We interpret and advance techniques from diffusion tensor imaging to extract and visualize geometric structures from XCT data of the woven carbon fiber reinforced composites.
-The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
+The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{complete unit cells} of woven fabrics.
 \begin{figure}[tb]
@@
 Figure~\ref{fig:flowchart} shows an over view of our approach.
-Starting from the XCT data, we first generate MetaTracts.
-Then we cluster the MetaTracts to generate and visualize fiber bundles.
-For a detailed description of the MetaTracts generation process see Sec.~\ref{sec:approach}.
-For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}.
-For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}.
-Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.
+Starting from XCT data, we first generate MetaTracts (see Sec.~\ref{sec:approach}).
+Then we cluster the MetaTracts to generate and visualize fiber bundles (see Sec.~\ref{subsec:fiber-bundles}).
+We discuss the visualizations we created for our domain experts in Sec.~\ref{sec:vis}.
+Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} finally summarizes the user evaluation of our method.
 
@@
 Its main clinical application is found in the study and treatment of neurological disorders.
-DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of white fibers in the human brain and  brain connectivity.
-A variety of algorithms have been proposed aiming for generating fiber-tract trajectories.
+DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of fibers in the human brain and brain connectivity.
+A variety of algorithms have been proposed for generating fiber-tract trajectories.
 In general, these reconstructions of fiber trajectories are clustered in bundles which are expected to be related anatomically or spatially.
-We broadly divide the related work on DTI into two parts that are of immediate relevance to our proposed solution: fiber tracking and fiber clustering.
+We broadly divide the related work on DTI into two parts of immediate relevance to our proposed solution: fiber tracking and fiber clustering.
 We discuss methods for analyzing the second order local structure and review the current state of the art in the visual analysis of fiber reinforced composites.
@@
 Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}.
-Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
+Clustering requires choosing a suitable proximity measure and a method for grouping ``close" fibers.
 Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}.
-Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge, are clustered separately.
-Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors.
-The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{ Brun2004,jonasson2005, ODonnell2007}.
-
+Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers which are close for certain distance and then diverge, are clustered separately.
+Brun et al.~\cite{Brun2004} use normalized cuts along with a pairwise distances measure computed using a 9D fiber shape descriptors.
+The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{Brun2004,jonasson2005, ODonnell2007}.
 Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique, Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}), can be applied to the problem of organizing fiber tracts data.
@@
 In the case of fiber bundles, the individual points are fiber tracts.
-In the ideal case fiber tracts that belong to the same bundle must remain ``close" to each other in the lower dimensional space.
+In the ideal case fiber tracts belonging to the same bundle must remain ``close" to each other in the lower dimensional space.
 Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of point wise minimum distances between two fibers.
@@
 %The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}. 
-
-The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
+The agglomerative hierarchical clustering method~\cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
 These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships.
@@
 There are a number of difficulties in hierarchical clustering.
-First, computing all pairs distances for tracts to generate the distance matrix is time consuming (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times).
+First, computing all pairs' distances for tracts to generate the distance matrix is time consuming~\cite{Garyfallidis2012}.
 Second, a ``correct'' distance measure to compare tracts must be chosen.
@@
 \subsection{Second order local structure}
-%!!!!! TODO REFINE SECTION !!!!!
 Unlike DTI, we do not have diffusion tensor data.
@@
 The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores and deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.
-Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
+Fritz et al.~\cite{Fritz2009} proposed workflows for NDT practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
 This approach allows analyzing fiber orientations based on direction transfer functions.
@@
 Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers.
-
-Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014} in recent papers, discussed the problems and current available solutions in geometric modeling of three dimensional composites.
-Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position?
+Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014}, discussed problems and current available solutions in geometric modeling of three dimensional composites.
+Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position.
 The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.
-
 %Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research.
 Recently, similar techniques have been applied to fiber tracing in composite materials.
-To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers.
+To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of CFRP laminates.
 
@@
 \includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
-\caption{Data characteristics: (a) Rendering of the data.
-(b) 2D slices along the X-axis and (c) 2D slice along the Z-axis.
-(d) Magnified region marked in green square (c).
+\caption{Data characteristics: (a) Rendering of dataset D1.
+(b, c) 2D slices along Z- and X-axis.
+(d) Zoom in of the green region marked in (c).
 Multiple fiber bundles cross and are indistinguishable by visual inspection alone.
@@
 \end{figure}
+
 \section {Data Characteristic and Assumptions}
 \label{sec:char_data}
-Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
-Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$.
-The data type is unsigned 8 bit integer.
-Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
-The unit cell is used as a basis structure for CFRP manufacturing.
-The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
+Figure~\ref{fig:data-char} shows dataset D1 with woven fiber bundles.
+The size of D1 is 450$\times$300$\times$500 voxels with isotropic resolution of 2 $\mu m$ and 8 bit unsigned integer scalars.
+Figure~\ref{fig:data-char} clearly shows the recurring fiber bundle weaving pattern of the composite unit cell used for manufacturing.
 Figure~\ref{fig:data-char}a shows a volume rendering of the dataset.
-Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis.
-Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest.
-The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images.
+Figure~\ref{fig:data-char}b and ~\ref{fig:data-char}c show 2D slices along the $X$- and the $Z$-axis respectively.
+Figure~\ref{fig:data-char}d zooms into the green region of interest.
+The resolution of D1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images.
 Figure~\ref{fig:data-char}d contains two bundles going in opposite directions.
-The separation between two fiber bundles is barely visible.
-The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization.
+Also the separation between two fiber bundles is barely visible.
 The fiber bundles may differ in terms of the amount of fibers in the bundle.
@@
 Depending on the weaving pattern, the fiber bundles cross each other in different orientations.
-In general, the number of orientations is limited by the weaving pattern.
-Nevertheless, the weaving pattern may cause the individual bundles to be curved.
-In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.
+The number of orientations is defined by the weaving pattern and typically consists of two main orientations.
+The weaving pattern may cause individual bundles to be curved.
+In consequence, individual fibers may be adjacent in Euclidean distance but belong to different bundles.
 We make the following \textit{assumptions} on our data:
+
 \begin{itemize}[noitemsep]
 \item The structure embedded in the data contains fiber bundles of indiscernible fibers.
-\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
-\item The local orientation may change gradually inside the fiber bundle.
-\item The local orientation may be noisy and not reliable.
+\item Local orientation: each point in a fiber bundle has a local orientation parallel to the corresponding area in the fiber bundle.
+\item Local orientation may gradually change in the fiber bundle.
+\item Local orientation may be noisy and not reliable.
 \item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
@@
 Extracting the MetaTracts consists of two  major steps.
-First we extract a local orientation vector at each grid location (Sec.~\ref{subsec:rh}).
+First, we extract a local orientation vector at each grid location (Sec.~\ref{subsec:rh}).
 Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (Sec.~\ref{subsec:mtprop}).
@@
 
-The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
-As output, each grid location becomes associated with an unit vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation.
+The input to the first stage of our pipeline is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
+As output, each grid location gets associated with an unit vector representing the local orientation and a real value [0,1] representing a measure of the reliability of the local orientation.
 We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
 The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.
-
 The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
@@
 Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular.
-They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.
-In order to determine reliable Hessians, we compute
-\color{red}{ the same metric} \color{blue}{a similar metric}\color{black}.
-We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details.
-Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude.
+They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the \textit{vesselness} criterion.
+In order to determine \textit{reliable Hessians} for our approach, we make use of the vesselness criterion as follows: Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude.
 Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix.
@@
 The following measures are defined in~\cite{Frangi1998}.
+
 \begin{equation}\label{RA}
@@
 \end{equation}
+
 \begin{equation}\label{RB}
@@
 Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors.
-Thus, we compute ``reliable Hessians" ($R_H$) to determine which locations in the volume provide reliable local orientation.
-($R_H$ is similar to the ``vesselness" in~\cite{Frangi1998}).
+Thus, we compute reliable Hessians ($R_H$) to determine which locations in the volume provide reliable local orientation.
+%($R_H$ is similar to the vesselness in~\cite{Frangi1998}).
 %Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
@@
 The unit vector representing the principal direction has been mapped to RGB color space.
-Figure~\ref{fig:reliable_hessian}a shows the entire data set.
+%Figure~\ref{fig:reliable_hessian}a shows the entire data set. 
 Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue.
-Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively.
+Figure~\ref{fig:reliable_hessian}b and c show 2D slices along the Z and X axes respectively.
 Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest.
-Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable.
+Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and therefore considered as unreliable.
 The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.
-
 % We note some intrinsic differences between the DTI and our XCT data.
@@
 Thus the local orientation at each voxel for XCT data is inherently more noisy (Fig.~\ref{fig:r_h} Right) than those computed from eigen analysis of diffusion tensors in DTI datasets.
-
-We note some intrinsic differences between the DTI and our XCT data.
+We also note here some intrinsic differences between the DTI and our XCT data.
 Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}.
@@
 This is not the case for our data.
-The local orientaion at each voxel is inherently more noisy.
-
-%
+The local orientation at each voxel is inherently more noisy.
+
 \begin{figure}[tb]
@@
  \vspace{-1.5em}
-\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in white square.}
+\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b,c) 2D slices along Z- and X axis. (d) Zoom in of the green region marked in (c).}
 \label{fig:reliable_hessian}
@@
 We do this by interpreting the underlying geometric structure of the fibers as a set of cylinders.
-
 MetaTacts are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data.
@@
 \begin{enumerate}[noitemsep]
-\item{Each MetaTract is associated with a continuous set of cylinders.}
-\item{Each MetaTract is associated with a start point which is a grid vertex.}
-\item{Each individual cylinder in a MetaTract has a constant length, a constant radius and a start point (which is also a grid vertex).}
-\item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
-\item{Each cylinder is parallel to the local orientation at its start point.}
+\item{MetaTract are associated with a continuous set of cylinders.}
+\item{MetaTract are associated with a start point at a grid vertex.}
+\item{Individual cylinders in a MetaTract have a constant length, a constant radius and a start point (which is also a grid vertex).}
+\item{Individual cylinders in a MetaTract (except the first one) are connected to their previous cylinders at their start point.}
+\item{Individual cylinders are parallel to the local orientation at their start points.}
 \end{enumerate}
@@
 \label{subsec:MetaTracts-generation}
- In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$.The procedure trivially extends to $\mathbb{R}^3$.
+In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$ (the procedure trivially extends to $\mathbb{R}^3$).
 In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles.
 The reliable Hessian step generates a local orientation vector in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1].
-In Figure~\ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue.
-Figure~\ref{fig:algo}c shows a single cylinder of a MetaTracts.
+In Figure~\ref{fig:algo}, regions with unreliable Hessian have been marked in blue.
+Figure~\ref{fig:algo}c shows an individual cylinder of a MetaTract.
 Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}).
@@
   \vspace{-1em}
- \caption{MetaTracts generation: (b) MetaTract generation (a) close up of marked region in (b). (c) Parts of a single MetaTract.}
+ \caption{MetaTracts generation: (b) Generation process. (a) Zoom in of the white marked region in (b). (c) Individual cylinder.}
   \label{fig:algo}
@@
 \end{figure}
-The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract.
-We call these vertices \textit{``candidate vertices"}.
+The set of vertices which are in the green region but not overlapped by blue are possible \textit{``candidate vertices"} for the start point of the next cylinder in a particular MetaTract.
 From these we select another grid vertex which will be the start point for the next cylinder.
@@
 This approach tackles noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation in an already noisy environment.
-% 
 In the example in Figure~\ref{fig:algo}b we select the orange grid vertex next and repeat the process.
 We do not select the purple vertex (Figure~\ref{fig:algo}c) because it is not a candidate vertex.
-If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices.
-These short tracts are removed.
+If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and are therefore removed.
 The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}a.
@@
 %
-MetaTracts in a given fiber bundle do not extend the full length of the bundle. MetaTracts may have different lengths and may partially overlap.
+MetaTracts in a given fiber bundle may extend the full length of the bundle and may feature different lengths as well as partial overlaps.
 
@@
 
-The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles. 
-Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering.
-Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with fibers which partially overlap (Figure~\ref{fig:length_distribution}b).
-Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
-We found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).
+The set of MetaTracts generated by the previous step needs to be clustered in order to extract the final fiber bundles. 
+Carbon fibers show both orientation and geometric proximity information, both of which can be used for clustering.
+Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with partially overlapping fibers (Figure~\ref{fig:length_distribution}b).
+We also found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).  
+So instead of creating a heuristic to artificially combine the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
 The two step clustering also made the problem more tractable and the parameters intuitive for the end users.
@@
 This step divides the individual MetaTracts into classes based on their major orientations.
-In order to cluster MetaTracts going in the same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01}.
+In order to cluster MetaTracts going in the same directions, we use a spectral embedding technique called Laplacian eigenmaps as originally introduced by Belkin and Niyogi \cite{Belkin01}.
 An eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.

### 2014-12-10 chapters/MetaTracts.tex
roll back 1 commit

--- 2014-12-08 chapters/MetaTracts.tex
@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components.
-The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of these continuously rising demands.
-%CFRPs have been successfully introduced in aeronautic and automotive %applications within recent years, replacing parts made of conventional %materials such as aluminum or steel.
-But the application areas of CFRPs %are not limited on these two industries.
-For a wide range of different %areas ranging from marine and civil engineering, wind energy, to sporting %goods manufacturing, CFRP is highly interesting.
-%The main reason therefore is found in CFRP materials having desirable %characteristics such as high specific stiffness, high specific strength %and high corrosion resistance. 
-CFRP materials are increasingly replacing conventional materials such as aluminum or steel, as they have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
-Moreover, CFRP materials show these characteristics at considerably lower weight.
-At the same time, highly complex and integrated components, which were previously impossible to manufacture may be produced from CFRPs.
+The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.
+CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing parts made of conventional materials such as aluminum or steel.
+But the application areas of CFRPs are not limited on these two industries.
+For a wide range of different areas ranging from marine and civil engineering, wind energy, to sporting goods manufacturing, CFRP is highly interesting.
+The main reason therefore is found in CFRP materials having desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
+CFRP materials show these characteristics at considerably lower weight compared to conventional materials.
+At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
 Primary structures and highly loaded components in aeronautics are one example.
 
-Typically, carbon fiber reinforced polymer components and more specifically CFRP laminates with endless carbon fibers as addressed in this work, consist of two main components:
-a matrix, which acts as a bonding component, and the reinforcements, which allow for achieving the desired characteristics.
-Various production processes are used to manufacture CFRP laminates.
+Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates we are addressing in this work, consist of two main components:
+a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.
+Various production processes are used to manufacture CFRP laminates with endless carbon fibers.
 Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern.
-These sheets of woven carbon fiber cloth are also referred to as fabric and form the geometrical structure of the final CFRP materials.
-Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth in similar or different orientation.
+These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric.
+Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth.
 Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate.
 Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
-The main function of the matrix is to act as a bonding between the individual carbon fiber bundles.
+The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles.
 After curing, the production process of the CFRP laminate is finished.
 
-The increasing share of CFRPs as well as the complexity of material system and final components has generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
-Ultrasonic testing (UT) is the most commonly used method for this purpose.
-While UT provides a quick and cost-efficient overview of the material, it also shows low resolution and may generate arbitrary results, e.g.
+The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
+Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
+On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material.
+On the other hand, it shows low resolution and may generate arbitrary results, e.g.
 due to the geometry of the final component.
 Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
-In contrast to UT, XCT generates a highly detailed 3D volumetric representation of the scanned specimen.
-In cone beam XCT geometry the specimen is placed on a rotary table between X-ray source and detector.
-The X-rays passing through the specimen get attenuated by the materials present.
-By transferring the X-rays in a scintillator layer into visible light, the detector records the corresponding 2D attenuation image (penetration image).
-The specimen is rotated stepwise recording at each position a 2D attenuation image to generate the full series of attenuation images of a $360^\circ$ rotation for a complete reconstruction of the data volume ~\cite{heinzl-2008-thesis}.
-Finally the series of penetration images is reconstructed to a 3D volumetric dataset of the specimen.
-While cone beam XCT can reach voxel sizes below 500 $nm$ generating high resolution volume data for comprehensive and detailed analyses, unfortunately there is still a trade-off between viewport and image resolution.
-The magnification reached within an XCT scan is determined by the used distances between source and specimen as well as source and detector.
-The magnification therefore directly influences both resolution and viewport: while higher resolutions decrease the viewport but show more details, lower resolutions allow for larger viewports and thus larger portions of the specimen.
+XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen.
+In the typical scanning geometry of industrial XCT, the specimen is placed on a rotary table between X-ray source and detector.
+When passing through the specimen the X-rays are attenuated by the material in the specimen.
+The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light.
+To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}.
+Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen.
+In state of the art devices, XCT can reach voxel sizes below 500 $nm$.
+XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens.
+Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry.
+The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector.
+The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.
 
@@
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Figure~\ref{fig:data-char} provides an example of our targeted dataset type. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
-Our work is motivated from the recent progress in two interrelated fields: Firstly, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
-Secondly, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
+Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
+Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
+On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
 %
@@
 Spatial context refers to how two or more bundles interact with each other.
-It provides answers to questions such as: Are these bundles in contact at a particular position in the data?
+For example, spatial context answers questions such as: Are these bundles in contact at a particular position in the data?
 What is the relative orientations of the contacting bundles?
 
-Providing answers to the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts.
-We present and evaluate a method which uses visualization techniques to gain insight into our data.
-We interpret and advance techniques from diffusion tensor imaging to extract and visualize geometric structures from XCT data of the woven carbon fiber reinforced composites.
+Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts.
+We detail and evaluate a method which uses visualization techniques to gain insight into our data.
+We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites.
 The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
@@
 Figure~\ref{fig:flowchart} shows an over view of our approach.
-Starting from the XCT data, we first generate MetaTracts.
-Then we cluster the MetaTracts to generate and visualize fiber bundles.
-For a detailed description of the MetaTracts generation process see Sec.~\ref{sec:approach}.
+Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters.
+For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}.
 For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}.
@@
 In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles.
-Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}.
+Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}.
 Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
@@
 
-Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014} in recent papers, discussed the problems and current available solutions in geometric modeling of three dimensional composites.
-Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position?
-The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.
+Lomov et al.~\cite{Lomov2010} in a recent paper, discussed the current available solutions in geometric modeling of three dimensional composites.
+Modeling of the composites starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position?
+The second builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.
 
@@
 Recently, similar techniques have been applied to fiber tracing in composite materials.
-To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers.
+To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers, which is the main scope of this work.
 
@@
 In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.
-We make the following \textit{assumptions} on our data:
+We make the following assumptions on our data:
 \begin{itemize}[noitemsep]
@@
 The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
-As output, each grid location becomes associated with an unit vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation.
+As output, each grid location becomes associated with a vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation.
 We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
-The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.
-
-The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
-This direction coincides with the direction of the tubular structure.
+The principal directions, in which the local second order structure of the image can be decomposed, are given by the eigen decomposition of the Hessian matrix which describes the local curvature.
+The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.
 
@@
 They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.
-In order to determine reliable Hessians, we compute
-\color{red}{ the same metric} \color{blue}{a similar metric}\color{black}.
+In order to determine reliable Hessians, we compute a similar metric.
 We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details.
@@
 Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors.
-Thus, we compute ``reliable Hessians" ($R_H$) to determine which locations in the volume provide reliable local orientation.
-($R_H$ is similar to the ``vesselness" in~\cite{Frangi1998}).
-%Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
+Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
 $$
@@
 The local orientaion at each voxel is inherently more noisy.
+Thus, techniques popular in DTI cannot be directly applied to XCT data.
 
@@
 We build on the key assumptions on the data, namely ``local orientation" and ``connectivity" while taking into account the noise and lack of resolution.
-We do this by interpreting the underlying geometric structure of the fibers as a set of cylinders.
-
+We do this by interpreting the underlying geometric structure of the fibers as a set of polylines (cylinders).
 MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data.
@@
 \item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
-\item{Each cylinder is parallel to the local orientation at its start point.}
+\item{Each cylinder is locally parallel to the local orientation at its start point.}
 \end{enumerate}
@@
 In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles.
-The reliable Hessian step  generates a local orientation vector in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1].
+The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1].
 In Figure~\ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue.
@@
 The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow.
-We generate a rectangle of length $L$ and radius $R$ (properties 3 and 5 in Section~\ref{subsec:mtprop}).
+We generate a rectangle of length $L$ and radius $R$, the green dashed rectangle shows the rectangle generated at $C_p$ (properties 3 and 5 in Section~\ref{subsec:mtprop}).
 %
@@
   \vspace{-1em}
- \caption{MetaTracts generation: (b) MetaTract generation (a) close up of marked region in (b). (c) Parts of a single MetaTract.}
+ \caption{MetaTracts generation: (b) MetaTract generation, region with $R_H$ less than threshold. (a) close up of marked region in (b). (c) Parts of a single MetaTract.}
   \label{fig:algo}
@@
 %\end{figure}
-Based on assumptions on our data (local orientation and connectivity), the order of the candidate vertices is based on the following \textit{characteristics}:
+Based on such assumptions on our data as local orientation and connectivity (Sec.~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:
 
@@
 We measure the distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$.
-For example, in Figure~\ref{fig:algo}b the distance is measured as the Euclidean distance between the green and the orange vertices projected on to $N_p$.
+For example, in Figure~\ref{fig:algo}b the distance is measured as the Euclidean distance between the purple and the orange vertices projected on to $N_p$.
 We refer to this perpendicular projection distance as \textit{`projected\_dist'}.
@@
 If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices.
-These short tracts are removed.
+Therefore these short tracts are removed.
 The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}a.
 The MetaTracts are colored with the mean orientation direction mapped to the RGB space.
-Figure~\ref{fig:length_distribution}a also shows a single slice of YZ plane.
-Consistent orientation is a key intrinsic feature in the data which becomes visually pronounced.
+Figure~\ref{fig:length_distribution}a shows a single slice of YZ plane.
+Consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.
@@
 Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
-We found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).
+We found that different clustering techniques performed preferably for different measures.
 The two step clustering also made the problem more tractable and  the parameters intuitive for the end users.
@@
 These MetaTracts were anomalies caused by overlapping fibers and did not represent true fiber bundles.
-Second, if two MetaTracts were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
+Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
 A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
@@
 \subsection{Choice of clustering techniques}
-\label{subsec:clus_choice}
 As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
@@
 Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
-As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large  incorrect clusters and small(low-cardinality) outliers.
+As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers.
 When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers.
 As $h$ increases, some appropriate bundles start to form.
-Even at $h=20$, the top fiber bundles incorrectly cluster together.
-%
-Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step into 10, 15, and 20 clusters.
+Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
+%
+On the other hand, Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step for 10, 15, and 20 clusters.
 In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
-Even at $h=20$, no correct bundles are identified.
 \begin{figure}[tb]
@@
 	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
-	\caption{Applying only (a)hierarchical and (b)dimensionality reduction followed by K-means clustering methods to MetaTracts for various numbers of clusters.
-Using min of directed Hausdorffs as distance measure.}
+	\caption{Applying only hierarchical (a) and K-means (b) clustering methods to MetaTracts for various numbers of clusters.}
 	\label{fig:comparison}
@@
  %glass
-We tested our method on three datasets with varying characteristics. We report results on two here and third in the supplemental section.  
+We tested our method on three datasets with varying characteristics. We report results on two due to lack of space.  
 Dataset 1 is described in Sec.~\ref{sec:char_data} (Figure~\ref{fig:data-char}).
@@
 With different cross-section sizes and varying degree of curvature of the bundles, dataset 1 is a complicated dataset.
-Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of dataset 1 decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization and mesh extraction.
+Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of dataset 1 decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization mesh extraction.
+MetaTracts correctly identified the fiber bundles.
 Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e, Figure~\ref{fig:crop-16-decomp}d).
 
-Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Figure~\ref{fig:orientation_clustering}d) clustered into 10 clusters, the ground truth being  5 clusters.
+The hierarchical clustering is robust against over segmentation. For example,
+Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Figure~\ref{fig:orientation_clustering}d).
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
@@
 The orientation clustering is robust and done as a part of preprocessing.
-The hierarchical clustering by itself, on the results of the orientation clusters can be done within an order of a few minutes on Intel Xeon E5-2667 workstation and can be repeated according to user input of $h$ (Sec.~\ref{sec:param_choices}).
+The hierarchical clustering itself, on the results of the orientation clusters can be done within an order of a few minutes on Intel Xeon E5-2667 workstation and can be repeated according to user input of $h$ (Sec.~\ref{sec:param_choices}).
 We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without any extra computations.
@@
 	\centering
-	\includegraphics[width=\linewidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval.pdf}
+	\includegraphics[width=\linewidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval_AMA.pdf}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
@@
 LS consists of 5 levels from ``strongly disagree" to ``strongly agree".
-Figure~\ref{fig:userstudy} shows the results in the sequential order they were asked.
+Figure~\ref{fig:userstudy} shows the results.
 Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Figure~\ref{fig:data-char}a).
-This task tests the perception of geometric structure and how common problems like uniform grayscale values and occlusion of the different fiber bundles affects the user.
+This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user.
 Figure~\ref{fig:userstudy} shows that, as expected the participants found it difficult to determine the number of bundles from volume rendering alone.
@@
 We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset.
-Due to the problems of occlusions and similar grayscale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts.
+Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts.
 In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
@@
 K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern.
-Our framework is robust to the choice of $h$.
+Our framework is robust to the choice of $h$ (Sec.~\ref{sec:results}).
 Large $h$ keeps the major bundles intact.
 For example, in  Figure~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Figure~\ref{fig:len_dist_crop16} all have less than 10 elements).
-This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown (avoiding over-segmentation).
+This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
 The following parameters were fixed for all the tests.
-We set the reliable Hessian threshold $R_{H}$ to be 0.3 for all our tests.
-A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.
-A very high $ R_{H}$ would lead to a decline in number of MetaTracts produced.
-The $\alpha$ and $\beta$ in $R_{H}$ are as explained in Frangi et al.
+We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests.
+A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.
+On the other hand, a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced.
+The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al.
 \cite{Frangi1998} and set to 0.5.
-The length and the radius parameters for the cylinders of MetaTracts decide how coarse the fiber cylinders are.
+The length and the radius parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are.
 These are dependent on the underlying fiber characteristic and the weaving pattern.
-Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber.
+Larger cylinders will also handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber.
 We used 10.00 and 2.00 for length and radius (measured in grid voxel size), respectively for all tests.
@@
 Thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well.
-We compared  $m=3-7$ experimentally without any change in results.
+We compared  $m=3-7$ experimentally without any dramatic change in results.
 %\begin{figure}[tb]
@@
 
-\section{Limitations and future work}
+\section{Conclusion and future work}
 \label{sec:conclusions}
-\color{red}
- A key assumption of the method is ``connectivity'' (Sec.~\ref{sec:char_data}).
-If the ``connectivity'' is not satisfied due to noise in the image, the then MetaTracts will be inaccurate. 
-The clustering process also assumes that the fiber bundles have a minimum width.
-\color{green}
 In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials.
@@
 We perform an user evaluation to indicate the usefulness of the method.
-\color{black} In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.
+In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up the extraction.
 %

### 2014-12-10 Rebuttal.tex

@@
 Reliable Hessians is computed exactly equivalently to Vesselness for a single scale (Firangi et al.~\cite{Frangi1998}).
+We have  updated the corresponding sentence in the paper to reflect the same.
+We explain further about the utility being different.
+
 A common approach and one utilized by the Frangi et al.~\cite{Frangi1998}, is to consider the Taylor series expansion in the neighborhood of point $x$ in an image as.
@@
 
-A2.
+Answer.
+Due to the manufacturing process, a vast majority of composites have bundles crossing at ~90 degree angles.
+With the separation factors being width, spacings and compactness of fiber bundles.
+These variations are well represented in our data sets.
+We maintain that the technique is useful and applicable to a large percentage of composite polymers.
+Nevertheless, we have added a third dataset in the supplementary materials which will be included in a comprehensive technical report which we cannot add currently due to space restrictions.
+
+Figure~\ref{fig:angle_bundles}f, bundles 7,10 and 6 shows a common example of variations in the datasets which we extract accurately. 
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{images_pvis/clustering.pdf}
+\caption{Fiber bundles extraction}
+\label{fig:angle_bundles}
+\end{figure}
 
@@
 \color{red}
+  It seems to me that the extracted fibers may not be real fibers. Is this
+  true? If so, how well can one trust the weaving patterns extracted based
+  on the imaginary fibers? 
+  \color{black}
+  Answer. 
+Yes, the extracted fibers cannot be mapped to real fibers.
+One of the goals  we  focused on was to move away from extraction of individual fibers, instead work on direct reconstruction of bundles.
+The design decisions adheres to extracting reliable bundles directly from the images.
+ Developing methods which can precisely compute the deviations from original bundles remains a future direction of the project.
+
+\color{red}
 The authors claim that their method can deal with low resolution data.
@@
 \color{black}
-Answer.
+
+Answer. Figure~\ref{fig:glass_fibers} shows an example where each fiber is clearly demarcated compared to the type of data we address in the paper (Figure~\ref{fig:data-char-rebuttal}).
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{images_pvis/glass_fibers}
+\caption{High Resolution fiber}
+\label{fig:glass_fibers}
+\end{figure}
 
@@
 Answer.
-Belkin  and Niyogi's~\cite{Belkin01} laplacian eigenmaps embeds the high dimensional data ( represented as a graph) into a low dimensional space where the euclidean distance approximates the distances in higher dimension space.
+Belkin  and Niyogi's~\cite{Belkin01} laplacian eigenmaps embeds the high dimensional data ( represented as a graph ) into a low dimensional space where the Euclidean distance approximates the distances in higher dimension space.
 For such reason, Eigen embedding for dimensionality reduction has become quiet popular and has found favor in tackling a variety of problems.
@@
 \color{black}
-Answer.
-A key assumptions is ``connectivity'', which states that moving along the direction of non-noisy local orientations in small increments, we reach another neighborhood with similar local orientation.
-If the data is noisy to the point where this does not hold true and in large neighborhoods the Hessian is completely inaccurate then the process will fail.
-We also assume the fiber bundles have a minimum width beyond which the data is too low for the clustering process to succeed.
+Answer. We have added a ``Limitations and future work" Section to the end of the paper. We copy it here for completeness.
 
+ A key assumption of the method is ``connectivity'' (Sec. Data Characteristic).
+ If the ``connectivity'' is not satisfied due to noise in the image, the then MetaTracts will be inaccurate. 
+ The clustering process also assumes that the fiber bundles have a minimum width. 
 \bibliographystyle{abbrv}

### 2014-12-10 metaTracts-pvis2014.tex
Rebuttals.
integrated acknowledgements, updated authors list
pvis.tex main page local commit

@@
 
-%\documentclass{vgtc}                          % final (conference style)
-\documentclass[review]{vgtc}                 % review
+\documentclass{vgtc}                          % final (conference style)
+%\documentclass[review]{vgtc}                 % review
 %\documentclass[widereview]{vgtc}             % wide-spaced review
@@
 \usepackage{enumitem}
+\usepackage{color}
 %% We encourage the use of mathptmx for consistent usage of times font
@@
 %% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
+\onlineid{181}
 
@@
 \author{Arindam Bhattacharya\thanks{e-mail: arindamb86@gmail.com}\\ %
-        \scriptsize The Ohio State University %
-\and Christoph Heinzl\thanks{e-mail:christoph.heinzl@fh-wels.at}\\ %
+     \scriptsize The Ohio State University\\ %
+\and Christoph Heinzl\thanks{e-mail: christoph.heinzl@fh-wels.at}\\ %
      \scriptsize University of Applied Sciences - Upper Austria%
-\and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
-    \scriptsize The Ohio State University}
+\and Artem Amirkhanov\thanks{e-mail: artem.amirkhanov@fh-wels.at}\\ %
+     \scriptsize University of Applied Sciences - Upper Austria%
+\and Johann Kastner\thanks{e-mail: johann.kastner@fh-wels.at}\\ %
+     \scriptsize University of Applied Sciences - Upper Austria%
+\and Rephael Wenger\thanks{e-mail: wenger@cse.ohio-state.edu}\\ %
+    \scriptsize The Ohio State University\\ %
+}
 
@@
 This work introduces MetaTracts, a novel method  for extracting and visualizing individual fiber bundles and weaving patterns from X-ray computed tomography (XCT) scans of endless carbon fiber reinforced  polymers (CFRP).
-The  proposed work flow is designed  to analyze large areas of CFRP material  such as complete unit cells integrating the recurring weaving pattern.
+The proposed work flow is designed to analyze unit cells of CFRP materials integrating the recurring weaving pattern.
 It is designed to handle XCT scans of low resolution, in which individual fibers are not visible or are barely visible.
@@
 
-\CCScatlist{ 
-  \CCScat{K.6.1}{Management of Computing and Information Systems}%
-{Project and People Management}{Life Cycle};
-  \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
-}
+%\CCScatlist{ 
+%  \CCScat{K.6.1}{Management of Computing and Information Systems}%
+%{Project and People Management}{Life Cycle};
+%  \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
+%}
 

### 2014-12-10 chapters/MetaTracts.tex
new commands for modified text (added, replaced, removed, replacedWith)

--- 2014-12-10 chapters/MetaTracts.tex
@@
 \label{sec:conclusions}
-\color{red}
+\added{
 A key assumption of the method is ``connectivity'' (Sec.~\ref{sec:char_data}).
@@
 The clustering process also assumes that the fiber bundles have a minimum width.
-\color{green}
+}
+\removed{
 In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials.
@@
 We perform an user evaluation to indicate the usefulness of the method.
-\color{black} In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.
+}
+In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.
 %

### 2014-12-10 chapters/MetaTracts.tex
color coded most of the changes

@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components.
-The material of choice for a growing number of applications is carbon fiber reinforced polymer (CFRP), which allows an integration of these continuously rising demands and increasingly replaces conventional materials such as aluminum or steel.
-%CFRPs have been successfully introduced in aeronautic and automotive %applications within recent years, replacing parts made of conventional %materials such as aluminum or steel.
-But the application areas of CFRPs %are not limited on these two industries.
-For a wide range of different %areas ranging from marine and civil engineering, wind energy, to sporting %goods manufacturing, CFRP is highly interesting.
-%The main reason therefore is found in CFRP materials having desirable %characteristics such as high specific stiffness, high specific strength %and high corrosion resistance. 
+The material of choice for a growing number of applications \replacedWith{are}{is} carbon fiber reinforced polymer (CFRP), which \replacedWith{allow}{allows} an integration of \replacedWith{those}{these} continuously rising demands \added{and increasingly replaces conventional materials such as aluminum or steel.}
+\replacedWith
+{
+CFRP materials show these characteristics at considerably lower weight compared to conventional materials.
+At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
+Primary structures and highly loaded components in aeronautics are one example.
+}
+{
 CFRP materials have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
@@
 After curing, the production process of the CFRP laminate is finished.
+}
 
@@
 
-In this work, we focus on datasets with larger viewports but lower resolutions where the individual carbon fibers (filaments) are indiscernible or barely visible.
+\replacedWith
+{
+In this work, we focus on datasets with larger viewports and
+}
+{
+In this work, we focus on datasets with larger viewports but
+}
+lower resolutions where the individual carbon fibers (filaments) are indiscernible or barely visible.
 Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of individual fibers.
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Figure~\ref{fig:data-char} depicts our targeted dataset type. It shows the recurring fiber bundle pattern in the final CFRP laminate, the \textit{unit cell}.
-Our work is motivated from the recent progress in two interrelated fields: Firstly, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
-Secondly, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of NDT for fiber reinforced components~\cite{Schilling2005}.
+Figure~\ref{fig:data-char} \replacedWith{provides an example of}{depicts} our targeted dataset \added{type}.
+It \removed{also} shows the recurring fiber bundle pattern in the final CFRP laminate, \replacedWith{called a}{the} \textit{unit cell}.
+Our work is motivated from the recent progress in two interrelated fields: \replacedWith{On one hand}{Firstly}, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
+\replacedWith{On the other hand}{Secondly}, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of \replacedWith{non destructive testing}{NDT} for fiber reinforced components~\cite{Schilling2005}.
 %
@@
 %
-While fiber bundles are now understood as highly important in determining component properties, the tools for visualizing the internal structure have not developed at the same pace.
+While fiber bundles are now understood as highly important in determining component properties, the tools for \replacedWith{visualization of}{visualizing} the internal structure have not developed at the same pace.
 To the best of our knowledge, there is no current work that can resolve simple queries such as:
@@
 \item{ Which fiber bundles show a particular orientation? }
-\item{ Which fiber bundles are of the same type of yarn? (i.e. which bundles show similar sizes or diameters, which is the largest or smallest fiber bundle?)}
+\item{ Which fiber bundles are of the same type of yarn\added{?} (i.e. which bundles show similar sizes or diameters, which is the largest or smallest fiber bundle?\added{)}}
 \end{itemize}
@@
 We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}.
-Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of them.
+Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of \replacedWith{bundles}{them}.
 Spatial context refers to how two or more bundles interact with each other.
-It provides answers to questions such as: Are these bundles in contact at a particular position in the data?
+\replacedWith{For example, spatial context}{It provides} answers \added{to} questions such as: Are these bundles in contact at a particular position in the data?
 What is the relative orientations of the contacting bundles?
-Providing answers to the above queries from the volume renderings of XCT datasets or from visual inspection of particular 2D slices is non-trivial, even for experts.
-We present and evaluate a method which uses visualization techniques to gain insight into our data.
-We interpret and advance techniques from diffusion tensor imaging to extract and visualize geometric structures from XCT data of the woven carbon fiber reinforced composites.
-The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{complete unit cells} of woven fabrics.
+\replacedWith{Answering}{Providing answers to} the above queries from the volume renderings of XCT datasets or from visual inspection of particular 2D slices is non-trivial, even for experts.
+We \replacedWith{detail}{present} and evaluate a method which uses visualization techniques to gain insight into our data.
+We \replacedWith{advance and interpret}{interpret and advance} techniques from diffusion tensor imaging to extract and visualize geometric structures from \replacedWith{3D X-ray computed tomography}{XCT} data of the woven carbon fiber reinforced composites.
+The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \replacedWith{lower resolutions and thus larger areas}{\textit{complete unit cells}} of woven fabrics.
 \begin{figure}[tb]
@@
 Figure~\ref{fig:flowchart} shows an over view of our approach.
+\replacedWith
+{
+Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters.
+For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}.
+For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}.
+For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}.
+Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.
+} 
+{
 Starting from XCT data, we first generate MetaTracts (see Sec.~\ref{sec:approach}).
@@
 Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} finally summarizes the user evaluation of our method.
+}
 
@@
 Its main clinical application is found in the study and treatment of neurological disorders.
-DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of fibers in the human brain and brain connectivity.
-A variety of algorithms have been proposed for generating fiber-tract trajectories.
+DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of \removed{white} fibers in the human brain and brain connectivity.
+A variety of algorithms have been proposed \removed{aiming} for generating fiber-tract trajectories.
 In general, these reconstructions of fiber trajectories are clustered in bundles which are expected to be related anatomically or spatially.
-We broadly divide the related work on DTI into two parts of immediate relevance to our proposed solution: fiber tracking and fiber clustering.
+We broadly divide the related work on DTI into two parts of \removed{that are of} immediate relevance to our proposed solution: fiber tracking and fiber clustering.
 We discuss methods for analyzing the second order local structure and review the current state of the art in the visual analysis of fiber reinforced composites.
@@
 Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}.
-Clustering requires choosing a suitable proximity measure and a method for grouping ``close" fibers.
+Clustering requires choosing a suitable proximity measure and \removed{choosing} a method for grouping ``close" fibers.
 Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}.
-Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers which are close for certain distance and then diverge, are clustered separately.
-Brun et al.~\cite{Brun2004} use normalized cuts along with a pairwise distances measure computed using a 9D fiber shape descriptors.
+Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers \replacedWith{that}{which} are close for certain distance and then diverge, are clustered separately.
+Brun et al.~\cite{Brun2004} use\removed{s} normalized cuts along with a pairwise distances measure computed using a 9\removed{-}D fiber shape descriptors.
 The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{Brun2004,jonasson2005, ODonnell2007}.
@@
 There are a number of difficulties in hierarchical clustering.
-First, computing all pairs' distances for tracts to generate the distance matrix is time consuming~\cite{Garyfallidis2012}.
+First, computing all pairs\added{'} distances for tracts to generate the distance matrix is time consuming~\cite{Garyfallidis2012}.
 Second, a ``correct'' distance measure to compare tracts must be chosen.
@@
 The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores and deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.
-Fritz et al.~\cite{Fritz2009} proposed workflows for NDT practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
+Fritz et al.~\cite{Fritz2009} proposed \removed{interactive} workflows for \replacedWith{non destructive testing}{NDT} practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
 This approach allows analyzing fiber orientations based on direction transfer functions.
@@
 Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers.
-Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014}, discussed problems and current available solutions in geometric modeling of three dimensional composites.
+Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014} \removed{in recent papers}, discussed problems and current available solutions in geometric modeling of three dimensional composites.
 Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position.
@@
 Recently, similar techniques have been applied to fiber tracing in composite materials.
-To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of CFRP laminates.
+To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of \replacedWith{composite fibers, which is the main scope of this work}{CFRP laminates}.
 
@@
 \includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
-\caption{Data characteristics: (a) Rendering of dataset D1.
+\caption
+{
+\replacedWith
+{
+Data characteristics: (a) Rendering of the data.
+(b) 2D slices along the X-axis and (c) 2D slice along the Z-axis.
+(d) Magnified region marked in green square (c).
+Multiple fiber bundles cross and are indistinguishable by visual inspection alone.
+}
+{
+Data characteristics: (a) Rendering of dataset D1.
 (b, c) 2D slices along Z- and X-axis.
@@
 }
+}
 \label{fig:data-char}
@@
 \label{sec:char_data}
+\replacedWith
+{
+Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
+Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$.
+The data type is unsigned 8 bit integer.
+Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern.
+The unit cell is used as a basis structure for CFRP manufacturing.
+The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
+Figure~\ref{fig:data-char}a shows a volume rendering of the dataset.
+Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis.
+Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest.
+The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images.
+Figure~\ref{fig:data-char}d contains two bundles going in opposite directions.
+The separation between two fiber bundles is barely visible.
+The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization.
+The fiber bundles may differ in terms of the amount of fibers in the bundle.
+Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
+Depending on the weaving pattern, the fiber bundles cross each other in different orientations.
+In general, the number of orientations is limited by the weaving pattern.
+Nevertheless, the weaving pattern may cause the individual bundles to be curved.
+In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.
+We make the following assumptions on our data:
+}
+{
 Figure~\ref{fig:data-char} shows dataset D1 with woven fiber bundles.
@@
 We make the following \textit{assumptions} on our data:
+}
 
@@
 \item The structure embedded in the data contains fiber bundles of indiscernible fibers.
-\item Local orientation: each point in a fiber bundle has a local orientation parallel to the corresponding area in the fiber bundle.
-\item Local orientation may gradually change in the fiber bundle.
-\item Local orientation may be noisy and not reliable.
+\item Local orientation: each point \removed{with}in a fiber bundle has a local orientation \removed{which is} parallel to the corresponding area in the fiber bundle.
+\item \replacedWith{The l}{L}ocal orientation may \removed{change} gradually \replacedWith{inside}{change in} the fiber bundle.
+\item \replacedWith{The l}{L}ocal orientation may be noisy and not reliable.
 \item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
@@
 Extracting the MetaTracts consists of two  major steps.
-First, we extract a local orientation vector at each grid location (Sec.~\ref{subsec:rh}).
+First\added{,} we extract a local orientation vector at each grid location (Sec.~\ref{subsec:rh}).
 Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (Sec.~\ref{subsec:mtprop}).
@@
 The input to the first stage of our pipeline is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
-As output, each grid location gets associated with an unit vector representing the local orientation and a real value [0,1] representing a measure of the reliability of the local orientation.
+As output, each grid location \replacedWith{becomes}{gets} associated with an unit vector representing the local orientation and a real value [0,1] representing a measure of the reliability of the local orientation.
 We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
 The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.
-The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
-This direction coincides with the direction of the tubular structure.
-
+\replacedWith
+{The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.}
+{The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
+This direction coincides with the direction of the tubular structure.}
+
+\replacedWith
+{
+Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular.
+They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.
+In order to determine reliable Hessians, we compute a similar metric.
+We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details.
+Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude.
+Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix.
+Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$).
+The sign indicates if the vessel is bright in a dark background or dark in a bright background.
+In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$).
+The following measures are defined in~\cite{Frangi1998}.
+}
+{
 Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular.
@@
 The following measures are defined in~\cite{Frangi1998}.
+}
 
@@
 Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors.
-Thus, we compute reliable Hessians ($R_H$) to determine which locations in the volume provide reliable local orientation.
-%($R_H$ is similar to the vesselness in~\cite{Frangi1998}).
-%Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
+Thus, we compute \replacedWith{the ``vesselness'' measure}{reliable Hessians ($R_H$)} to determine which locations in the volume provide reliable local orientation.
 $$
@@
 Thus the local orientation at each voxel for XCT data is inherently more noisy (Fig.~\ref{fig:r_h} Right) than those computed from eigen analysis of diffusion tensors in DTI datasets.
-We also note here some intrinsic differences between the DTI and our XCT data.
+
+\replacedWith
+{We note some intrinsic differences between the DTI and our XCT data.
 Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}.
@@
 This is not the case for our data.
-The local orientation at each voxel is inherently more noisy.
+The local orientaion at each voxel is inherently more noisy.
+Thus, techniques popular in DTI cannot be directly applied to XCT data.}
+{We also note here some intrinsic differences between the DTI and our XCT data.
+Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}.
+The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding.
+This is not the case for our data.
+The local orientation at each voxel is inherently more noisy.}
 
@@
  \vspace{-1.5em}
-\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b,c) 2D slices along Z- and X axis. (d) Zoom in of the green region marked in (c).}
+\caption{Reliable Hessians.
+(a) Colored according to Orientation vector mapped to RGB.
+(b\added{,c}) 2D slice\added{s} along Z\replacedWith{ and (c) along}{-} and X axis.
+(d) \replacedWith{Magnified}{Zoom in of the green} region marked in \replacedWith{green square}{(c)}.}
 \label{fig:reliable_hessian}
@@
 We build on the key assumptions on the data, namely ``local orientation" and ``connectivity" while taking into account the noise and lack of resolution.
-We do this by interpreting the underlying geometric structure of the fibers as a set of cylinders.
+We do this by interpreting the underlying geometric structure of the fibers as a set of \removed{polylines(cylinders)} cylinders.
 MetaTacts are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data.
@@
 \begin{enumerate}[noitemsep]
-\item{MetaTract are associated with a continuous set of cylinders.}
-\item{MetaTract are associated with a start point at a grid vertex.}
+\item{\replacedWith{Each MetaTract is}{MetaTract are} associated with a continuous set of cylinders.}
+\item{\replacedWith{Each MetaTract is}{MetaTract are} associated with a start point at a grid vertex.}
 \item{Individual cylinders in a MetaTract have a constant length, a constant radius and a start point (which is also a grid vertex).}
@@
 In Figure~\ref{fig:algo}, regions with unreliable Hessian have been marked in blue.
-Figure~\ref{fig:algo}c shows an individual cylinder of a MetaTract.
+Figure~\ref{fig:algo}c shows \replacedWith{a single}{an individual} cylinder of a MetaTract.
 Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}).
@@
   \vspace{-1em}
- \caption{MetaTracts generation: (b) Generation process. (a) Zoom in of the white marked region in (b). (c) Individual cylinder.}
+ \caption{MetaTracts generation: (b) \replacedWith{MetaTract generation, region with $R_H$ less than threshold}{Generation process}.
+(a) \replacedWith{Close up of}{Zoom in of the white} marked region in (b).
+(c) \replacedWith{Parts of a single MetaTract}{Individual cylinder}.}
   \label{fig:algo}
@@
 \label{subsec:fiber-bundles}
-
-The set of MetaTracts generated by the previous step needs to be clustered in order to extract the final fiber bundles. 
-Carbon fibers show both orientation and geometric proximity information, both of which can be used for clustering.
-Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with partially overlapping fibers (Figure~\ref{fig:length_distribution}b).
+\replacedWith{The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles.}
+{The set of MetaTracts generated by the previous step needs to be clustered in order to extract the final fiber bundles. }
+\replacedWith{Carbon fibers hav}{Carbon fibers show} both orientation and geometric proximity information, both of which can be used for clustering.
+Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with \replacedWith{fibers which partially overlap}{partially overlapping fibers} (Figure~\ref{fig:length_distribution}b).
 We also found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).  
-So instead of creating a heuristic to artificially combine the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
+\added{So} instead of creating a heuristic \replacedWith{which}{to} artificially combine the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
 The two step clustering also made the problem more tractable and the parameters intuitive for the end users.
@@
 This step divides the individual MetaTracts into classes based on their major orientations.
-In order to cluster MetaTracts going in the same directions, we use a spectral embedding technique called Laplacian eigenmaps as originally introduced by Belkin and Niyogi \cite{Belkin01}.
+In order to cluster MetaTracts going in the same directions, we use a spectral embedding technique called Laplacian eigenmaps \replacedWith{which was}{as} originally introduced by Belkin and Niyogi \cite{Belkin01}.
 An eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
@@
 These MetaTracts were anomalies caused by overlapping fibers and did not represent true fiber bundles.
-Second, if two MetaTracts were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
+Second, if two \replacedWith{fiber bundles}{MetaTracts} were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
 A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
@@
 	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
-	\caption{Applying only (a)hierarchical and (b)dimensionality reduction followed by K-means clustering methods to MetaTracts for various numbers of clusters.
-Using min of directed Hausdorffs as distance measure.}
+	\caption{Applying only (a)hierarchical and (b)\added{dimensionality reduction followed by} K-means clustering methods to MetaTracts for various numbers of clusters \added{(distance measure is minimal directed Hausdorffs.}}
 	\label{fig:comparison}
@@
  %glass
-We tested our method on three datasets with varying characteristics. We report results on two here and third in the supplemental section.  
+We tested our method on three datasets with varying characteristics. We report results on two \replacedWith{due to lack of space}{here and third in the supplemental section}.  
 Dataset 1 is described in Sec.~\ref{sec:char_data} (Figure~\ref{fig:data-char}).
@@
 Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of dataset 1 decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization and mesh extraction.
-Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e, Figure~\ref{fig:crop-16-decomp}d).
-
-Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Figure~\ref{fig:orientation_clustering}d) clustered into 10 clusters, the ground truth being  5 clusters.
+\removed{MetaTracts correctly identified the fiber bundles.} Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e, Figure~\ref{fig:crop-16-decomp}d).
+
+\removed{The hierarchical clustering is robust against over segmentation. For example,}
+Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster in orientation 1 (Figure~\ref{fig:orientation_clustering}d) \added{clustered into 10 clusters, the ground truth being  5 clusters}.
 We generate the correct clusters and outlier bundles with very few elements which can be discarded.
@@
 LS consists of 5 levels from ``strongly disagree" to ``strongly agree".
-Figure~\ref{fig:userstudy} shows the results in the sequential order they were asked.
+Figure~\ref{fig:userstudy} shows the results \added{in the sequential order they were asked}.
 Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Figure~\ref{fig:data-char}a).
@@
 For example, in  Figure~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Figure~\ref{fig:len_dist_crop16} all have less than 10 elements).
-This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown (avoiding over-segmentation).
+This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown \added{(avoiding over-segmentation)}.
 The following parameters were fixed for all the tests.
@@
 A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.
-A very high $ R_{H}$ would lead to a decline in number of MetaTracts produced.
+\replacedWith{On the other hand, a}{A} very high $ R_{H}$ would lead to a decline in number of MetaTracts produced.
 The $\alpha$ and $\beta$ in $R_{H}$ are as explained in Frangi et al.
@@
 
+\added
+{
 \acknowledgements{
 The research leading to these results has received funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM.
-Our special thanks go to Ilya Ilya Straumit (KU Leuven) who provided us with dataset1.
-}
+Our special thanks go to Ilya Straumit (KU Leuven) who provided us with dataset1.
+}
+}

### 2014-12-10 chapters/MetaTracts.tex
shortened the paper to fit within 8 pages

@@
 	\label{fig:flowchart}
+	\vskip-0.2cm
 \end{figure}
@@
 \label{fig:data-char}
+\vskip-0.2cm
 \end{figure}
@@
 {
-Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
-Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$.
+Figure \ref{fig:data-char} shows \replacedWith{dataset 1}{D1} with woven fiber bundles.
+\replacedWith{Dataset 1}{D1} is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$.
 The data type is unsigned 8 bit integer.
@@
 Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest.
-The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images.
+The resolution of \replacedWith{dataset 1}{D1} renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images.
 Figure~\ref{fig:data-char}d contains two bundles going in opposite directions.
@@
 \label{fig:reliable_hessian}
+\vskip-0.2cm
 \end{figure}
@@
   \label{fig:algo}
-    \vspace{-1.5em}
+    %\vspace{-1.5em}
+    \vskip-0.2cm
 \end{figure}
@@
 	\label{fig:length_distribution}
+	\vskip-0.2cm
 \end{figure}
@@
   \label{fig:orientation_clustering}
+  \vskip-0.2cm
   \end{figure}
@@
 	\label{fig:comparison}
+	\vskip-0.2cm
 \end{figure}
@@
 	\label{fig:crop-16-decomp}
+	\vskip-0.2cm
 \end{figure}
@@
 	\label{fig:prepreg}
+	\vskip-0.2cm
 \end{figure}
@@
 We tested our method on three datasets with varying characteristics. We report results on two \replacedWith{due to lack of space}{here and third in the supplemental section}.  
-Dataset 1 is described in Sec.~\ref{sec:char_data} (Figure~\ref{fig:data-char}).
-Dataset 1 has eleven separate fiber bundles (5 along one orientation and 6 along the other).
+Dataset \added{D}1 is described in Sec.~\ref{sec:char_data} (Figure~\ref{fig:data-char}).
+\replacedWith{Dataset 1}{D1} has eleven separate fiber bundles (5 along one orientation and 6 along the other).
 All of them were identified correctly by MetaTracts.
-With different cross-section sizes and varying degree of curvature of the bundles, dataset 1 is a complicated dataset.
-Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of dataset 1 decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization and mesh extraction.
+With different cross-section sizes and varying degree of curvature of the bundles, \replacedWith{dataset 1}{D1} is a complicated dataset.
+Figure~\ref{fig:orientation_clustering} and Figure~\ref{fig:crop-16-decomp} shows the result of \replacedWith{dataset 1}{D1} decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters followed by voxelization and mesh extraction.
 \removed{MetaTracts correctly identified the fiber bundles.} Note how the thin and curved purple cluster bundle 7 is extracted well (Figure~\ref{fig:orientation_clustering}e, Figure~\ref{fig:crop-16-decomp}d).
@@
 Figure~\ref{fig:length_distribution}b shows the length per MetaTract distribution for a particular fiber bundle.
-
-Figure~\ref{fig:prepreg} shows the results on our dataset 2.
+Figure~\ref{fig:prepreg} shows the results on our dataset 2 \added{(D2)}.
 This dataset has different characteristics, and is dense with \textit{flat and thin} bundles, it consists of 4 bundles along one direction and 5 along the other.
-The dimensions are 300$\times$350$\times$300 and the datatype is uint16 (dataset 1 is uint8).
+The dimensions are 300$\times$350$\times$300 and the datatype is uint16 (\replacedWith{dataset 1}{D1} is uint8).
 The bundles are indistinguishable in the original data set Figure~\ref{fig:prepreg}a.
@@
 
-The fiber bundle extraction using MetaTracts was implemented in C++ using ITK.
-All clustering was done in R.
+\removed{The fiber bundle extraction using }MetaTracts \replacedWith{was}{are} implemented in C++ using ITK.
+\removed{All }Clustering \replacedWith{was}{is} done in R.
 The preprocessing is time consuming and may take several hours.
@@
 	\label{fig:userstudy}
+	\vskip-0.2cm
 \end{figure}
-Four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization participated in the user evaluation of our work.
+Four NDT practitioners \replacedWith{who are familiar}{working} with \replacedWith{X-Ray computed tomography}{XCT}\removed{ and its visual representations}, four material scientists \removed{who are} experienced in analyzing \replacedWith{fiber reinforced polymers}{CFRPs} and four \replacedWith{practitioners of visualization}{visualization practitioners} participated in the user evaluation of \replacedWith{our work}{MetaTracts}.
 We setup a questionnaire with questions on solving given tasks as well as questions using the Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques.
@@
 Figure~\ref{fig:userstudy} shows the results \added{in the sequential order they were asked}.
-Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Figure~\ref{fig:data-char}a).
-This task tests the perception of geometric structure and how common problems like uniform grayscale values and occlusion of the different fiber bundles affects the user.
-Figure~\ref{fig:userstudy} shows that, as expected the participants found it difficult to determine the number of bundles from volume rendering alone.
-Another common approach is to color map the top eigenvector of Hessian matrix to RGB space.
-We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Figure~\ref{fig:reliable_hessian}a) and asked the same query.
-Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test.
-The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective.
-Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Figure~\ref{fig:orientation_clustering}d,e).
-All participants could correctly identify the number of fiber bundles.
-All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering.
-They especially liked the two dimensional slices of the MetaTracts (Figure~\ref{fig:crop-16-decomp}b, compared to Figure~\ref{fig:data-char}c).
-The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts.
-We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset.
-Due to the problems of occlusions and similar grayscale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts.
+\replacedWith{Task 1}{First task}\removed{ of the questionnaire} was to \removed{``}identify the number of fiber bundles\removed{"} using a video\replacedWith{ showing}{of} 360$\degree$ rotation of the\replacedWith{ volume rendering of dataset 1}{D1 volume rendering} (Figure~\ref{fig:data-char}a).\removed{ This task tests the perception of geometric structure and how common problems like uniform grayscale values and occlusion of the different fiber bundles affects the user.}\replacedWith{ Figure~\ref{fig:userstudy} shows that, as expected}{Results show that} the participants found it difficult to determine the number of bundles from volume rendering alone.\replacedWith{ Another common approach is to color map the top eigenvector of Hessian matrix to RGB space.
+We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors}{Second task evaluates RGB color mapping of the top Hessian matrix eigenvector.
+We again show a full 360$\degree$ rotation video where eigenvectors are color-mapped and rendered using volume rendering }\replacedWith{ (similar to}{(as in} Figure~\ref{fig:reliable_hessian}a) and asked the same query.
+Compared to the volume rendering, the Hessian color map provides more information about the major orientations and \replacedWith{we see}{results demonstrate} that the evaluators performed better than simple volume rendering in this test.
+The majority of our participants agreed or strongly agreed \removed{(LS)} that Hessian color map was more effective.
+Finally, we showed the video of bundles extracted with MetaTracts (for orientation 1 and 2 separately Figure~\ref{fig:orientation_clustering}d,e).
+All participants could correctly identify the number of fiber bundles\replacedWith{.
+All participants}{and} agreed that MetaTracts \replacedWith{was}{are} more effective in perceiving geometric structure than simple volume rendering or Hessian color map.
+They especially liked the \replacedWith{two dimensional}{2D} slices of the MetaTracts (Figure~\ref{fig:crop-16-decomp}b, compared to Figure~\ref{fig:data-char}c).
+
+The next set of tasks evaluated the spatial context perceived from MetaTracts.\replacedWith{ We achieve this by asking the evaluators}{The participants were asked} to select the fiber bundle which is largest, smallest, or the most similar to a given one from a \removed{sub}set of fiber bundles.
+Due to the problems of occlusions and similar grayscale values these tests proven to be more difficult to perform successfully with volume rendering of the plain data, compared to MetaTracts.
 In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
-Finally, the majority participants agreed that the extracted meshes provides better spatial context than volume rendering.
+Finally, the majority of participants agreed that the extracted meshes provides better spatial context than volume rendering.
 
@@
 	\label{fig:len_dist_crop16}
+	\vskip-0.2cm
 \end{figure}
@@
 We used 10.00 and 2.00 for length and radius (measured in grid voxel size), respectively for all tests.
-Simpler geometry (example dataset 2) was experimentally found to handle larger cylinders better.
+Simpler geometry (\replacedWith{example dataset 2}{e.g., D2}) was experimentally found to handle larger cylinders better.
 $\eta$ in sec.~\ref{subsec:dist_clustering} decides how quickly the hierarchical clustering converges, experimentally 0.3 to 0.6 gave similar results.
@@
 The research leading to these results has received funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM.
-Our special thanks go to Ilya Straumit (KU Leuven) who provided us with dataset1.
-}
-}
+We especially thank Ilya Straumit (KU Leuven) for providing us with dataset 1.
+}
+}

### 2014-12-11 chapters/MetaTracts.tex
Major changes fits in 8 pages

--- 2014-12-10 chapters/MetaTracts.tex
@@
 Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components.
-The material of choice for a growing number of applications is carbon fiber reinforced polymer (CFRP), which allows an integration of these continuously rising demands and increasingly replaces conventional materials such as aluminum or steel.
+The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of these continuously rising demands.
 %CFRPs have been successfully introduced in aeronautic and automotive %applications within recent years, replacing parts made of conventional %materials such as aluminum or steel.
@@
 %The main reason therefore is found in CFRP materials having desirable %characteristics such as high specific stiffness, high specific strength %and high corrosion resistance. 
-CFRP materials have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
+CFRP materials are increasingly replacing conventional materials such as aluminum or steel, as they have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance.
 Moreover, CFRP materials show these characteristics at considerably lower weight.
@@
 Primary structures and highly loaded components in aeronautics are one example.
-Typically, carbon fiber reinforced polymer components and more specifically CFRP laminates with endless carbon fibers as addressed in this work consist of two main components:
+
+Typically, carbon fiber reinforced polymer components and more specifically CFRP laminates with endless carbon fibers as addressed in this work, consist of two main components:
 a matrix, which acts as a bonding component, and the reinforcements, which allow for achieving the desired characteristics.
 Various production processes are used to manufacture CFRP laminates.
-Most of these processes start with the reinforcement component, weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern.
+Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern.
 These sheets of woven carbon fiber cloth are also referred to as fabric and form the geometrical structure of the final CFRP materials.
-Depending on the requirements of the final component, fabrics may be stacked in multiple layers in similar or different orientation.
+Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth in similar or different orientation.
 Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate.
-Resins are then integrated in the material system to fill the gaps in the fabric forming the matrix component.
+Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component.
 The main function of the matrix is to act as a bonding between the individual carbon fiber bundles.
@@
 
-The increasing share of CFRPs as well as the complexity of both material system and final component has generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
+The increasing share of CFRPs as well as the complexity of material system and final components has generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}.
 Ultrasonic testing (UT) is the most commonly used method for this purpose.
-While UT provides a quick and cost-efficient overview of the material, it also lacks resolution and may generate arbitrary results, e.g., due to the geometry of the component.
+While UT provides a quick and cost-efficient overview of the material, it also shows low resolution and may generate arbitrary results, e.g.
+due to the geometry of the final component.
 Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}.
@@
 By transferring the X-rays in a scintillator layer into visible light, the detector records the corresponding 2D attenuation image (penetration image).
-The specimen is rotated recording at each angular step a 2D attenuation image to generate the full series of attenuation images of a $360^\circ$ rotation required for a complete reconstruction of the data volume ~\cite{heinzl-2008-thesis}.
+The specimen is rotated stepwise recording at each position a 2D attenuation image to generate the full series of attenuation images of a $360^\circ$ rotation for a complete reconstruction of the data volume ~\cite{heinzl-2008-thesis}.
+Finally the series of penetration images is reconstructed to a 3D volumetric dataset of the specimen.
 While cone beam XCT can reach voxel sizes below 500 $nm$ generating high resolution volume data for comprehensive and detailed analyses, unfortunately there is still a trade-off between viewport and image resolution.
@@
 
-In this work, we focus on datasets with larger viewports but lower resolutions where the individual carbon fibers (filaments) are indiscernible or barely visible.
+In this work, we focus on datasets with larger viewports and lower resolutions where the individual carbon fibers (filaments) are indiscernible or barely visible.
 Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of individual fibers.
 %of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
-Figure~\ref{fig:data-char} depicts our targeted dataset type. It shows the recurring fiber bundle pattern in the final CFRP laminate, the \textit{unit cell}.
+Figure~\ref{fig:data-char} provides an example of our targeted dataset type. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
 Our work is motivated from the recent progress in two interrelated fields: Firstly, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}.
-Secondly, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of NDT for fiber reinforced components~\cite{Schilling2005}.
+Secondly, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.
 %
@@
 %
-While fiber bundles are now understood as highly important in determining component properties, the tools for visualizing the internal structure have not developed at the same pace.
+While fiber bundles are now understood as highly important in determining component properties, the tools for visualization of the internal structure have not developed at the same pace.
 To the best of our knowledge, there is no current work that can resolve simple queries such as:
@@
 \item{ Which fiber bundles show a particular orientation? }
-\item{ Which fiber bundles are of the same type of yarn? (i.e. which bundles show similar sizes or diameters, which is the largest or smallest fiber bundle?)}
+\item{ Which fiber bundles are of the same type of yarn (i.e. which bundles show similar sizes or diameters), which is the largest or smallest fiber bundle?}
 \end{itemize}
@@
 We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}.
-Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of them.
+Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles.
 Spatial context refers to how two or more bundles interact with each other.
@@
 What is the relative orientations of the contacting bundles?
-Providing answers to the above queries from the volume renderings of XCT datasets or from visual inspection of particular 2D slices is non-trivial, even for experts.
+
+Providing answers to the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts.
 We present and evaluate a method which uses visualization techniques to gain insight into our data.
 We interpret and advance techniques from diffusion tensor imaging to extract and visualize geometric structures from XCT data of the woven carbon fiber reinforced composites.
-The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{complete unit cells} of woven fabrics.
+The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
 \begin{figure}[tb]
@@
 Figure~\ref{fig:flowchart} shows an over view of our approach.
-Starting from XCT data, we first generate MetaTracts (see Sec.~\ref{sec:approach}).
-Then we cluster the MetaTracts to generate and visualize fiber bundles (see Sec.~\ref{subsec:fiber-bundles}).
-We discuss the visualizations we created for our domain experts in Sec.~\ref{sec:vis}.
-Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} finally summarizes the user evaluation of our method.
+Starting from the XCT data, we first generate MetaTracts.
+Then we cluster the MetaTracts to generate and visualize fiber bundles.
+For a detailed description of the MetaTracts generation process see Sec.~\ref{sec:approach}.
+For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}.
+For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}.
+Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.
 
@@
 Its main clinical application is found in the study and treatment of neurological disorders.
-DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of fibers in the human brain and brain connectivity.
-A variety of algorithms have been proposed for generating fiber-tract trajectories.
+DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of white fibers in the human brain and  brain connectivity.
+A variety of algorithms have been proposed aiming for generating fiber-tract trajectories.
 In general, these reconstructions of fiber trajectories are  clustered in bundles which are expected to be related anatomically or spatially.
-We broadly divide the related work on DTI into two parts of immediate relevance to our proposed solution: fiber tracking and fiber clustering.
+We broadly divide the related work on DTI into two parts that are of immediate relevance to our proposed solution: fiber tracking and fiber clustering.
 We discuss  methods for analyzing the second order local structure and review the current state of the art in the visual analysis of fiber reinforced composites.
@@
 \label{subsec:fiberEx} 
-A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Basser2000, Mori1999, Mori2002}.
+A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Mori1999, Mori2002}.
 Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion.
@@
 Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers.
-Basser et al.~\cite{Basser2002,Basser2000} proposed that white matter tracts could be represented as 3D curves in space.
+Basser et al.~\cite{Basser2002} proposed that white matter tracts could be represented as 3D curves in space.
 They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data.
@@
 \label {subsec:fiberClus}
-In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles.
+In DTI, similarity measures are used to group fiber tracts into bundles.
 Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}.
-Clustering requires choosing a suitable proximity measure and a method for grouping ``close" fibers.
+Clustering requires a suitable proximity measure and a method for grouping ``close" fibers.
 Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}.
-Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers which are close for certain distance and then diverge, are clustered separately.
-Brun et al.~\cite{Brun2004} use normalized cuts along with a pairwise distances measure computed using a 9D fiber shape descriptors.
-The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{Brun2004,jonasson2005, ODonnell2007}.
+Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge, are clustered separately.
+Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors.
+The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{ Brun2004,jonasson2005, ODonnell2007}.
+
 Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}), can be applied to the problem of organizing fiber tracts data.
@@
 In the case of fiber bundles, the individual points are fiber tracts.
-In the ideal case fiber tracts belonging to the same bundle must remain ``close" to each other in the lower dimensional space.
+In the ideal case fiber tracts that belong to the same bundle must remain ``close" to each other in the lower dimensional space.
 Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of point wise minimum distances between two fibers.
@@
 %The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}.
-The agglomerative hierarchical clustering method~\cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
+
+The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
 These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships.
@@
 There are a number of difficulties in hierarchical clustering.
-First, computing all pairs' distances for tracts to generate the distance matrix is time consuming~\cite{Garyfallidis2012}.
+First, computing all pairs distances for tracts to generate the distance matrix is time consuming.
 Second, a ``correct'' distance measure to compare tracts must be chosen.
@@
 \subsection{Second order local structure}
+%!!!!! TODO REFINE SECTION !!!!!
 Unlike DTI, we do not have diffusion tensor data.
 Instead, we have a scalar volume with tubular structures embedded in them.
-Analyzing curvilinear structures in volumetric images has been utilized for a variety of purposes including center line extraction~\cite{Bouix2005} and vascular image enhancement as proposed by Frangi et al.~\cite{Frangi1998} and Sato et al.~\cite{Sato1997}.
+Analyzing curvilinear structures in volumetric images has been utilized for a variety of purposes including center line extraction~\cite{Bouix2005} and vascular image enhancement~\cite{Frangi1998,Sato1997}.
 Frangi et al.~\cite{Frangi1998} introduced a method based on studying the eigenvalues of the  Hessian matrix specifically for the purposes of developing vessel enhancement filters.
@@
 \subsection {Visual Analysis and Modeling of Fiber Reinforced Composites}
-The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores and deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.
-Fritz et al.~\cite{Fritz2009} proposed workflows for NDT practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
-This approach allows analyzing fiber orientations based on direction transfer functions.
-Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites.
-They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
+The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fiber extraction from high resolution data where the individual fibers are clearly discernible.
+Fritz et al.~\cite{Fritz2009} proposed interactive workflows%for non destructive testing practitioners
+to explore and quantify steel fibers in reinforced sprayed concrete. %This approach allows analyzing fiber orientations based on direction transfer functions. 
+Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites.%They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
 Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
-Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014}, discussed problems and current available solutions in geometric modeling of three dimensional composites.
-Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position.
+Lomov et al.~\cite{Lomov2010} discusses the problems and current available solutions in geometric modeling of three dimensional composites.
+%Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position?
 The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.
+
 %Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research.
 Recently, similar techniques have been applied to fiber tracing in composite materials.
-To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of CFRP laminates.
+%To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers.
 
@@
 \includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
-\caption{Data characteristics: (a) Rendering of dataset D1.
-(b, c) 2D slices along Z- and X-axis.
-(d) Zoom in of the green region marked in (c).
+\caption{Data characteristics: (a) Rendering of the data.
+(b) 2D slices along the X-axis and (c) 2D slice along the Z-axis.
+(d) Magnified region marked in green square (c).
 Multiple fiber bundles cross and are indistinguishable by visual inspection alone.
@@
 \end{figure}
-
 \section {Data Characteristic and Assumptions}
 \label{sec:char_data}
-Figure~\ref{fig:data-char} shows dataset D1 with woven fiber bundles.
-The size of D1 is 450$\times$300$\times$500 voxels with isotropic resolution of 2 $\mu m$ and 8 bit unsigned integer scalars.
-Figure~\ref{fig:data-char} clearly shows the recurring fiber bundle weaving pattern of the composite unit cell used for manufacturing.
+Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles.
+Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$.
+The data type is unsigned 8 bit integer.
+Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as a regional subset of the recurring fiber bundle weaving pattern.
+The unit cell is used as a basis structure for CFRP manufacturing.
+The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material.
 Figure~\ref{fig:data-char}a shows a volume rendering of the dataset.
-Figure~\ref{fig:data-char}b and ~\ref{fig:data-char}c show 2D slices along the $X$- and the $Z$-axis respectively.
-Figure~\ref{fig:data-char}d zooms into the green region of interest.
-The resolution of D1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images.
+Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis.
+Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest.
+The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images.
 Figure~\ref{fig:data-char}d contains two bundles going in opposite directions.
-Also the separation between two fiber bundles is barely visible.
+The separation between two fiber bundles is barely visible.
+The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization.
 The fiber bundles may differ in terms of the amount of fibers in the bundle.
@@
 Depending on the weaving pattern, the fiber bundles cross each other in different orientations.
-The number of orientations is defined by the weaving pattern and typically consists of two main orientations.
-The weaving pattern may cause individual bundles to be curved.
-In consequence, individual fibers may be adjacent in Euclidean distance but belong to different bundles.
+In general, the number of orientations is limited by the weaving pattern.
+Nevertheless, the weaving may cause the individual bundles to be curved.
+In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.
 We make the following \textit{assumptions} on our data:
-
 \begin{itemize}[noitemsep]
 \item The structure embedded in the data contains fiber bundles of indiscernible fibers.
-\item Local orientation: each point in a fiber bundle has a local orientation parallel to the corresponding area in the fiber bundle.
-\item Local orientation may gradually change in the fiber bundle.
-\item Local orientation may be noisy and not reliable.
+\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
+The local orientation may be noisy and not reliable.
+The local orientation may change gradually inside the fiber bundle.
+%\item The local orientation may change gradually inside the fiber bundle. 
+%\item The local orientation may be noisy and not reliable.
 \item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
@@
 Extracting the MetaTracts consists of two  major steps.
-First, we extract a local orientation vector at each grid location (Sec.~\ref{subsec:rh}).
+First we extract a local orientation vector at each grid location (Sec.~\ref{subsec:rh}).
 Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (Sec.~\ref{subsec:mtprop}).
@@
 
-The input to the first stage of our pipeline is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
-As output, each grid location gets associated with an unit vector representing the local orientation and a real value [0,1] representing a measure of the reliability of the local orientation.
+The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
+As output, each grid location becomes associated with an unit vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation.
 We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
-The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.
+The Hessian matrix represents the local curvature.
+The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.
 The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
 This direction coincides with the direction of the tubular structure.
-
 Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular.
-They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the \textit{vesselness} criterion.
-In order to determine \textit{reliable Hessians} for our approach, we make use of the vesselness criterion as follows: Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude.
+They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.
+In order to determine reliable Hessians, we compute
+ the same metric.
+We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details.
+Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude.
 Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix.
@@
 The following measures are defined in~\cite{Frangi1998}.
-
 \begin{equation}\label{RA}
@@
 \end{equation}
-
 \begin{equation}\label{RB}
@@
 Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors.
-Thus, we compute reliable Hessians ($R_H$) to determine which locations in the volume provide reliable local orientation.
-%($R_H$ is similar to the vesselness in~\cite{Frangi1998}).
+Thus, we compute ``reliable Hessians" ($R_H$) to determine which locations in the volume provide reliable local orientation.
 %Thus, we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.
@@
 The $R_{H}$ is interpreted as a reliability measure of the local orientation.
-Grid locations where the $R_{H}$ is above a cutoff threshold are marked as regions with reliable orientations (see Sec.~\ref{sec:param_choices} on parameter choices).
+Grid locations where the $R_{H}$ is above a cutoff threshold are marked as regions with reliable orientations (see Sec.~\ref{sec:param_choices}).
 %\begin{figure}[h]
@@
 The unit vector representing the principal direction has been mapped to RGB color space.
-%Figure~\ref{fig:reliable_hessian}a shows the entire data set. 
-Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue.
-Figure~\ref{fig:reliable_hessian}b and c show 2D slices along the Z and X axes respectively.
+Figure~\ref{fig:reliable_hessian}a shows the entire data set.
+%Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue.
+Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively.
 Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest.
-Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and therefore considered as unreliable.
+Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable.
 The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.
+
 % We note some intrinsic differences between the DTI and our XCT data.
@@
 Thus the local orientation at each voxel for XCT data is inherently more noisy (Fig.~\ref{fig:r_h} Right) than those computed from eigen analysis of diffusion tensors in DTI datasets.
-We also note here some intrinsic differences between the DTI and our XCT data.
+
+We note some intrinsic differences between the DTI and our XCT data.
 Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}.
-The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding.
-This is not the case for our data.
-The local orientation at each voxel is inherently more noisy.
+The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding.The local orientation at each voxel in XCT data is inherently more noisy.
 
+%
 \begin{figure}[tb]
@@
  \vspace{-1.5em}
-\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b,c) 2D slices along Z- and X axis. (d) Zoom in of the green region marked in (c).}
+\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in white square.}
 \label{fig:reliable_hessian}
@@
 We do this by interpreting the underlying geometric structure of the fibers as a set of cylinders.
+
  MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data.
@@
 \begin{enumerate}[noitemsep]
-\item{MetaTract are associated with a continuous set of cylinders.}
-\item{MetaTract are associated with a start point at a grid vertex.}
-\item{Individual cylinders in a MetaTract have a constant length, a constant radius and a start point (which is also a grid vertex).}
-\item{Individual cylinders in a MetaTract (except the first one) are connected to their previous cylinders at their start point.}
-\item{Individual cylinders are parallel to the local orientation at their start points.}
+\item{Each MetaTract is associated with a continuous set of cylinders.}
+\item{Each MetaTract is associated with a start point which is a grid vertex.}
+\item{Each individual cylinder in a MetaTract has a constant length, a constant radius and a start point (which is also a grid vertex).}
+\item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
+\item{Each cylinder is parallel to the local orientation at its start point.}
 \end{enumerate}
@@
 \label{subsec:MetaTracts-generation}
-In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$ (the procedure trivially extends to $\mathbb{R}^3$).
+ In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$.The procedure trivially extends to $\mathbb{R}^3$.
 In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles.
 The reliable Hessian step  generates a local orientation vector in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1].
-In Figure~\ref{fig:algo}, regions with unreliable Hessian have been marked in blue.
-Figure~\ref{fig:algo}c shows an individual cylinder of a MetaTract.
+In Figure~\ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue.
+Figure~\ref{fig:algo}c shows a single cylinder of a MetaTracts.
 Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}).
@@
   \vspace{-1em}
- \caption{MetaTracts generation: (b) Generation process. (a) Zoom in of the white marked region in (b). (c) Individual cylinder.}
+ \caption{MetaTracts generation: (b) MetaTract generation (a) close up of marked region in (b). (c) Parts of a single MetaTract.}
   \label{fig:algo}
@@
 \end{figure}
-The set of vertices which are in the green region but not overlapped by blue are possible \textit{``candidate vertices"} for the start point of the next cylinder in a particular MetaTract.
+The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract.
+We call these vertices \textit{``candidate vertices"}.
 From these we select another grid vertex which will be the start point for the next cylinder.
@@
 This approach tackles noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation in an already noisy environment.
+% 
 In the example in Figure~\ref{fig:algo}b we select the orange grid vertex next and repeat the process.
 We do not select the purple vertex (Figure~\ref{fig:algo}c) because it is not a candidate vertex.
-If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and are therefore removed.
+If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices.
+These short tracts are removed.
 The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}a.
@@
 %
-MetaTracts in a given fiber bundle may extend the full length of the bundle and may feature different lengths as well as partial overlaps.
+MetaTracts in a given fiber bundle do not extend the full length of the bundle. MetaTracts may have different lengths and may partially overlap.
 
@@
 
-The set of MetaTracts generated by the previous step needs to be clustered in order to extract the final fiber bundles. 
-Carbon fibers show both orientation and geometric proximity information, both of which can be used for clustering.
-Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with partially overlapping fibers (Figure~\ref{fig:length_distribution}b).
-We also found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).  
-So instead of creating a heuristic to artificially combine the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
+The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles. 
+Carbon fibers have both orientation and geometric proximity information, which can be used for clustering.
+Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with fibers which partially overlap (Figure~\ref{fig:length_distribution}b).
+Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
+We found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).
 The two step clustering also made the problem more tractable and  the parameters intuitive for the end users.
@@
 This step divides the individual MetaTracts into classes based on their major orientations.
-In order to cluster MetaTracts going in the same directions, we use a spectral embedding technique called Laplacian eigenmaps as originally introduced by Belkin and Niyogi \cite{Belkin01}.
+In order to cluster MetaTracts going in the same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01}.
 An eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
@@
 The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an $m$-dimensional space (see Sec.~\ref{sec:param_choices} for values of m).
-In our case, each MetaTract is a data point.
-We introduce a simple \textit{orientation based similarity measure}.
-There is no spatial information involved, just a partition of the MetaTracts based on orientation which is already inherent in the data.
-Following this, given a pair of  MetaTracts, we define the edge weights between two tracts as the cosine of the maximum angle between the local orientations ($N_P$) of all pairs of start points ($C_P$) between the two MetaTracts.
+ In our case, each MetaTract is a data point. We introduce a simple \textit{orientation based similarity measure} to partition the  MetaTracts based on orientation already inherent in the data.
+Given a pair of  MetaTracts, we define the edge weights between two tracts as the cosine of the maximum angle between the local orientations ($N_P$) of all pairs of start points ($C_P$) between the two MetaTracts.
 The edge weights give a distance matrix representing the distance between each pair of points.
 Using the Belkin and Niyogi algorithm we ``embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix.
-%
-%August 13 2014
-%For all our tests the fibers were then embedded in a $\eta$ low dimensional space.
-The number of dimensions in the high dimensional space is equal to the number of datapoints or the number of MetaTracts in our case.
-%For all our test cases we set $\eta$ to be five (see sec~\ref{sec:param_choices} for parameter choices for $\eta$.).
-%  
  We used conventional K-means for clustering this lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
-For our test case there are two major directions of the fiber bundles and $K$ was set to 2.
+For our test case there are two major directions of the fiber bundles ($K=2$).
 Due to the advantages offered by the dimensionality reduction, even if there are curved fiber bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
-Figrue.~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
+Figure.~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
 As is expected there is a clear distinction based on fiber orientation.
@@
 %  
-Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering.
-As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large  incorrect clusters and small(low-cardinality) outliers.
-When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers.
+Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters without first performing orientation clustering (ground truth is 11 clusters).
+As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large  incorrect clusters and small (low-cardinality) outliers.
+When number of clusters ($h$) is 10, two large incorrect clusters are generated and the rest are outliers.
 As $h$ increases, some appropriate bundles start to form.
@@
 %
-Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step into 10, 15, and 20 clusters.
+Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering by proximity alone, after being embedded in a $m$-dimensional (lower) space, without a second hierarchical clustering step into 10, 15, and 20 clusters.
 In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
@@
 \label {sec:results}
-%\begin{figure*}
-%\centering
-%\captionsetup[subfigure]{labelformat=empty}
-%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-G.PNG}}
-%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-H.PNG}}
-%	%\subfloat[]{\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6-I.PNG}}
-%	
-%	\subfloat[(a) Dat set 2]{\includegraphics[width=0.7\linewidth]{imagesMT2014/prepreg/Capture8.PNG}}
-%	\subfloat[(b) Data set 3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/dataset203.png}}	
-%	\caption{(a) Shows in order, the volume rendering of the second data set, a single (2D) slice, the orientation clustering bundles and the final bundles followed by proximity clustering.
-(b) Shows in order the third data set, a single (2D) slice and finally the extracted fiber bundles.}
-%	\label{fig:crop-6}
-%\end{figure*}
-
 \begin{figure}[tb]
@@
  %glass
-We tested our method on three datasets with varying characteristics. We report results on two here and third in the supplemental section.  
+We tested our method on three datasets with varying characteristics. We report results on two here.  
 Dataset 1 is described in Sec.~\ref{sec:char_data} (Figure~\ref{fig:data-char}).
@@
 The hierarchical clustering by itself, on the results of the orientation clusters can be done within an order of a few minutes on Intel Xeon E5-2667 workstation and can be repeated according to user input of $h$ (Sec.~\ref{sec:param_choices}).
-We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without any extra computations.
-
+%We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without any extra computations.
 \section {User Evaluation}
@@
 	\centering
-	\includegraphics[width=\linewidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval_AMA.pdf}
+	\includegraphics[width=\linewidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval.pdf}
 	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
@@
 We setup a questionnaire with questions on solving given tasks as well as questions using the Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques.
-LS consists of 5 levels from ``strongly disagree" to ``strongly agree".
-Figure~\ref{fig:userstudy} shows the results in the sequential order they were asked.
-Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Figure~\ref{fig:data-char}a).
+ %LS consists of 5 levels from ``strongly disagree" to ``strongly agree".
+Figure~\ref{fig:userstudy} shows the results of the queries in sequential order.
+Task 1 was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Figure~\ref{fig:data-char}a).
 This task tests the perception of geometric structure and how common problems like uniform grayscale values and occlusion of the different fiber bundles affects the user.
-Figure~\ref{fig:userstudy} shows that, as expected the participants found it difficult to determine the number of bundles from volume rendering alone.
-Another common approach is to color map the top eigenvector of Hessian matrix to RGB space.
-We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Figure~\ref{fig:reliable_hessian}a) and asked the same query.
-Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test.
+As expected the participants found it difficult to determine the number of bundles from volume rendering alone.
+Another commonly used  approach is to color map the top eigenvector of Hessian matrix to RGB space.
+We showed a video of the volume rendering of the color mapped eigenvectors (similar to Figure~\ref{fig:reliable_hessian}a) and asked the same query.
+Compared to the volume rendering, the Hessian color map provides more information about the major orientations and the evaluators performed better in this test than in the direct volume rendering test.
 The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective.
 Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Figure~\ref{fig:orientation_clustering}d,e).
-All participants could correctly identify the number of fiber bundles.
-All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering.
+All participants could correctly identify the number of fiber bundles and agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering.
 They especially liked the two dimensional slices of the MetaTracts (Figure~\ref{fig:crop-16-decomp}b, compared to Figure~\ref{fig:data-char}c).
-The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts.
-We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset.
-Due to the problems of occlusions and similar grayscale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts.
-In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
+The next set of tasks, tested the spatial context perceived from MetaTracts.
+We asked the evaluators to select the largest fiber bundle from a subset of fiber bundles, similarly the smallest from a subset and given a fiber bundle, to select the most similar one from a subset.
+The majority evaluators strongly agreed that the tasks were straightforward compared to volume rendering.
 Finally, the majority participants agreed that the extracted meshes provides better spatial context than volume rendering.
@@
 Large $h$ keeps the major bundles intact.
-For example, in  Figure~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Figure~\ref{fig:len_dist_crop16} all have less than 10 elements).
+For example, in  Figure~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters are removed (clusters 2,3,4,5 and 7 in Figure~\ref{fig:len_dist_crop16} all have less than 10 elements).
 This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown (avoiding over-segmentation).
 The following parameters were fixed for all the tests.
-We set the reliable Hessian threshold $R_{H}$ to be 0.3 for all our tests.
-A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.
-A very high $ R_{H}$ would lead to a decline in number of MetaTracts produced.
+We set the reliable Hessian threshold $R_{H}$ to be 0.3.
+A $R_{H}$ of 0.0 means all points have reliable local orientation which might cause spurious MetaTracts detection.
+A very high $ R_{H}$ leads to a decline in number of MetaTracts produced.
 The $\alpha$ and $\beta$ in $R_{H}$ are as explained in Frangi et al.
@@
 %\end{figure}
-
-%\begin{figure}[htp]
-%\centering
-% \includegraphics[width=\linewidth]{imagesMetaTracts/glass-clus-2.eps}
-%  \caption{ Fig  A and C shows the results fiber extraction. B and D shows the results of clustering with clipping planes.}\label{fig:closeup-full}
-%\end{figure}
-
 \section{Limitations and future work}
 \label{sec:conclusions}
-\added{
 A key assumption of the method is ``connectivity'' (Sec.~\ref{sec:char_data}).
-If the ``connectivity'' is not satisfied due to noise in the image, the then MetaTracts will be inaccurate. 
+If the ``connectivity'' is not satisfied due to noise in the image, then the generated MetaTracts will be inaccurate. 
 The clustering process also assumes that the fiber bundles have a minimum width.
-}
-\removed{
+%\color{green}
 In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials.
@@
 We perform an user evaluation to indicate the usefulness of the method.
-}
+\color{black} 
 In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.
@@
 \acknowledgements{
-The research leading to these results has received funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM.
-Our special thanks go to Ilya Ilya Straumit (KU Leuven) who provided us with dataset1.
+This work was supported by funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM.
+We thank Ilya Straumit (KU Leuven) for providing us dataset1.
 }

### 2014-12-11 metaTracts-pvis2014.tex
new commands for modified text (added, replaced, removed, replacedWith)
color coded most of the changes
shortened the paper to fit within 8 pages
Major changes fits in 8 pages

--- 2014-09-25 metaTracts-pvis2014.tex
@@
 \and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
-    \scriptsize The Ohio State University}
+\scriptsize The Ohio State University
+\and Artem Amirkhanov\thanks{e-mail:artem.amirkhanov@fh-wels.at}\\ %
+     \scriptsize University of Applied Sciences - Upper Austria%
+\and Johann Kastner\thanks{e-mail:johann.kastner@fh-wels.at}\\ %
+	     \scriptsize University of Applied Sciences - Upper Austria%
+	    }
 
@@
 \input {./chapters/MetaTracts.tex}
-
 \bibliographystyle{abbrv}

### 2014-12-11 Rebuttal.tex
Major changes fits in 8 pages

@@
 We maintain that the technique is useful and applicable to a large percentage of composite polymers.
-Nevertheless, we have added a third dataset in the supplementary materials which will be included in a comprehensive technical report which we cannot add currently due to space restrictions.
+We have demonstrated the technique on three datasets, two of them are in the paper and the third is in the supplemental section due space  constraints.
+Getting additional datasets at this time is not possible
 
-Figure~\ref{fig:angle_bundles}f, bundles 7,10 and 6 shows a common example of variations in the datasets which we extract accurately. 
+In dataset 1, Figure~\ref{fig:angle_bundles}f, bundles 7,10 and 6 shows a common example of variations in the datasets which we extract accurately. 
 \begin{figure}
@@
 
+\begin{figure}[htp]
+\centering
+ \includegraphics[width=0.4\linewidth]{imagesMT2014/glass-cluster-a.eps}
+  \includegraphics[width=0.4\linewidth]{imagesMT2014/glass-cluster-b.eps}
+   \includegraphics[width=0.4\linewidth]{imagesMT2014/glass-scalar.eps}
+  \caption{ Fig  A and C shows the results fiber extraction. B and D shows the results of clustering with clipping planes.}\label{fig:closeup-full}
+\end{figure}
+
 \section{Reviewer 4}

### 2014-12-11 chapters/MetaTracts.tex

--- 2014-12-10 chapters/MetaTracts.tex
@@
 \label {subsec:fiberClus}
-In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles.
+In DTI, similarity measures such as proximity between fibers are used to group fiber tracts into bundles.
 Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}.
@@
 \subsection {Visual Analysis and Modeling of Fiber Reinforced Composites}
-The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores and deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.
-Fritz et al.~\cite{Fritz2009} proposed \removed{interactive} workflows for \replacedWith{non destructive testing}{NDT} practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
-This approach allows analyzing fiber orientations based on direction transfer functions.
-Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites.
-They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
+The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fiber extraction from high resolution data where the individual fibers are clearly discernible.
+Fritz et al.~\cite{Fritz2009} proposed interactive workflows%for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
+%This approach allows analyzing fiber orientations based on direction transfer functions.
+Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites.%They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
 Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
-Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014} \removed{in recent papers}, discussed problems and current available solutions in geometric modeling of three dimensional composites.
-Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position.
+Lomov et al.~\cite{Lomov2010} discusses the problems and current available solutions in geometric modeling of three dimensional composites.
+%Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position?
 The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.
+
 %Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research.
 Recently, similar techniques have been applied to fiber tracing in composite materials.
-To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of \replacedWith{composite fibers, which is the main scope of this work}{CFRP laminates}.
+%To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers.
 
@@
 
-The input to the first stage of our pipeline is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
-As output, each grid location \replacedWith{becomes}{gets} associated with an unit vector representing the local orientation and a real value [0,1] representing a measure of the reliability of the local orientation.
+The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$.
+As output, each grid location \replacedWith{becomes}{gets} associated with an unit vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation.
 We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
-The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.
+The Hessian matrix represents the local curvature.
+The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
+This direction coincides with the direction of the tubular structure.
 \replacedWith
@@
 
-\replacedWith
-{
+The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
+This direction coincides with the direction of the tubular structure.
+
 Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular.
 They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.
-In order to determine reliable Hessians, we compute a similar metric.
+In order to determine reliable Hessians, we compute
+\color{red}{ the same metric} \color{blue}{a similar metric}\color{black}.
 We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details.
@@
 The following measures are defined in~\cite{Frangi1998}.
-}
-{
-Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular.
-They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the \textit{vesselness} criterion.
-In order to determine \textit{reliable Hessians} for our approach, we make use of the vesselness criterion as follows: Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude.
-Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix.
-Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$).
-The sign indicates if the vessel is bright in a dark background or dark in a bright background.
-In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$).
-The following measures are defined in~\cite{Frangi1998}.
-}
-
 \begin{equation}\label{RA}
@@
 
+%
 \begin{figure}[tb]
@@
 The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an $m$-dimensional space (see Sec.~\ref{sec:param_choices} for values of m).
-In our case, each MetaTract is a data point.
-We introduce a simple \textit{orientation based similarity measure}.
-There is no spatial information involved, just a partition of the MetaTracts based on orientation which is already inherent in the data.
-Following this, given a pair of  MetaTracts, we define the edge weights between two tracts as the cosine of the maximum angle between the local orientations ($N_P$) of all pairs of start points ($C_P$) between the two MetaTracts.
+ In our case, each MetaTract is a data point. We introduce a simple \textit{orientation based similarity measure} to partition the  MetaTracts based on orientation already inherent in the data.
+Given a pair of  MetaTracts, we define the edge weights between two tracts as the cosine of the maximum angle between the local orientations ($N_P$) of all pairs of start points ($C_P$) between the two MetaTracts.
 The edge weights give a distance matrix representing the distance between each pair of points.
 Using the Belkin and Niyogi algorithm we ``embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix.
-%
-%August 13 2014
-%For all our tests the fibers were then embedded in a $\eta$ low dimensional space.
-The number of dimensions in the high dimensional space is equal to the number of datapoints or the number of MetaTracts in our case.
-%For all our test cases we set $\eta$ to be five (see sec~\ref{sec:param_choices} for parameter choices for $\eta$.).
-%  
 We used conventional K-means for clustering this lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
-For our test case there are two major directions of the fiber bundles and $K$ was set to 2.
+For our test case there are two major directions of the fiber bundles ($K=2$).
 Due to the advantages offered by the dimensionality reduction, even if there are curved fiber bundles the user has to provide just the major fiber bundle directions of the weaving pattern.
-Figrue.~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
+Figure.~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes.
 As is expected there is a clear distinction based on fiber orientation.
@@
 %\end{figure}
-
-%\begin{figure}[htp]
-%\centering
-% \includegraphics[width=\linewidth]{imagesMetaTracts/glass-clus-2.eps}
-%  \caption{ Fig  A and C shows the results fiber extraction. B and D shows the results of clustering with clipping planes.}\label{fig:closeup-full}
-%\end{figure}
-
 \section{Limitations and future work}
@@
 
-\added
-{
 \acknowledgements{
-The research leading to these results has received funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM.
+This work was supported by funding from the European Union’s Seventh Framework Programme (FP7/2007-2013) under grant agreement ACP2-GA-2012-314562-QUICOM.
 We especially thank Ilya Straumit (KU Leuven) for providing us with dataset 1.
 }
-}

### 2014-12-11 metaTracts_supplymental_pvis2014.tex

@@
 		Supplementary Material}
-	\author{Online submission ID: \\ 181}
+	\author{Arindam Bhattacharya\thanks{e-mail: arindamb86@gmail.com}\\ %
+		\scriptsize The Ohio State University %
+		\and Christoph Heinzl\thanks{e-mail:christoph.heinzl@fh-wels.at}\\ %
+		\scriptsize University of Applied Sciences - Upper Austria%
+		\and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
+		\scriptsize The Ohio State University
+		\and Artem Amirkhanov\thanks{e-mail:artem.amirkhanov@fh-wels.at}\\ %
+		\scriptsize University of Applied Sciences - Upper Austria%
+		\and Johann Kastner\thanks{e-mail:johann.kastner@fh-wels.at}\\ %
+		\scriptsize University of Applied Sciences - Upper Austria%
+		}
 	\maketitle

### 2014-12-11 Rebuttal.tex
rebuttal update

--- 2014-12-10 Rebuttal.tex
@@
 \maketitle
+We thank the reviewers for their reviews of the submitted manuscript, we have addressed the queries,  re-worked and improved the manuscript. The summary reviews had three issues, 
+\begin{enumerate}
+	   \item  Visualization of more datasets, particularly, the ones, as R1 has
+	   mentioned with crossings at other angles.
+	   \item Effects of choice of length in MetaTract generation.
+	   \item Performance studies and limitations.
+\end{enumerate}
+We incorporated the following changes
+\begin{enumerate}
+	\item {Due to the manufacturing process, a vast majority of composites have bundles crossing at ~90 degree angles.
+With the separation factors being width, spacings and compactness of fiber bundles.
+These variations are well represented in our data sets.
+We maintain that the technique is useful and applicable to a large percentage of composite polymers.
+We have demonstrated the technique on three datasets, two of them are in the paper and the third is in the supplemental section due space  constraints.
+Getting additional datasets is time consuming.
+
+		In dataset 1, Figure~\ref{fig:angle_bundles}f, bundles 7,10 and 6 show  bundles which  though parallel differ in angle and bundle width.It is a common example of variations in the datasets which we extract accurately.
+
+		During additional experiments, we did apply our technique to a different class of fibers where the individual fibers are clearly visible, namely the glass-fibers.
+We have added the results to the supplementary section.
+		\begin{figure}
+			\centering
+			\includegraphics[width=0.7\linewidth]{images_pvis/clustering.pdf}
+			\caption{Fiber bundles extraction}
+			\label{fig:angle_bundles}
+		\end{figure}}
+		\item {
+			As mentioned in the paper, 
+			``The length and the radius parameters for the cylinders of MetaTracts decide how coarse the fiber cylinders are.
+These are dependent on the underlying fiber characteristic and the weaving pattern.
+Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber.
+We used 10.00 and 2.00 for length and radius (measured in grid voxel size), respectively for all tests.
+			Simpler geometry (example dataset 2) was experimentally found to handle larger cylinders better.''
+
+			Reviewer 3 and  4, asked  about the effect of the length threshold ``$\eta$".
+``$\eta$" decides how quickly the hierarchical clustering converges.
+As mentioned in the paper,  experimentally 0.3 to 0.6 gave qualitatively similar results.
+Quantitatively in our  tests $1.2\% - 5\%$ of fibers (total number of fibers $\sim$10000)were removed, while quantitatively small, without removing these clusters are difficult to separate.
+			}
+			\item{
+				We have added a limitation and future work section to the paper.}
+\end{enumerate}
 
@@
 \color{black}
-
-Answer.
-Due to the manufacturing process, a vast majority of composites have bundles crossing at ~90 degree angles.
-With the separation factors being width, spacings and compactness of fiber bundles.
-These variations are well represented in our data sets.
-We maintain that the technique is useful and applicable to a large percentage of composite polymers.
-Nevertheless, we have added a third dataset in the supplementary materials which will be included in a comprehensive technical report which we cannot add currently due to space restrictions.
-
-Figure~\ref{fig:angle_bundles}f, bundles 7,10 and 6 shows a common example of variations in the datasets which we extract accurately. 
-\begin{figure}
-\centering
-\includegraphics[width=0.7\linewidth]{images_pvis/clustering.pdf}
-\caption{Fiber bundles extraction}
-\label{fig:angle_bundles}
-\end{figure}
+Please see summary answer 1, also see supplemental sections.
 
@@
 \color{black}
-Answer.
+We modified the related works section.
 
@@
  Developing methods which can precisely compute the deviations from original bundles remains a future direction of the project.
+We did experiment with glass fibers (see supplemental section) where the individual fibers are clearly separated, and proved as a sanity check that the MetaTract technique does extract correct fibers.
 
@@
 
-Answer. Figure~\ref{fig:glass_fibers} shows an example where each fiber is clearly demarcated compared to the type of data we address in the paper (Figure~\ref{fig:data-char-rebuttal}).
-\begin{figure}
-\centering
-\includegraphics[width=0.7\linewidth]{images_pvis/glass_fibers}
-\caption{High Resolution fiber}
-\label{fig:glass_fibers}
-\end{figure}
+Answer.  Please look at the supplementary section on glass fibers.
 
@@
 \color{black}
+Please see summary answer 2.
 
 \color{red}
-
 What are the limitations of the approach? Are there data sets where the

### 2014-12-11 metaTracts-pvis2014.tex
after merge fixes
rebuttal update

@@
 \usepackage{color}
+\usepackage{gensymb}
+
+%uncomment to hide higlighting the revised text
+\newcommand{\removed}[1]{}
+\newcommand{\added}[1]{{#1}}
+\newcommand{\replaced}[1]{}
+
+%comment to hide higlighting the revised text
+%\newcommand{\removed}[1]{\textcolor{green}{#1}}
+%\newcommand{\added}[1]{\textcolor{red}{#1}}
+%\newcommand{\replaced}[1]{\textcolor{blue}{#1}}
+
+\newcommand{\replacedWith}[2]{\replaced{#1} \added{#2}}
+
 %% We encourage the use of mathptmx for consistent usage of times font
@@
      \scriptsize University of Applied Sciences - Upper Austria%
-\and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
-\scriptsize The Ohio State University
 \and Artem Amirkhanov\thanks{e-mail:artem.amirkhanov@fh-wels.at}\\ %
@@
      \scriptsize University of Applied Sciences - Upper Austria%
+     \and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
+     \scriptsize The Ohio State University
 	    }
@@
 \input {./chapters/MetaTracts.tex}
+
 \bibliographystyle{abbrv}

### 2014-12-11 supplemental/supplementaryMaterial.tex
rebuttal update

@@
 In the voxelized MetaTracts we can clearly see where and how the bundles interact.
+\subsection{Glass fibers}
+Figure~\ref{fig:glass_fibers}a shows an example of glass fibers embedded in an epoxy matrix.
+Unlike the fibers in the other datasets (example Figure~\ref{fig:dataset3}a) the glass fibers are clearly separated from each other Figure~\ref{fig:glass_fibers}a.
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{images_pvis/glass-fibers}
+	\caption{MetaTracts being applied to glass fibers. (a) Scalar data consisting of glass fibers. (b) single slice, (c) Extracted MetaTracts colored mapped to direction.}
+	\label{fig:glass_fibers}
+	\end{figure}

### 2014-12-11 chapters/MetaTracts.tex
after merge fixes
local commit

@@
 The results of clustering the orientation clusters (see Figure~\ref{fig:orientation_clustering}a,c) are shown in Figure~\ref{fig:orientation_clustering}d,e respectively.
-
 \subsection{Choice of clustering techniques}

### 2014-12-11 chapters/MetaTracts.tex
minor changes

@@
 They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.
-In order to determine reliable Hessians, we compute a similar metric.
+In order to determine reliable Hessians, we compute the same metric.
 We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details.
@@
 
-The next set of tasks evaluated the spatial context perceived from MetaTracts.\replacedWith{ We achieve this by asking the evaluators}{The participants were asked} to select the fiber bundle which is largest, smallest, or the most similar to a given one from a \removed{sub}set of fiber bundles.
+The next set of tasks evaluated the spatial context perceived from MetaTracts.\replacedWith{ We achieve this by asking the evaluators}{The participants were asked} to select the fiber bundle which was largest, smallest, or the most similar to a given one from a \removed{sub}set of fiber bundles.
 Due to the problems of occlusions and similar grayscale values these tests proven to be more difficult to perform successfully with volume rendering of the plain data, compared to MetaTracts.
@@
 The following parameters were fixed for all the tests.
-We set the reliable Hessian threshold $R_{H}$ to be 0.3 for all our tests.
-A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.
-\replacedWith{On the other hand, a}{A} very high $ R_{H}$ would lead to a decline in number of MetaTracts produced.
+We set the reliable Hessian threshold $R_{H}$ to be 0.3.
+A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.\replacedWith{On the other hand, a}{A} very high $ R_{H}$ would lead to a decline in number of MetaTracts produced.
 The $\alpha$ and $\beta$ in $R_{H}$ are as explained in Frangi et al.
@@
 We used 10.00 and 2.00 for length and radius (measured in grid voxel size), respectively for all tests.
-Simpler geometry (\replacedWith{example dataset 2}{e.g., D2}) was experimentally found to handle larger cylinders better.
-$\eta$ in sec.~\ref{subsec:dist_clustering} decides how quickly the hierarchical clustering converges, experimentally 0.3 to 0.6 gave similar results.
+Simpler geometry (\replacedWith{example dataset 2}{e.g., D2} ) was experimentally found to handle larger cylinders better.
+$\eta$ in sec.~\ref{subsec:dist_clustering} decides how quickly the hierarchical clustering converges, experimentally values 0.3 to 0.6 removed $1.2\% - 5\%$ of fibers (total number of fibers $\sim$10000) and gave similar results.
  %Short tracts of less than 0.01 times the smallest axes size are prone to have erroneous local orientation and are therefore removed. 
@@
 %\end{figure}
-\section{Limitations and future work}
-\label{sec:conclusions}
+\subsection{Limitations}
+\label{sec:limitations}
 \added{
 A key assumption of the method is ``connectivity'' (Sec.~\ref{sec:char_data}).
-If the ``connectivity'' is not satisfied due to noise in the image, the then MetaTracts will be inaccurate. 
+If the ``connectivity'' is not satisfied due to noise in the image, then the  generated MetaTracts will be inaccurate. 
 The clustering process also assumes that the fiber bundles have a minimum width.
@@
 }
+
+\section{Conclusions and Future Work}
+\label{sec:conclusions}
+In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials.
+We show that our framework works at comparatively low resolution and with dense fiber arrangements (when extracting single fibers might not be possible).
+It handles complex fiber patterns such as ``cross overs'' and ``braiding''.
 In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.

### 2014-12-11 Rebuttal.tex
Performance

@@
 			\item{
-				We have added a limitation and future work section to the paper.}
+				We have added a limitations section to the paper.
+We provide the approximate times of MetaTracts computation of our datasets in the Experimental Results section of the paper.
+Unfortunately the full scale performance study was not possible due to the time constraints of the minor revision.}
 \end{enumerate}

### 2014-12-11 Rebuttal.tex
rebuttal revision (wip)

--- 2014-12-11 Rebuttal.tex
@@
 \maketitle
-We thank the reviewers for their reviews of the submitted manuscript, we have addressed the queries,  re-worked and improved the manuscript. The summary reviews had three issues, 
+We thank the reviewers for their reviews of the submitted manuscript, in order to address the raised queries we carefully re-worked and improved the manuscript, improved some figures, included additional explanations, and extended supplementary material.
+This document describes a summary of the introduced improvements.
+The summary review mentioned three main issues:
 \begin{enumerate}
@@
 \end{enumerate}
-We incorporated the following changes
+We fully dedicated ourselves to address these comments and incorporated the following changes:
 \begin{enumerate}
 	\item {Due to the manufacturing process, a vast majority of composites have bundles crossing at ~90 degree angles.
-With the separation factors being width, spacings and compactness of fiber bundles.
-These variations are well represented in our data sets.
-We maintain that the technique is useful and applicable to a large percentage of composite polymers.
-We have demonstrated the technique on three datasets, two of them are in the paper and the third is in the supplemental section due space  constraints.
-Getting additional datasets is time consuming.
+With the separation factors being widths, spacings and compactness of fiber bundles.
+These variations are well represented in our data sets: total of three in the original submission.
+We present evaluation of MetaTracts for datasets D1 and D2 in the main manuscript.
+Additionally applicability for dataset D3 and glass fiber reinforced dataset is presented in the supplimentary material due to space constraints.
+Therefore we argue that the technique is useful and applicable to a range of various composite polymers.
+Providing additional datasets is unfortunately not possible given the time constraints of the minor revision.
 
-		In dataset 1, Figure~\ref{fig:angle_bundles}f, bundles 7,10 and 6 show  bundles which  though parallel differ in angle and bundle width.It is a common example of variations in the datasets which we extract accurately.
+		In dataset 1, Figure~\ref{fig:angle_bundles}f, bundles 7,10 and 6 show  bundles which  though parallel differ in angle and bundle width.
+It is a common example of variations in the datasets which we extract accurately.
 
@@
 		\item {
-			As mentioned in the paper, 
+			We clarified the effects of choice of length and radius of MetaTracts cylinders in the paper: 
 			``The length and the radius parameters for the cylinders of MetaTracts decide how coarse the fiber cylinders are.
@@
 			Reviewer 3 and  4, asked  about the effect of the length threshold ``$\eta$".
-``$\eta$" decides how quickly the hierarchical clustering converges.
-As mentioned in the paper,  experimentally 0.3 to 0.6 gave qualitatively similar results.
+``$\eta$" on how quickly the hierarchical clustering converges.
+We mention in the manuscript that our experimental studies concluded that the length threshold choice of 0.3 to 0.6 gave qualitatively similar results.
 Quantitatively in our  tests $1.2\% - 5\%$ of fibers (total number of fibers $\sim$10000)were removed, while quantitatively small, without removing these clusters are difficult to separate.
+These experimental results were included in the paper.
 			}
 			\item{
-				We have added a limitation and future work section to the paper.}
+				We have added a limitations section to the paper.
+We also detailed the approximate times of MetaTracts computation for our datasets in the Experimental Results section of the paper.
+Unfortunately the full scale performance study was not possible due to the time constraints of the minor revision.
+Additional difficulty of obtaining the detailed performance numbers is that it would require the access to workstations which are actively used and regularly occupied by other researchers and developers.
+Therefore only the reasonable round-ups of the performance and computation run-times are presented.
+				}
 \end{enumerate}

### 2014-12-11 Rebuttal.tex
rebuttal reviewed

@@
 \usepackage{color}
+\usepackage[parfill]{parskip}
+
 %opening
@@
 \begin{document}
+\noindent
 
@@
 The summary review mentioned three main issues:
+\textbf{\textit{
 \begin{enumerate}
@@
 \end{enumerate}
+}}
 We fully dedicated ourselves to address these comments and incorporated the following changes:
@@
 
+ The remainder of this document addresses comments and remarks raised by individual reviewers.
+
 \section{Reviewer 1}
-\color{red}
+
+\textbf{\textit{
 Q1.
@@
 The authors are encouraged to state clearly the originality and contribution of $R_{H}$.
-\color{black}
+}}
 
@@
 Reliable Hessians is computed exactly equivalently to Vesselness for a single scale (Firangi et al.~\cite{Frangi1998}).
-We have  updated the corresponding sentence in the paper to reflect the same.
+We have  updated the corresponding sentence in the paper to reflect this information.
 We explain further about the utility being different.
@@
 We use this orientation vector to create MetaTracts.
+\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
 
@@
 
-\color{red}
+\textbf{\textit{
 Q2.
@@
 The authors are encouraged to provide more than one fiber configurations.
-\color{black}
-Please see summary answer 1, also see supplemental sections.
+}}
 
-\color{red}
+A2. Please see answer 1 to the summary review and the extended supplementary material.
+
+\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
+
+\textbf{\textit{
 Q3. All clustering was done in R. The letter R is not explained.
-\color{black}
+}}
 
-A3. R is a free software environment for statistical computing and graphics.
+A3. R is a free software environment for statistical computing and graphics. This explanation is now added to the paper.
+
+\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
 
 \section{Reviewer 2}
-\color{red}
+
+\textbf{\textit{
+Q1.
 I feel the background section could be improved.
@@
 The paper gives some relatively high-level description of this but more background would be appreciated.
+}}
 
-\color{black}
-We modified the related works section.
+A1. We modified the related works section.
 
-\color{red}
-  It seems to me that the extracted fibers may not be real fibers. Is this
+\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
+
+\textbf{\textit{
+Q2. It seems to me that the extracted fibers may not be real fibers. Is this
 true? If so, how well can one trust the weaving patterns extracted based
 on the imaginary fibers?
-  \color{black}
-  Answer. 
+}}
+
+A2.
 Yes, the extracted fibers cannot be mapped to real fibers.
-One of the goals  we  focused on was to move away from extraction of individual fibers, instead work on direct reconstruction of bundles.
-The design decisions adheres to extracting reliable bundles directly from the images.
+One of our main goals is to move away from extraction of individual fibers and work instead on direct reconstruction of bundles.
+These design decisions adheres to extracting reliable bundles directly from the images.
 Developing methods which can precisely compute the deviations from original bundles remains a future direction of the project.
@@
 
-\color{red}
-The authors claim that their method can deal with low resolution data.
+\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
+
+\textbf{\textit{
+Q3. The authors claim that their method can deal with low resolution data.
 But this might be in the eyes of the beholders, since in my opinion the example data in the paper all have fairly high resolutions.
 Perhaps a comparison with a really high resolution data using this method and other existing methods could diffuse the doubt.
-\color{black}
+}}
 
-Answer.  Please look at the supplementary section on glass fibers.
+A3.
+In the context of this work by low resolution we intend situations when resolving individual fibers is not possible.
+We state this in the paper.
+Please look at the supplementary section on glass fibers as an example of high-resolution data.
+
+\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
 
 \section{Reviewer 4}
-\color{red}
-In Section 5.1, k-means is applied to the m-dimensional space using
+
+\textbf{\textit{
+Q1. In Section 5.1, k-means is applied to the m-dimensional space using
 Euclidean distance, i.e., each dimension is considered equally important.
 Is that really the case when they represent eigenvectors?
+}}
 
-\color{black}
-Answer.
+A1.
 Belkin  and Niyogi's~\cite{Belkin01} laplacian eigenmaps embeds the high dimensional data ( represented as a graph ) into a low dimensional space where the Euclidean distance approximates the distances in higher dimension space.
 For such reason, Eigen embedding for dimensionality reduction has become quiet popular and has found favor in tackling a variety of problems.
-Empirically the K-means for orientation clustering provided us excellent results.
+Empirically the K-means for orientation clustering provided us with excellent results.
 The major orientations are limited by the inherent characteristics of our data set and this also plays a role in Laplacian embedding followed by k-means giving us such good results.
-A formal proof is not provided.
+A formal proof is not provided due to the strict space constraints.
 
-\color{red}
-In Section 5.2, the authors mention that they remove short Meta Tracks.
+\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
+
+\textbf{\textit{
+Q2. In Section 5.2, the authors mention that they remove short Meta Tracks.
 How much data is thrown away here? Can examples be provided to get a
 better understanding?
-\color{black}
-Please see summary answer 2.
+}}
 
-\color{red}
-What are the limitations of the approach? Are there data sets where the
+A2. Please see the answer 2 to the summary review.
+
+\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
+
+\textbf{\textit{
+Q3. What are the limitations of the approach? Are there data sets where the
 method fails?
-\color{black}
-Answer. We have added a ``Limitations and future work" Section to the end of the paper. We copy it here for completeness.
+}
+}
+
+A3. We have added a ``Limitations and future work" Section to the end of the paper. We cite it here for completeness:
 

### 2014-12-11 Rebuttal.tex
removed section numbering
minor ruler fix
bold answer indices

@@
 
-\section{Reviewer 1}
+\section*{Reviewer 1}
 
@@
 
-A1.
-Reliable Hessians is computed exactly equivalently to Vesselness for a single scale (Firangi et al.~\cite{Frangi1998}).
+\textbf{A1.} Reliable Hessians is computed exactly equivalently to Vesselness for a single scale (Firangi et al.~\cite{Frangi1998}).
 We have  updated the corresponding sentence in the paper to reflect this information.
@@
 
-A2. Please see answer 1 to the summary review and the extended supplementary material.
+\textbf{A2.} Please see answer 1 to the summary review and the extended supplementary material.
 
@@
 
-A3. R is a free software environment for statistical computing and graphics. This explanation is now added to the paper.
+\textbf{A3.} R is a free software environment for statistical computing and graphics. This explanation is now added to the paper.
 
-\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
-
-\section{Reviewer 2}
+\section*{Reviewer 2}
 
@@
 
-A1. We modified the related works section.
+\textbf{A1.} We modified the related works section.
 
@@
 
-A2.
-Yes, the extracted fibers cannot be mapped to real fibers.
+\textbf{A2.} Yes, the extracted fibers cannot be mapped to real fibers.
 One of our main goals is to move away from extraction of individual fibers and work instead on direct reconstruction of bundles.
@@
 
-A3.
-In the context of this work by low resolution we intend situations when resolving individual fibers is not possible.
+\textbf{A3.} In the context of this work by low resolution we intend situations when resolving individual fibers is not possible.
 We state this in the paper.
@@
 
-\makebox[\linewidth]{\rule{0.25\textwidth}{0.4pt}}
-
-\section{Reviewer 4}
+\section*{Reviewer 4}
 
@@
 
-A1.
-Belkin  and Niyogi's~\cite{Belkin01} laplacian eigenmaps embeds the high dimensional data ( represented as a graph ) into a low dimensional space where the Euclidean distance approximates the distances in higher dimension space.
+\textbf{A1.} Belkin  and Niyogi's~\cite{Belkin01} laplacian eigenmaps embeds the high dimensional data ( represented as a graph ) into a low dimensional space where the Euclidean distance approximates the distances in higher dimension space.
 For such reason, Eigen embedding for dimensionality reduction has become quiet popular and has found favor in tackling a variety of problems.
@@
 
-A2. Please see the answer 2 to the summary review.
+\textbf{A2.} Please see the answer 2 to the summary review.
 
@@
 
-A3. We have added a ``Limitations and future work" Section to the end of the paper. We cite it here for completeness:
+\textbf{A3.} We have added a ``Limitations and future work" Section to the end of the paper. We cite it here for completeness:
 

### 2014-12-11 chapters/MetaTracts.tex
rebuttal reviewed
spell check pass

@@
 \item{ Which fiber bundles show a particular orientation? }
-\item{ Which fiber bundles are of the same type of yarn\added{?} (i.e. which bundles show similar sizes or diameters, which is the largest or smallest fiber bundle?\added{)}}
+\item{ Which fiber bundles are of the same type of yarn? I.e. which bundles show similar sizes or diameters, which is the largest or smallest fiber bundle\added{?}}
 \end{itemize}
@@
 \replacedWith{For example, spatial context}{It provides} answers \added{to} questions such as: Are these bundles in contact at a particular position in the data?
-What is the relative orientations of the contacting bundles?
+What are the relative orientations of the contacting bundles?
 \replacedWith{Answering}{Providing answers to} the above queries from the volume renderings of XCT datasets or from visual inspection of particular 2D slices is non-trivial, even for experts.
@@
 Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}.
-Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers \replacedWith{that}{which} are close for certain distance and then diverge, are clustered separately.
+Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the closest distances mean, so that fibers \replacedWith{that}{which} are close for certain distance and then diverge, are clustered separately.
 Brun et al.~\cite{Brun2004} use\removed{s} normalized cuts along with a pairwise distances measure computed using a 9\removed{-}D fiber shape descriptors.
@@
 \subsection {Visual Analysis and Modeling of Fiber Reinforced Composites}
-The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fiber extraction from high resolution data where the individual fibers are clearly discernible.
-Fritz et al.~\cite{Fritz2009} proposed interactive workflows%for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
+The approaches presented in visualization and analysis of composites mainly focuses on individual objects such as fiber extraction from high resolution data where the individual fibers are clearly discernible.
+Fritz et al.~\cite{Fritz2009} proposed interactive workflows.
+%for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete.
 %This approach allows analyzing fiber orientations based on direction transfer functions.
-Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites.%They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
+Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites.
+%They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
 Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
@@
 The Hessian matrix represents the local curvature.
-The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
+The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.
+The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest.
 This direction coincides with the direction of the tubular structure.
@@
 Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix.
-Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$).
+Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign ($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$).
 The sign indicates if the vessel is bright in a dark background or dark in a bright background.
@@
 The utility of the vesselness is a little different in our framework than Frangi et al.~\cite{Frangi1998}.
-First, vesselness in biology is computed for different scales because the vessels can be of different sizes. In our case, usually the width of individual fibers are known a priori.
+First, vesselness in biology is computed for different scales because the vessels can be of different sizes. In our case, usually the widths of individual fibers are known a priori.
 Second, and more importantly, we do not have clear tubular structures embedded in a dark contrast matrix such as in blood vessels.
@@
 Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest.
-Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable.
-The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.
+Note, the dark regions within bundles are regions where the $R_H$ is less than the threshold and are unreliable.
+The bundles are not uniformly colored and the Hessians and the corresponding directions are noisy.
 
@@
 This is not the case for our data.
-The local orientation at each voxel is inherently more noisy.}
+The local orientations at each voxel are inherently noisier.}
 
@@
 We do this by interpreting the underlying geometric structure of the fibers as a set of \removed{polylines(cylinders)} cylinders.
- MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data.
+ MetaTracts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data.
 MetaTracts share the following \textit{properties}:
@@
 \begin{enumerate}[noitemsep]
-\item{\replacedWith{Each MetaTract is}{MetaTract are} associated with a continuous set of cylinders.}
-\item{\replacedWith{Each MetaTract is}{MetaTract are} associated with a start point at a grid vertex.}
-\item{Individual cylinders in a MetaTract have a constant length, a constant radius and a start point (which is also a grid vertex).}
-\item{Individual cylinders in a MetaTract (except the first one) are connected to their previous cylinders at their start point.}
+\item{\replacedWith{Each MetaTract is}{MetaTracts are} associated with a continuous set of cylinders.}
+\item{\replacedWith{Each MetaTract is}{MetaTracts are} associated with a start point at a grid vertex.}
+\item{Individual cylinders in MetaTracts have constant lengths, radii, and start points (which are also grid's vertices).}
+\item{Individual cylinders in MetaTracts (except the first one) are connected to their previous cylinders at their start point.}
 \item{Individual cylinders are parallel to the local orientation at their start points.}
@@
 \end{figure}
-The set of vertices which are in the green region but not overlapped by blue are possible \textit{``candidate vertices"} for the start point of the next cylinder in a particular MetaTract.
+The set of vertices which are located in the green region but not in the blue one are potential \textit{``candidate vertices"} for the start point of the next cylinder in a MetaTract.
 From these we select another grid vertex which will be the start point for the next cylinder.
@@
 \added{So} instead of creating a heuristic \replacedWith{which}{to} artificially combine the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
-The two step clustering also made the problem more tractable and the parameters intuitive for the end users.
+The two-step clustering also made the problem more tractable and the parameters intuitive for the end users.
 For the orientation clustering, we use dimension reduction followed by K-means clustering.
@@
 To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
-shortest tracts (tracts less than $\eta$ times the median length, set to 0.6) in each bundle. We then reapplied hierarchical clustering.
+shortest tracts (length less than $\eta$ times the median length, set to 0.6) in each bundle. We then reapplied hierarchical clustering.
 We repeated both until a steady state of clusters is reached and no new small fibers can be removed.
@@
 Figure~\ref{fig:prepreg} shows the results on our dataset 2 \added{(D2)}.
-This dataset has different characteristics, and is dense with \textit{flat and thin} bundles, it consists of 4 bundles along one direction and 5 along the other.
-The dimensions are 300$\times$350$\times$300 and the datatype is uint16 (\replacedWith{dataset 1}{D1} is uint8).
+This dataset has different characteristics, and is dense with \textit{flat and thin} bundles; it consists of 4 bundles along one direction and 5 along the other.
+The dimensions are 300$\times$350$\times$300 and the data type is uint16 (\replacedWith{dataset 1}{D1} is uint8).
 The bundles are indistinguishable in the original data set Figure~\ref{fig:prepreg}a.
@@
 \removed{The fiber bundle extraction using }MetaTracts \replacedWith{was}{are} implemented in C++ using ITK.
-\removed{All }Clustering \replacedWith{was}{is} done in R.
+\removed{All }Clustering \replacedWith{was}{is} done in R \added{(free software environment for statistical computing and graphics)}.
 The preprocessing is time consuming and may take several hours.
@@
 All participants could correctly identify the number of fiber bundles\replacedWith{.
-All participants}{and} agreed that MetaTracts \replacedWith{was}{are} more effective in perceiving geometric structure than simple volume rendering or Hessian color map.
+All participants}{ and} agreed that MetaTracts \replacedWith{was}{are} more effective in perceiving geometric structure than simple volume rendering or Hessian color map.
 They especially liked the \replacedWith{two dimensional}{2D} slices of the MetaTracts (Figure~\ref{fig:crop-16-decomp}b, compared to Figure~\ref{fig:data-char}c).
 
-The next set of tasks evaluated the spatial context perceived from MetaTracts.\replacedWith{ We achieve this by asking the evaluators}{The participants were asked} to select the fiber bundle which was largest, smallest, or the most similar to a given one from a \removed{sub}set of fiber bundles.
+The next set of tasks evaluated the spatial context perceived from MetaTracts.\replacedWith{ We achieve this by asking the evaluators}{ The participants were asked} to select the fiber bundle which was largest, smallest, or the most similar to a given one from a \removed{sub}set of fiber bundles.
 Due to the problems of occlusions and similar grayscale values these tests proven to be more difficult to perform successfully with volume rendering of the plain data, compared to MetaTracts.
@@
 We set the reliable Hessian threshold $R_{H}$ to be 0.3.
-A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.\replacedWith{On the other hand, a}{A} very high $ R_{H}$ would lead to a decline in number of MetaTracts produced.
-The $\alpha$ and $\beta$ in $R_{H}$ are as explained in Frangi et al.
+A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.\replacedWith{On the other hand, a}{ A} very high $ R_{H}$ would lead to a decline in number of MetaTracts produced.
+Coefficients $\alpha$ and $\beta$ in $R_{H}$ are as explained in Frangi et al.
 \cite{Frangi1998} and set to 0.5.
@@
  $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays; we have used [7-10] and half the length of cylinder (5.0).
-Our number of fiber bundle directions are limited.
+Our number of fiber bundle directions is limited.
 Thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well.

### 2014-12-11 metaTracts-pvis2014.tex
minor changes
spell check pass

@@
 
-\newcommand{\replacedWith}[2]{\replaced{#1} \added{#2}}
+\newcommand{\replacedWith}[2]{\replaced{#1}\added{#2}}
 
@@
 First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
-In the second step, these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach: first by orientation, then by proximity.
+In the second step, these extracted fiber bundle sections (MetaTracts) are clustered using a two-step approach: first by orientation, then by proximity.
 The tool can generate volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
